### 2020-05-18 analyses/07-ANOVA/table_posthoc.tex
full upload

\begin{tabular}{lllllrrl}
\toprule
nesting & variable & test & contrast1 & contrast2 & statistic & pvalue & signif\\
\midrule
Abaco & PC2 & tukey & coastal & coppice & -0.71 & 0.8690 & \\
Abaco & PC2 & tukey & coastal & mangrove & -8.28 & 0.0000 & ***\\
Abaco & PC2 & tukey & coppice & mangrove & -6.82 & 0.0000 & ***\\
Abaco & PC3 & tukey & coastal & coppice & 4.89 & 0.0025 & **\\
Abaco & PC3 & tukey & coastal & mangrove & 7.52 & 0.0000 & ***\\
\addlinespace
Abaco & PC3 & tukey & coppice & mangrove & 2.55 & 0.1746 & \\
Bimini & PC1 & tukey & coastal & coppice & -2.57 & 0.1726 & \\
Bimini & PC1 & tukey & coastal & mangrove & -3.49 & 0.0428 & *\\
Bimini & PC1 & tukey & coppice & mangrove & -0.70 & 0.8741 & \\
Bimini & PC2 & tukey & coastal & coppice & 2.10 & 0.3044 & \\
\addlinespace
Bimini & PC2 & tukey & coastal & mangrove & 3.91 & 0.0200 & *\\
Bimini & PC2 & tukey & coppice & mangrove & 1.44 & 0.5674 & \\
Bimini & PC3 & dunnett & coastal & coppice & 3.59 & 0.0067 & **\\
Bimini & PC3 & dunnett & coastal & mangrove & -0.21 & 0.9954 & \\
Bimini & PC3 & dunnett & coppice & mangrove & -3.03 & 0.0162 & *\\
\addlinespace
Cayman Brac & PC1 & dunnett & coastal & coppice & -2.93 & 0.0225 & *\\
Cayman Brac & PC1 & dunnett & coastal & mangrove & -0.74 & 0.8436 & \\
Cayman Brac & PC1 & dunnett & coppice & mangrove & 3.60 & 0.0032 & **\\
Cayman Brac & PC2 & tukey & coastal & coppice & 2.28 & 0.2491 & \\
Cayman Brac & PC2 & tukey & coastal & mangrove & 4.15 & 0.0140 & *\\
\addlinespace
Cayman Brac & PC2 & tukey & coppice & mangrove & 1.98 & 0.3476 & \\
Cayman Brac & PC3 & tukey & coastal & coppice & 3.46 & 0.0467 & *\\
Cayman Brac & PC3 & tukey & coastal & mangrove & 8.16 & 0.0000 & ***\\
Cayman Brac & PC3 & tukey & coppice & mangrove & 4.97 & 0.0028 & **\\
Little Cayman & PC1 & tukey & coastal & coppice & -4.08 & 0.0167 & *\\
\addlinespace
Little Cayman & PC1 & tukey & coastal & mangrove & -1.51 & 0.5389 & \\
Little Cayman & PC1 & tukey & coppice & mangrove & 2.65 & 0.1593 & \\
Little Cayman & PC2 & dunnett & coastal & coppice & 0.88 & 0.7637 & \\
Little Cayman & PC2 & dunnett & coastal & mangrove & -4.02 & 0.0014 & **\\
Little Cayman & PC2 & dunnett & coppice & mangrove & -6.99 & 0.0000 & ***\\
\addlinespace
Little Cayman & PC3 & tukey & coastal & coppice & -4.28 & 0.0114 & *\\
Little Cayman & PC3 & tukey & coastal & mangrove & -7.03 & 0.0000 & ***\\
Little Cayman & PC3 & tukey & coppice & mangrove & -2.18 & 0.2822 & \\
Long Island & PC3 & tukey & coastal & coppice & 3.75 & 0.0285 & *\\
Long Island & PC3 & tukey & coastal & mangrove & 4.19 & 0.0127 & *\\
\addlinespace
Long Island & PC3 & tukey & coppice & mangrove & 0.47 & 0.9408 & \\
Long Island & PC4 & tukey & coastal & coppice & 0.29 & 0.9766 & \\
Long Island & PC4 & tukey & coastal & mangrove & 3.47 & 0.0454 & *\\
Long Island & PC4 & tukey & coppice & mangrove & 2.81 & 0.1257 & \\
\bottomrule
\end{tabular}

### 2020-05-18 analyses/07-ANOVA/table_postkw.tex
full upload

\begin{tabular}{lllllrrl}
\toprule
island & variable & test & contrast1 & contrast2 & statistic & pvalue & signif\\
\midrule
Abaco & PC2 & nemenyi & coastal & coppice & 0.43 & 0.9511 & \\
Abaco & PC2 & nemenyi & coastal & mangrove & 6.50 & 0.0000 & ***\\
Abaco & PC2 & nemenyi & coppice & mangrove & 5.47 & 0.0003 & ***\\
Bimini & PC1 & nemenyi & coastal & coppice & 2.62 & 0.1523 & \\
Bimini & PC1 & nemenyi & coastal & mangrove & 3.42 & 0.0411 & *\\
\addlinespace
Bimini & PC1 & nemenyi & coppice & mangrove & 0.60 & 0.9048 & \\
Bimini & PC3 & nemenyi & coastal & coppice & 4.97 & 0.0013 & **\\
Bimini & PC3 & nemenyi & coastal & mangrove & 0.92 & 0.7924 & \\
Bimini & PC3 & nemenyi & coppice & mangrove & 4.93 & 0.0014 & **\\
Little Cayman & PC3 & nemenyi & coastal & coppice & 3.73 & 0.0227 & *\\
\addlinespace
Little Cayman & PC3 & nemenyi & coastal & mangrove & 6.23 & 0.0000 & ***\\
Little Cayman & PC3 & nemenyi & coppice & mangrove & 2.00 & 0.3345 & \\
Long Island & PC1 & nemenyi & coastal & coppice & 4.31 & 0.0065 & **\\
Long Island & PC1 & nemenyi & coastal & mangrove & 0.26 & 0.9816 & \\
Long Island & PC1 & nemenyi & coppice & mangrove & 3.94 & 0.0148 & *\\
\bottomrule
\end{tabular}

### 2020-05-18 R/save_table.R

@@ -1,0 +1,13 @@
+#' Save tables for this study
+
+save_table <- function(tab, nm, digits = 0) {
+
+  library(knitr)
+
+  write.csv(tab, paste0(nm, ".csv"), row.names = FALSE)
+  tab <- knitr::kable(tab, "latex", digits = digits, booktabs = TRUE)
+  texfile <- file(paste0(nm, ".tex"))
+  writeLines(tab, texfile)
+  close(texfile)
+
+}

### 2020-05-18 analyses/03-PCA/PCA.R

@@ -1,0 +1,107 @@
+rm(list = ls())
+
+# Exploration of the principal component analysis
+
+library(nmgc)
+library(tidyverse)
+library(knitr)
+
+# Here we are checking first the correlation between PC1 and brightness
+# and second, the variance explained by each PC on each island
+
+data <- read.csv("data/reflectance.csv", header = TRUE)
+
+# Perform PCA on each island
+pca <- npcomp(data, paste0("wl", 300:700), nesting = "island", combine = TRUE, reduce = 1:4)
+
+# And on the whole archipelago
+pca_pooled <- npcomp(data, paste0("wl", 300:700))
+
+####  1. Correlation between brightness and PC1 ####
+
+# Calculate brightness (mean reflectance)
+data$brightness <- data %>% select_at(paste0("wl", 300:700)) %>% apply(1, mean)
+
+# Attach PC1
+data <- data %>% select_at(c("island", "habitat", "brightness"))
+data$PC1 <- pca$x$PC1
+data$PC1pooled <- pca_pooled$x[, 1]
+data <- data %>%
+  gather_("variable", "PC1", c("PC1", "PC1pooled")) %>%
+  mutate(island = ifelse(variable == "PC1pooled", "Archipelago", as.character(island))) %>%
+  dplyr::select(-variable)
+
+# Measure correlation
+res_cor <- data %>%
+  split(f = .[["island"]]) %>%
+  map_dfr(
+    ~ cor.test(.x$PC1, .x$brightness) %>%
+      .[c("estimate", "p.value")] %>%
+      data.frame %>%
+      rename(r2 = "estimate", pvalue = "p.value") %>%
+      mutate(r2 = r2^2),
+    .id = "island"
+  )
+
+# Reorder
+res_cor <- res_cor %>%
+  filter(island != "Archipelago") %>%
+  rbind(., res_cor %>% filter(island == "Archipelago"))
+
+# Save table
+t1 <- res_cor
+t1 <- latex %>%
+  add_signif() %>%
+  mutate(pvalue = ifelse(pvalue < 0.0001, "< 0.0001", pvalue))
+colnames(t1)[colnames(t1) == "signif"] <- ""
+save_table(t1, "table_brightness", digits = c(0, 3, 0))
+
+
+# Plot the correlation
+p <- data %>%
+  filter(island != "Archipelago") %>%
[…]

### 2020-05-18 analyses/04-machine learning/MachineLearning.R

@@ -1,0 +1,319 @@
+rm(list = ls())
+
+library(nmgc)
+library(tidyverse)
+library(cowplot)
+library(knitr)
+
+data <- read.csv("data/reflectance.csv", header = TRUE)
+
+#### Analysis within each island ####
+
+# Principal components
+
+variables <- paste0("PC", 1:4)
+
+# SVM-classification on PCs
+
+res <- classify(
+  data, variables, grouping = "habitat", nesting = "island",
+  to_pcomp = paste0("wl", 300:700), digest = TRUE, test = TRUE, k = 5,
+  nrep = 100, seed = 24, method = "SVM", importance = TRUE
+)
+
+# Prepare a confusion matrix plot to show on the side
+
+conf <- res$avg$`Long Island`
+conf <- conf %>% apply(2, function(x) x / sum(x))
+conf <- conf %>%
+  data.frame %>%
+  rownames_to_column("predicted") %>%
+  gather_("true", "freq", colnames(conf))
+colnames(conf) <- c("predicted", "true", "freq")
+
+confplot <- ggplot(conf, aes(x = true, y = predicted, fill = freq)) +
+  geom_tile() +
+  scale_fill_gradient(low = "white", high = "darkgreen", limits = c(0, 1)) +
+  theme_bw() +
+  labs(x = "True habitat", y = "Predicted habitat", fill = "Frequency") +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1))
+
+blank <- ggplot() + theme_void()
+
+p <- plot_classif(res, facets = "island")
+plot_grid(p, plot_grid(blank, confplot, blank, nrow = 3), ncol = 2, rel_widths = c(2, 1))
+ggsave("classif_svm_pca.png", height = 6, width = 9, dpi = 300)
+
+write.csv(res$mean, "table_classif_svm_pca.csv", row.names = FALSE)
+
+# Latex table
+latex <- res$mean
+head(latex)
+latex <- latex %>%
+  mutate(signif = ifelse(pvalue < 0.05, "*", "")) %>%
+  mutate(signif = ifelse(pvalue < 0.01, "**", signif))
+colnames(latex)[colnames(latex) == "signif"] <- ""
+latex <- kable(latex, digits = c(0, 3, 0, 1, 0, 4), format = "latex")
+texfile <- file("table_classif_svm_pca.tex")
+writeLines(latex, texfile)
+close(texfile)
+
+res$imp %>%
[…]

### 2020-05-18 analyses/05-assumptions/Assumptions.R

@@ -1,0 +1,77 @@
+rm(list = ls())
+
+# Here we test for the assumptions of (M)ANOVA across different levels
+
+library(nmgc)
+library(dewlap)
+
+data <- read.csv("data/reflectance.csv", header = TRUE)
+wl <- paste0("wl", 300:700)
+variables <- paste0("PC", 1:4)
+
+#### 1. Pooled data -- across the whole archipelago ####
+
+data <- cbind(data, data.frame(npcomp(data, wl)$x)[, variables])
+
+# 1.1. Multivariate normality
+t1 <- test_multinorm(data, variables, grouping = "habitat")[, -1]
+
+# 1.2. Heterogeneity of covariance matrices
+t2 <- test_covariance(data, variables, grouping = "habitat")[, -1]
+
+# 1.3. Univariate normality
+t3 <- test_multinorm(data, variables, grouping = "habitat", univariate = TRUE)[, -1]
+
+# 1.4. Heterogeneity of variances
+t4 <- test_covariance(data, variables, grouping = "habitat", univariate = TRUE)[, -1]
+
+save_table(t1, "table_multinorm_pooled", digits = c(0, 2, 0, 0))
+save_table(t2, "table_covariance_pooled", digits = c(1, 0, 4, 0))
+save_table(t3, "table_normality_pooled", digits = c(0, 0, 3, 4, 0))
+save_table(t4, "table_variance_pooled", digits = c(0, 2, 0, 4))
+
+#### 2. Within islands ####
+
+# 2.1. Multivariate normality
+t5 <- test_multinorm(data, variables, grouping = "habitat", nesting = "island")
+
+# 2.2. Homogeneity of covariance matrices
+t6 <- test_covariance(data, variables, grouping = "habitat", nesting = "island")
+
+save_table(t5, "table_multinorm", digits = c(0, 0, 2, 4, 0))
+save_table(t6, "table_covariance", digits = c(0, 1, 0, 4, 0))
+
+# 2.3. Univariate normality
+t7 <- test_multinorm(data, variables, grouping = "habitat", nesting = "island", univariate = TRUE)
+
+# Reduce to the islands with evidence for deviations from multivariate normality
+t7 <- t7 %>%
+  group_by(island, habitat) %>%
+  nest() %>%
+  right_join(t5) %>%
+  filter(pvalue < 0.05) %>%
+  select(island, habitat, data) %>%
+  unnest(cols = c(data))
+
[…]

### 2020-05-18 analyses/06-pooled/Pooled.R

@@ -1,0 +1,159 @@
+rm(list = ls())
+
+# Here we explore dewlap coloration between habitats across the whole archipelago
+
+library(nmgc)
+library(GGally)
+library(MANOVA.RM)
+library(heplots)
+library(cowplot)
+
+data <- read.csv("data/reflectance.csv", header = TRUE)
+wl <- paste0("wl", 300:700)
+variables <- paste0("PC", 1:4)
+
+data <- cbind(data, data.frame(npcomp(data, wl)$x)[, variables])
+data <- cbind(data, pca)
+
+#### 1. Eyeball ####
+
+# We plot the color variables across habitats at the scale of the whole archipelago
+
+colors <- c(coastal = "darkgoldenrod1", coppice = "forestgreen", mangrove = "mediumseagreen")
+alpha <- 0.2
+
+lowerplot <- function(data, mapping, ...) {
+  ggplot(data = data, mapping = mapping) +
+    geom_point(..., alpha = alpha) +
+    stat_ellipse() +
+    scale_color_manual(values = colors)
+}
+
+diagplot <- function(data, mapping, ...) {
+  ggplot(data = data, mapping = mapping) +
+    geom_density(..., alpha = alpha) +
+    scale_fill_manual(values = colors) +
+    scale_color_manual(values = colors)
+}
+
+diagplot2 <- function(data, mapping, ...) {
+  ggplot(data = data, mapping = mapping) +
+    geom_boxplot(..., alpha = alpha) +
+    scale_color_manual(values = colors) +
+    coord_flip()
+}
+
+p <- ggpairs(
+  data, columns = variables, mapping = aes(color = habitat),
+  upper = NULL, lower = list(continuous = lowerplot), diag = list(continuous = diagplot2)
+) +
+  theme_bw()
+
+p
+
+# There seems to be no difference between habitats.
+# To confirm this we run a two-way MANOVA with islands and habitats as effects
+
+#### 2. 2-way MANOVA ####
+
+# 2.1. Parametric MANOVA
+
+res_manova <- manova(cbind(PC1, PC2, PC3, PC4) ~ island * habitat, data = data)
+res_manova %>% summary(test = "Pillai")
+
+# This test suggests that all effects are significant, including an archipelago-wide
+# habitat-effect
+# How big is this effect?
+
+# 2.2. Effect sizes
+
+etasq(res_manova)
+
+# Very small archipelago-wide effect of habitat
+
[…]

### 2020-05-18 analyses/08-spatial correlation/SpatialAutocorrelation.R

@@ -1,0 +1,27 @@
+# Spatial autocorrelation by permutation test
+
+rm(list = ls())
+
+library(tidyverse)
+library(nmgc)
+
+data <- read.csv("data/reflectance.csv", header = TRUE)
+
+# Spatial autocorrelation test
+res_cortest <- nspcortest(
+  data, variables = paste0("PC", 1:4), nesting = "island", nperm = 1000, seed = 42,
+  to_pcomp = paste0("wl", 300:400), keep = "habitat"
+)
+
+# Save table
+t1 <- res_cortest$res
+t1 <- t1 %>% add_signif()
+save_table(t1, "table_spatial", digits = c(0, 3, 3, 0))
+
+# Table with multivariate means per site
+res_cortest$sites
+write.csv(res_cortest$sites, "table_sites.csv", row.names = FALSE)
+
+# Make a Latex table
+t2 <- res_cortest$sites
+save_table(t2, "table_sites", digits = c(0, 4, 4, 0, 4, 4, 4, 4))

### 2020-05-18 analyses/09-parallelism/Parallelism.R

@@ -1,0 +1,84 @@
+rm(list = ls())
+
+# Analysis of parallel divergence across islands
+
+library(tidyverse)
+library(nmgc)
+library(RColorBrewer)
+
+# Are differences in coloration similar on different islands?
+# To test this, perform a PCA on the whole archipelago
+# Then, measure and plot the pairwise differences between the habitats
+# in color space, along each PC
+# Re-scale these contrasts if needed, because some islands may be more
+# variable than others
+# If islands undergo the same divergence process, then they should cluster
+# together in this contrast space
+
+data <- read.csv("data/reflectance.csv", header = TRUE)
+variables <- paste0("PC", 1:4)
+wl <- paste0("wl", 300:700)
+
+# Perform PCA at the scale of the whole archipelago
+data <- data <- cbind(data, data.frame(npcomp(data, wl)$x)[, variables])
+
+# Compute contrast scores across islands and variables
+newdata <- data %>%
+  group_by(island, habitat) %>%
+  summarize_at(variables, mean) %>%
+  gather_("variable", "score", variables) %>%
+  spread_("habitat", "score") %>%
+  mutate(
+    coa_cop = coastal - coppice,
+    coa_man = coastal - mangrove,
+    cop_man = coppice - mangrove
+  ) %>%
+  ungroup() %>%
+  group_by(variable) %>%
+  mutate_at(
+    c("coa_cop", "coa_man", "cop_man"),
+    function(x) scale(x, center = FALSE, scale = TRUE)
+  ) %>%
+  filter(island %in% c("Abaco", "Bimini", "Cayman Brac", "Little Cayman", "Long Island")) %>%
+  ungroup()
+
+newdata <- newdata %>% mutate(
+  x = coa_cop,
+  y = cop_man
+)
+
+newdata <- newdata %>% rename(vbl = "variable")
+
+p <- ggplot(newdata, aes(x = x, y = y, fill = vbl, shape = island)) +
+  geom_point(size = 5) +
+  xlab("Coastal versus coppice") +
+  ylab("Coppice versus mangrove") +
+  theme_bw() +
+  labs(shape = "Island", fill = "Variable") +
+  scale_fill_manual(values = rev(brewer.pal(length(variables), "YlOrRd"))) +
+  scale_shape_manual(values = 21:25) +
[…]

### 2020-05-18 analyses/10-distances/Distances.R

@@ -1,0 +1,144 @@
+# Welcome to hell...
+
+rm(list = ls())
+
+# Here we explore the distances at which differences between sites are detected
+
+library(tidyverse)
+library(nmgc)
+
+data <- read.csv("data/reflectance.csv", header = TRUE)
+wl <- paste0("wl", 300:700)
+variables <- paste0("PC", 1:4)
+data <- cbind(data, npcomp(data, wl, nesting = "island", reduce = 1:4)$x)
+
+# Kepp only the most significant islands
+keep <- c("Abaco", "Bimini", "Cayman Brac", "Little Cayman", "Long Island")
+data <- data %>% filter(island %in% keep)
+data <- data %>% select(island, habitat, site, longitude, latitude, PC1, PC2, PC3, PC4)
+
+# Rearrange the data
+data <- data %>% gather_("variable", "score", variables)
+
+# Function to compare sites within an island for a given variable
+compare_sites <- function(data) {
+
+  data <- data %>%
+    group_by(site, habitat, longitude, latitude) %>%
+    nest()
+
+  ii <- 1:(nrow(data) - 1)
+  names(ii) <- ii
+  out <- map_dfr(ii, function(i) {
+    jj <- (i + 1):(nrow(data))
+    names(jj) <- jj
+    map_dfr(jj, function(j) {
+      res <- wilcox.test(data$data[[i]]$score, data$data[[j]]$score)
+      data.frame(U = res$statistic, pvalue = res$p.value)
+    }, .id = "j")
+  }, .id = "i")
+
+  out <- out %>% mutate_at(c("i", "j"), as.numeric)
+  out <- out %>% mutate(
+    habitat_i = map_chr(i, ~ as.character(data$habitat[.x])),
+    habitat_j = map_chr(j, ~ as.character(data$habitat[.x]))
+  )
+  out <- out %>% filter(habitat_i != habitat_j)
+  out <- out %>% mutate(
+    lon_i = map_dbl(i, ~ data$longitude[.x]),
+    lat_i = map_dbl(i, ~ data$latitude[.x]),
+    lon_j = map_dbl(j, ~ data$longitude[.x]),
+    lat_j = map_dbl(j, ~ data$latitude[.x])
+  )
+  out <- out %>% mutate(
+    distance = pmap_dbl(list(lon_i, lat_i, lon_j, lat_j), function(x1, y1, x2, y2) {
+      distm(x = c(x1, y1), y = c(x2, y2))
[…]

### 2020-05-18 maps/make_maps.R

@@ -1,0 +1,109 @@
+# Make maps of the archipelago
+
+rm(list = ls())
+
+library(tidyverse)
+library(sf)
+library(rnaturalearth)
+library(rnaturalearthdata)
+library(tmap)
+library(tmaptools)
+
+# Load geographical data
+world <- ne_countries(scale = "large", returnclass = "sf")
+
+# Filter the Caribbean only
+island_map_data <- world %>% filter(subregion == "Caribbean")
+
+# Read the bounding boxes of all islands
+island_coordinates <- read_csv("data/island_coordinates.csv")
+
+# Make them a nested list and rremove islands we do not want
+island_coordinates <- island_coordinates %>%
+  group_by(island) %>%
+  filter(!island %in% c("Conception Island", "Rum Cay")) %>%
+  nest
+
+islands <- island_coordinates$island
+
+# Intersect the Caribbean map with each ot the islands' bounding boxes to get one map per island
+island_map_data <- map(island_coordinates$data, function(curr_island_data) {
+
+  extent <- st_bbox(c(
+    xmin = curr_island_data$lonmin,
+    xmax = curr_island_data$lonmax,
+    ymin = curr_island_data$latmin,
+    ymax = curr_island_data$latmax
+  ), crs = 4326)
+
+  extent <- st_as_sfc(extent)
+
+  return(st_intersection(island_map_data, extent))
+
+})
+
+island_map_data <- do.call(rbind, island_map_data)
+
+# Load sampling sites
+data <- readRDS("data/locations.rds")
+
+# Add island names for facetting
+island_map_data$island <- levels(data$island)
+
+# Map the whole archipelago
+p <- tm_shape(world, bbox = bb(matrix(c(-85, 18, -70, 30), 2, 2))) +
+  tm_style("classic") +
+  tm_polygons(col = "palegoldenrod", border.col = "gray50") +
+  tm_grid(alpha = 0.5, n.x = 5, labels.size = 1) +
+  tm_scale_bar(position = c(0.9, 0.85), just = "right", lwd = 1, text.size = 0.7) +
+  tm_compass(position = c(0.05, 0.05), just = "left", size = 2, type = "arrow") +
+  tm_shape(island_map_data) +
+  tm_polygons(col = "gray10", border.col = "gray10") +
+  tm_shape(island_map_data) +
[…]

### 2020-05-18 analyses/07-ANOVA/ANOVA.R
issue fixed

@@ -128,9 +128,17 @@ p <- data %>%
 # 3.2. Add ANOVA P-values
 
 # Prepare P-value dataset
-p_anova <- res_anova$res %>%
-  rename(island = "nesting") %>%
-  mutate(plabel = round(pvalue, 4) %>% paste("P =", .)) %>%
+p_anova <- res_anova$res %>% rename(island = "nesting")
+
+# Replace with KW where applicable
+p_kw <- t3 %>% select(island, variable, pvalue)
+
+p_anova <- p_anova %>%
+  full_join(p_kw %>% group_by(island, variable) %>% nest()) %>%
+  mutate(pvalue = map2_dbl(pvalue, data, ~ if (is.null(.y)) .x else .y$pvalue)) %>%
+  select(-data)
+
+p_anova <- p_anova %>% mutate(plabel = round(pvalue, 4) %>% paste("P =", .)) %>%
   mutate(plabel = ifelse(pvalue < 0.0001, "P < 0.0001", plabel)) %>%
   mutate(plabel = ifelse(pvalue < 0.05, str_replace(plabel, "$", "*"), plabel))
 

### 2020-05-23 ms/main.tex
added used packages to the DESCRIPTION

\documentclass{article}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{verbatim} % multiline comments

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{float}
\usepackage{rotating}
\usepackage{booktabs}
%\restylefloat{figure} % for force placing figures
%\floatplacement{figure}{H}
%\floatplacement{table}{H}
%\floatplacement{longtable}{H}
%\usepackage[markers,nolists]{endfloat}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
%\usepackage{apalike}
\usepackage[round]{natbib}
\usepackage[switch]{lineno} % line numbers
\usepackage{longtable}
\usepackage{multirow}
% Supp mat
\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\title{Dewlap color variation in \textit{Anolis sagrei} is maintained between habitats within islands of the West Indies}
\author{
    \textsc{Rapha\"{e}l Scherrer$^{1,3}$ \thanks{Corresponding author: r.scherrer@rug.nl}, Colin M. Donihue$^{1,4}$, }\\
	\textsc{R. Graham Reynolds$^2$, Jonathan B. Losos$^{1,4}$ and Anthony J. Geneva$^1$} \\[1ex]
	%\thanks{A thank you} \\[1ex] % Your name
	\normalsize $^1$ Department of Organismic and Evolutionary Biology and Museum of Comparative Zoology \\ \normalsize Harvard University, Cambridge, MA, USA \\
	\normalsize $^2$ Department of Biology, University of North Carolina Asheville, Asheville, NC, USA\\ % Your institutionhttps://www.overleaf.com/project/5a0705e5f4573c1ace38dbdf
	\normalsize $^3$ Current address: Groningen Institute for Evolutionary Life Sciences,\\
	\normalsize Groningen, The Netherlands\\
	\normalsize $^4$ Current address: Department of Biology, Washington University, St. Louis, MO, USA\\
}
\date{} % Leave empty to omit a date

\begin{document}

\linenumbers

\maketitle

\begin{abstract}
    \input{sections/abstract.tex}
\end{abstract}

\textbf{Keywords} --- \textit{Anolis}, reflectance, local adaptation, sexual signal, supervised machine learning

\section*{Introduction}

\begin{table}
	\input{../analyses/01-counts/table_counts.tex}
\end{table}

\input{sections/introduction.tex}

\section*{Methods}

\input{sections/methods.tex}

\section*{Results}

\input{sections/results.tex}

\section*{Discussion}

\input{sections/discussion.tex}

\subsection*{Acknowledgements}

\input{sections/acknowledgments.tex}

\pagebreak

\section*{Figures}

\input{sections/figures.tex}

\pagebreak

\section*{Tables}

\input{sections/tables.tex}

\pagebreak

\bibliographystyle{apalike}
\bibliography{library.bib}

\beginsupplement

%\section*{Supplementary Figures}

%\input{sections/suppfigures.tex}

\pagebreak

\section*{Supplementary Figures}

\input{sections/suppfigures}

\pagebreak

\section*{Supplementary Tables}

\input{sections/supptables.tex}

\end{document}

### 2020-05-23 ms/sections/abstract.tex
added used packages to the DESCRIPTION

Animal signals evolve in an ecological context. Moreover, locally adapting animal sexual signals can be especially important for initiating or reinforcing reproductive isolation during the early stages of speciation. Dewlap color in \textit{Anolis} lizards can be highly variable between populations in relation to both biotic and abiotic adaptive drivers, albeit at relatively large geographical scales. Here, we investigated local adaptation of the dewlap across habitat-types at a small spatial scale, as this may give an indication of how conditions for the early stages of speciation may be met. We explored variation in dewlap coloration in one widespread species, \textit{Anolis sagrei}, across three characteristic habitats spanning the Bahamas and the Cayman Islands. Using reflectance spectrometry as well as supervised machine learning, we found some consistent differences in spectral properties of the dewlap between habitats within small islands. Passive divergence in dewlap phenotype associated with isolation-by-distance did not explain our results. Instead, the observed patterns in dewlap coloration are more consistent with an adaptive explanation in these \textit{A. sagrei} populations, as one would otherwise expect differences within islands to be erased by gene flow at such small geographical scales. Although these habitat-specific dewlap differences vary in magnitude and direction across islands, and islands themselves differ substantially, we found a suite of consistent archipelago-wide differences between habitat types, suggesting parallel responses to similar selective pressures. While at present, populations from these different habitats probably experience too much gene flow to follow distinct evolutionary lineages, should additional barriers arise between habitat-specific populations, the observed disruptive selection on dewlap coloration may facilitate ecological speciation.

### 2020-05-23 ms/sections/acknowledgments.tex
added used packages to the DESCRIPTION

Collection permission was granted by the Bahamas Environment, Science and Technology Commission, the Bahamas National Trust, and the Cayman Islands Department of the Environment. The authors thank Sofia Prado-Irwin, Pavitra Muralidhar, Nicholas Herrmann, Richard E. Glor, Alberto R. Puente-Rol\'{o}n, Kevin Aviles-Rodriguez, Kristin Winchell, and Melissa Kemp for assistance in the field. Funding for this work was provided by the Templeton Foundation (to JBL), NSF DEB~\#1500761 (to AJG), NSF DBI~\#1609284 (to CMD), and a Harvard Museum of Comparative Zoology Putnam Grant (to RGR).

### 2020-05-23 ms/sections/discussion.tex
added used packages to the DESCRIPTION

% Extra results to get: minimum distance at which we found differences and range-wide results

% There are inconsistent differences that are not explained by distance

\paragraph{Dewlap coloration differs between habitat-types} We found that male dewlap coloration in \textit{A. sagrei} significantly varied between habitat-types (beach scrub bush, primary coppice forest and mangrove forest) on five islands of the West Indies: Abaco, Bimini, Cayman Brac, Little Cayman and Long Island. However, the habitat-specific variation in dewlaps was not consistent between these islands. Although those results are consistent with adaptation at a very local scale, other evolutionary drivers could be at work, including phenotypic plasticity, random drift, or historical contingency including multiple colonization events. We reject this last explanation because all of the island populations in this study are strictly monophyletic, implying a single colonization event per island (van de Schoot, unpublished thesis; \citealt{Driessens2017}). While random drift cannot be completely ruled out, we see little evidence for a role of phenotypic isolation-by-distance (spatial autocorrelation) in explaining the differences we report. While we found a significant signal of isolation-by-distance on Eleuthera, we did not detect differences in dewlap coloration between habitats on this island.

% Plasticity is unlikely

\paragraph{A role of phenotypic plasticity is unlikely} Phenotypic plasticity could explain differences in coloration between habitat populations because of diet differences. The yellow, orange and red coloration in anoline dewlaps are produced by pterins and carotenoids \citep{Ortiz1962, Ortiz1962a, Ortiz1963, Ortiz1966, Macedonia2000, Steffen2007, Steffen2009}. Animals lack the ability to synthesize carotenoids, and those must therefore be found in the diet, while pterins are synthesized from nucleotides \citep{Goodwin1984, Hill2002, Hill2006}. Experimental manipulation of dietary carotenoid content showed no effect on dewlap coloration in \textit{A. sagrei} \citep{Steffen2010} nor in \textit{A. distichus} \citep{Ng2013}, another species with an orange-based dewlap. Plasticity due to differences in development (e.g. egg rearing conditions) is unlikely because dewlap coloration develops at sexual maturity \citep{Ng2013}. \citet{Cox2017} further found a high degree of heritability of dewlap coloration in \textit{A. sagrei}. These studies suggest that dewlap coloration is not a plastic trait, although transgenerational plastic effects cannot be completely ruled out by these one or two-generation common garden experiments \citep{Tariel2020}. That leaves an adaptive explanation, where dewlap color could be under differential natural and/or sexual selection in these different habitats.

% The maintenance of divergence in the face of gene flow

\paragraph{Divergence is maintained in the face of gene flow} The small spatial scale (the median distance between island sample sites was 11.2km + mention the minimum distance at which we found differences) and the lack of geographic barriers within islands, together with the high mobility \textit{A. sagrei} individuals \citep{Kamath2018}, imply ample opportunity for extensive gene flow between the populations in this study. In agreement with this, populations from different habitats were found to not be monophyletic based on <insert marker name here> (van de Schoot et al. unpublished thesis). While gene flow is expected to erase any differences acquired by drift without selection maintaining such polymorphism \citep{Dieckmann1999}, these divergent patterns remain, which may support an adaptive explanation. Maintenance of dewlap color divergence despite gene flow has been found in \textit{A. distichus} in Hispaniola \citep{Ng2012, Ng2016} and proposed as a mechanism of reproductive isolation in the early stages of speciation \citep{Ng2011, Lambert2013, Ng2017}. \citet{Stapley2011} found that dewlap color polymorphism was maintained in the absence of genetic structure between populations of \textit{A. apletophallus} from Panama. \citet{Thorpe2002a} found that divergence in dewlap coloration matched habitat-type better than mitochondrial lineage in \textit{A. roquet} on Martinique, and a convergent pattern was found in \textit{A. trinitatis} on the featureless island of St Vincent \citep{Thorpe2002b}. Divergence in body coloration, but not dewlap coloration, was also reported in \textit{A. conspersus} on another small island, Grand Cayman \citep{Macedonia2001}. Those results suggest a high adaptive potential of the anoline dewlap.

% The evidence for sexual selection

\paragraph{Sexual selection could be at play} Substantial levels of promiscuity in \textit{A. sagrei} suggest ample opportunity for female mate choice and sexual selection \citep{Kamath2018}. A number of studies suggest that characteristics of the dewlap are indicators of male quality and may therefore act as a cue in mate choice, according to the "good genes" model of sexual selection \citep{Andersson1994}. \citet{Cook2013} found lower orange reflectance in dewlaps with heavily parasitized \textit{A. brevirostris}, suggesting a trade-off in carotenoid use between the immune response and pigment deposition. \citet{Steffen2014} found that lower UV and orange-red reflectance predict contest-winning success between males. \citet{Driessens2015} further found that more yellow and red dewlaps (relative to UV) predict better body condition, and that higher yellow and UV reflectance at the margin of the dewlap predict higher hematocrit (the concentration of red blood cells), indicating a better health. Other aspects of the dewlap may be important for sexual selection, for example, dewlap size correlates with bite force \citep{Vanhooydonck2005} and sexual size dimorphism (a proxy for sexual selection; \citealt{Vanhooydonck2009}). Even though display frequency increases in the presence of females \citep{Driessens2014}, there seems to be no link between dewlap display frequency and mating success \citep{Tokarz2002, Tokarz2005} or individual quality \citep{Driessens2015}. Our report of inconsistent divergence between islands, however, is at odds with the good-genes model of sexual selection, which would predict the evolution of dewlap coloration in the same direction across the archipelago. This is consistent with \citet{Baeckens2018}, who found no link between the average island-dewlap coloration and sexual size dimorphism in \textit{A. sagrei} across the West Indies. If sexual selection is at play in our system, it more likely involves divergent female preferences that are not linked to male quality but rather to components of the environment, and may well be arbitrary (Fisherian sexual selection; \citealt{Andersson1994}).

% Could there be local adaptation? To what?

\paragraph{Dewlap coloration could be locally adapted} Presently, we do not know the adaptive drivers of dewlap  color divergence observed in this study. Some degree of parallel evolution is usually a good indicator for an adaptive process \citep{Losos2011}, and convergent patterns of dewlap color evolution in similar environments across islands and species have been documented \citep{Thorpe2002a, Thorpe2002b}. Here, we found evidence for a few consistencies in within-island divergence across the nine islands studied (show some archipelago-wide results?); dewlap brightness was lower in the primary coppice habitat on three islands, and the primary coppice habitat was intermediate in coloration between mangrove and beach scrub on two islands. Those patterns could reflect adaptation to components of the habitat \citep{Endler1988}. However, it is not clear what those components of the habitat might be. Previous studies have found that dewlap coloration maximizes detectability given the light conditions in the local habitat, primarily through UV contrast -- with UV-brighter dewlaps in UV-dark, mesic habitats and UV-darker dewlaps in UV-bright, xeric habitats -- in \textit{A. cristatellus} and \textit{A. cooki} on Puerto Rico \citep{Leal2002, Leal2004}. We found no such pattern in \textit{A. sagrei}, where instead, we found the darkest dewlaps in the darkest, mesic habitat -- primary coppice forest -- on three islands, and dewlaps often differed the most between beach scrub and mangrove forest, two xeric habitats with similar, high irradiance levels \citep{Howard1950, Schoener1968}. The inconsistent and idiosyncratic patterns we observed suggest that dewlap color variation between habitats cannot be predicted by habitat identity alone. Studies of Jamaican and Hispaniolan anoles similarly found between-habitat differences in dewlap coloration but no evidence for higher detectability \citep{Fleishman2009, Ng2012}. Habitats on different islands may also differ in other aspects than light conditions such as densities of predators or other anole species, which have been shown to affect among-island dewlap diversity \citep{Vanhooydonck2009, Baeckens2018}. In particular, \citet{Baeckens2018} recently showed that dewlaps with spotted patterns occurred more often in \textit{A. sagrei} on islands with more coexisting species of anoles.

% Conclusion, and a word about speciation

\paragraph{Implications in the context of speciation} Local adaptation can be a precursor to ecological speciation, a process that may have given rise to the adaptive radiation of \textit{Anolis} lizards \citep{Harmon2003, Gavrilets2009}. Ecologically-mediated divergence of a sexual signal may be a potent path to the evolution of reproductive isolation through divergent sexual selection \citep{Reynolds2007, Servedio2011}. Evidence suggests that dewlap coloration could take this role in anoles \citep{Ng2011, Lambert2013, Geneva2015, Ng2017}, or at least that it is frequently involved in species recognition \citep{Williams1969, Williams1977, Losos1985, Macedonia1994, Fleishman2000, Macedonia2013, Ingram2016, Baeckens2018}. Although this signal is not detected at the phylogenetic scale of the whole genus \citep{Nicholson2007, Harrison2012, Ingram2016}, sexual signals are often evolutionarily very labile \citep{Kraaijeveld2011}, and the anole dewlap in particular is capable of rapid macroevolution; for example, \textit{A. conspersus} on Grand Cayman evolved a UV-blue dewlap from an ancestral orange dewlap in 2 to 3 million years \citep{Macedonia2001}. We present evidence of multiple cases of potentially adaptive maintenance of habitat-associated dewlap divergence over small geographical scale in \textit{A. sagrei} across the West Indies. While these populations do not appear to be in the process of speciation, our results strongly suggest that the anoline dewlap has enough micro-scale, local adaptive potential to participate in the build-up of reproductive isolation, should it be recruited for assortative mating.

### 2020-05-23 ms/sections/figures.tex
added used packages to the DESCRIPTION

% Map of the West Indies
\begin{figure}[H]
    \centering
	\includegraphics[width=0.8\textwidth]{figures/map.pdf}
	\caption{Map of the West Indies with sampled islands highlighted in black.}
	\label{fig:map}
\end{figure}

% Classification accuracy of SVMs
\begin{figure}[H]
    \centering
	\includegraphics[width=\textwidth]{figures/classif_svm_pca.png}
	\caption{Distributions of classification accuracy across all SVM machines (100 replicates of 5 cross-validation bins each). The dashed line represents the density of a corresponding null binomial distribution, which would be expected under random guessing (testing sets with 20\% of the observations for each island and success probability of $1/3$). Inset plots show the corresponding average confusion matrices and represent the proportion of lizards from each habitat (columns) reassigned in each other habitat (rows), with an interpretation guide in the right panel. P-values indicate deviations of the mean classification accuracy to the expected null binomial distribution. *, P < 0.05}
	\label{fig:classif_svm_pca}
\end{figure}

% Boxplots and analyses of variance
\begin{sidewaysfigure}
    \centering
	\includegraphics[width=18cm]{figures/figure_anova.png}
	\caption{Dewlap coloration varies between habitat-types along different dimensions on different islands. (A) Boxplots show the distribution of the data for the five islands with significant differences in dewlap coloration between habitat (P-values in Fig. \ref{fig:classif_svm_pca}), along the first four principal components. We report the P-values of univariate analyses of variance conducted for each island (see Methods). Horizontal bars indicate \textit{post hoc} significant contrasts at a 0.05 error rate (see Methods). *, P < 0.05. (B) Mapping of the reflectance at various wavelengths onto the principal components. Note that PC1 largely represents brightness.}
	\label{fig:boxplots}
\end{sidewaysfigure}

### 2020-05-23 ms/sections/introduction.tex
added used packages to the DESCRIPTION

% Animal signals may be shaped by their environment

The staggering diversity of animal communication signals has long been of interest to evolutionary biologists. Animals use chemical, mechanical, electromagnetic, and visual signals to communicate in a wide variety of contexts, including competition for mates, species recognition, aposematism, cooperation, etc. \citep{Bradbury2011}. A primary evolutionary factor shaping communication signals is the sensory system and behavior of their recipient(s) (the sensory drive hypothesis; \citealt{Endler1988,Endler1992,Endler1998}). Over the past decades, scientists have established that signals evolve in an ecological context and are dependent on environmental conditions \citep{Endler1992,Endler1993,Endler1993a}. Just as different habitats may favor different combinations of eco-morphological traits to maximize performance and fitness \citep{Arnold1983}, they may also shape different forms of a signal, so as to maximize its transmission and detection (e.g. \citealt{Seehausen1997}), or reduce its detection by unintended recipients such as predators \citep{Endler1984,Endler1990,Endler1991,Halfwerk2014}. This selective pressure may drive the local adaptation of communication signals.\\

% One main barrier to the maintenance of diversity is gene flow

One potential barrier to the maintenance of localized signal divergence is the homogenizing effect of gene flow. Population genetics theory suggests that gene flow may counteract local adaptation between localities and prevent divergence altogether, especially at small spatial scales, because of the inflow of maladapted alleles or because of the breaking of linkage between coevolving loci \citep{Felsenstein1976, Garcia-Ramos1997, Dieckmann1999, Lenormand2002, Hendry2007}. This has been confirmed empirically in systems such as stick-insects \citep{Nosil2004} and sticklebacks \citep{Hendry2007a}. Yet, examples of microgeographic adaptation, i.e. adaptation at smaller scales than the range of dispersal, exist, highlighting a high potential of some organisms to respond to selection in the face of gene flow (see \citealt{Richardson2014} and references therein). Examples include small scale adaptation in fragmented areas in Australian fruit flies \citep{Willi2012}, or local adaptation to predation pressure in North American salamanders \citep{Richardson2013}. Therefore, despite evidence that local adaptation may be particularly difficult at small spatial scales where gene flow tends to cause adjoining populations to remain genetically homogeneous, the potential adaptive response of species traits, in particular communication signals, to localized differences in habitats remains relatively unknown \citep{Richardson2014}.\\

% Anolis lizards and their dewlap

Lizards of the neotropical genus \textit{Anolis} are a model system for studying the eco-evolutionary dynamics of local adaptation and natural selection \citep{Losos2009}. A particularly conspicuous trait of anoles is their dewlap; an extensible flap of skin that is typically sexually dimorphic and used as a communication signal in courtship \citep{Sigmund1983, Driessens2014, Driessens2015}, competition \citep{Losos1985, Macedonia1994, Macedonia2013} as well as in predator deterrence \citep{Leal1995, Leal1997, Leal1997b}. Dewlap characteristics vary widely among the approximately $400$ species of the genus \citep{Nicholson2007}. Interspecific variation in dewlap coloration is implicated in species recognition \citep{Rand1970b, Williams1969, Williams1977, Losos1985, Macedonia1994, Fleishman2000, Macedonia2013}, and possibly involved in speciation \citep{Lambert2013, Geneva2015, Ng2017}.\\

Within species, studies have shown a link between variation in dewlap coloration and differences in habitats or climatic conditions \citep{Macedonia2001, Leal2002, Thorpe2002a, Thorpe2002b, Leal2004, Vanhooydonck2009, Ng2012, Ng2013, Ng2016, Vanhooydonck2009, Driessens2017}. Some studies suggest that those differences may be adaptive, and that dewlaps may have evolved to maximize detectability given local light conditions \citep{Fleishman2001, Leal2002, Leal2004}. Other studies testing this hypothesis, however, found no pattern \citep{Fleishman2009, Ng2012, Macedonia2014a}.\\

% So what? Introducing Anolis sagrei as a model species

% I am re-citing the same papers here, is that okay?
% This is the room for sagrei in particular

Previous studies investigating variation in anole dewlaps compared populations at relatively large geographical scales, e.g. between islands \citep{Vanhooydonck2009, Driessens2017} or within large islands such as Puerto Rico \citep{Leal2002, Leal2004} or Hispaniola \citep{Ng2012, Ng2016}. These large scales should reduce gene flow \citep{Ng2011, Lambert2013, Richardson2014, Ng2017}. That said, examples do exist of divergence in dewlap coloration at smaller scales or between populations with high degrees of gene flow \citep{Thorpe2002a, Thorpe2002b, Stapley2011, Ng2016}.\\

The species \textit{Anolis sagrei} is widespread across islands of the West Indies \citep{Reynolds2020}. It is a model organism in studies of local adaptation \citep{Losos1994, Losos1997a, Losos2001, Kolbe2012}, biological invasion \citep{Kolbe2008} and sexual selection \citep{Tokarz2002, Tokarz2005, Tokarz2006, Driessens2014, Steffen2014, Driessens2015}. Between-island variation in the mainly orange-red color of its dewlap was shown to be better explained by climatic variables \citep{Driessens2017} than biotic factors such as sexual selection or predation pressure \citep{Vanhooydonck2009, Baeckens2018}. How intra-island differences in habitat may contribute to the diversity of dewlap coloration, however, remains unexplored, and may reveal new insights into the scale of local adaptation despite gene flow.\\

The island bank systems of the Bahamas and Cayman Islands comprise relatively small islands, with no major geographic barriers within islands limiting dispersal for this promiscuous species \citep{Kamath2018}. These islands all share three characteristic native West Indian habitat-types -- beach scrub bush, closed-canopy primary coppice forest, and mangrove forest -- that are often spatially intermingled. These habitats contrast in environmental parameters including vegetation community, light irradiance, humidity and temperature \citep{Howard1950, Schoener1968}. Each of these islands has been colonized independently by \textit{A. sagrei} (\citealt{Driessens2017, Reynolds2020}?, van de Schoot et al. unpubl.), such that these archipelagos constitute an ideal suite of natural replicates to explore within-island dewlap diversity across multiple islands.\\

% What we did

Here, we analyzed the color characteristics \textit{A. sagrei} dewlaps within nine islands in the Bahamas and Cayman Islands, combining reflectance spectrometry and supervised machine learning. Our sampling design included sites in close proximity (the median distance between two sites within an island was $11.2$km). We tested the hypothesis that the spatial scale was too small for phenotypic divergence to build up. If this was not the case, we predicted that if light conditions in the environment indeed drive color evolution, dewlaps should be most similar between beach scrub and mangrove forest, which both have high levels of light irradiance, contrary to the darker, closed-canopy coppice forest. Similar, if detectability is maximized given the local conditions, we expected darker and more contrasting dewlaps in high irradiance habitats. Finally, if habitat characteristics are strong determinants of dewlap color variation, similar patterns should be observed across multiple islands \citep{Losos2011}. We found strong support for fine-scale, within-island differences in coloration between lizards inhabiting the three habitat-types in several color space dimensions, suggesting a potentially strong effect of divergent selection. However, the divergence patterns we observed did not match our expectations and were highly variable between islands. We found no evidence of isolation-by-distance as an explanation for the observed differences. Our results are nevertheless consistent with small-scale adaptive maintenance of signal polymorphism despite presumed considerable opportunity for gene flow.

% Bibliography notes:

% Anolis sagrei was shown to be able to rapidly invade available ecological niches upon colonization of new islands through changes in limb proportions (Losos1997a; Losos2001)

% Convergence

% Besides, \textit{Anolis} lizards are well-known examples of eco-morphological evolutionary convergence, where multiple ``ecomorphs'' adapted independently, but in similar ways, to similar ecological niches across islands of the Greater Antilles (Losos2004; Losos2009). This extent of evolutionary determinism seems to act within species at present too. Populations of \textit{A. sagrei} were shown to undergo the same eco-morphological divergence upon colonization of new islands as ecomorphs did millions of years before (Losos1997a; Losos2001; Calsbeek2007; Losos2007). Strong developmental constraints on morphology seem to explain this tendency (Sanger2012; McGlothlin2018). However, despite the evidence suggesting potential for evolutionary convergence at the intra-specific level in anoles, this pattern is unknown for communication signals such as dewlap coloration. Populations of a same species, because they share a common genetic architecture, are expected to evolve convergent adaptations more often than distantly related species (Ord2015), and this could apply to the dewlap of \textit{A. sagrei}. On the other hand, sexual signals are often evolutionarily labile due to sexual selection (Kraaijeveld2011), but also selection mediated by other recipients of the signal (Endler1981; Endler1988; Endler1984; Endler1993; Endler1993a; Endler1998). This is consistent with the high diversity of dewlaps across anole species (Nicholson2007) and could imply evolutionary contingency for this colorful trait.\\

### 2020-05-23 ms/sections/methods.tex
added used packages to the DESCRIPTION

% Citation for classifiers as a good group comparison tool
% Yang and Yang on LDA being performed on PC

\subsection*{Data collection}

We sampled 466 lizards from seven islands in the Bahamas Archipelago -- Abaco, North Andros, South Andros, South Bimini, Eleuthera, Long Island, Ragged Island -- and two in the Cayman Islands -- Cayman Brac and Little Cayman (Figure \ref{fig:map}). These islands and island banks were chosen to span the West Indian range of \textit{Anolis sagrei}. Three habitats were sampled on each island based on characterizations by \citet{Howard1950} and \citet{Schoener1968}. Each habitat is clearly distinguishable by their dominant vegetation type --- xeric coastal scrub (open, relatively dry habitat consisting of low vegetation or isolated trees), primary coppice forest (closed-canopy forest) and mangrove forest (wet coastal habitat with trees growing in brackish water and high light penetration). Sample sizes are given in Table \ref{suptab:counts}. Our sampling design enabled us to test for differences between habitats at a coarse and fine geographical scale. The median distance between two localities within an island was $11.18$km, with some islands being sampled at smaller or larger scales (Figure \ref{supfig:map}, Table \ref{suptab:locations}). $80.3$\% of all pairwise distances within islands were below $50$km. Additionally, there are no major barriers to dispersal (such as mountains) on any of the islands that we sampled.

\subsection*{Reflectance measurements}

We measured reflectance between 300 and 700nm wavelength, a range that encompasses the colors visible to most lizards and vertebrates in general \citep{Lazareva2012}. Measurements were taken with an Ocean Optics USB4000 spectrometer, a pulsed Xenon light source (PX-2, Ocean Optics, Largo, FL, USA) and a reflectance probe protected by a black anodized aluminum sheath. Measurements were taken with a 45-degree inclination to prevent specular reflection \citep{Endler1990}. The device was regularly standardized with a Spectralon white standard (Labsphere, North Sutton, NH, USA). Reflectance was measured at the center of the dewlap.

\subsection*{Analysis}

All analyses in this study were performed in R 3.6.1 \citep{RCoreTeam2019}.

\subsubsection*{Dimensionality reduction}

Reflectance curves were smoothened using the R package pavo \citep{Maia2013} as well as custom R functions (where are those?) to results into one reflectance value at each nanometer in wavelength from 300 to 700nm. Because neighboring wavelengths are highly collinear in reflectance, we reduced the dimensionality of the data using principal component analysis (PCA), as per \citet{Cuthill1999, Leal2002}. We performed PCA on each island separately and systematically retained the first four principal components (PC), which together always explained more than $88.8\%$ of the variance across islands (Table \ref{suptab:pcavariances}). PC1 explained between $40$ and $56$\% of the variance across islands; PC2 explained $17.4$--$27.9$\%; PC3 $12.7$--$17.6$\% and PC4 $4.3$--$10.5$\%. The first four PCs explained similar proportions of variance when calculated for all islands together (Table \ref{suptab:pcavariances}). PCs need not represent the same wavelengths across islands because they are fitted on different datasets. Nevertheless, PC1 was very collinear with brightness for all islands (Table \ref{suptab:brightness}). PC2 correlated highly with the red and ultraviolet ends of the spectrum, which were inversely correlated with each other (Fig. \ref{fig:boxplots}A). Higher PCs corresponded to various combinations of wavelengths. Because PC1 correlated uniformly with all wavelengths across the spectrum  we considered PC2 onwards to capture the chromatic dimensions of color space, i.e. the relative contributions of the wavelengths regardless of brightness.

\subsubsection*{Machine learning}

Our data violated the multivariate analysis of variance (MANOVA) assumption of homogeneity of covariance matrices across groups for all islands but Ragged Island (Box's M-test, \citealt{Box1949, Morrison1988}, implemented in the R package heplots, \citealt{Fox2018},  Table \ref{suptab:covariances}). We also detected within-habitat deviations from multivariate normality, primarily on Abaco, Bimini and Eleuthera (Henze-Zirkler's test, \citealt{Henze1990}, implemented in the R package MVN, \citealt{Korkmaz2014}, Table \ref{suptab:multinorm}). For these reasons and to reduce the chances of false discovery, we conducted multivariate group comparisons using support vector machines (SVMs), a model-free, nonparametric supervised machine learning technique.\\

% Box M is sensitive to non-normality \citep{Fox2018}

Machine learning for group comparison has become more popular in ecology and evolution in the recent years (e.g. \citet{Pigot2020}). In particular, SVMs are designed to find the best possible nonlinear boundaries between labelled groups of points in multidimensional spaces, without assumptions about the distribution of the data \citep{Cortes1995, Cristianini2000, Kim2018}. This makes them well suited to field biological data, which often violate the assumptions of classical linear modeling \citep{Kim2018} and can be, as in the case of coloration, inherently highly multivariate \citep{Cuthill1999}. The significance and robustness of a classification is assessed by training a machine to recognize differences between groups on part of the data, and using that fitted machine to predict the group-labels of data points from a testing set that were not included in the training, based solely on their multivariate coordinates. This cross-validation procedure results in a proportion of correctly classified points, or generalization accuracy score, which can be compared to that expected under random guessing using a binomial test.\\

In this study, we performed SVM classifications on each island separately. We used a standard five-fold cross-validation procedure, where the data were randomly split into five bins of approximately equal sizes, and each bin was in turn taken as the testing set while the rest was used as a training set, thus resulting in five trained machines per cross-validation. We replicated this procedure 100 times for each island to account for stochastic outcomes. We performed binomial tests to evaluate the significance of the deviations in observed mean generalization accuracy per island to null expectations under random guessing. Each training data set was downsampled to the size of its least represented habitat to ensure balanced training samples. We ensured that each habitat was represented by at least five data points in the training set. All classification analyses were repeated using the more classical linear discriminant analysis (LDA), a supervised machine learning technique finding linear boundaries that maximize the differences between groups, albeit assuming multivariate normality and homogeneity of covariance matrices \citep{Ripley1996a}. We used the R package rminer \citep{Cortez2010, Cortez2016} for SVMs, and MASS (citation) for LDAs. We used rminer's default heuristic search option to automatically tune the Gaussian kernel parameter $\sigma$ and the complexity parameter $C$ for the SVMs.\\

\subsubsection*{Univariate analyses}

For each island where significant differences in multivariate dewlap coloration were detected between habitats, we used multiple univariate analyses of variance (ANOVA) to identify which variables were responsible for the observed differences. We constructed our ANOVA models in two steps, as per \citet{Zuur2009}. In a first step, we accounted for heterogeneity of variances across groups by systematically comparing the goodness-of-fit of an ANOVA model estimated with ordinary least squares (OLS) with that of a model estimated with generalized least squares (GLS), which allowed one estimate of residual variance per habitat (using the R package nlme, \citealt{Pinheiro2000, Pinheiro2020}). Both models were fitted with restricted maximum likelihood (REML). Goodness-of-fit was estimated using Akaike's Information Criterion corrected for small sample sizes (AICc, from the R package MuMIn, \citealt{Barton2019}), and the estimation method yielding the lowest AICc was retained. In a second step, we re-fitted the retained model with maximum likelihood (ML) to test for the effect of habitat-type using likelihood ratio tests (LRT) between a model including a habitat-term and a null model lacking it. Significant \textit{post hoc} contrasts were assessed using Tukey's Honest Significant Difference (HSD) test whenever the OLS-model was the best fit, and pairwise Wilcoxon tests whenever there was evidence for heterogeneity of variances, and the GLS-model was the best fit.\\

We further tested the robustness of our results to non-normality. For each habitat-island combination where multivariate non-normality was detected (Table \ref{suptab:multinorm}), we assessed the univariate normality of each variable (Shapiro-Wilk's test, Table \ref{suptab:normality}) and backed our ANOVA results with nonparametric Kruskal-Wallis tests every time significant deviations were detected.\\

\subsubsection*{Spatial autocorrelation}

We tested for within-island spatial autocorrelation between the geographical distances among sampling sites and their euclidean distances in multivariate color space (mean PC1 to PC4 per site, Table \ref{suptab:locations}), regardless of habitat-type. Because often only a few sites were sampled per island, we could not get meaningful results from tests using sites as units of observation. Instead, we designed a permutation test where we randomly reshuffled individual lizards across sites within islands 1,000 times each, and systematically recalculated Pearson's correlation coefficient between geographic distances (computed as geodesic distances using the R package geosphere, \citealt{Hijmans2019}) and phenotypic distances. We used the resulting null distributions of correlation coefficients to assess the significance of the observed spatial autocorrelation for each island (Table \ref{suptab:autocor}).

### 2020-05-23 ms/sections/results.tex
added used packages to the DESCRIPTION

 % Little but detectable differences on five islands

We tested for variation in \textit{A sagrei} dewlap coloration between populations living in three characteristic habitat types across nine islands that span the West Indian range of the brown anole (Fig \ref{fig:map}, \ref{supfig:map}). We found that most of the variation in coloration partitioned between islands (semi-parametric MANOVA, modified ANOVA-type statistic (MATS) = 2009.6, P < 0.001, Fig. \ref{supfig:reflectance}, explained variance $\eta^2 = 44.3$\%, MANOVA approximation). Nonetheless, we did find evidence for differences in dewlap coloration between habitat-types, and those were mostly island-specific (habitat-by-island interaction term, MATS = 384.4, P < 0.001, explained variance $\eta^2 = 11.4$\%), leaving a small but significant portion of the variation explained by an archipelago-wide habitat effect (MATS = 42.5, P = .001, $\eta^2 = 4.8$\%).\\

The small archipelago-wide effect of habitat-type was detected for PC1, PC2 and PC3 (mixed ANOVA with island as a random effect, Table \ref{suptab:anova_pooled}), but this effect was too small for post hoc tests to find which habitats differed. Because we identified deviations from univariate normality for each group (Table \ref{suptab:norm_pooled}, but no deviations from homoskedasticity, Table \ref{suptab:}), we repeated the analysis with a Kruskal-Wallis test

SVM classifiers correctly assigned individuals to their habitat of origin based solely upon dewlap coloration on five islands: Abaco, Bimini, Cayman Brac, Little Cayman, and Long island (Fig. \ref{fig:classif_svm_pca}). An LDA approach yielded similar success rates (Fig. \ref{supfig:classif_lda_pca}), suggesting robust differences between these populations. Of the five islands, Little Cayman was the best discriminated with a mean SVM generalization success of 73.4\% (Table \ref{suptab:classif_svm_pca}). The results of the classification analyses on PCA data were very similar to results from SVMs and LDAs trained on reflectance values at 50nm-spaced wavelengths from 300 to 700nm (Fig. \ref{supfig:classif_svm_refl} and \ref{supfig:classif_lda_refl}).\\

Differentiation in dewlap coloration occurred in multiple dimensions of color space. Moreover, the differences in dewlaps between habitats were not always consistent between islands, thus, we will discuss the habitat-specific variation in dewlap coloration for each island where significant differences were detected in turn (Fig. \ref{fig:boxplots}, Table \ref{tab:anova}). Figure \ref{fig:boxplots}A provides a key to map principal component scores to the underlying wavelengths. Throughout, all reported significant differences in analyses of variance (ANOVA, Table \ref{tab:anova}) where significant deviations from normality were detected (Table \ref{suptab:normality}) were also significant in nonparametric Kruskal-Wallis tests (Table \ref{suptab:kruskal}).\\

On Abaco, dewlaps did not differ in PC1, which represents brightness. Mangrove lizards had significantly lower PC2 scores, corresponding to higher ultraviolet reflectance and lower red reflectance. Coastal beach scrub lizards had lower scores on PC3, corresponding to lower ultraviolet reflectance and higher blue reflectance.\\

On Bimini, coastal beach scrub lizards had significantly brighter dewlaps than lizards from mangroves (PC1), but mangrove lizards had higher PC2 scores than beach scrub lizards, indicating higher violet and blue reflectance, and lower red reflectance. Lizards from primary coppice had higher PC3 scores overall, which correlated very positively with ultraviolet reflectance.\\

On Cayman Brac, coppice-lizard dewlaps were significantly less bright than lizards from the other habitats. Coastal beach scrub lizards had dewlaps that scored low on PC2, corresponding to lower violet-blue and more red, while the mangrove lizards exhibited the opposite: relatively higher levels of violet-blue and less red. In PC3 space we found that dewlaps from lizards in the coastal habitat had high ultraviolet reflectance, coppice lizards had intermediate levels, and mangrove lizards had relatively low levels.\\

On Little Cayman, the dewlaps of coppice lizards were significantly darker (PC1) than coastal-lizards. Mangrove lizards had less ultraviolet and redder dewlaps (PC2). The dewlaps of the coastal beach scrub lizards had higher levels of red and ultraviolet reflectance and less blue reflectance than the dewlaps of the other habitat-populations (PC3).\\

On Long Island, lizards from the coppice habitat had darker dewlaps than lizards from the other habitats (PC1, Kruskal-Wallis test accounting for the presence of one outlier with very high brightness, Fig. \ref{fig:boxplots}A, Table \ref{suptab:kruskal}). Coastal lizards had relatively more ultraviolet and less blue-green reflectance in their dewlaps (PC3). These coastal-habitat lizards also scored lower on PC4, corresponding to slightly more violet and green-yellow dewlaps, and less blue dewlaps, than the mangrove lizards on the island.\\

Sensitivity analyses on classifiers suggested an overall higher relative importance for PC2 and PC3 in determining between-group differences on Abaco, both in SVM and LDA classifiers (Fig. \ref{supfig:importance_svm_pca}, \ref{supfig:importance_lda_pca}), consistent with our ANOVA results (Fig. \ref{fig:boxplots}B). There was no strong signal of differences in relative importance among principal components on the other islands. Sensitivy analyses of SVMs trained on reflectance scores rather than principal components revealed, however, a consistently higher importance of ultraviolet reflectance in between-group differences on all islands (Fig. \ref{fig:importance_svm_refl}). This pattern was not recovered for LDAs trained on reflectance scores (Fig. \ref{fig:importance_lda_refl}). Another way to get insights into the importance of variables is to perform a single LDA on the significant islands and see what variables correlate with the LDs (on PCA and on reflectance data).\\

% No spatial autocorrelation

We did not find significant spatial autocorrelation between the sampling sites on the islands where we detected a significant habitat effect. We did, however, detect a significant positive signal of autocorrelation on Eleuthera ($P = 0.02$, Table \ref{suptab:autocor}), suggesting possible color differentiation through isolation-by-distance on this island.\\

In contrast, differences in dewlap coloration between habitats were often detected in close geographical proximity. For example, Bimini, Cayman Brac and Little Cayman were among the smallest islands in our study (Fig. \ref{supfig:map}). Besides, for pairs of habitats where significant differences in dewlap coloration were detected along some principal components, comparisons of the actual sampling sites indicate that the most detectable differences involve sites that are mostly 5-10km apart, with significant differences in PC3 detected between the beach scrub and coppice habitat at little more than 100m distance from each other on Bimini (Fig. \ref{supfig:distances}, Table \ref{suptab:distances}).\\

% Patterns of differentiation were inconsistent across islands

Patterns of differentiation were inconsistent across the five most significant islands. No pattern of variation was shared by all five islands, along any dimension. Some patterns did seem more common however, such as darker dewlaps among coppice lizards (Cayman Brac, Little
Cayman, and Long Island, Fig. \ref{fig:boxplots}) or the intermediate position of coppice lizards in chromatic color space (Cayman Brac and Long Island). In other cases, patterns of differentiation were reversed from one island to another, with more ultraviolet reflecting dewlaps in mangroves than in coastal habitat on Abaco and Cayman Brac, but the opposite on Little Cayman and Long Island. Overall, it seemed that patterns of heterogeneity of variance were often driven by higher variances in coloration within beach scrub lizards (Fig. \ref{fig:boxplots}, Table \ref{tab:anova}). Yet other patterns were idiosyncratic, such as the combination of higher red and ultraviolet reflectance in coastal lizards on Little Cayman, where the rule seemed to be a negative correlation between ultraviolet and red reflectance across every other island.\\

That said, archipelago-wide differences in dewlap coloration between habitats were detected by a set of SVMs trained on data pooled from the whole archipelago, regardless of island identity, both for PCA data and reflectance scores (Fig. \ref{supfig:classif_svm_pca_pooled}, \ref{supfig:classif_svm_refl_pooled}). This pattern seemed to be driven by mangrove lizards being correctly reassigned more often than predicted by chance. Sensitivity analyses on these machines suggest a relatively small role of long wavelengths (red reflectance) in driving this pattern (Fig. \ref{supfig:importance_svm_refl_pooled}), but did not reveal strong differences between the PCs in relative importance (Fig. \ref{supfig:importance_svm_pca_pooled}).  Archipelago-wide differences were not detected by LDA classifiers at all (Fig. \ref{supfig:classif_lda_pca_pooled}, \ref{supfig:classif_lda_refl_pooled}), nor by a MANOVA (Wilk's $\lambda = 0.98$, df = 2, $F(8,920)=1.17$, P = 0.316) or a semi-parametric MANOVA-equivalent (R package MANOVA.RM, citation, Wald-type statistic (df = 8) = 12.58, parametric bootstrap with 1,000 iterations, P = 0.131, Fig. \ref{supfig:pooled}).\\

% Mention the MANOVA table somewhere
% LDA to get the importance of each variable whenever MANOVA is significant, or restrict to SVM significance ---> this could be a useful function for the package

### 2020-05-23 ms/sections/suppfigures.tex
added used packages to the DESCRIPTION

% Map of each island
\begin{figure}[H]
\centering
	\includegraphics[width=\textwidth]{../maps/detailed_map.pdf}
	\caption{Map of the sampling sites and corresponding habitats across 9 islands of the West Indies.}
	\label{supfig:map}
\end{figure}

% Reflectance curves
\begin{figure}[H]
    \centering
	\includegraphics[width=\textwidth]{../analyses/02-reflectance/figure_reflectance.png}
	\caption{5-95th percentile distributions of lizard dewlap reflectance values (in \% of incoming light) across wavelengths for each island and each habitat.}
	\label{supfig:reflectance}
\end{figure}

% Classification accuracy of LDAs
\begin{figure}[H]
    \centering
	\includegraphics[width=\textwidth]{../analyses/04-machine learning/plots/classif_lda_pca.png}
	\caption{Distributions of classification accuracy across all LDA machines (100 replicates of 5 cross-validation bins each). The dashed line represents the density of a corresponding null binomial distribution, which would be expected under random guessing (testing sets with 20\% of the observations for each island and success probability of $1/3$). Inset plots show the corresponding average confusion matrices and represent the proportion of lizards from each habitat (columns) reassigned in each other habitat (rows), with an interpretation guide in the right panel. P-values indicate deviations of the mean classification accuracy to the expected null binomial distribution. *, P < 0.05}
	\label{supfig:classif_lda_pca}
\end{figure}

\begin{figure}[H]
    \centering
	\includegraphics[width=\textwidth]{../analyses/04-machine learning/plots/classif_svm_refl.png}
	\caption{Distributions of classification accuracy across all SVM machines (100 replicates of 5 cross-validation bins each) trained directly on reflectance data from 300 to 700nm, at 50nm intervals. The dashed line represents the density of a corresponding null binomial distribution, which would be expected under random guessing (testing sets with 20\% of the observations for each island and success probability of $1/3$). Inset plots show the corresponding average confusion matrices and represent the proportion of lizards from each habitat (columns) reassigned in each other habitat (rows), with an interpretation guide in the right panel. P-values indicate deviations of the mean classification accuracy to the expected null binomial distribution. *, P < 0.05}
	\label{supfig:classif_svm_refl}
\end{figure}

\begin{figure}[H]
    \centering
	\includegraphics[width=\textwidth]{../analyses/04-machine learning/plots/classif_lda_refl.png}
	\caption{Distributions of classification accuracy across all LDA machines (100 replicates of 5 cross-validation bins each) trained directly on reflectance data from 300 to 700nm, at 50nm intervals. The dashed line represents the density of a corresponding null binomial distribution, which would be expected under random guessing (testing sets with 20\% of the observations for each island and success probability of $1/3$). Inset plots show the corresponding average confusion matrices and represent the proportion of lizards from each habitat (columns) reassigned in each other habitat (rows), with an interpretation guide in the right panel. P-values indicate deviations of the mean classification accuracy to the expected null binomial distribution. *, P < 0.05}
	\label{supfig:classif_lda_refl}
\end{figure}

### 2020-05-23 ms/tables/table_anova.tex
added used packages to the DESCRIPTION

% ANOVA table
\begin{table}[H]
    \centering
    \caption{Analyses of variance (ANOVA) testing the effect of habitat-type on multiple response variables across the most significant islands according to the SVM classification (Figure \ref{fig:classif_svm_pca}). Best fit refers to the fitting method, i.e. OLS, ordinary least squares (favored when $\Delta$AICc $> 0$) or GLS, generalized least squares (favored when $\Delta$AICc $< 0$). AICc values are presented for the best-fitting model. df, degrees of freedom of the model; df$_{\mbox{LRT}}$, degrees of freedom of the likelihood ratio test (LRT) used to test the habitat effect; Log-lik., log-likelihood of the full model, i.e. including the habitat effect; $\chi^2$, likelihood-ratio between the full model and a null model excluding the habitat effect. *, P < 0.05; **, P < 0.01; ***, P < 0.001.}
    \begin{tabular}{lllrrrrrrll}
        \hline
        Island & Variable & Best fit & df & AICc & $\Delta$AICc & df$_{\mbox{LRT}}$ & Log-lik. & $\chi^2$ & P-value & \\
        \hline
        Abaco & PC1 & OLS & 4 & 710.40 & 2.16 & 2 & -356.96 & 0.14 & 0.9308 & \\
        Abaco & PC2 & OLS & 4 & 620.06 & 4.02 & 2 & -310.16 & 31.74 & < 0.0001 &  ***\\
        Abaco & PC3 & OLS & 4 & 517.80 & 2.01 & 2 & -257.18 & 27.37 & < 0.0001 &  ***\\
        Abaco & PC4 & OLS & 4 & 440.59 & 0.78 & 2 & -217.18 & 1.36 & 0.5070 &  \\
        Bimini & PC1 & OLS & 4 & 561.34 & 0.77 & 2 & -283.06 & 7.40 & 0.0248 &  *\\
        Bimini & PC2 & OLS & 4 & 448.06 & 1.29 & 2 & -223.77 & 8.09 & 0.0175 &  *\\
        Bimini & PC3 & GLS & 6 & 405.30 & -0.23 & 2 & -199.18 & 10.39 & 0.0056 &  **\\
        Bimini & PC4 & OLS & 4 & 274.15 & 3.54 & 2 & -132.74 & 0.33 & 0.8499 &  \\
        Cayman Brac & PC1 & GLS & 6 & 402.81 & -4.05 & 2 & -200.86 & 13.81 & 0.0010 &  **\\
        Cayman Brac & PC2 & OLS & 4 & 332.14 & 3.51 & 2 & -165.91 & 8.41 & 0.0149 &  *\\
        Cayman Brac & PC3 & OLS & 4 & 295.78 & 2.77 & 2 & -146.57 & 27.16 & < 0.0001 &  ***\\
        Cayman Brac & PC4 & OLS & 4 & 279.24 & 4.33 & 2 & -137.77 & 5.63 & 0.0600 &  \\
        Little Cayman & PC1 & OLS & 4 & 367.17 & 2.50 & 2 & -185.99 & 8.18 & 0.0167 &  *\\
        Little Cayman & PC2 & GLS & 6 & 287.61 & -3.61 & 2 & -140.46 & 29.76 & < 0.0001 &  ***\\
        Little Cayman & PC3 & OLS & 4 & 277.69 & 1.41 & 2 & -138.06 & 21.34 & < 0.0001 &  ***\\
        Little Cayman & PC4 & OLS & 4 & 226.70 & 2.53 & 2 & -110.74 & 2.85 & 0.2410 &  \\
        Long Island & PC1 & GLS & 6 & 442.34 & -2.09 & 2 & -221.16 & 2.91 & 0.2331 &  \\
        Long Island & PC2 & GLS & 6 & 351.37 & -3.08 & 2 & -172.57 & 4.52 & 0.1043 &  \\
        Long Island & PC3 & OLS & 4 & 322.06 & 3.67 & 2 & -159.98 & 11.24 & 0.0036 &  **\\
        Long Island & PC4 & OLS & 4 & 195.45 & 2.38 & 2 & -92.88 & 6.46 & 0.0395 &  *\\
        \hline
    \end{tabular}
    \label{tab:anova}
\end{table}

### 2020-05-23 R/save_table.R

@@ -1,11 +1,25 @@
 #' Save tables for this study
+#'
+#' @param tab The table
+#' @param nm File name (without the extension)
+#' @param digits,col_names Parameters for `knitr::kable`
+#'
+#' @return Nothing. Just saves.
+#'
+#' @export
 
-save_table <- function(tab, nm, digits = 0) {
+save_table <- function(tab, nm, digits = 0, col_names = NULL) {
 
   library(knitr)
 
+  # Save table as CSV
   write.csv(tab, paste0(nm, ".csv"), row.names = FALSE)
-  tab <- knitr::kable(tab, "latex", digits = digits, booktabs = TRUE)
+
+  # Save LaTeX table
+  if (is.null(col_names)) col_names <- NA
+  tab <- knitr::kable(
+    tab, "latex", digits = digits, booktabs = TRUE, col.names = col_names
+  )
   texfile <- file(paste0(nm, ".tex"))
   writeLines(tab, texfile)
   close(texfile)

### 2020-05-23 analyses/01-counts/Counts.R

@@ -2,9 +2,9 @@
 
 rm(list = ls())
 
-library(knitr)
+library(dewlap)
 
 data <- read.csv("data/reflectance.csv", header = TRUE)
 tab <- table(data$island, data$habitat)
 
-save_table(tab, "table_counts")
+save_table(tab, "analyses/01-counts/table_counts")

### 2020-05-24 ms/sections/figures.tex
working with tables

@@
     \centering
-	\includegraphics[width=0.8\textwidth]{figures/map.pdf}
+	\includegraphics[width=0.8\textwidth]{../maps/map.pdf}
 	\caption{Map of the West Indies with sampled islands highlighted in black.}
 	\label{fig:map}
-\end{figure}
-
-% Classification accuracy of SVMs
-\begin{figure}[H]
-    \centering
-	\includegraphics[width=\textwidth]{figures/classif_svm_pca.png}
-	\caption{Distributions of classification accuracy across all SVM machines (100 replicates of 5 cross-validation bins each).
-The dashed line represents the density of a corresponding null binomial distribution, which would be expected under random guessing (testing sets with 20\% of the observations for each island and success probability of $1/3$).
-Inset plots show the corresponding average confusion matrices and represent the proportion of lizards from each habitat (columns) reassigned in each other habitat (rows), with an interpretation guide in the right panel.
-P-values indicate deviations of the mean classification accuracy to the expected null binomial distribution.
-*, P < 0.05}
-	\label{fig:classif_svm_pca}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=18cm]{figures/figure_anova.png}
-	\caption{Dewlap coloration varies between habitat-types along different dimensions on different islands.
-(A) Boxplots show the distribution of the data for the five islands with significant differences in dewlap coloration between habitat (P-values in Fig.
-\ref{fig:classif_svm_pca}), along the first four principal components.
-We report the P-values of univariate analyses of variance conducted for each island (see Methods).
-Horizontal bars indicate \textit{post hoc} significant contrasts at a 0.05 error rate (see Methods).
-*, P < 0.05.
-(B) Mapping of the reflectance at various wavelengths onto the principal components.
-Note that PC1 largely represents brightness.}
-	\label{fig:boxplots}
+	\includegraphics[width=18cm]{../analyses/07-ANOVA/figure_anova.png}
+	\caption{Dewlap color variation between habitat-types across the most significant islands.
+(A) Mapping of reflectance at various wavelengths onto the principal components (loadings from the PCA rotation matrix).
+(B) Distribution of PC scores between habitats along the first four PCs on each island where significant between-habitat differences were detected using SVMs.
+P-values are reported for univariate ANOVA (or Kruskal-Wallis tests when applicable, see Methods).
+Post hoc significant differences at a 0.05 error rate are indicated with horizontal bars.
+*, P < 0.05.}
+	\label{fig:anova}
 \end{sidewaysfigure}
+
+% Classification accuracy of SVMs
+\begin{figure}[H]
+    \centering
+	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_svm_pca"}
+	\caption{SVM classification accuracy across islands based on principal component data.
+Histograms show accuracy distributions over 100 replicates for each five cross-validation bins per island.
+The dashed line is the density of a corresponding null binomial distribution, which would be expected under random guessing (testing sets with 20\% of the observations for each island and success probability of $1/3$).
+Inset plots show the corresponding average confusion matrices and represent the proportion of lizards from each habitat (columns) reassigned in each other habitat (rows), with an interpretation guide in the right panel.
+Binomial test P-values indicate deviations of the mean classification accuracy to the null distribution.
+*, P < 0.05.}
+	\label{fig:classif-svm-pca}
+\end{figure}

### 2020-05-24 ms/sections/methods.tex
working with tables

@@
 PCs need not represent the same wavelengths across islands because they are fitted on different datasets.
-Nevertheless, PC1 was very collinear with brightness for all islands (Table \ref{suptab:brightness}).
+Nevertheless, PC1 was very collinear with brightness for all islands (Figure..., Table \ref{suptab:brightness}).
 PC2 correlated highly with the red and ultraviolet ends of the spectrum, which were inversely correlated with each other (Fig.
@@
 
+\subsubsection*{Pooled analyses}
+
+In addition to within-island PCA, we performed a PCA on pooled data from the whole archipelago.
+Explain what wavelengths mapped onto what PCs (Table suptab:pcavariances).
+We used this dataset to partition the variance in dewlap coloration among islands, habitats and habitats within islands, using a two-way multivariate analysis of variance (MANOVA) with an interaction term.
+However, because the assumptions of parametric MANOVA were violated for all islands but Ragged Island (multivariate normality and homogeneity of covariance matrices, Tables suptab:multinorm, suptab:covariance), we used a semi-parametric MANOVA instead (R package MANOVA.RM, citation), with P-values calculated from a bootstrap procedure with 1,000 iterations.
+We calculated the proportion of variance explained by islands, habitats and the habitat-by-island interaction using partial effect sizes $\eta^2$ on a MANOVA-approximation of the analysis (R package heplots).
+
 \subsubsection*{Machine learning}
 
-Our data violated the multivariate analysis of variance (MANOVA) assumption of homogeneity of covariance matrices across groups for all islands but Ragged Island (Box's M-test, \citealt{Box1949, Morrison1988}, implemented in the R package heplots, \citealt{Fox2018},  Table \ref{suptab:covariances}).
+% move this to the previous section?
+
+Our data violated the multivariate analysis of variance (MANOVA) assumption of homogeneity of covariance matrices across groups for all islands but Ragged Island (see previous section, \citealt{Box1949, Morrison1988}, implemented in the R package heplots, \citealt{Fox2018},  Table \ref{suptab:covariances}).
 We also detected within-habitat deviations from multivariate normality, primarily on Abaco, Bimini and Eleuthera (Henze-Zirkler's test, \citealt{Henze1990}, implemented in the R package MVN, \citealt{Korkmaz2014}, Table \ref{suptab:multinorm}).
@@
 This makes them well suited to field biological data, which often violate the assumptions of classical linear modeling \citep{Kim2018} and can be, as in the case of coloration, inherently highly multivariate \citep{Cuthill1999}.
-The significance and robustness of a classification is assessed by training a machine to recognize differences between groups on part of the data, and using that fitted machine to predict the group-labels of data points from a testing set that were not included in the training, based solely on their multivariate coordinates.
+First, a machine is trained to recognize differences between groups within a subset of the data called the training set.
+Significance of differences is then assessed by testing the accuracy of that fitted machine in predicting the group-labels of data points that were not included in the training, called a testing set, based solely on their multivariate coordinates.
 This cross-validation procedure results in a proportion of correctly classified points, or generalization accuracy score, which can be compared to that expected under random guessing using a binomial test.\\
@@
 In this study, we performed SVM classifications on each island separately.
-We used a standard five-fold cross-validation procedure, where the data were randomly split into five bins of approximately equal sizes, and each bin was in turn taken as the testing set while the rest was used as a training set, thus resulting in five trained machines per cross-validation.
+We used a standard five-fold cross-validation procedure, where the data were randomly split into five bins of approximately equal sizes.
+Each bin was in turn taken as the testing set while the rest was used as a training set, thus resulting in five trained machines per cross-validation.
 We replicated this procedure 100 times for each island to account for stochastic outcomes.
-We performed binomial tests to evaluate the significance of the deviations in observed mean generalization accuracy per island to null expectations under random guessing.
+We performed binomial tests to evaluate the significance of deviations in observed mean generalization accuracy per island to null expectations under random guessing.
 Each training data set was downsampled to the size of its least represented habitat to ensure balanced training samples.
-We ensured that each habitat was represented by at least five data points in the training set.
+We ensured that each habitat was represented by at least five data points in the training set.\\
+
 All classification analyses were repeated using the more classical linear discriminant analysis (LDA), a supervised machine learning technique finding linear boundaries that maximize the differences between groups, albeit assuming multivariate normality and homogeneity of covariance matrices \citep{Ripley1996a}.
@@
 We used rminer's default heuristic search option to automatically tune the Gaussian kernel parameter $\sigma$ and the complexity parameter $C$ for the SVMs.\\
+
+The same procedure was repeated on principal components from the whole archipelago (see Pooled analyses) to evaluate the significance of archipelago-wide differences in dewlap coloration across habitats.\\
+
+All machine learning classifications performed on principal components were also repeated on the original reflectance datasets reduced to 50-nm spaced wavelengths from 300 to 700nm.\\
+
+We conducted one-dimensional sensitivity analyses (R pacakge rminer) to determine the relative importance of the different input variables during classification where significant differences were detected, both on machines trained on principal components and machines trained on non-transformed reflectance at various wavelengths.
+In parallel, we conducted univariate analyses of variance to independently test the importance of different variables in between-habitat variation, on islands where the machines detected significant differences based on binomial tests (next section).
 
@@
 Both models were fitted with restricted maximum likelihood (REML).
-Goodness-of-fit was estimated using Akaike's Information Criterion corrected for small sample sizes (AICc, from the R package MuMIn, \citealt{Barton2019}), and the estimation method yielding the lowest AICc was retained.
-In a second step, we re-fitted the retained model with maximum likelihood (ML) to test for the effect of habitat-type using likelihood ratio tests (LRT) between a model including a habitat-term and a null model lacking it.
-Significant \textit{post hoc} contrasts were assessed using Tukey's Honest Significant Difference (HSD) test whenever the OLS-model was the best fit, and pairwise Wilcoxon tests whenever there was evidence for heterogeneity of variances, and the GLS-model was the best fit.\\
+Goodness-of-fit was estimated using Akaike's Information Criterion corrected for small sample sizes (AICc, R package MuMIn, \citealt{Barton2019}), and the estimation method yielding the lowest AICc was retained.
+In a second step, we re-fitted the retained model with maximum likelihood (ML) to test for the effect of habitat-type using likelihood ratio tests (LRT) between a model including a habitat-term and a null model lacking the habitat-term.\\
 
-We further tested the robustness of our results to non-normality.
-For each habitat-island combination where multivariate non-normality was detected (Table \ref{suptab:multinorm}), we assessed the univariate normality of each variable (Shapiro-Wilk's test, Table \ref{suptab:normality}) and backed our ANOVA results with nonparametric Kruskal-Wallis tests every time significant deviations were detected.\\
+We tested the assumptions of the parametric ANOVA for each island included in the univariate analyses.
+For all islands where deviations from multivariate normality were detected in at least one habitat (Table multinorm), we assessed univariate normality for each principal component (Shapiro's test, Table normality).
+For skewed PCs that deviated significantly from normality, we repeated the analysis using a nonparametric Kruskal-Wallis tests.
+We found no multivariate outliers based on the Mahalanobis distance (package).
+We used the cases of better fit of the GLS model relative to the OLS model as evidence for heterogeneity of variances, which were then accounted for by the GLS approach (Table anova).\\
+
+Significant \textit{post hoc} contrasts were assessed using Tukey's Honest Significant Difference (HSD) test whenever the assumptions of normality and homogeneity of variances was met, Dunnett's T3 method when only homogeneity of variances was violated but not normality, and Nemenyi's test when normality was violated.\\
+
+We used the same procedure to investigate which variables, if any, were involved in archipelago-wide multivariate differences between habitats detected in our two-way MANOVA design (see Pooled analyses).
+However, in the first step or our model comparison procedure, we added mixed-effect equivalents of our OLS and GLS models, this time with island as a random effect.
+The resulting four models were compared and the best fitting variance structure was retained as explained above.
 
@@
 We tested for within-island spatial autocorrelation between the geographical distances among sampling sites and their euclidean distances in multivariate color space (mean PC1 to PC4 per site, Table \ref{suptab:locations}), regardless of habitat-type.
-Because often only a few sites were sampled per island, we could not get meaningful results from tests using sites as units of observation.
+Because often only a few sites were sampled per island, we could not get meaningful results from tests that use sites as units of observation, such as Moran's I test.
 Instead, we designed a permutation test where we randomly reshuffled individual lizards across sites within islands 1,000 times each, and systematically recalculated Pearson's correlation coefficient between geographic distances (computed as geodesic distances using the R package geosphere, \citealt{Hijmans2019}) and phenotypic distances.
-We used the resulting null distributions of correlation coefficients to assess the significance of the observed spatial autocorrelation for each island (Table \ref{suptab:autocor}).
+We used the resulting null distributions of correlation coefficients to assess the significance of the observed spatial autocorrelation for each island.
+
+\subsubsection*{Site differences}
+
+In this study, we were interested in the minimum spatial scale at which significant differences between habitats could be detected within islands.
+We performed multiple pairwise nonparametric Wilcoxon tests to compare dewlap coloration between sites with different habitat-types, for each pair of habitats and each variable where significant differences were detected with our analyses of variances.

### 2020-05-24 ms/sections/results.tex
working with tables

@@
 We tested for variation in \textit{A sagrei} dewlap coloration between populations living in three characteristic habitat types across nine islands that span the West Indian range of the brown anole (Fig \ref{fig:map}, \ref{supfig:map}).
-We found that most of the variation in coloration partitioned between islands (semi-parametric MANOVA, modified ANOVA-type statistic (MATS) = 2009.6, P < 0.001, Fig.
+We found that most of the variation in coloration partitioned between islands (two-way semi-parametric MANOVA, modified ANOVA-type statistic (MATS) = 2009.6, P < 0.001, Fig.
 \ref{supfig:reflectance}, explained variance $\eta^2 = 44.3$\%, MANOVA approximation).
@@
 
-The small archipelago-wide effect of habitat-type was detected for PC1, PC2 and PC3 (mixed ANOVA with island as a random effect, Table \ref{suptab:anova_pooled}), but this effect was too small for post hoc tests to find which habitats differed.
-Because we identified deviations from univariate normality for each group (Table \ref{suptab:norm_pooled}, but no deviations from homoskedasticity, Table \ref{suptab:}), we repeated the analysis with a Kruskal-Wallis test
+The small archipelago-wide effect of habitat-type was detected for PC1, PC2 and PC3 (mixed-effect ANOVA with island as a random effect, Table \ref{suptab:anova_pooled}, backed up by KW), but this effect was too small for post hoc tests to find which habitats differed.
+Archipelago-wide differences in dewlap coloration between habitats were also detected by SVMs trained on pooled data regardless of island identity, both for PCA data and reflectance scores (Fig.
+\ref{supfig:classif_svm_pca_pooled}, \ref{supfig:classif_svm_refl_pooled}).
+This pattern seemed to be driven by mangrove lizards being correctly reassigned more often than predicted by chance.
+Sensitivity analyses on these machines suggest a relatively small role of long wavelengths (red reflectance) in driving this pattern (Fig.
+\ref{supfig:importance_svm_refl_pooled}), but did not reveal strong differences between the PCs in relative importance (Fig.
+\ref{supfig:importance_svm_pca_pooled}).
+Archipelago-wide differences were not detected by LDA classifiers at all (Fig.
+\ref{supfig:classif_lda_pca_pooled}, \ref{supfig:classif_lda_refl_pooled}).
+So the difference could be in something else than means.\\
 
-SVM classifiers correctly assigned individuals to their habitat of origin based solely upon dewlap coloration on five islands: Abaco, Bimini, Cayman Brac, Little Cayman, and Long island (Fig.
+Within islands, SVM classifiers correctly assigned individuals to their habitat of origin based solely upon dewlap coloration on five islands: Abaco, Bimini, Cayman Brac, Little Cayman, and Long island (Fig.
 \ref{fig:classif_svm_pca}).
@@
 Moreover, the differences in dewlaps between habitats were not always consistent between islands, thus, we will discuss the habitat-specific variation in dewlap coloration for each island where significant differences were detected in turn (Fig.
-\ref{fig:boxplots}, Table \ref{tab:anova}).
-Figure \ref{fig:boxplots}A provides a key to map principal component scores to the underlying wavelengths.
-Throughout, all reported significant differences in analyses of variance (ANOVA, Table \ref{tab:anova}) where significant deviations from normality were detected (Table \ref{suptab:normality}) were also significant in nonparametric Kruskal-Wallis tests (Table \ref{suptab:kruskal}).\\
+\ref{fig:boxplots}, Table \ref{tab:anova}, Table KW).
+Figure \ref{fig:boxplots}A provides a key to map principal component scores to the underlying wavelengths.\\
 
@@
 
-On Long Island, lizards from the coppice habitat had darker dewlaps than lizards from the other habitats (PC1, Kruskal-Wallis test accounting for the presence of one outlier with very high brightness, Fig.
-\ref{fig:boxplots}A, Table \ref{suptab:kruskal}).
+On Long Island, lizards from the coppice habitat had darker dewlaps than lizards from the other habitats (PC1).
 Coastal lizards had relatively more ultraviolet and less blue-green reflectance in their dewlaps (PC3).
@@
 There was no strong signal of differences in relative importance among principal components on the other islands.
-Sensitivy analyses of SVMs trained on reflectance scores rather than principal components revealed, however, a consistently higher importance of ultraviolet reflectance in between-group differences on all islands (Fig.
+Sensitivity analyses of SVMs trained on reflectance scores rather than principal components revealed, however, a consistently higher importance of ultraviolet reflectance in between-group differences on all islands (Fig.
 \ref{fig:importance_svm_refl}).
 This pattern was not recovered for LDAs trained on reflectance scores (Fig.
-\ref{fig:importance_lda_refl}).
-Another way to get insights into the importance of variables is to perform a single LDA on the significant islands and see what variables correlate with the LDs (on PCA and on reflectance data).\\
+\ref{fig:importance_lda_refl}).\\
 
@@
 \ref{supfig:map}).
-Besides, for pairs of habitats where significant differences in dewlap coloration were detected along some principal components, comparisons of the actual sampling sites indicate that the most detectable differences involve sites that are mostly 5-10km apart, with significant differences in PC3 detected between the beach scrub and coppice habitat at little more than 100m distance from each other on Bimini (Fig.
+Besides, for pairs of habitats where significant differences in dewlap coloration were detected along some principal components, comparisons of the actual sampling sites indicate that the most detectable differences involved sites that were mostly 5-10km apart.
+Our most extreme case of significant differences occurred for PC3 between the beach scrub site and a coppice site, separated from each other by a a few hundreds of meters at most on Bimini (Wilcoxon test, Fig.
 \ref{supfig:distances}, Table \ref{suptab:distances}).\\
@@
 Patterns of differentiation were inconsistent across the five most significant islands.
-No pattern of variation was shared by all five islands, along any dimension.
-Some patterns did seem more common however, such as darker dewlaps among coppice lizards (Cayman Brac, Little
-Cayman, and Long Island, Fig.
+Contrasts in principal components between habitats, calculated on pooled data from the whole archipelago, were not similar, for any component, among islands (contrast figure, MANOVA, P > 0.05).
+No pattern of variation was shared by all five significant islands, along any dimension.
+Some patterns did seem more common however, such as darker dewlaps among coppice lizards (Cayman Brac, Little Cayman, and Long Island, Fig.
 \ref{fig:boxplots}) or the intermediate position of coppice lizards in chromatic color space (Cayman Brac and Long Island).
@@
 Yet other patterns were idiosyncratic, such as the combination of higher red and ultraviolet reflectance in coastal lizards on Little Cayman, where the rule seemed to be a negative correlation between ultraviolet and red reflectance across every other island.\\
-
-That said, archipelago-wide differences in dewlap coloration between habitats were detected by a set of SVMs trained on data pooled from the whole archipelago, regardless of island identity, both for PCA data and reflectance scores (Fig.
-\ref{supfig:classif_svm_pca_pooled}, \ref{supfig:classif_svm_refl_pooled}).
-This pattern seemed to be driven by mangrove lizards being correctly reassigned more often than predicted by chance.
-Sensitivity analyses on these machines suggest a relatively small role of long wavelengths (red reflectance) in driving this pattern (Fig.
-\ref{supfig:importance_svm_refl_pooled}), but did not reveal strong differences between the PCs in relative importance (Fig.
-\ref{supfig:importance_svm_pca_pooled}).
-Archipelago-wide differences were not detected by LDA classifiers at all (Fig.
-\ref{supfig:classif_lda_pca_pooled}, \ref{supfig:classif_lda_refl_pooled}), nor by a MANOVA (Wilk's $\lambda = 0.98$, df = 2, $F(8,920)=1.17$, P = 0.316) or a semi-parametric MANOVA-equivalent (R package MANOVA.RM, citation, Wald-type statistic (df = 8) = 12.58, parametric bootstrap with 1,000 iterations, P = 0.131, Fig.
-\ref{supfig:pooled}).\\
-
-% Mention the MANOVA table somewhere
-% LDA to get the importance of each variable whenever MANOVA is significant, or restrict to SVM significance ---> this could be a useful function for the package

### 2020-05-24 ms/sections/suppfigures.tex
working with tables

@@
 	\includegraphics[width=\textwidth]{../maps/detailed_map.pdf}
-	\caption{Map of the sampling sites and corresponding habitats across 9 islands of the West Indies.}
+	\caption{Map of the sampling sites and corresponding habitats across nine islands of the West Indies.}
 	\label{supfig:map}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics{"../analyses/03-PCA/figure_brightness"}
+	\caption{Correlation between dewlap brightness (as measured by the mean reflectance from 300 to 700nm in wavelength) and PC1 score for each island.
+Pearson's squared correlation coefficients are reported.
+*, P < 0.05.}
+	\label{supfig:brightness}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics{"../analyses/03-PCA/figure_brightness_pooled"}
+	\caption{Correlation between dewlap brightness (as measured by the mean reflectance from 300 to 700nm in wavelength) and PC1 score across the whole archipelago.
+Pearson's squared correlation coefficient is reported.
+*, P < 0.05.}
+	\label{supfig:brightness_pooled}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=\textwidth]{../analyses/02-reflectance/figure_reflectance.png}
-	\caption{5-95th percentile distributions of lizard dewlap reflectance values (in \% of incoming light) across wavelengths for each island and each habitat.}
+	\includegraphics[width=\textwidth]{"../analyses/02-reflectance/figure_reflectance"}
+	\caption{5-95th percentile range of lizard dewlap reflectance values (in \% of incoming light) across wavelengths for each island and each habitat.}
 	\label{supfig:reflectance}
-\end{figure}
-
-% Classification accuracy of LDAs
-\begin{figure}[H]
-    \centering
-	\includegraphics[width=\textwidth]{../analyses/04-machine learning/plots/classif_lda_pca.png}
-	\caption{Distributions of classification accuracy across all LDA machines (100 replicates of 5 cross-validation bins each).
-The dashed line represents the density of a corresponding null binomial distribution, which would be expected under random guessing (testing sets with 20\% of the observations for each island and success probability of $1/3$).
-Inset plots show the corresponding average confusion matrices and represent the proportion of lizards from each habitat (columns) reassigned in each other habitat (rows), with an interpretation guide in the right panel.
-P-values indicate deviations of the mean classification accuracy to the expected null binomial distribution.
-*, P < 0.05}
-	\label{supfig:classif_lda_pca}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=\textwidth]{../analyses/04-machine learning/plots/classif_svm_refl.png}
-	\caption{Distributions of classification accuracy across all SVM machines (100 replicates of 5 cross-validation bins each) trained directly on reflectance data from 300 to 700nm, at 50nm intervals.
-The dashed line represents the density of a corresponding null binomial distribution, which would be expected under random guessing (testing sets with 20\% of the observations for each island and success probability of $1/3$).
-Inset plots show the corresponding average confusion matrices and represent the proportion of lizards from each habitat (columns) reassigned in each other habitat (rows), with an interpretation guide in the right panel.
-P-values indicate deviations of the mean classification accuracy to the expected null binomial distribution.
-*, P < 0.05}
-	\label{supfig:classif_svm_refl}
+	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_svm_pca_pooled"}
+	\caption{Archipelago-wide SVM classification accuracy based on principal component data.
+Machines were trained on individual dewlaps regardless of island identity.
+The histogram shows the accuracy distribution over 100 replicates for each five cross-validation bins.
+The legend is the same as in Figure \ref{fig:classif-svm-pca}.}
+	\label{supfig:classif-svm-pca-pooled}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=\textwidth]{../analyses/04-machine learning/plots/classif_lda_refl.png}
-	\caption{Distributions of classification accuracy across all LDA machines (100 replicates of 5 cross-validation bins each) trained directly on reflectance data from 300 to 700nm, at 50nm intervals.
-The dashed line represents the density of a corresponding null binomial distribution, which would be expected under random guessing (testing sets with 20\% of the observations for each island and success probability of $1/3$).
-Inset plots show the corresponding average confusion matrices and represent the proportion of lizards from each habitat (columns) reassigned in each other habitat (rows), with an interpretation guide in the right panel.
-P-values indicate deviations of the mean classification accuracy to the expected null binomial distribution.
-*, P < 0.05}
-	\label{supfig:classif_lda_refl}
+	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_svm_refl_pooled"}
+	\caption{Archipelago-wide SVM classification accuracy based on reflectance data at 50nm-intervals in wavelength (see Methods).
+Machines were trained on individual dewlaps regardless of island identity.
+The histogram shows the accuracy distribution over 100 replicates for each five cross-validation bins.
+The legend is the same as in Figure \ref{fig:classif-svm-pca}.}
+	\label{supfig:classif-svm-refl-pooled}
 \end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/importance_svm_pca_pooled"}
+	\caption{Sensitivity analyses of the different input variables in the archipelago-wide SVM classification on principal component data (Figure \ref{supfig:classif-svm-pca-pooled}), with relative importance computed for every machine.}
+	\label{supfig:importance-svm-pca-pooled}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/importance_svm_refl_pooled"}
+	\caption{Sensitivity analyses of the different input variables in the archipelago-wide SVM classification on reflectance data at 50nm-intervals in wavelength (Figure \ref{supfig:classif-svm-refl-pooled}), with relative importance computed for every machine.}
+	\label{supfig:importance-svm-refl-pooled}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_lda_pca_pooled"}
+	\caption{Archipelago-wide LDA classification accuracy based on principal component data.
+Machines were trained on individual dewlaps regardless of island identity.
+The histogram shows the accuracy distribution over 100 replicates for each five cross-validation bins.
+The legend is the same as in Figure \ref{fig:classif-svm-pca}.}
+	\label{supfig:classif-lda-pca-pooled}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_lda_refl_pooled"}
+	\caption{Archipelago-wide LDA classification accuracy based on reflectance data at 50nm-intervals in wavelength (see Methods).
+Machines were trained on individual dewlaps regardless of island identity.
+The histogram shows the accuracy distribution over 100 replicates for each five cross-validation bins.
+The legend is the same as in Figure \ref{fig:classif-svm-pca}.}
+	\label{supfig:classif-lda-refl-pooled}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_lda_pca"}
+	\caption{LDA classification accuracy across islands based on principal component data.
+Histograms show accuracy distributions over 100 replicates for each five cross-validation bins per island.
+The legend is the same as in Figure \ref{fig:classif-svm-pca}.}
+	\label{supfig:classif-lda-pca}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_svm_refl"}
+	\caption{SVM classification accuracy across islands based on reflectance data at 50nm-intervals in wavelength (see Methods).
+Histograms show accuracy distributions over 100 replicates for each five cross-validation bins per island.
+The legend is the same as in Figure \ref{fig:classif-svm-pca}.}
+	\label{supfig:classif-svm-refl}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_lda_refl"}
+	\caption{LDA classification accuracy across islands based on reflectance data at 50nm-intervals in wavelength (see Methods).
+Histograms show accuracy distributions over 100 replicates for each five cross-validation bins per island.
+The legend is the same as in Figure \ref{fig:classif-svm-pca}.}
+	\label{supfig:classif-lda-refl}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/importance_svm_pca"}
+	\caption{Sensitivity analyses of the different input variables in the within-island SVM classification on principal component data (Figure \ref{supfig:classif-svm-pca}), with relative importance computed for every machine.}
+	\label{supfig:importance-svm-pca}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/importance_lda_pca"}
+	\caption{Sensitivity analyses of the different input variables in the within-island LDA classification on principal component data (Figure \ref{supfig:classif-lda-pca}), with relative importance computed for every machine.}
+	\label{supfig:importance-lda-pca}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/importance_svm_pca"}
+	\caption{Sensitivity analyses of the different input variables in the archipelago-wide SVM classification on reflectance at 50nm-intervals in wavelength (Figure \ref{supfig:classif-svm-refl}), with relative importance computed for every machine.}
+	\label{supfig:importance-svm-refl}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/importance_lda_pca"}
+	\caption{Sensitivity analyses of the different input variables in the archipelago-wide LDA classification on reflectance at 50nm-intervals in wavelength (Figure \ref{supfig:classif-lda-refl}), with relative importance computed for every machine.}
+	\label{supfig:importance-lda-refl}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{"../analyses/10-distances/figure_distances2"}
+	\caption{Spatial scale of between-habitat differences in dewlap coloration.
+For each variable and each pair of habitats where significant differences were detected (Figure \ref{fig:anova}), we performed multiple post hoc pairwise comparisons between the sites involved (Figure \ref{supfig:map}, Table \ref{suptab:sites}), using nonparametric Wilcoxon-Mann-Whitney tests.
+Here we report, for each pair of habitats, the distances between sites that significantly differed in dewlap coloration at an error rate of 0.05 (P-values corrected with the Benjamini-Hochberg procedure for multiple testing).}
+	\label{supfig:distances}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{"../analyses/09-parallelism/figure_contrasts"}
+	\caption{Test of parallel divergence between islands.
+Differences in habitat-means, or contrasts, are shown for two pairs of habitats for each principal component on each island, rescaled so the standard deviation of the means along each principal component is one.
+The contrasts represent the patterns of between-habtiat variation on each island, for a given principal component.
+The absence of clustering of islands by variable indicates that islands differ in their between-habitat divergence patterns.
+This is confirmed by a non-significant MANOVA test of the between versus within-variable variance in contrasts.}
+	\label{supfig:contrasts}
+\end{figure}

### 2020-05-24 ms/sections/tables.tex
working with tables

\begin{table}
	\label{tab:anova}
	\caption{}
	\centering
	\input{"../analyses/07-ANOVA/table_anova.tex"}
\end{table}

### 2020-05-24 R/save_table.R

@@ -2,13 +2,13 @@
 #'
 #' @param tab The table
 #' @param nm File name (without the extension)
-#' @param digits,col_names Parameters for `knitr::kable`
+#' @param ... Extra parameters for `knitr::kable`
 #'
 #' @return Nothing. Just saves.
 #'
 #' @export
 
-save_table <- function(tab, nm, digits = 0, col_names = NULL) {
+save_table <- function(tab, nm, ...) {
 
   library(knitr)
 
@@ -18,7 +18,7 @@ save_table <- function(tab, nm, digits = 0, col_names = NULL) {
   # Save LaTeX table
   if (is.null(col_names)) col_names <- NA
   tab <- knitr::kable(
-    tab, "latex", digits = digits, booktabs = TRUE, col.names = col_names
+    tab, "latex", booktabs = TRUE, linesep = "", escape = FALSE, ...
   )
   texfile <- file(paste0(nm, ".tex"))
   writeLines(tab, texfile)

### 2020-05-24 analyses/03-PCA/PCA.R

@@ -4,7 +4,6 @@ rm(list = ls())
 
 library(nmgc)
 library(tidyverse)
-library(knitr)
 
 # Here we are checking first the correlation between PC1 and brightness
 # and second, the variance explained by each PC on each island
@@ -50,11 +49,14 @@ res_cor <- res_cor %>%
 
 # Save table
 t1 <- res_cor
-t1 <- latex %>%
+t1 <- t1 %>%
   add_signif() %>%
   mutate(pvalue = ifelse(pvalue < 0.0001, "< 0.0001", pvalue))
-colnames(t1)[colnames(t1) == "signif"] <- ""
-save_table(t1, "table_brightness", digits = c(0, 3, 0))
+col.names <- c("Island", "$R^2$", "$P$", "")
+save_table(
+  t1, "analyses/03-PCA/table_brightness", digits = c(0, 3, 0),
+  col.names = col.names, align = "lrrl"
+)
 
 
 # Plot the correlation
@@ -104,4 +106,6 @@ expvar <- expvar %>% mutate(total = expvar %>% dplyr::select(PC1:PC4) %>% apply(
 
 # Save table
 t2 <- expvar
-save_table(t2, "table_expvar", digits = c(0, 3, 3, 3, 3, 3))
+col.names <- c("Island", "PC1", "PC2", "PC3", "PC4", "Total")
+fname <- "analyses/03-PCA/table_expvar"
+save_table(t2, fname, digits = c(0, 3, 3, 3, 3, 3), col.names = col.names)

### 2020-05-24 analyses/05-assumptions/Assumptions.R

@@ -25,10 +25,21 @@ t3 <- test_multinorm(data, variables, grouping = "habitat", univariate = TRUE)[,
 # 1.4. Heterogeneity of variances
 t4 <- test_covariance(data, variables, grouping = "habitat", univariate = TRUE)[, -1]
 
-save_table(t1, "table_multinorm_pooled", digits = c(0, 2, 0, 0))
-save_table(t2, "table_covariance_pooled", digits = c(1, 0, 4, 0))
-save_table(t3, "table_normality_pooled", digits = c(0, 0, 3, 4, 0))
-save_table(t4, "table_variance_pooled", digits = c(0, 2, 0, 4))
+t1_names <- c("Habitat", "$HZ$", "$P$", "")
+t2_names <- c("$\\chi^2$", "df", "$P$", "")
+t3_names <- c("Habitat", "Variable", "$W$", "$P$", "")
+t4_names <- c("Variable", "$K^2$", "df", "$P$", "")
+
+fname <- "analyses/05-assumptions/table_%s"
+t1_fname <- sprintf(fname, "multinorm_pooled")
+t2_fname <- sprintf(fname, "covariance_pooled")
+t3_fname <- sprintf(fname, "normality_pooled")
+t4_fname <- sprintf(fname, "variance_pooled")
+
+save_table(t1, t1_fname, digits = c(0, 2, 0, 0), col.names = t1_names)
+save_table(t2, t2_fname, digits = c(1, 0, 4, 0), col.names = t2_names)
+save_table(t3, t3_fname, digits = c(0, 0, 3, 4, 0), col.names = t3_names)
+save_table(t4, t4_fname, digits = c(0, 2, 0, 4), col.names = t4_names)
 
 #### 2. Within islands ####
 
@@ -38,8 +49,14 @@ t5 <- test_multinorm(data, variables, grouping = "habitat", nesting = "island")
 # 2.2. Homogeneity of covariance matrices
 t6 <- test_covariance(data, variables, grouping = "habitat", nesting = "island")
 
-save_table(t5, "table_multinorm", digits = c(0, 0, 2, 4, 0))
-save_table(t6, "table_covariance", digits = c(0, 1, 0, 4, 0))
+t5_names <- c("Island", "Habitat", "$HZ$", "$P$", "")
+t6_names <- c("Island", "$\\chi^2$", "df", "$P$", "")
+
+t5_fname <- sprintf(fname, "multinorm")
[…]

### 2020-05-24 analyses/07-ANOVA/table_posthoc.tex
regenerated tables

@@
 \toprule
-nesting & variable & test & contrast1 & contrast2 & statistic & pvalue & signif\\
+Island & Variable & Test & Contrast &  & Statistic & $P$ & \\
 \midrule
@@
 Abaco & PC3 & tukey & coastal & mangrove & 7.52 & 0.0000 & ***\\
-\addlinespace
 Abaco & PC3 & tukey & coppice & mangrove & 2.55 & 0.1746 & \\
@@
 Bimini & PC2 & tukey & coastal & coppice & 2.10 & 0.3044 & \\
-\addlinespace
 Bimini & PC2 & tukey & coastal & mangrove & 3.91 & 0.0200 & *\\
 Bimini & PC2 & tukey & coppice & mangrove & 1.44 & 0.5674 & \\
-Bimini & PC3 & dunnett & coastal & coppice & 3.59 & 0.0067 & **\\
+Bimini & PC3 & dunnett & coastal & coppice & 3.59 & 0.0065 & **\\
 Bimini & PC3 & dunnett & coastal & mangrove & -0.21 & 0.9954 & \\
 Bimini & PC3 & dunnett & coppice & mangrove & -3.03 & 0.0162 & *\\
-\addlinespace
 Cayman Brac & PC1 & dunnett & coastal & coppice & -2.93 & 0.0225 & *\\
@@
 Cayman Brac & PC2 & tukey & coastal & mangrove & 4.15 & 0.0140 & *\\
-\addlinespace
 Cayman Brac & PC2 & tukey & coppice & mangrove & 1.98 & 0.3476 & \\
@@
 Little Cayman & PC1 & tukey & coastal & coppice & -4.08 & 0.0167 & *\\
-\addlinespace
 Little Cayman & PC1 & tukey & coastal & mangrove & -1.51 & 0.5389 & \\
@@
 Little Cayman & PC2 & dunnett & coppice & mangrove & -6.99 & 0.0000 & ***\\
-\addlinespace
 Little Cayman & PC3 & tukey & coastal & coppice & -4.28 & 0.0114 & *\\
@@
 Long Island & PC3 & tukey & coastal & mangrove & 4.19 & 0.0127 & *\\
-\addlinespace
 Long Island & PC3 & tukey & coppice & mangrove & 0.47 & 0.9408 & \\

### 2020-05-24 analyses/07-ANOVA/table_postkw.tex
regenerated tables

@@
 \toprule
-island & variable & test & contrast1 & contrast2 & statistic & pvalue & signif\\
+Island & Variable & Test & Contrast &  & Statistic & $P$ & \\
 \midrule
@@
 Bimini & PC1 & nemenyi & coastal & mangrove & 3.42 & 0.0411 & *\\
-\addlinespace
 Bimini & PC1 & nemenyi & coppice & mangrove & 0.60 & 0.9048 & \\
@@
 Little Cayman & PC3 & nemenyi & coastal & coppice & 3.73 & 0.0227 & *\\
-\addlinespace
 Little Cayman & PC3 & nemenyi & coastal & mangrove & 6.23 & 0.0000 & ***\\

### 2020-05-24 R/save_table.R

@@ -16,7 +16,6 @@ save_table <- function(tab, nm, ...) {
   write.csv(tab, paste0(nm, ".csv"), row.names = FALSE)
 
   # Save LaTeX table
-  if (is.null(col_names)) col_names <- NA
   tab <- knitr::kable(
     tab, "latex", booktabs = TRUE, linesep = "", escape = FALSE, ...
   )

### 2020-05-24 analyses/04-machine learning/MachineLearning.R

@@ -3,7 +3,6 @@ rm(list = ls())
 library(nmgc)
 library(tidyverse)
 library(cowplot)
-library(knitr)
 
 data <- read.csv("data/reflectance.csv", header = TRUE)
 
@@ -44,19 +43,12 @@ p <- plot_classif(res, facets = "island")
 plot_grid(p, plot_grid(blank, confplot, blank, nrow = 3), ncol = 2, rel_widths = c(2, 1))
 ggsave("classif_svm_pca.png", height = 6, width = 9, dpi = 300)
 
-write.csv(res$mean, "table_classif_svm_pca.csv", row.names = FALSE)
+fname <- "analyses/04-machine learning/table_classif_%s"
+colnames <- c("Island", "Accuracy", "N", "$p_{\\mbox{test}}$", "$n_{\\mbox{test}}$", "$P$", "")
 
-# Latex table
-latex <- res$mean
-head(latex)
-latex <- latex %>%
-  mutate(signif = ifelse(pvalue < 0.05, "*", "")) %>%
-  mutate(signif = ifelse(pvalue < 0.01, "**", signif))
-colnames(latex)[colnames(latex) == "signif"] <- ""
-latex <- kable(latex, digits = c(0, 3, 0, 1, 0, 4), format = "latex")
-texfile <- file("table_classif_svm_pca.tex")
-writeLines(latex, texfile)
-close(texfile)
+t1 <- res$mean
+t1_fname <- sprintf(fname, "svm_pca")
+save_table(t1, t1_fname, digits = c(0, 3, 0, 1, 0, 4, 0), col.names = colnames)
 
 res$imp %>%
   gather_("variable", "importance", variables) %>%
@@ -82,19 +74,9 @@ p <- plot_classif(res, facets = "island", fill = "coral")
 plot_grid(p, plot_grid(blank, confplot, blank, nrow = 3), ncol = 2, rel_widths = c(2, 1))
 ggsave("classif_lda_pca.png", height = 6, width = 9, dpi = 300)
 
-write.csv(res$mean, "table_classif_lda_pca.csv", row.names = FALSE)
-
-# Latex table
-latex <- res$mean
-head(latex)
-latex <- latex %>%
-  add_signif() %>%
-  mutate(pvalue = ifelse(pvalue < 0.0001, "< 0.0001", pvalue))
-colnames(latex)[colnames(latex) == "signif"] <- ""
[…]

### 2020-05-24 analyses/05-assumptions/Assumptions.R

@@ -67,10 +67,12 @@ t7 <- t7 %>%
   nest() %>%
   right_join(t5) %>%
   filter(pvalue < 0.05) %>%
-  select(island, habitat, data) %>%
+  dplyr::select(island, habitat, data) %>%
   unnest(cols = c(data))
 
-save_table(t7, "table_normality", digits = c(0, 0, 0, 3, 4, 0))
+t7_names <- c("Island", "Habitat", "Variable", "$W$", "$P$", "")
+t7_fname <- sprintf(fname, "normality")
+save_table(t7, t7_fname, digits = c(0, 0, 0, 3, 4, 0), col.names = t7_names)
 
 # 2.4. Heterogeneity of variances
 t8 <- test_covariance(data, variables, grouping = "habitat", nesting = "island", univariate = TRUE)
@@ -82,10 +84,12 @@ t8 <- t8 %>%
   nest() %>%
   right_join(t6) %>%
   filter(pvalue < 0.05) %>%
-  select(island, data) %>%
+  dplyr::select(island, data) %>%
   unnest(cols = c(data))
 
-save_table(t8, "table_homoskedasticity", digits = c(0, 0, 3, 0, 4, 0))
+t8_names <- c("Island", "Variable", "$K^2$", "df", "$P$", "")
+t8_fname <- sprintf(fname, "homoskedasticity")
+save_table(t8, t8_fname, digits = c(0, 0, 0, 3, 4, 0), col.names = t8_names)
 
 # 3. Outliers
 

### 2020-05-24 analyses/06-pooled/Pooled.R

@@ -13,7 +13,6 @@ wl <- paste0("wl", 300:700)
 variables <- paste0("PC", 1:4)
 
 data <- cbind(data, data.frame(npcomp(data, wl)$x)[, variables])
-data <- cbind(data, pca)
 
 #### 1. Eyeball ####
 
@@ -74,10 +73,10 @@ etasq(res_manova)
 # 2.3. Semi-parametric MANOVA
 
 # We know, however that there are deviations from multivariate normality...
-read.csv("analyses/assumptions/table_multinorm_pooled.csv", header = TRUE)
+read.csv("analyses/05-assumptions/table_multinorm_pooled.csv", header = TRUE)
 
 # and homogeneity of covariance matrices
-read.csv("analyses/assumptions/table_covariance_pooled.csv", header = TRUE)
+read.csv("analyses/05-assumptions/table_covariance_pooled.csv", header = TRUE)
 
 # So we repeat the analysis with a semi-parametric MANOVA
 
@@ -93,10 +92,10 @@ res_manova2
 # But first we assess normality and homogeneity of variances
 
 # There is evidence for non-normality for all variables ...
-read.csv("analyses/assumptions/table_normality_pooled.csv", header = TRUE)
+read.csv("analyses/05-assumptions/table_normality_pooled.csv", header = TRUE)
 
 # and no evidence for heterogeneity of variance
-read.csv("analyses/assumptions/table_variance_pooled.csv", header = TRUE)
+read.csv("analyses/05-assumptions/table_variance_pooled.csv", header = TRUE)
 
 # 3.1 Univariate ANOVAs
 
@@ -107,14 +106,21 @@ res_nanova <- nanova(
   parametric = TRUE
 )
 t1 <- res_nanova$res[, -1]
-save_table(t1, "table_anova_pooled", digits = c(0, 0, 0, 1, 1, 3, 0, 1, 2, 4, 0))
+t1_fname <- "analyses/06-pooled/table_anova_pooled"
+t1_names <- c(
+  "Variable", "Best fit", "df", "AICc", "$\\Delta$AICc", "AICcw",
+  "df$_{\\mbox{LRT}}$", "Log-lik.", "$\\chi^2$", "$P$", ""
+)
+save_table(t1, t1_fname, digits = c(0, 0, 0, 1, 1, 3, 0, 1, 2, 4, 0), col.names = t1_names)
 
[…]

### 2020-05-24 analyses/07-ANOVA/ANOVA.R

@@ -44,7 +44,9 @@ res_smanova <- nanova(
 )
 
 t1 <- res_smanova$res
-save_table(t1, "table_smanova", digits = c(0, 1, 4, 0))
+t1_fname <- "analyses/07-ANOVA/table_smanova"
+t1_names <- c("Island", "MATS", "$P$", "")
+save_table(t1, t1_fname, digits = c(0, 1, 4, 0), col.names = t1_names)
 
 #### 2. Multiple ANOVAs on significant islands ####
 
@@ -64,10 +66,17 @@ res_anova <- nanova(
 # colinear with our grouping variable
 
 t2 <- res_anova$res
-save_table(t2, "table_anova", digits = c(0, 0, 0, 0, 1, 1, 3, 0, 1, 2, 4, 0))
+t2_fname <- "analyses/07-ANOVA/table_anova"
+t2_names <- c(
+  "Island", "Variable", "Best fit", "df", "AICc", "$\\Delta$AICc", "AICcw",
+  "df$_{\\mbox{LRT}}$", "Log-lik.", "$\\chi^2$", "$P$", ""
+)
+save_table(t2, t2_fname, digits = c(0, 0, 0, 0, 1, 1, 3, 0, 1, 2, 4, 0), col.names = t2_names)
 
 t2. <- res_anova$ph
-save_table(t2., "table_posthoc", digits = c(0, 0, 0, 0, 0, 2, 4, 0))
+t2._fname <- "analyses/07-ANOVA/table_posthoc"
+t2._names <- c("Island", "Variable", "Test", "Contrast", "", "Statistic", "$P$", "")
+save_table(t2., t2._fname, digits = c(0, 0, 0, 0, 0, 2, 4, 0), col.names = t2._names)
 
 # 2.2. Kruskal-Wallis tests wherever normality is not met
 
@@ -92,7 +101,9 @@ t3 <- res_kw$res %>%
   select(-dev) %>%
   drop_na()
 
-save_table(t3, "table_kw", digits = c(0, 0, 2, 0, 4, 0))
+t3_fname <- "analyses/07-ANOVA/table_kw"
+t3_names <- c("Island", "Variable", "$\\chi^2$", "df", "$P$", "")
+save_table(t3, t3_fname, digits = c(0, 0, 2, 0, 4, 0), col.names = t3_names)
 
 t3. <- res_kw$ph %>%
   rename(island = "nesting") %>%
[…]

### 2020-05-24 analyses/08-spatial correlation/SpatialAutocorrelation.R

@@ -16,12 +16,12 @@ res_cortest <- nspcortest(
 # Save table
 t1 <- res_cortest$res
 t1 <- t1 %>% add_signif()
-save_table(t1, "table_spatial", digits = c(0, 3, 3, 0))
+t1_fname <- "analyses/08-spatial correlation/table_spatial"
+t1_names <- c("Island", "$r^2$", "$P$", "N", "")
+save_table(t1, t1_fname, digits = c(0, 3, 3, 0), col.names = t1_names)
 
 # Table with multivariate means per site
-res_cortest$sites
-write.csv(res_cortest$sites, "table_sites.csv", row.names = FALSE)
-
-# Make a Latex table
 t2 <- res_cortest$sites
-save_table(t2, "table_sites", digits = c(0, 4, 4, 0, 4, 4, 4, 4))
+t2_fname <- "analyses/08-spatial correlation/table_sites"
+t2_names <- c("Island", "Longitude", "Latitude", "Habitat", "PC1", "PC2", "PC3", "PC4")
+save_table(t2, t2_fname, digits = c(0, 4, 4, 0, 4, 4, 4, 4), col.names = t2_names)

### 2020-05-24 analyses/10-distances/Distances.R

@@ -1,16 +1,16 @@
-# Welcome to hell...
-
 rm(list = ls())
 
 # Here we explore the distances at which differences between sites are detected
 
 library(tidyverse)
 library(nmgc)
+library(RColorBrewer)
 
 data <- read.csv("data/reflectance.csv", header = TRUE)
 wl <- paste0("wl", 300:700)
 variables <- paste0("PC", 1:4)
 data <- cbind(data, npcomp(data, wl, nesting = "island", reduce = 1:4)$x)
+data <- data[, -grep("island", colnames(data))[2]]
 
 # Kepp only the most significant islands
 keep <- c("Abaco", "Bimini", "Cayman Brac", "Little Cayman", "Long Island")
@@ -141,4 +141,6 @@ label <- sprintf(label, res_spearman$rho, res_spearman$pvalue)
 t1 <- res_wilcox %>% select(-U, -pvalue)
 t1 <- t1 %>% bind_cols(res_wilcox %>% ungroup %>% select(U, pvalue))
 t1 <- t1 %>% add_signif()
-save_table(t1, "table_wilcoxon", digits = c(0, 0, 0, 0, 4, 4, 4, 4, 0, 0, 4))
+t1_fname <- "analyses/10-distances/table_wilcoxon"
+t1_names <- c("Island", "Variable", "Contrast", "", "Lon. 1", "Lat. 1", "Lon. 2", "Lat. 2", "Distance (m)", "$U$", "$P$", "")
+save_table(t1, t1_fname, digits = c(0, 0, 0, 0, 4, 4, 4, 4, 0, 0, 4), col.names = t1_names)

### 2020-05-24 ms/sections/suppfigures.tex
tables and figures look good

@@
 	\centering
-	\includegraphics{"../analyses/03-PCA/figure_brightness"}
+	\includegraphics[width=0.8\textwidth]{"../analyses/03-PCA/figure_brightness"}
 	\caption{Correlation between dewlap brightness (as measured by the mean reflectance from 300 to 700nm in wavelength) and PC1 score for each island.
@@
 	\centering
-	\includegraphics{"../analyses/03-PCA/figure_brightness_pooled"}
+	\includegraphics[width=0.5\textwidth]{"../analyses/03-PCA/figure_brightness_pooled"}
 	\caption{Correlation between dewlap brightness (as measured by the mean reflectance from 300 to 700nm in wavelength) and PC1 score across the whole archipelago.
@@
 	\centering
-	\includegraphics[width=\textwidth]{"../analyses/02-reflectance/figure_reflectance"}
+	\includegraphics[width=0.8\textwidth]{"../analyses/02-reflectance/figure_reflectance"}
 	\caption{5-95th percentile range of lizard dewlap reflectance values (in \% of incoming light) across wavelengths for each island and each habitat.}
@@
 	\centering
-	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_svm_pca_pooled"}
+	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/classif_svm_pca_pooled"}
 	\caption{Archipelago-wide SVM classification accuracy based on principal component data.
@@
 	\centering
-	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_svm_refl_pooled"}
+	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/classif_svm_refl_pooled"}
 	\caption{Archipelago-wide SVM classification accuracy based on reflectance data at 50nm-intervals in wavelength (see Methods).
@@
 	\centering
-	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/importance_svm_pca_pooled"}
+	\includegraphics[width=0.4\textwidth]{"../analyses/04-machine learning/plots/importance_svm_pca_pooled"}
 	\caption{Sensitivity analyses of the different input variables in the archipelago-wide SVM classification on principal component data (Figure \ref{supfig:classif-svm-pca-pooled}), with relative importance computed for every machine.}
@@
 	\centering
-	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/importance_svm_refl_pooled"}
+	\includegraphics[width=0.6\textwidth]{"../analyses/04-machine learning/plots/importance_svm_refl_pooled"}
 	\caption{Sensitivity analyses of the different input variables in the archipelago-wide SVM classification on reflectance data at 50nm-intervals in wavelength (Figure \ref{supfig:classif-svm-refl-pooled}), with relative importance computed for every machine.}
@@
 	\centering
-	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_lda_pca_pooled"}
+	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/classif_lda_pca_pooled"}
 	\caption{Archipelago-wide LDA classification accuracy based on principal component data.
@@
 	\centering
-	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_lda_refl_pooled"}
+	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/classif_lda_refl_pooled"}
 	\caption{Archipelago-wide LDA classification accuracy based on reflectance data at 50nm-intervals in wavelength (see Methods).
@@
 	\centering
-	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/importance_svm_pca"}
+	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/importance_svm_pca"}
 	\caption{Sensitivity analyses of the different input variables in the within-island SVM classification on principal component data (Figure \ref{supfig:classif-svm-pca}), with relative importance computed for every machine.}
@@
 	\centering
-	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/importance_lda_pca"}
+	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/importance_lda_pca"}
 	\caption{Sensitivity analyses of the different input variables in the within-island LDA classification on principal component data (Figure \ref{supfig:classif-lda-pca}), with relative importance computed for every machine.}
@@
 	\centering
-	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/importance_svm_pca"}
+	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/importance_svm_pca"}
 	\caption{Sensitivity analyses of the different input variables in the archipelago-wide SVM classification on reflectance at 50nm-intervals in wavelength (Figure \ref{supfig:classif-svm-refl}), with relative importance computed for every machine.}
@@
 	\centering
-	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/importance_lda_pca"}
+	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/importance_lda_pca"}
 	\caption{Sensitivity analyses of the different input variables in the archipelago-wide LDA classification on reflectance at 50nm-intervals in wavelength (Figure \ref{supfig:classif-lda-refl}), with relative importance computed for every machine.}
@@
 	\centering
-	\includegraphics[width=\textwidth]{"../analyses/09-parallelism/figure_contrasts"}
+	\includegraphics[width=0.7\textwidth]{"../analyses/09-parallelism/figure_contrasts"}
 	\caption{Test of parallel divergence between islands.

### 2020-05-24 analyses/03-PCA/PCA.R

@@ -2,6 +2,7 @@ rm(list = ls())
 
 # Exploration of the principal component analysis
 
+library(dewlap)
 library(nmgc)
 library(tidyverse)
 
@@ -52,7 +53,7 @@ t1 <- res_cor
 t1 <- t1 %>%
   add_signif() %>%
   mutate(pvalue = ifelse(pvalue < 0.0001, "< 0.0001", pvalue))
-col.names <- c("Island", "$R^2$", "$P$", "")
+col.names <- c("Island", "$r^2$", "$P$", "")
 save_table(
   t1, "analyses/03-PCA/table_brightness", digits = c(0, 3, 0),
   col.names = col.names, align = "lrrl"
@@ -76,7 +77,7 @@ res_cor <- res_cor %>%
   mutate(plabel = ifelse(pvalue < 0.05, str_replace(plabel, "$", "*"), plabel))
 p + geom_text(
   data = res_cor %>% filter(island != "Archipelago"),
-  aes(label = paste0("R\U000B2 = ", r2, "\n", plabel)), x = 40, y = -25, hjust = 1
+  aes(label = paste0("r\U000B2 = ", r2, "\n", plabel)), x = 40, y = -25, hjust = 1
 )
 
 ggsave("figure_brightness.png", width = 6, height = 6, dpi = 300)
@@ -90,7 +91,7 @@ data %>%
   xlab("Brightness (mean reflectance, %)") +
   geom_text(
     data = res_cor %>% filter(island == "Archipelago"),
-    aes(label = paste0("R\U000B2 = ", r2, "\n", plabel)), x = 40, y = -25, hjust = 1
+    aes(label = paste0("r\U000B2 = ", r2, "\n", plabel)), x = 40, y = -25, hjust = 1
   )
 
 ggsave("figure_brightness_pooled.png", width = 4, height = 4, dpi = 300)

### 2020-05-24 analyses/04-machine learning/MachineLearning.R

@@ -1,5 +1,6 @@
 rm(list = ls())
 
+library(dewlap)
 library(nmgc)
 library(tidyverse)
 library(cowplot)
@@ -44,7 +45,7 @@ plot_grid(p, plot_grid(blank, confplot, blank, nrow = 3), ncol = 2, rel_widths =
 ggsave("classif_svm_pca.png", height = 6, width = 9, dpi = 300)
 
 fname <- "analyses/04-machine learning/table_classif_%s"
-colnames <- c("Island", "Accuracy", "N", "$p_{\\mbox{test}}$", "$n_{\\mbox{test}}$", "$P$", "")
+colnames <- c("Island", "Accuracy", "$N$", "$p_{\\mbox{test}}$", "$n_{\\mbox{test}}$", "$P$", "")
 
 t1 <- res$mean
 t1_fname <- sprintf(fname, "svm_pca")

### 2020-05-24 analyses/06-pooled/Pooled.R

@@ -2,6 +2,7 @@ rm(list = ls())
 
 # Here we explore dewlap coloration between habitats across the whole archipelago
 
+library(dewlap)
 library(nmgc)
 library(GGally)
 library(MANOVA.RM)

### 2020-05-24 analyses/08-spatial correlation/SpatialAutocorrelation.R

@@ -2,6 +2,7 @@
 
 rm(list = ls())
 
+library(dewlap)
 library(tidyverse)
 library(nmgc)
 
@@ -17,7 +18,7 @@ res_cortest <- nspcortest(
 t1 <- res_cortest$res
 t1 <- t1 %>% add_signif()
 t1_fname <- "analyses/08-spatial correlation/table_spatial"
-t1_names <- c("Island", "$r^2$", "$P$", "N", "")
+t1_names <- c("Island", "$r$", "$P$", "$N$", "")
 save_table(t1, t1_fname, digits = c(0, 3, 3, 0), col.names = t1_names)
 
 # Table with multivariate means per site

### 2020-05-24 analyses/09-parallelism/Parallelism.R

@@ -2,6 +2,7 @@ rm(list = ls())
 
 # Analysis of parallel divergence across islands
 
+library(dewlap)
 library(tidyverse)
 library(nmgc)
 library(RColorBrewer)

### 2020-05-24 analyses/10-distances/Distances.R

@@ -2,6 +2,7 @@ rm(list = ls())
 
 # Here we explore the distances at which differences between sites are detected
 
+library(dewlap)
 library(tidyverse)
 library(nmgc)
 library(RColorBrewer)

### 2020-05-25 ms/sections/discussion.tex
nearly there

@@
 However, the habitat-specific variation in dewlaps was not consistent between these islands.
-Although those results are consistent with adaptation at a very local scale, other evolutionary drivers could be at work, including phenotypic plasticity, random drift, or historical contingency including multiple colonization events.
+Although those results are consistent with adaptation at a very local scale, other evolutionary drivers could be at work, including phenotypic plasticity, random drift, or historical contingency, including multiple colonization events.
 We reject this last explanation because all of the island populations in this study are strictly monophyletic, implying a single colonization event per island (van de Schoot, unpublished thesis; \citealt{Driessens2017}).
 While random drift cannot be completely ruled out, we see little evidence for a role of phenotypic isolation-by-distance (spatial autocorrelation) in explaining the differences we report.
-While we found a significant signal of isolation-by-distance on Eleuthera, we did not detect differences in dewlap coloration between habitats on this island.
+We did detect a significant signal of isolation-by-distance on Eleuthera, but we did not find differences in dewlap coloration between habitats on this island.
+
+% The maintenance of divergence in the face of gene flow
+
+\paragraph{Divergence is maintained in the face of gene flow} We detected differences between habitats at relatively small spatial scales, most of the time between sites 5-10km apart, and sometimes a few hundred meters away.
+This, together with the lack of geographic barriers within islands and the high mobility of \textit{A.
+sagrei} individuals \citep{Kamath2018}, imply ample opportunity for extensive gene flow between our study populations.
+In agreement with this, populations from different habitats were found to not be monophyletic based on <insert marker name here> (van de Schoot et al.
+unpublished thesis).
+The maintenance of phenotypic differences at such a small spatial scales, in the face of gene flow, may support an adaptive explanation \citep{Richardson2014}.\\
+
+The persistence of dewlap color divergence despite gene flow has been documented in multiple species of anoles, sometimes in relation to environmental conditions.
+\citet{Ng2012} and \citet{Ng2016} found divergent dewlap coloration in the face of gene flow between subspecies of \textit{A.
+distichus} across Hispaniola, and proposed this as a mechanism of reproductive isolation in the early stages of speciation \citep{Ng2011, Lambert2013, Ng2017}.
+\citet{Stapley2011} found that dewlap color polymorphism was maintained in the absence of genetic structure between populations of \textit{A.
+apletophallus} from Panama.
+\citet{Thorpe2002a} found that divergence in dewlap coloration matched habitat-type better than mitochondrial lineage in \textit{A.
+roquet} on Martinique, and a convergent pattern was found in \textit{A.
+trinitatis} on the featureless island of St Vincent \citep{Thorpe2002b}.
+Finally, divergence in body coloration, but not dewlap coloration, was also reported in \textit{A.
+conspersus} on another small island, Grand Cayman \citep{Macedonia2001}.
+Altogether, these results are consistent with a generally high adaptive potential of dewlap coloration in anoles over small distances, which could apply to $A.
+sagrei$.
+
+% Could there be local adaptation? To what?
+
+\paragraph{Dewlap coloration could be locally adapted} Presently, we do not know the adaptive drivers of dewlap color divergence observed in this study.
+Some degree of parallel evolution is usually a good indicator for an adaptive process \citep{Losos2011}, and convergent patterns of dewlap color evolution in similar environments across islands and species have been documented in other species \citep{Thorpe2002a, Thorpe2002b}.
+Here, however, we found little consistency in between-habitat divergence across the nine islands studied; except maybe for dewlap brightness being lower in the primary coppice habitat on three islands, and the primary coppice habitat being intermediate in coloration between mangrove and beach scrub on two islands.
+Those patterns could reflect adaptation to components of the habitat \citep{Endler1988}, but it is not clear what those components of the habitat might be.\\
+
+Previous studies have suggested that dewlap coloration may have evolved to be maximally detectable under local light conditions imposed by the environment, primarily through UV contrast (i.e.
+UV-brighter dewlaps in UV-dark, mesic habitats and UV-darker dewlaps in UV-bright, xeric habitats), in \textit{A.
+cristatellus} and \textit{A.
+cooki} from Puerto Rico \citep{Leal2002, Leal2004}.
+Although UV reflectance was commonly involved in beetween-habitat divergence, we found no such patterns in \textit{A.
+sagrei}, where instead, we found the darkest dewlaps in the darkest, mesic habitat -- primary coppice forest -- on three islands, and dewlaps often differed the most between beach scrub and mangrove forest, two xeric habitats with similar, high irradiance levels \citep{Howard1950, Schoener1968}.
+The inconsistent and idiosyncratic patterns we observed suggest that dewlap color variation between habitats cannot be predicted by habitat identity alone.
+Studies of Jamaican and Hispaniolan anoles similarly found between-habitat differences in dewlap coloration but no evidence for higher detectability \citep{Fleishman2009, Ng2012}.
+Habitats on different islands may also differ in other aspects than light conditions such as densities of predators or other anole species, which have been shown to affect among-island dewlap diversity \citep{Vanhooydonck2009, Baeckens2018}.
+In particular, \citet{Baeckens2018} recently showed that dewlaps with spotted patterns occurred more often in \textit{A.
+sagrei} on islands with more coexisting species of anoles.
+Therefore, dewlap coloration in $A.
+sagrei$ may be under the influence of a mosaic of interacting local environmental factors, which need not be the same across islands.
 
@@
 Animals lack the ability to synthesize carotenoids, and those must therefore be found in the diet, while pterins are synthesized from nucleotides \citep{Goodwin1984, Hill2002, Hill2006}.
-Experimental manipulation of dietary carotenoid content showed no effect on dewlap coloration in \textit{A.
+However, experimental manipulation of dietary carotenoid content showed no effect on dewlap coloration in \textit{A.
 sagrei} \citep{Steffen2010} nor in \textit{A.
-distichus} \citep{Ng2013}, another species with an orange-based dewlap.
-Plasticity due to differences in development (e.g.
-egg rearing conditions) is unlikely because dewlap coloration develops at sexual maturity \citep{Ng2013}.
-\citet{Cox2017} further found a high degree of heritability of dewlap coloration in \textit{A.
+distichus} \citep{Ng2013}, which also has an orange-based dewlap.
+This makes a plastic response to differences in diet across habitats unlikely.
+Developmental plasticity (e.g.
+due to differences egg rearing conditions) is also unlikely because dewlap coloration develops at sexual maturity in anoles \citep{Ng2013}.
+The differences we observed could therefore be heritable.
+This hypothesis is further supported by \citet{Cox2017}, who found a high degree of heritability of dewlap coloration in \textit{A.
 sagrei}.
-These studies suggest that dewlap coloration is not a plastic trait, although transgenerational plastic effects cannot be completely ruled out by these one or two-generation common garden experiments \citep{Tariel2020}.
-That leaves an adaptive explanation, where dewlap color could be under differential natural and/or sexual selection in these different habitats.
-
-% The maintenance of divergence in the face of gene flow
-
-\paragraph{Divergence is maintained in the face of gene flow} The small spatial scale (the median distance between island sample sites was 11.2km + mention the minimum distance at which we found differences) and the lack of geographic barriers within islands, together with the high mobility \textit{A.
-sagrei} individuals \citep{Kamath2018}, imply ample opportunity for extensive gene flow between the populations in this study.
-In agreement with this, populations from different habitats were found to not be monophyletic based on <insert marker name here> (van de Schoot et al.
-unpublished thesis).
-While gene flow is expected to erase any differences acquired by drift without selection maintaining such polymorphism \citep{Dieckmann1999}, these divergent patterns remain, which may support an adaptive explanation.
-Maintenance of dewlap color divergence despite gene flow has been found in \textit{A.
-distichus} in Hispaniola \citep{Ng2012, Ng2016} and proposed as a mechanism of reproductive isolation in the early stages of speciation \citep{Ng2011, Lambert2013, Ng2017}.
-\citet{Stapley2011} found that dewlap color polymorphism was maintained in the absence of genetic structure between populations of \textit{A.
-apletophallus} from Panama.
-\citet{Thorpe2002a} found that divergence in dewlap coloration matched habitat-type better than mitochondrial lineage in \textit{A.
-roquet} on Martinique, and a convergent pattern was found in \textit{A.
-trinitatis} on the featureless island of St Vincent \citep{Thorpe2002b}.
-Divergence in body coloration, but not dewlap coloration, was also reported in \textit{A.
-conspersus} on another small island, Grand Cayman \citep{Macedonia2001}.
-Those results suggest a high adaptive potential of the anoline dewlap.
+Although most studies used one or two-generation common garden experiments and thus could not rule out transgenerational plastic effects \citep{Tariel2020}, dewlap coloration generally seems to not be a plastic trait.
+This further reinforces an adaptive explanation, where dewlap color could be under differential natural and/or sexual selection in these different habitats.
 
@@
 sagrei} suggest ample opportunity for female mate choice and sexual selection \citep{Kamath2018}.
-A number of studies suggest that characteristics of the dewlap are indicators of male quality and may therefore act as a cue in mate choice, according to the "good genes" model of sexual selection \citep{Andersson1994}.
+A number of studies have suggested that characteristics of the dewlap are indicators of male quality and may therefore act as a cue in mate choice, according to the "good genes" model of sexual selection \citep{Andersson1994}.
 \citet{Cook2013} found lower orange reflectance in dewlaps with heavily parasitized \textit{A.
@@
 \citet{Steffen2014} found that lower UV and orange-red reflectance predict contest-winning success between males.
-\citet{Driessens2015} further found that more yellow and red dewlaps (relative to UV) predict better body condition, and that higher yellow and UV reflectance at the margin of the dewlap predict higher hematocrit (the concentration of red blood cells), indicating a better health.
-Other aspects of the dewlap may be important for sexual selection, for example, dewlap size correlates with bite force \citep{Vanhooydonck2005} and sexual size dimorphism (a proxy for sexual selection; \citealt{Vanhooydonck2009}).
-Even though display frequency increases in the presence of females \citep{Driessens2014}, there seems to be no link between dewlap display frequency and mating success \citep{Tokarz2002, Tokarz2005} or individual quality \citep{Driessens2015}.
+\citet{Driessens2015} further found that more yellow and red dewlaps (relative to UV) predict better body condition, and that higher yellow and UV reflectance at the margin of the dewlap predict higher hematocrit (the concentration of red blood cells), indicating a better health.\\
+
+Other aspects of the dewlap may be important for sexual selection.
+For example, dewlap size in $A.
+sagrei$ was found to correlate with bite force \citep{Vanhooydonck2005} and sexual size dimorphism (a proxy for sexual selection; \citealt{Vanhooydonck2009}).
+Even though display frequency increases in the presence of females \citep{Driessens2014}, no link was found between dewlap display frequency and mating success \citep{Tokarz2002, Tokarz2005} or individual quality \citep{Driessens2015}.
 Our report of inconsistent divergence between islands, however, is at odds with the good-genes model of sexual selection, which would predict the evolution of dewlap coloration in the same direction across the archipelago.
-This is consistent with \citet{Baeckens2018}, who found no link between the average island-dewlap coloration and sexual size dimorphism in \textit{A.
+This is in agreement with \citet{Baeckens2018}, who found no link between the average island-dewlap coloration and sexual size dimorphism in \textit{A.
 sagrei} across the West Indies.
-If sexual selection is at play in our system, it more likely involves divergent female preferences that are not linked to male quality but rather to components of the environment, and may well be arbitrary (Fisherian sexual selection; \citealt{Andersson1994}).
-
-% Could there be local adaptation? To what?
-
-\paragraph{Dewlap coloration could be locally adapted} Presently, we do not know the adaptive drivers of dewlap  color divergence observed in this study.
-Some degree of parallel evolution is usually a good indicator for an adaptive process \citep{Losos2011}, and convergent patterns of dewlap color evolution in similar environments across islands and species have been documented \citep{Thorpe2002a, Thorpe2002b}.
-Here, we found evidence for a few consistencies in within-island divergence across the nine islands studied (show some archipelago-wide results?); dewlap brightness was lower in the primary coppice habitat on three islands, and the primary coppice habitat was intermediate in coloration between mangrove and beach scrub on two islands.
-Those patterns could reflect adaptation to components of the habitat \citep{Endler1988}.
-However, it is not clear what those components of the habitat might be.
-Previous studies have found that dewlap coloration maximizes detectability given the light conditions in the local habitat, primarily through UV contrast -- with UV-brighter dewlaps in UV-dark, mesic habitats and UV-darker dewlaps in UV-bright, xeric habitats -- in \textit{A.
-cristatellus} and \textit{A.
-cooki} on Puerto Rico \citep{Leal2002, Leal2004}.
-We found no such pattern in \textit{A.
-sagrei}, where instead, we found the darkest dewlaps in the darkest, mesic habitat -- primary coppice forest -- on three islands, and dewlaps often differed the most between beach scrub and mangrove forest, two xeric habitats with similar, high irradiance levels \citep{Howard1950, Schoener1968}.
-The inconsistent and idiosyncratic patterns we observed suggest that dewlap color variation between habitats cannot be predicted by habitat identity alone.
-Studies of Jamaican and Hispaniolan anoles similarly found between-habitat differences in dewlap coloration but no evidence for higher detectability \citep{Fleishman2009, Ng2012}.
-Habitats on different islands may also differ in other aspects than light conditions such as densities of predators or other anole species, which have been shown to affect among-island dewlap diversity \citep{Vanhooydonck2009, Baeckens2018}.
-In particular, \citet{Baeckens2018} recently showed that dewlaps with spotted patterns occurred more often in \textit{A.
-sagrei} on islands with more coexisting species of anoles.
+If sexual selection is at play in our system, it more likely involves divergent female preferences that are not linked to male quality but rather to components of the environment, or may well be arbitrary (Fisherian sexual selection; \citealt{Andersson1994}).
 

### 2020-05-25 ms/sections/results.tex
nearly there

@@
 
-The small archipelago-wide effect of habitat-type was detected for PC1, PC2 and PC3 (mixed-effect ANOVA with island as a random effect, Table \ref{suptab:anova_pooled}, backed up by KW), but this effect was too small for post hoc tests to find which habitats differed.
+The small archipelago-wide effect of habitat-type was detected for PC1, PC2 and PC3 (mixed-effect ANOVA with island as a random effect, Table \ref{suptab:anova-pooled}), but this effect was too small for post hoc tests to find which habitats differed.
 Archipelago-wide differences in dewlap coloration between habitats were also detected by SVMs trained on pooled data regardless of island identity, both for PCA data and reflectance scores (Fig.
-\ref{supfig:classif_svm_pca_pooled}, \ref{supfig:classif_svm_refl_pooled}).
+\ref{supfig:classif-svm-pca-pooled}, \ref{supfig:classif-svm-refl-pooled}).
 This pattern seemed to be driven by mangrove lizards being correctly reassigned more often than predicted by chance.
 Sensitivity analyses on these machines suggest a relatively small role of long wavelengths (red reflectance) in driving this pattern (Fig.
-\ref{supfig:importance_svm_refl_pooled}), but did not reveal strong differences between the PCs in relative importance (Fig.
-\ref{supfig:importance_svm_pca_pooled}).
+\ref{supfig:importance-svm-refl-pooled}), but did not reveal strong differences between the PCs in relative importance (Fig.
+\ref{supfig:importance-svm-pca-pooled}).
 Archipelago-wide differences were not detected by LDA classifiers at all (Fig.
-\ref{supfig:classif_lda_pca_pooled}, \ref{supfig:classif_lda_refl_pooled}).
+\ref{supfig:classif-lda-pca-pooled}, \ref{supfig:classif-lda-refl-pooled}).
 So the difference could be in something else than means.\\
@@
 Within islands, SVM classifiers correctly assigned individuals to their habitat of origin based solely upon dewlap coloration on five islands: Abaco, Bimini, Cayman Brac, Little Cayman, and Long island (Fig.
-\ref{fig:classif_svm_pca}).
+\ref{fig:classif-svm-pca}).
 An LDA approach yielded similar success rates (Fig.
-\ref{supfig:classif_lda_pca}), suggesting robust differences between these populations.
-Of the five islands, Little Cayman was the best discriminated with a mean SVM generalization success of 73.4\% (Table \ref{suptab:classif_svm_pca}).
+\ref{supfig:classif-lda-pca}), suggesting robust differences between these populations.
+Of the five islands, Little Cayman was the best discriminated with a mean SVM generalization success of 73.4\% (Table \ref{suptab:classif-svm-pca}).
 The results of the classification analyses on PCA data were very similar to results from SVMs and LDAs trained on reflectance values at 50nm-spaced wavelengths from 300 to 700nm (Fig.
-\ref{supfig:classif_svm_refl} and \ref{supfig:classif_lda_refl}).\\
+\ref{supfig:classif-svm-refl} and \ref{supfig:classif-lda-refl}).\\
 
@@
 Moreover, the differences in dewlaps between habitats were not always consistent between islands, thus, we will discuss the habitat-specific variation in dewlap coloration for each island where significant differences were detected in turn (Fig.
-\ref{fig:boxplots}, Table \ref{tab:anova}, Table KW).
-Figure \ref{fig:boxplots}A provides a key to map principal component scores to the underlying wavelengths.\\
+\ref{fig:anova}, Tables \ref{tab:anova}, \ref{suptab:kruskal}).
+Figure \ref{fig:anova}A provides a key to map principal component scores to the underlying wavelengths.\\
 
@@
 Sensitivity analyses on classifiers suggested an overall higher relative importance for PC2 and PC3 in determining between-group differences on Abaco, both in SVM and LDA classifiers (Fig.
-\ref{supfig:importance_svm_pca}, \ref{supfig:importance_lda_pca}), consistent with our ANOVA results (Fig.
-\ref{fig:boxplots}B).
+\ref{supfig:importance-svm-pca}, \ref{supfig:importance-lda-pca}), consistent with our ANOVA results (Fig.
+\ref{fig:anova}B).
 There was no strong signal of differences in relative importance among principal components on the other islands.
 Sensitivity analyses of SVMs trained on reflectance scores rather than principal components revealed, however, a consistently higher importance of ultraviolet reflectance in between-group differences on all islands (Fig.
-\ref{fig:importance_svm_refl}).
+\ref{supfig:importance-svm-refl}).
 This pattern was not recovered for LDAs trained on reflectance scores (Fig.
-\ref{fig:importance_lda_refl}).\\
+\ref{supfig:importance-lda-refl}).\\
 
@@
 We did not find significant spatial autocorrelation between the sampling sites on the islands where we detected a significant habitat effect.
-We did, however, detect a significant positive signal of autocorrelation on Eleuthera ($P = 0.02$, Table \ref{suptab:autocor}), suggesting possible color differentiation through isolation-by-distance on this island.\\
+We did, however, detect a significant positive signal of autocorrelation on Eleuthera (Table \ref{suptab:autocor}), suggesting possible color differentiation through isolation-by-distance on this island.\\
 
@@
 Some patterns did seem more common however, such as darker dewlaps among coppice lizards (Cayman Brac, Little Cayman, and Long Island, Fig.
-\ref{fig:boxplots}) or the intermediate position of coppice lizards in chromatic color space (Cayman Brac and Long Island).
+\ref{fig:anova}) or the intermediate position of coppice lizards in chromatic color space (Cayman Brac and Long Island).
 In other cases, patterns of differentiation were reversed from one island to another, with more ultraviolet reflecting dewlaps in mangroves than in coastal habitat on Abaco and Cayman Brac, but the opposite on Little Cayman and Long Island.
 Overall, it seemed that patterns of heterogeneity of variance were often driven by higher variances in coloration within beach scrub lizards (Fig.
-\ref{fig:boxplots}, Table \ref{tab:anova}).
+\ref{fig:anova}, Table \ref{tab:anova}).
 Yet other patterns were idiosyncratic, such as the combination of higher red and ultraviolet reflectance in coastal lizards on Little Cayman, where the rule seemed to be a negative correlation between ultraviolet and red reflectance across every other island.\\

### 2020-05-25 ms/sections/results.tex
done?

@@
 \ref{supfig:reflectance}, explained variance $\eta^2 = 44.3$\%, MANOVA approximation).
-Nonetheless, we did find evidence for differences in dewlap coloration between habitat-types, and those were mostly island-specific (habitat-by-island interaction term, MATS = 384.4, P < 0.001, explained variance $\eta^2 = 11.4$\%), leaving a small but significant portion of the variation explained by an archipelago-wide habitat effect (MATS = 42.5, P = .001, $\eta^2 = 4.8$\%).\\
+Nonetheless, we did find evidence for differences in dewlap coloration between habitat-types, and those were mostly island-specific (habitat-by-island interaction term, MATS = 384.4, P < 0.001, explained variance $\eta^2 = 11.4$\%), leaving a small but significant portion of the variation explained by an archipelago-wide habitat effect (MATS = 42.5, P = 0.001, $\eta^2 = 4.8$\%).\\
 
@@
 Besides, for pairs of habitats where significant differences in dewlap coloration were detected along some principal components, comparisons of the actual sampling sites indicate that the most detectable differences involved sites that were mostly 5-10km apart.
-Our most extreme case of significant differences occurred for PC3 between the beach scrub site and a coppice site, separated from each other by a a few hundreds of meters at most on Bimini (Wilcoxon test, Fig.
-\ref{supfig:distances}, Table \ref{suptab:distances}).\\
+Our most extreme case of significant differences occurred for PC3 between the beach scrub site and a coppice site, separated from each other by a a few hundreds of meters at most on Bimini (multiple pairwise Wilcoxon-Mann-Whitney tests, Fig.
+\ref{supfig:distances}).\\
 
@@
 Patterns of differentiation were inconsistent across the five most significant islands.
-Contrasts in principal components between habitats, calculated on pooled data from the whole archipelago, were not similar, for any component, among islands (contrast figure, MANOVA, P > 0.05).
+Contrasts in principal components between habitats, calculated on pooled data from the whole archipelago, were not similar, for any component, among islands (Fig.
+\ref{supfig:contrasts}; MANOVA, Pillai's trace = 0.354, $F(6, 32) = 1$, P = 0.36).
 No pattern of variation was shared by all five significant islands, along any dimension.

### 2020-05-30 ms/sections/discussion.tex
bbl

@@
 conspersus} on another small island, Grand Cayman \citep{Macedonia2001}.
-Altogether, these results are consistent with a generally high adaptive potential of dewlap coloration in anoles over small distances, which could apply to $A.
-sagrei$.
+Altogether, our results align with previous studies suggesting a generally high adaptive potential of dewlap coloration in anoles over small distances, which could apply to \textit{A.
+sagrei}.
 
@@
 cooki} from Puerto Rico \citep{Leal2002, Leal2004}.
-Although UV reflectance was commonly involved in beetween-habitat divergence, we found no such patterns in \textit{A.
+Although UV reflectance was commonly involved in between-habitat divergence, we found no such patterns in \textit{A.
 sagrei}, where instead, we found the darkest dewlaps in the darkest, mesic habitat -- primary coppice forest -- on three islands, and dewlaps often differed the most between beach scrub and mangrove forest, two xeric habitats with similar, high irradiance levels \citep{Howard1950, Schoener1968}.

### 2020-05-30 ms/sections/methods.tex
bbl

@@
 Our sampling design enabled us to test for differences between habitats at a coarse and fine geographical scale.
-The median distance between two localities within an island was $11.18$km, with some islands being sampled at smaller or larger scales (Figure \ref{supfig:map}, Table \ref{suptab:locations}).
+The median distance between two localities within an island was $11.18$km, with some islands being sampled at smaller or larger scales (Figure \ref{supfig:map}, Table \ref{suptab:sites}).
 $80.3$\% of all pairwise distances within islands were below $50$km.
@@
 
-Reflectance curves were smoothened using the R package pavo \citep{Maia2013} as well as custom R functions (where are those?) to results into one reflectance value at each nanometer in wavelength from 300 to 700nm.
-Because neighboring wavelengths are highly collinear in reflectance, we reduced the dimensionality of the data using principal component analysis (PCA), as per \citet{Cuthill1999, Leal2002}.
+Reflectance curves were smoothened using the R package pavo \citep{Maia2013} as well as custom R functions, down to one reflectance value at each nanometer in wavelength from 300 to 700nm.
+Because neighboring wavelengths are highly collinear in reflectance, we reduced the dimensionality of the data using principal component analysis (PCA), as per \citet{Cuthill1999} and \citet{Leal2002}.
 We performed PCA on each island separately and systematically retained the first four principal components (PC), which together always explained more than $88.8\%$ of the variance across islands (Table \ref{suptab:pcavariances}).
@@
 PCs need not represent the same wavelengths across islands because they are fitted on different datasets.
-Nevertheless, PC1 was very collinear with brightness for all islands (Figure..., Table \ref{suptab:brightness}).
+Nevertheless, PC1 was very collinear with brightness for all islands (Figure \ref{supfig:brightness}, Table \ref{suptab:brightness}).
 PC2 correlated highly with the red and ultraviolet ends of the spectrum, which were inversely correlated with each other (Fig.
-\ref{fig:boxplots}A).
+\ref{fig:anova}A).
 Higher PCs corresponded to various combinations of wavelengths.
@@
 In addition to within-island PCA, we performed a PCA on pooled data from the whole archipelago.
-Explain what wavelengths mapped onto what PCs (Table suptab:pcavariances).
+The first four principal components explained 91.3\% of the variance (Table \ref{suptab:pcavariances}).
+Again PC1 strongly correlated with brightness (Fig.
+\ref{supfig:brightness_pooled}, Table \ref{suptab:brightness}).
+PC2 was positively correlated to short wavelengths (ultraviolet to blue) and negatively correlated to long wavelengths (green to red, Fig.
+\ref{supfig:pooled}B).
+PC3 was strongly negatively correlated with UV reflectance and positively correlated with blue-green.
+PC4 was made of a mosaic of wavelengths, correlating positively with blue and red but negatively with ultraviolet and yellow.\\
+
 We used this dataset to partition the variance in dewlap coloration among islands, habitats and habitats within islands, using a two-way multivariate analysis of variance (MANOVA) with an interaction term.
-However, because the assumptions of parametric MANOVA were violated for all islands but Ragged Island (multivariate normality and homogeneity of covariance matrices, Tables suptab:multinorm, suptab:covariance), we used a semi-parametric MANOVA instead (R package MANOVA.RM, citation), with P-values calculated from a bootstrap procedure with 1,000 iterations.
-We calculated the proportion of variance explained by islands, habitats and the habitat-by-island interaction using partial effect sizes $\eta^2$ on a MANOVA-approximation of the analysis (R package heplots).
+However, because the assumptions of parametric MANOVA were violated for all islands but Ragged Island (multivariate normality, Henze-Zirkler's test, \citealt{Henze1990}, R package MVN, \citealt{Korkmaz2014}, Table \ref{suptab:multinorm}; and homogeneity of covariance matrices, Box's M-test, \citealt{Box1949, Morrison1988}, R package heplots, \citealt{Fox2018}, Table \ref{suptab:covariance}), we used a semi-parametric MANOVA instead (R package MANOVA.RM, \citealt{Friedrich2018}), with P-values calculated from a bootstrap procedure with 1,000 iterations.
+We calculated the proportion of variance explained by islands, habitats and the habitat-by-island interaction using partial effect sizes $\eta^2$ on a MANOVA-approximation of the analysis (R package heplots, \citealt{Fox2018}).
 
@@
 
-Our data violated the multivariate analysis of variance (MANOVA) assumption of homogeneity of covariance matrices across groups for all islands but Ragged Island (see previous section, \citealt{Box1949, Morrison1988}, implemented in the R package heplots, \citealt{Fox2018},  Table \ref{suptab:covariances}).
-We also detected within-habitat deviations from multivariate normality, primarily on Abaco, Bimini and Eleuthera (Henze-Zirkler's test, \citealt{Henze1990}, implemented in the R package MVN, \citealt{Korkmaz2014}, Table \ref{suptab:multinorm}).
-For these reasons and to reduce the chances of false discovery, we conducted multivariate group comparisons using support vector machines (SVMs), a model-free, nonparametric supervised machine learning technique.\\
+Because of the aforementioned violations of the MANOVA assumptions, and to reduce the chances of false discovery, we conducted multivariate group comparisons using support vector machines (SVMs), a model-free, powerful nonparametric supervised machine learning technique.\\
 
@@
 Machine learning for group comparison has become more popular in ecology and evolution in the recent years (e.g.
-\citet{Pigot2020}).
+\citealt{Pigot2020}).
 In particular, SVMs are designed to find the best possible nonlinear boundaries between labelled groups of points in multidimensional spaces, without assumptions about the distribution of the data \citep{Cortes1995, Cristianini2000, Kim2018}.
@@
 All classification analyses were repeated using the more classical linear discriminant analysis (LDA), a supervised machine learning technique finding linear boundaries that maximize the differences between groups, albeit assuming multivariate normality and homogeneity of covariance matrices \citep{Ripley1996a}.
-We used the R package rminer \citep{Cortez2010, Cortez2016} for SVMs, and MASS (citation) for LDAs.
+We used the R package rminer \citep{Cortez2010, Cortez2016} for SVMs, and MASS \citep{Venables2002} for LDAs.
 We used rminer's default heuristic search option to automatically tune the Gaussian kernel parameter $\sigma$ and the complexity parameter $C$ for the SVMs.\\
@@
 
-We conducted one-dimensional sensitivity analyses (R pacakge rminer) to determine the relative importance of the different input variables during classification where significant differences were detected, both on machines trained on principal components and machines trained on non-transformed reflectance at various wavelengths.
+We conducted one-dimensional sensitivity analyses using rminer \citep{Cortez2013} to determine the relative importance of the different input variables during classification where significant differences were detected, both on machines trained on principal components and machines trained on non-transformed reflectance at various wavelengths.
 In parallel, we conducted univariate analyses of variance to independently test the importance of different variables in between-habitat variation, on islands where the machines detected significant differences based on binomial tests (next section).
@@
 We tested the assumptions of the parametric ANOVA for each island included in the univariate analyses.
-For all islands where deviations from multivariate normality were detected in at least one habitat (Table multinorm), we assessed univariate normality for each principal component (Shapiro's test, Table normality).
-For skewed PCs that deviated significantly from normality, we repeated the analysis using a nonparametric Kruskal-Wallis tests.
-We found no multivariate outliers based on the Mahalanobis distance (package).
-We used the cases of better fit of the GLS model relative to the OLS model as evidence for heterogeneity of variances, which were then accounted for by the GLS approach (Table anova).\\
+For all islands where deviations from multivariate normality were detected in at least one habitat (Table \ref{suptab:multinorm}), we assessed univariate normality for each principal component (Shapiro-Wilk's test, Table \ref{suptab:normality}).
+For skewed PCs that deviated significantly from normality, we repeated the analysis using a nonparametric Kruskal-Wallis tests \citep{Hollander2013}.
+We found no multivariate outliers based on the Mahalanobis distance (package MVN, \citealt{Korkmaz2014}).
+We used the cases of better fit of the GLS model relative to the OLS model as evidence for heterogeneity of variances, which were then accounted for by the GLS approach (Table \ref{tab:anova}).\\
 
-Significant \textit{post hoc} contrasts were assessed using Tukey's Honest Significant Difference (HSD) test whenever the assumptions of normality and homogeneity of variances was met, Dunnett's T3 method when only homogeneity of variances was violated but not normality, and Nemenyi's test when normality was violated.\\
+Significant post hoc contrasts were assessed using Tukey's Honest Significant Difference (HSD) test whenever the assumptions of normality and homogeneity of variances was met \citep{Tukey1949}, Dunnett's T3 method when only homogeneity of variances was violated but not normality \citep{Dunnett1980}, and Nemenyi's test when normality was violated \citep{Nemenyi1963}.
+All post hoc tests were performed with the R package PMCMRplus \citep{Pohlert2020}.\\
 
@@
 
-We tested for within-island spatial autocorrelation between the geographical distances among sampling sites and their euclidean distances in multivariate color space (mean PC1 to PC4 per site, Table \ref{suptab:locations}), regardless of habitat-type.
-Because often only a few sites were sampled per island, we could not get meaningful results from tests that use sites as units of observation, such as Moran's I test.
-Instead, we designed a permutation test where we randomly reshuffled individual lizards across sites within islands 1,000 times each, and systematically recalculated Pearson's correlation coefficient between geographic distances (computed as geodesic distances using the R package geosphere, \citealt{Hijmans2019}) and phenotypic distances.
+We tested for within-island spatial autocorrelation between the geographical distances among sampling sites and their euclidean distances in multivariate color space (mean PC1 to PC4 per site, Table \ref{suptab:sites}), regardless of habitat-type.
+Because often only a few sites were sampled per island, we could not get meaningful results from tests that use sites as units of observation, such as Moran's I test \citep{Gittleman1990}.
+Instead, we designed a permutation test where we randomly reshuffled individual lizards across sites within islands 1,000 times each, and systematically recalculated Pearson's correlation coefficient between geographic distances (computed as geodesic distances, R package geosphere, \citealt{Hijmans2019}) and phenotypic distances.
 We used the resulting null distributions of correlation coefficients to assess the significance of the observed spatial autocorrelation for each island.
@@
 In this study, we were interested in the minimum spatial scale at which significant differences between habitats could be detected within islands.
-We performed multiple pairwise nonparametric Wilcoxon tests to compare dewlap coloration between sites with different habitat-types, for each pair of habitats and each variable where significant differences were detected with our analyses of variances.
+We performed multiple pairwise nonparametric Wilcoxon-Mann-Whitney's tests \citep{Hollander2013} to compare dewlap coloration between sites with different habitat-types, for each pair of habitats and each variable where significant differences were detected with our analyses of variances.
+The P-values were adjusted using a Benjamini-Hochberg correction for multiple testing \citep{Benjamini1995}.

### 2020-05-30 ms/sections/suppfigures.tex
nearly there
bbl

@@
 
-\begin{figure}
+\begin{figure}[H]
 	\centering
@@
 	\label{supfig:brightness_pooled}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width = \textwidth]{"../analyses/06-pooled/figure_pooled"}
+	\caption{(A) Principal component scores and 5\% confidence ellipses across habitats for the whole archipelago.
+The principal component analysis was performed on reflectance data from all islands pooled together.
+(B) PCA rotation matrix showing the loadings of each wavelength from 300 to 700nm onto the principal components.}
+	\label{supfig:pooled}
 \end{figure}

### 2020-07-04 ms/sections/discussion.tex
changing location content for the ms

@@
 Although those results are consistent with adaptation at a very local scale, other evolutionary drivers could be at work, including phenotypic plasticity, random drift, or historical contingency, including multiple colonization events.
-We reject this last explanation because all of the island populations in this study are strictly monophyletic, implying a single colonization event per island (van de Schoot, unpublished thesis; \citealt{Driessens2017}).
+We reject this last explanation because all of the island populations in this study are strictly monophyletic, implying a single colonization event per island (van de Schoot, unpublished thesis; \citealt{Driessens2017, Reynolds2020}).
 While random drift cannot be completely ruled out, we see little evidence for a role of phenotypic isolation-by-distance (spatial autocorrelation) in explaining the differences we report.
@@
 Here, however, we found little consistency in between-habitat divergence across the nine islands studied; except maybe for dewlap brightness being lower in the primary coppice habitat on three islands, and the primary coppice habitat being intermediate in coloration between mangrove and beach scrub on two islands.
-Those patterns could reflect adaptation to components of the habitat \citep{Endler1988}, but it is not clear what those components of the habitat might be.\\
+Those patterns could reflect adaptation to components of the habitat \citep{Endler1988}, but it is not clear what those components of the habitat might be.
+Although those patterns could reflect adaptation to components of the habitat \citep{Endler1988}, it is not clear what those components might be.
+Alternatively, neutral processes could produce patterns similar to local adaptation, such that this possibility cannot be completely ruled out \citep{Miles2019} (full subsection for this?).\\
 

### 2020-07-04 ms/sections/introduction.tex
bbl
changing location content for the ms

@@
 The staggering diversity of animal communication signals has long been of interest to evolutionary biologists.
-Animals use chemical, mechanical, electromagnetic, and visual signals to communicate in a wide variety of contexts, including competition for mates, species recognition, aposematism, cooperation, etc.
-\citep{Bradbury2011}.
+Animals use chemical, mechanical, electromagnetic, and visual signals to communicate in a wide variety of contexts, including, for example, competition for mates, species recognition, aposematism, and cooperation \citep{Bradbury2011}.
 A primary evolutionary factor shaping communication signals is the sensory system and behavior of their recipient(s) (the sensory drive hypothesis; \citealt{Endler1988,Endler1992,Endler1998}).
@@
 Each of these islands has been colonized independently by \textit{A.
-sagrei} (\citealt{Driessens2017, Reynolds2020}?, van de Schoot et al.
+sagrei} (\citealt{Driessens2017, Reynolds2020}, van de Schoot et al.
 unpubl.), such that these archipelagos constitute an ideal suite of natural replicates to explore within-island dewlap diversity across multiple islands.\\
@@
 Our sampling design included sites in close proximity (the median distance between two sites within an island was $11.2$km).
-We tested the hypothesis that the spatial scale was too small for phenotypic divergence to build up.
-If this was not the case, we predicted that if light conditions in the environment indeed drive color evolution, dewlaps should be most similar between beach scrub and mangrove forest, which both have high levels of light irradiance, contrary to the darker, closed-canopy coppice forest.
+We tested for divergence in dewlap phenotype between habitats within islands and between islands across the range of A.
+sagrei.
+We predicted that if light conditions in the environment indeed drive color evolution, dewlaps should be most similar between beach scrub and mangrove forest, which both have high levels of light irradiance, contrary to the darker, closed-canopy coppice forest.
 Similar, if detectability is maximized given the local conditions, we expected darker and more contrasting dewlaps in high irradiance habitats.
@@
 We found strong support for fine-scale, within-island differences in coloration between lizards inhabiting the three habitat-types in several color space dimensions, suggesting a potentially strong effect of divergent selection.
-However, the divergence patterns we observed did not match our expectations and were highly variable between islands.
+However, the divergence patterns we observed did not match our \textit{a priori} predictions and were inconsistent between islands.
 We found no evidence of isolation-by-distance as an explanation for the observed differences.

### 2020-07-04 ms/fetch_content.sh

@@ -1,0 +1,45 @@
+# Fetch figures
+cp ../analyses/.../classif_svm_pca.png figures
+cp ../analyses/.../figure_anova.png figures
+cp ../analyses/.../map.pdf figures
+
+# Fetch supplementary figures
+cp ../analyses/.../classif_lda_pca.png suppfigures
+cp ../analyses/.../classif_lda_pca_pooled.png suppfigures
+cp ../analyses/.../classif_lda_refl.png suppfigures
+cp ../analyses/.../classif_lda_refl_pooled.png suppfigures
+cp ../analyses/.../classif_svm_pca_pooled.png suppfigures
+cp ../analyses/.../classif_svm_refl.png suppfigures
+cp ../analyses/.../classif_svm_refl_pooled.png suppfigures
+cp ../analyses/.../detailed_map.pdf suppfigures
+cp ../analyses/.../figure_brightness.png suppfigures
+cp ../analyses/.../figure_brightness_pooled.png suppfigures
+cp ../analyses/.../figure_contrasts.png suppfigures
+cp ../analyses/.../figure_distances2.png suppfigures
+cp ../analyses/.../figure_distances.png suppfigures
+cp ../analyses/.../figure_pooled.png suppfigures
+cp ../analyses/.../figure_reflectance.png suppfigures
+cp ../analyses/.../importance_lda_pca.png suppfigures
+cp ../analyses/.../importance_lda_pca_pooled.png suppfigures
+cp ../analyses/.../importance_lda_refl.png suppfigures
+cp ../analyses/.../importance_lda_refl_pooled.png suppfigures
+cp ../analyses/.../importance_svm_pca.png suppfigures
+cp ../analyses/.../importance_svm_pca_pooled.png suppfigures
+cp ../analyses/.../importance_svm_refl.png suppfigures
+cp ../analyses/.../importance_svm_refl_pooled.png suppfigures
+
+# Fetch tables
+cp ../analyses/07-ANOVA/table_anova.tex tables
+
+# Fetch supplementary tables
+cp ../analyses/01-counts/table_counts.tex supptables
+cp ../analyses/08-spatial-correlation/table_sites.tex supptables
+cp ../analyses/03-PCA/table_expvar.tex supptables
+cp ../analyses/03-PCA/table_brightness.tex supptables
+cp ../analyses/05-assumptions/table_multinorm.tex supptables
+cp ../analyses/05-assumptions/table_covariance.tex supptables
[…]

### 2020-07-04 ms/sections/discussion.tex
implemented colin's comments and refactored the discussion to put more emphasis on neutral processes

@@
 We reject this last explanation because all of the island populations in this study are strictly monophyletic, implying a single colonization event per island (van de Schoot, unpublished thesis; \citealt{Driessens2017, Reynolds2020}).
-While random drift cannot be completely ruled out, we see little evidence for a role of phenotypic isolation-by-distance (spatial autocorrelation) in explaining the differences we report.
-We did detect a significant signal of isolation-by-distance on Eleuthera, but we did not find differences in dewlap coloration between habitats on this island.
 
-% The maintenance of divergence in the face of gene flow
+\paragraph{A role of neutral drift is unlikely} Differences in organismal traits between environments are not necessarily proof of adaptation or selection, and genetic drift may result in patterns similar to local adaptation \citep{Miles2019}.
+Nevertheless, two lines of evidence from our data suggest that this scenario may be implausible.
+First, we found little evidence for a role of phenotypic isolation-by-distance (spatial autocorrelation) in explaining the differences we report.
+We did detect a significant signal of isolation-by-distance on Eleuthera, but there were no differences in dewlap coloration between habitats on this island.
+Second, we detected differences between habitats at relatively small spatial scales, most of the time between sites 5-10km apart, sometimes a few hundred meters away, on Bimini for example.
+Such small-scale differences would be unlikely under strong gene flow \citep{Richardson2014}.
+Our study islands lack geographic barriers to the movement of $A.
+sagrei$, which have been shown to be highly mobile on these islands \citep{Kamath2018}, implying widespread gene flow across sites and habitats.
+Moreover, habitat-populations were found to be non-monophyletic with respect to <insert marker name here> (van de Schoot et al.
+unpublished thesis), which is also suggestive of gene flow.\\
 
-\paragraph{Divergence is maintained in the face of gene flow} We detected differences between habitats at relatively small spatial scales, most of the time between sites 5-10km apart, and sometimes a few hundred meters away.
-This, together with the lack of geographic barriers within islands and the high mobility of \textit{A.
-sagrei} individuals \citep{Kamath2018}, imply ample opportunity for extensive gene flow between our study populations.
-In agreement with this, populations from different habitats were found to not be monophyletic based on <insert marker name here> (van de Schoot et al.
-unpublished thesis).
-The maintenance of phenotypic differences at such a small spatial scales, in the face of gene flow, may support an adaptive explanation \citep{Richardson2014}.\\
-
-The persistence of dewlap color divergence despite gene flow has been documented in multiple species of anoles, sometimes in relation to environmental conditions.
+Our results align with previous documented cases of persistence of dewlap color divergence despite gene flow in multiple species of anoles, sometimes in relation to environmental conditions.
 \citet{Ng2012} and \citet{Ng2016} found divergent dewlap coloration in the face of gene flow between subspecies of \textit{A.
@@
 Finally, divergence in body coloration, but not dewlap coloration, was also reported in \textit{A.
-conspersus} on another small island, Grand Cayman \citep{Macedonia2001}.
-Altogether, our results align with previous studies suggesting a generally high adaptive potential of dewlap coloration in anoles over small distances, which could apply to \textit{A.
-sagrei}.
+conspersus} on another small island, Grand Cayman \citep{Macedonia2001}.\\
 
-% Could there be local adaptation? To what?
+\paragraph{Dewlap coloration could be locally adapted} Although phenotypic divergence at small spatial scales may be an indicator of local adaptation, the inconsistency of the between-habitat divergence patterns we observed across islands implies that neutral processes cannot be definitely ruled out \citep{Losos2011}.
+Conversely, the absence of parallel divergence across islands does not rule out local adaptation either \citep{Losos2011}.
+If selection has an effect, it may be dependent on aspects of the environment that are not encompassed by our broad habitat-type classification into coastal dry scrub, primary coppice forest and mangrove forest, and dewlap coloration may be influenced by a mosaic of interacting local environmental factors, which need not be the same across islands.\\
 
-\paragraph{Dewlap coloration could be locally adapted} Presently, we do not know the adaptive drivers of dewlap color divergence observed in this study.
-Some degree of parallel evolution is usually a good indicator for an adaptive process \citep{Losos2011}, and convergent patterns of dewlap color evolution in similar environments across islands and species have been documented in other species \citep{Thorpe2002a, Thorpe2002b}.
-Here, however, we found little consistency in between-habitat divergence across the nine islands studied; except maybe for dewlap brightness being lower in the primary coppice habitat on three islands, and the primary coppice habitat being intermediate in coloration between mangrove and beach scrub on two islands.
-Those patterns could reflect adaptation to components of the habitat \citep{Endler1988}, but it is not clear what those components of the habitat might be.
-Although those patterns could reflect adaptation to components of the habitat \citep{Endler1988}, it is not clear what those components might be.
-Alternatively, neutral processes could produce patterns similar to local adaptation, such that this possibility cannot be completely ruled out \citep{Miles2019} (full subsection for this?).\\
-
-Previous studies have suggested that dewlap coloration may have evolved to be maximally detectable under local light conditions imposed by the environment, primarily through UV contrast (i.e.
+Previous studies have described convergent patterns of dewlap color evolution in similar environments across islands and species \citep{Thorpe2002a, Thorpe2002b}.
+Others have suggested that dewlap coloration may have evolved to be maximally detectable under local light conditions imposed by the environment, primarily through UV contrast (i.e.
 UV-brighter dewlaps in UV-dark, mesic habitats and UV-darker dewlaps in UV-bright, xeric habitats), in \textit{A.

### 2020-07-04 ms/sections/suppfigures.tex
implemented colin's comments and refactored the discussion to put more emphasis on neutral processes

@@
 \centering
-	\includegraphics[width=\textwidth]{../maps/detailed_map.pdf}
+	\includegraphics[width=\textwidth]{suppfigures/detailed_map.pdf}
 	\caption{Map of the sampling sites and corresponding habitats across nine islands of the West Indies.}
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/03-PCA/figure_brightness"}
+	\includegraphics[width=0.8\textwidth]{suppfigures/figure_brightness.png}
 	\caption{Correlation between dewlap brightness (as measured by the mean reflectance from 300 to 700nm in wavelength) and PC1 score for each island.
@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{"../analyses/03-PCA/figure_brightness_pooled"}
+	\includegraphics[width=0.5\textwidth]{suppfigures/figure_brightness_pooled.png}
 	\caption{Correlation between dewlap brightness (as measured by the mean reflectance from 300 to 700nm in wavelength) and PC1 score across the whole archipelago.
@@
 	\centering
-	\includegraphics[width = \textwidth]{"../analyses/06-pooled/figure_pooled"}
+	\includegraphics[width = \textwidth]{suppfigures/figure_pooled.png}
 	\caption{(A) Principal component scores and 5\% confidence ellipses across habitats for the whole archipelago.
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/02-reflectance/figure_reflectance"}
+	\includegraphics[width=0.8\textwidth]{suppfigures/figure_reflectance.png}
 	\caption{5-95th percentile range of lizard dewlap reflectance values (in \% of incoming light) across wavelengths for each island and each habitat.}
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/classif_svm_pca_pooled"}
+	\includegraphics[width=0.8\textwidth]{suppfigures/classif_svm_pca_pooled.png}
 	\caption{Archipelago-wide SVM classification accuracy based on principal component data.
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/classif_svm_refl_pooled"}
+	\includegraphics[width=0.8\textwidth]{suppfigures/classif_svm_refl_pooled.png}
 	\caption{Archipelago-wide SVM classification accuracy based on reflectance data at 50nm-intervals in wavelength (see Methods).
@@
 	\centering
-	\includegraphics[width=0.4\textwidth]{"../analyses/04-machine learning/plots/importance_svm_pca_pooled"}
+	\includegraphics[width=0.4\textwidth]{suppfigures/importance_svm_pca_pooled.png}
 	\caption{Sensitivity analyses of the different input variables in the archipelago-wide SVM classification on principal component data (Figure \ref{supfig:classif-svm-pca-pooled}), with relative importance computed for every machine.}
@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{"../analyses/04-machine learning/plots/importance_svm_refl_pooled"}
+	\includegraphics[width=0.6\textwidth]{suppfigures/importance_svm_refl_pooled.png}
 	\caption{Sensitivity analyses of the different input variables in the archipelago-wide SVM classification on reflectance data at 50nm-intervals in wavelength (Figure \ref{supfig:classif-svm-refl-pooled}), with relative importance computed for every machine.}
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/classif_lda_pca_pooled"}
+	\includegraphics[width=0.8\textwidth]{suppfigures/classif_lda_pca_pooled.png}
 	\caption{Archipelago-wide LDA classification accuracy based on principal component data.
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/classif_lda_refl_pooled"}
+	\includegraphics[width=0.8\textwidth]{suppfigures/classif_lda_refl_pooled.png}
 	\caption{Archipelago-wide LDA classification accuracy based on reflectance data at 50nm-intervals in wavelength (see Methods).
@@
 	\centering
-	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_lda_pca"}
+	\includegraphics[width=\textwidth]{suppfigures/classif_lda_pca.png}
 	\caption{LDA classification accuracy across islands based on principal component data.
@@
 	\centering
-	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_svm_refl"}
+	\includegraphics[width=\textwidth]{suppfigures/classif_svm_refl.png}
 	\caption{SVM classification accuracy across islands based on reflectance data at 50nm-intervals in wavelength (see Methods).
@@
 	\centering
-	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_lda_refl"}
+	\includegraphics[width=\textwidth]{suppfigures/classif_lda_refl.png}
 	\caption{LDA classification accuracy across islands based on reflectance data at 50nm-intervals in wavelength (see Methods).
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/importance_svm_pca"}
+	\includegraphics[width=0.8\textwidth]{suppfigures/importance_svm_pca.png}
 	\caption{Sensitivity analyses of the different input variables in the within-island SVM classification on principal component data (Figure \ref{supfig:classif-svm-pca}), with relative importance computed for every machine.}
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/importance_lda_pca"}
+	\includegraphics[width=0.8\textwidth]{suppfigures/importance_lda_pca.png}
 	\caption{Sensitivity analyses of the different input variables in the within-island LDA classification on principal component data (Figure \ref{supfig:classif-lda-pca}), with relative importance computed for every machine.}
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/importance_svm_pca"}
+	\includegraphics[width=0.8\textwidth]{suppfigures/importance_svm_pca.png}
 	\caption{Sensitivity analyses of the different input variables in the archipelago-wide SVM classification on reflectance at 50nm-intervals in wavelength (Figure \ref{supfig:classif-svm-refl}), with relative importance computed for every machine.}
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/importance_lda_pca"}
+	\includegraphics[width=0.8\textwidth]{suppfigures/importance_lda_pca.png}
 	\caption{Sensitivity analyses of the different input variables in the archipelago-wide LDA classification on reflectance at 50nm-intervals in wavelength (Figure \ref{supfig:classif-lda-refl}), with relative importance computed for every machine.}
@@
 	\centering
-	\includegraphics[width=\textwidth]{"../analyses/10-distances/figure_distances2"}
+	\includegraphics[width=\textwidth]{suppfigures/figure_distances2.png}
 	\caption{Spatial scale of between-habitat differences in dewlap coloration.
@@
 	\centering
-	\includegraphics[width=0.7\textwidth]{"../analyses/09-parallelism/figure_contrasts"}
+	\includegraphics[width=0.7\textwidth]{suppfigures/figure_contrasts.png}
 	\caption{Test of parallel divergence between islands.

### 2020-07-04 analyses/04-machine learning/MachineLearning.R

@@ -1,289 +1,0 @@
-rm(list = ls())
-
-library(dewlap)
-library(nmgc)
-library(tidyverse)
-library(cowplot)
-
-data <- read.csv("data/reflectance.csv", header = TRUE)
-
-#### Analysis within each island ####
-
-# Principal components
-
-variables <- paste0("PC", 1:4)
-
-# SVM-classification on PCs
-
-res <- classify(
-  data, variables, grouping = "habitat", nesting = "island",
-  to_pcomp = paste0("wl", 300:700), digest = TRUE, test = TRUE, k = 5,
-  nrep = 100, seed = 24, method = "SVM", importance = TRUE
-)
-
-# Prepare a confusion matrix plot to show on the side
-
-conf <- res$avg$`Long Island`
-conf <- conf %>% apply(2, function(x) x / sum(x))
-conf <- conf %>%
-  data.frame %>%
-  rownames_to_column("predicted") %>%
-  gather_("true", "freq", colnames(conf))
-colnames(conf) <- c("predicted", "true", "freq")
-
-confplot <- ggplot(conf, aes(x = true, y = predicted, fill = freq)) +
-  geom_tile() +
-  scale_fill_gradient(low = "white", high = "darkgreen", limits = c(0, 1)) +
-  theme_bw() +
-  labs(x = "True habitat", y = "Predicted habitat", fill = "Frequency") +
-  theme(axis.text.x = element_text(angle = 60, hjust = 1))
-
-blank <- ggplot() + theme_void()
-
-p <- plot_classif(res, facets = "island")
-plot_grid(p, plot_grid(blank, confplot, blank, nrow = 3), ncol = 2, rel_widths = c(2, 1))
-ggsave("classif_svm_pca.png", height = 6, width = 9, dpi = 300)
-
-fname <- "analyses/04-machine learning/table_classif_%s"
-colnames <- c("Island", "Accuracy", "$N$", "$p_{\\mbox{test}}$", "$n_{\\mbox{test}}$", "$P$", "")
-
-t1 <- res$mean
-t1_fname <- sprintf(fname, "svm_pca")
-save_table(t1, t1_fname, digits = c(0, 3, 0, 1, 0, 4, 0), col.names = colnames)
-
-res$imp %>%
-  gather_("variable", "importance", variables) %>%
-  filter(island %in% c("Abaco", "Bimini", "Cayman Brac", "Little Cayman", "Long Island")) %>%
[…]

### 2020-07-04 analyses/04-machine-learning/MachineLearning.R

@@ -1,0 +1,289 @@
+rm(list = ls())
+
+library(dewlap)
+library(nmgc)
+library(tidyverse)
+library(cowplot)
+
+data <- read.csv("data/reflectance.csv", header = TRUE)
+
+#### Analysis within each island ####
+
+# Principal components
+
+variables <- paste0("PC", 1:4)
+
+# SVM-classification on PCs
+
+res <- classify(
+  data, variables, grouping = "habitat", nesting = "island",
+  to_pcomp = paste0("wl", 300:700), digest = TRUE, test = TRUE, k = 5,
+  nrep = 100, seed = 24, method = "SVM", importance = TRUE
+)
+
+# Prepare a confusion matrix plot to show on the side
+
+conf <- res$avg$`Long Island`
+conf <- conf %>% apply(2, function(x) x / sum(x))
+conf <- conf %>%
+  data.frame %>%
+  rownames_to_column("predicted") %>%
+  gather_("true", "freq", colnames(conf))
+colnames(conf) <- c("predicted", "true", "freq")
+
+confplot <- ggplot(conf, aes(x = true, y = predicted, fill = freq)) +
+  geom_tile() +
+  scale_fill_gradient(low = "white", high = "darkgreen", limits = c(0, 1)) +
+  theme_bw() +
+  labs(x = "True habitat", y = "Predicted habitat", fill = "Frequency") +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1))
+
+blank <- ggplot() + theme_void()
+
+p <- plot_classif(res, facets = "island")
+plot_grid(p, plot_grid(blank, confplot, blank, nrow = 3), ncol = 2, rel_widths = c(2, 1))
+ggsave("classif_svm_pca.png", height = 6, width = 9, dpi = 300)
+
+fname <- "analyses/04-machine learning/table_classif_%s"
+colnames <- c("Island", "Accuracy", "$N$", "$p_{\\mbox{test}}$", "$n_{\\mbox{test}}$", "$P$", "")
+
+t1 <- res$mean
+t1_fname <- sprintf(fname, "svm_pca")
+save_table(t1, t1_fname, digits = c(0, 3, 0, 1, 0, 4, 0), col.names = colnames)
+
+res$imp %>%
+  gather_("variable", "importance", variables) %>%
+  filter(island %in% c("Abaco", "Bimini", "Cayman Brac", "Little Cayman", "Long Island")) %>%
[…]

### 2020-07-04 analyses/08-spatial-correlation/SpatialAutocorrelation.R

@@ -1,0 +1,28 @@
+# Spatial autocorrelation by permutation test
+
+rm(list = ls())
+
+library(dewlap)
+library(tidyverse)
+library(nmgc)
+
+data <- read.csv("data/reflectance.csv", header = TRUE)
+
+# Spatial autocorrelation test
+res_cortest <- nspcortest(
+  data, variables = paste0("PC", 1:4), nesting = "island", nperm = 1000, seed = 42,
+  to_pcomp = paste0("wl", 300:400), keep = "habitat"
+)
+
+# Save table
+t1 <- res_cortest$res
+t1 <- t1 %>% add_signif()
+t1_fname <- "analyses/08-spatial correlation/table_spatial"
+t1_names <- c("Island", "$r$", "$P$", "$N$", "")
+save_table(t1, t1_fname, digits = c(0, 3, 3, 0), col.names = t1_names)
+
+# Table with multivariate means per site
+t2 <- res_cortest$sites
+t2_fname <- "analyses/08-spatial correlation/table_sites"
+t2_names <- c("Island", "Longitude", "Latitude", "Habitat", "PC1", "PC2", "PC3", "PC4")
+save_table(t2, t2_fname, digits = c(0, 4, 4, 0, 4, 4, 4, 4), col.names = t2_names)

### 2020-07-04 ms/fetch_content.sh

@@ -1,32 +1,34 @@
+# Run this script to copy selected files from the analysis folders to the manuscript folder
+
 # Fetch figures
-cp ../analyses/.../classif_svm_pca.png figures
-cp ../analyses/.../figure_anova.png figures
-cp ../analyses/.../map.pdf figures
+cp ../analyses/04-machine-learning/plots/classif_svm_pca.png figures
+cp ../analyses/07-ANOVA/figure_anova.png figures
+cp ../maps/map.pdf figures
 
 # Fetch supplementary figures
-cp ../analyses/.../classif_lda_pca.png suppfigures
-cp ../analyses/.../classif_lda_pca_pooled.png suppfigures
-cp ../analyses/.../classif_lda_refl.png suppfigures
-cp ../analyses/.../classif_lda_refl_pooled.png suppfigures
-cp ../analyses/.../classif_svm_pca_pooled.png suppfigures
-cp ../analyses/.../classif_svm_refl.png suppfigures
-cp ../analyses/.../classif_svm_refl_pooled.png suppfigures
-cp ../analyses/.../detailed_map.pdf suppfigures
-cp ../analyses/.../figure_brightness.png suppfigures
-cp ../analyses/.../figure_brightness_pooled.png suppfigures
-cp ../analyses/.../figure_contrasts.png suppfigures
-cp ../analyses/.../figure_distances2.png suppfigures
-cp ../analyses/.../figure_distances.png suppfigures
-cp ../analyses/.../figure_pooled.png suppfigures
-cp ../analyses/.../figure_reflectance.png suppfigures
-cp ../analyses/.../importance_lda_pca.png suppfigures
-cp ../analyses/.../importance_lda_pca_pooled.png suppfigures
-cp ../analyses/.../importance_lda_refl.png suppfigures
-cp ../analyses/.../importance_lda_refl_pooled.png suppfigures
-cp ../analyses/.../importance_svm_pca.png suppfigures
-cp ../analyses/.../importance_svm_pca_pooled.png suppfigures
-cp ../analyses/.../importance_svm_refl.png suppfigures
-cp ../analyses/.../importance_svm_refl_pooled.png suppfigures
+cp ../analyses/04-machine-learning/plots/classif_lda_pca.png suppfigures
+cp ../analyses/04-machine-learning/plots/classif_lda_pca_pooled.png suppfigures
+cp ../analyses/04-machine-learning/plots/classif_lda_refl.png suppfigures
[…]

### 2020-07-04 ms/run_pandoc.sh

@@ -1,0 +1,4 @@
+# Run this script to convert LaTeX files into ODT and DOCX documents
+# Requires pandoc and pandoc-crossref being installed
+
+pandoc main.tex --filter=pandoc-crossref --bibliography=library.bib --csl=journal-of-evolutionary-biology.csl -o scherrer2020.odt -o scherrer2020.docx

### 2020-07-12 ms/sections/acknowledgments.tex
second review

@@
 Glor, Alberto R.
-Puente-Rol\'{o}n, Kevin Aviles-Rodriguez, Kristin Winchell, and Melissa Kemp for assistance in the field.
-Funding for this work was provided by the Templeton Foundation (to JBL), NSF DEB~\#1500761 (to AJG), NSF DBI~\#1609284 (to CMD), and a Harvard Museum of Comparative Zoology Putnam Grant (to RGR).
+Puente-Rol\'{o}n, Kevin Aviles-Rodriguez, Kristin Winchell, Jason Fredette and Melissa Kemp for assistance in the field and Max Lambert and James Stroud for helpful discussions.
+Funding for this work was provided by the Templeton Foundation (to JBL), NSF DEB~\#1927194 (to JBL and AJG), NSF DEB~\#1500761 (to AJG), NSF DBI~\#1609284 (to CMD), and a Harvard Museum of Comparative Zoology Putnam Expedition Grant (to RGR).

### 2020-07-12 ms/sections/discussion.tex
second review

@@
-% Extra results to get: minimum distance at which we found differences and range-wide results
-
 % There are inconsistent differences that are not explained by distance
@@
 \paragraph{Dewlap coloration differs between habitat-types} We found that male dewlap coloration in \textit{A.
-sagrei} significantly varied between habitat-types (beach scrub bush, primary coppice forest and mangrove forest) on five islands of the West Indies: Abaco, Bimini, Cayman Brac, Little Cayman and Long Island.
+sagrei} significantly varied between fine-scale habitat-types (beach scrub bush, primary coppice forest and mangrove forest) on five islands of the West Indies: Abaco, Bimini, Cayman Brac, Little Cayman and Long Island.
 However, the habitat-specific variation in dewlaps was not consistent between these islands.
@@
 sagrei$, which have been shown to be highly mobile on these islands \citep{Kamath2018}, implying widespread gene flow across sites and habitats.
-Moreover, habitat-populations were found to be non-monophyletic with respect to <insert marker name here> (van de Schoot et al.
-unpublished thesis), which is also suggestive of gene flow.\\
+Moreover, habitat-populations within each island were found to be non-monophyletic admixtures of multiple haplotypes, based on phylogenetic analysis of mitochondrial DNA sequences (van de Schoot et al.
+unpublished thesis), suggesting widespread gene flow.\\
 
@@
 sagrei} on islands with more coexisting species of anoles.
-Therefore, dewlap coloration in $A.
-sagrei$ may be under the influence of a mosaic of interacting local environmental factors, which need not be the same across islands.
+Our data is consistent with those previous results in suggesting that adaptation to local light conditions, or at least broad habitat types, may not be a major driver of the variation in dewlap coloration in \textit{A.
+sagrei}.\\
 
-% Plasticity is unlikely
+% The evidence for sexual selection
 
-\paragraph{A role of phenotypic plasticity is unlikely} Phenotypic plasticity could explain differences in coloration between habitat populations because of diet differences.
+\paragraph{Sexual selection could be at play} Sexual selection is an important component of an adaptive, sensory drive hypothesis of dewlap color divergence across environmental conditions, because in this scenario habitat-dependent selection is executed by the signal recipients \citep{Endler1988}, such as mating partners \citep{Driessens2014}.
+An alternative yet non-mutually exclusive explanation for the observed idiosyncratic patterns of within-island divergence is that of "Fisherian" sexual selection \citep{Andersson1994}.
+Under the Fisherian model of sexual selection, arbitrary female preferences (i.e.
+independent of the environment) for certain male ornaments may drive divergent evolution of male signals, such as dewlap coloration, if female preferences differ between localities for other reasons than environment-dependent perception abilities (a situation that could have arisen e.g.
+in an initial phase of genetic drift).
+Substantial levels of promiscuity in \textit{A.
+sagrei} suggest ample opportunity for female mate choice \citep{Kamath2018}, and are in line with this explanation.
+Therefore, arbitrary, habitat-independent female preferences could further explain the inconsistent patterns of divergence across islands.
+In contrast with this, however, \citet{Baeckens2018} found no link between \textit{A.
+sagrei} dewlap coloration and size dimorphism (a proxy for sexual selection) in an among-island study of the same archipelagos.\\
+
+Besides, although this process is well-established in some systems such as African cichlids (e.g.
+\citealt{Seehausen1997}), previous work suggests that dewlap coloration in \textit{Anolis} may act as a mating cue according to a "good genes" model of sexual selection \citep{Andersson1994}, rather than a Fisherian process.
+In the good genes model, the cue under sexual selection is an indicator of individual quality (e.g.
+better immune response to disease) and of indirect benefit to the offspring.
+For example, \citet{Cook2013} found lower orange reflectance in dewlaps with heavily parasitized \textit{A.
+brevirostris}, suggesting a trade-off in carotenoid use between the immune response and pigment deposition.
+\citet{Steffen2014} found that lower UV and orange-red reflectance predict contest-winning success between \textit{A.
+sagrei} males, while \citet{Driessens2015} further found that more yellow and red dewlaps (relative to UV) predict better body condition, and that higher yellow and UV reflectance at the margin of the dewlap predict higher hematocrit (the concentration of red blood cells), indicating a better health.
+Other aspects of the dewlap than color have also been found to be indicators of individual quality, such as dewlap size \citep{Vanhooydonck2005, Vanhooydonck2009}, but not dewlap display frequency \citep{Tokarz2002, Tokarz2005, Driessens2015}.\\
+
+\paragraph{A role of phenotypic plasticity is unlikely} Differences in coloration between habitat populations may not be genetically determined, and may be influenced by environmental factors such as parasite load, as mentioned above \citep{Cook2013}.
 The yellow, orange and red coloration in anoline dewlaps are produced by pterins and carotenoids \citep{Ortiz1962, Ortiz1962a, Ortiz1963, Ortiz1966, Macedonia2000, Steffen2007, Steffen2009}.
@@
 This makes a plastic response to differences in diet across habitats unlikely.
-Developmental plasticity (e.g.
-due to differences egg rearing conditions) is also unlikely because dewlap coloration develops at sexual maturity in anoles \citep{Ng2013}.
+Furthermore, developmental plasticity during the ontogeny is also unlikely because dewlap coloration develops at sexual maturity in anoles \citep{Ng2013}.
 The differences we observed could therefore be heritable.
@@
 Although most studies used one or two-generation common garden experiments and thus could not rule out transgenerational plastic effects \citep{Tariel2020}, dewlap coloration generally seems to not be a plastic trait.
-This further reinforces an adaptive explanation, where dewlap color could be under differential natural and/or sexual selection in these different habitats.
-
-% The evidence for sexual selection
-
-\paragraph{Sexual selection could be at play} Substantial levels of promiscuity in \textit{A.
-sagrei} suggest ample opportunity for female mate choice and sexual selection \citep{Kamath2018}.
-A number of studies have suggested that characteristics of the dewlap are indicators of male quality and may therefore act as a cue in mate choice, according to the "good genes" model of sexual selection \citep{Andersson1994}.
-\citet{Cook2013} found lower orange reflectance in dewlaps with heavily parasitized \textit{A.
-brevirostris}, suggesting a trade-off in carotenoid use between the immune response and pigment deposition.
-\citet{Steffen2014} found that lower UV and orange-red reflectance predict contest-winning success between males.
-\citet{Driessens2015} further found that more yellow and red dewlaps (relative to UV) predict better body condition, and that higher yellow and UV reflectance at the margin of the dewlap predict higher hematocrit (the concentration of red blood cells), indicating a better health.\\
-
-Other aspects of the dewlap may be important for sexual selection.
-For example, dewlap size in $A.
-sagrei$ was found to correlate with bite force \citep{Vanhooydonck2005} and sexual size dimorphism (a proxy for sexual selection; \citealt{Vanhooydonck2009}).
-Even though display frequency increases in the presence of females \citep{Driessens2014}, no link was found between dewlap display frequency and mating success \citep{Tokarz2002, Tokarz2005} or individual quality \citep{Driessens2015}.
-Our report of inconsistent divergence between islands, however, is at odds with the good-genes model of sexual selection, which would predict the evolution of dewlap coloration in the same direction across the archipelago.
-This is in agreement with \citet{Baeckens2018}, who found no link between the average island-dewlap coloration and sexual size dimorphism in \textit{A.
-sagrei} across the West Indies.
-If sexual selection is at play in our system, it more likely involves divergent female preferences that are not linked to male quality but rather to components of the environment, or may well be arbitrary (Fisherian sexual selection; \citealt{Andersson1994}).
+This further reinforces an adaptive explanation, where dewlap color could be under differential natural and/or sexual selection in these different habitats.\\
 
@@
 sagrei} across the West Indies.
-While these populations do not appear to be in the process of speciation, our results strongly suggest that the anoline dewlap has enough micro-scale, local adaptive potential to participate in the build-up of reproductive isolation, should it be recruited for assortative mating.
+While these intra-island populations do not appear to be in the process of speciation, our results suggest that the anoline dewlap has enough micro-scale, local adaptive potential to contribute of reproductive isolation, should it be recruited for assortative mating.

### 2020-07-12 ms/sections/introduction.tex
second review

@@
 between islands \citep{Vanhooydonck2009, Driessens2017} or within large islands such as Puerto Rico \citep{Leal2002, Leal2004} or Hispaniola \citep{Ng2012, Ng2016}.
-These large scales should reduce gene flow \citep{Ng2011, Lambert2013, Richardson2014, Ng2017}.
+These large scales or marine barriers should reduce gene flow \citep{Ng2011, Lambert2013, Richardson2014, Ng2017}.
 That said, examples do exist of divergence in dewlap coloration at smaller scales or between populations with high degrees of gene flow \citep{Thorpe2002a, Thorpe2002b, Stapley2011, Ng2016}.\\
@@
 
-The island bank systems of the Bahamas and Cayman Islands comprise relatively small islands, with no major geographic barriers within islands limiting dispersal for this promiscuous species \citep{Kamath2018}.
+The island bank systems of the Bahamas and Cayman Islands presently, if not historically, comprise relatively small islands, with no major geographic barriers within islands limiting dispersal for this promiscuous species \citep{Kamath2018}.
 These islands all share three characteristic native West Indian habitat-types -- beach scrub bush, closed-canopy primary coppice forest, and mangrove forest -- that are often spatially intermingled.
 These habitats contrast in environmental parameters including vegetation community, light irradiance, humidity and temperature \citep{Howard1950, Schoener1968}.
-Each of these islands has been colonized independently by \textit{A.
-sagrei} (\citealt{Driessens2017, Reynolds2020}, van de Schoot et al.
+Each of these island groups (Caymans and Bahamas) has been colonized independently by \textit{A.
+sagrei} from Cuba (\citealt{Driessens2017, Reynolds2020}, van de Schoot et al.
 unpubl.), such that these archipelagos constitute an ideal suite of natural replicates to explore within-island dewlap diversity across multiple islands.\\
@@
 Our sampling design included sites in close proximity (the median distance between two sites within an island was $11.2$km).
-We tested for divergence in dewlap phenotype between habitats within islands and between islands across the range of A.
-sagrei.
+We tested for divergence in dewlap phenotype between habitats within islands and between islands across the range of \textit{A.
+sagrei}.
 We predicted that if light conditions in the environment indeed drive color evolution, dewlaps should be most similar between beach scrub and mangrove forest, which both have high levels of light irradiance, contrary to the darker, closed-canopy coppice forest.
@@
 We found no evidence of isolation-by-distance as an explanation for the observed differences.
-Our results are nevertheless consistent with small-scale adaptive maintenance of signal polymorphism despite presumed considerable opportunity for gene flow.
+Our results are nevertheless consistent with small-scale adaptive maintenance of signal polymorphism despite presumed considerable opportunity for gene flow, and might suggest that idiosyncracies of local drift, selection, and gene flow contribute to differing outcomes in different populations.
 

### 2020-07-12 ms/sections/methods.tex
second review

@@
 
-We sampled 466 lizards from seven islands in the Bahamas Archipelago -- Abaco, North Andros, South Andros, South Bimini, Eleuthera, Long Island, Ragged Island -- and two in the Cayman Islands -- Cayman Brac and Little Cayman (Figure \ref{fig:map}).
-These islands and island banks were chosen to span the West Indian range of \textit{Anolis sagrei}.
+We sampled 466 male \textit{Anolis sagrei} from seven islands in the Bahamas Archipelago -- Abaco, North Andros, South Andros, South Bimini, Eleuthera, Long Island, Ragged Island -- and two in the Cayman Islands -- Cayman Brac and Little Cayman (Figure \ref{fig:map}).
+These islands and island banks were chosen to span the breadth of the West Indian range of \textit{Anolis sagrei}, because they have highly similar habitat types,  and because the \textit{A.
+sagrei} on each island group are derived from ancient and distinct colonization events from Cuba (i.e.
+relatively evolutionarily independent, \citealt{Reynolds2020}).
 Three habitats were sampled on each island based on characterizations by \citet{Howard1950} and \citet{Schoener1968}.
@@
 $80.3$\% of all pairwise distances within islands were below $50$km.
-Additionally, there are no major barriers to dispersal (such as mountains) on any of the islands that we sampled.
+Additionally, there are no major barriers to dispersal (such as mountains or grassland) on any of the islands that we sampled.
 
@@
 
-Reflectance curves were smoothened using the R package pavo \citep{Maia2013} as well as custom R functions, down to one reflectance value at each nanometer in wavelength from 300 to 700nm.
+Reflectance curves were smoothed using the R package pavo \citep{Maia2013} as well as with custom R functions, down to one reflectance value at each nanometer in wavelength from 300 to 700nm.
 Because neighboring wavelengths are highly collinear in reflectance, we reduced the dimensionality of the data using principal component analysis (PCA), as per \citet{Cuthill1999} and \citet{Leal2002}.
@@
 Because often only a few sites were sampled per island, we could not get meaningful results from tests that use sites as units of observation, such as Moran's I test \citep{Gittleman1990}.
-Instead, we designed a permutation test where we randomly reshuffled individual lizards across sites within islands 1,000 times each, and systematically recalculated Pearson's correlation coefficient between geographic distances (computed as geodesic distances, R package geosphere, \citealt{Hijmans2019}) and phenotypic distances.
+Instead, we designed a permutation test where we randomly reshuffled individual lizards across sites within islands 1,000 times each, and systematically recalculated Pearson's correlation coefficient between geographic distances (computed as geodesic distances in the R package geosphere; \citealt{Hijmans2019}) and phenotypic distances.
 We used the resulting null distributions of correlation coefficients to assess the significance of the observed spatial autocorrelation for each island.

### 2020-07-12 ms/sections/results.tex
second review

@@
 We tested for variation in \textit{A sagrei} dewlap coloration between populations living in three characteristic habitat types across nine islands that span the West Indian range of the brown anole (Fig \ref{fig:map}, \ref{supfig:map}).
-We found that most of the variation in coloration partitioned between islands (two-way semi-parametric MANOVA, modified ANOVA-type statistic (MATS) = 2009.6, P < 0.001, Fig.
+We found that most of the variation in coloration is partitioned between islands (two-way semi-parametric MANOVA, modified ANOVA-type statistic (MATS) = 2009.6, P < 0.001, Fig.
 \ref{supfig:reflectance}, explained variance $\eta^2 = 44.3$\%, MANOVA approximation).
@@
 Archipelago-wide differences were not detected by LDA classifiers at all (Fig.
-\ref{supfig:classif-lda-pca-pooled}, \ref{supfig:classif-lda-refl-pooled}).
-So the difference could be in something else than means.\\
+\ref{supfig:classif-lda-pca-pooled}, \ref{supfig:classif-lda-refl-pooled}).\\ %So, the differences could be in something other than means.\\
 
@@
 \ref{supfig:map}).
-Besides, for pairs of habitats where significant differences in dewlap coloration were detected along some principal components, comparisons of the actual sampling sites indicate that the most detectable differences involved sites that were mostly 5-10km apart.
+Indeed, most significant differences in dewlap coloration involved sites that were 5-10km apart.
 Our most extreme case of significant differences occurred for PC3 between the beach scrub site and a coppice site, separated from each other by a a few hundreds of meters at most on Bimini (multiple pairwise Wilcoxon-Mann-Whitney tests, Fig.

### 2020-07-12 ms/run_pandoc.sh

@@ -1,4 +1,5 @@
 # Run this script to convert LaTeX files into ODT and DOCX documents
 # Requires pandoc and pandoc-crossref being installed
 
-pandoc main.tex --filter=pandoc-crossref --bibliography=library.bib --csl=journal-of-evolutionary-biology.csl -o scherrer2020.odt -o scherrer2020.docx
+pandoc main.tex --filter=pandoc-crossref --bibliography=library.bib --csl=journal-of-evolutionary-biology.csl -o scherrer2020.odt
+pandoc main.tex --filter=pandoc-crossref --bibliography=library.bib --csl=journal-of-evolutionary-biology.csl -o scherrer2020.docx

### 2020-07-13 ms/sections/abstract.tex
stuff

@@
 Here, we investigated local adaptation of the dewlap across habitat-types at a small spatial scale, as this may give an indication of how conditions for the early stages of speciation may be met.
-We explored variation in dewlap coloration in one widespread species, \textit{Anolis sagrei}, across three characteristic habitats spanning the Bahamas and the Cayman Islands.
+We explored variation in dewlap coloration in the most widespread species of anole, \textit{Anolis sagrei}, across three characteristic habitats spanning the Bahamas and the Cayman Islands.
 Using reflectance spectrometry as well as supervised machine learning, we found some consistent differences in spectral properties of the dewlap between habitats within small islands.

### 2020-07-13 ms/sections/discussion.tex
almost ready to go

@@
 Our study islands lack geographic barriers to the movement of $A.
-sagrei$, which have been shown to be highly mobile on these islands \citep{Kamath2018}, implying widespread gene flow across sites and habitats.
-Moreover, habitat-populations within each island were found to be non-monophyletic admixtures of multiple haplotypes, based on phylogenetic analysis of mitochondrial DNA sequences (van de Schoot et al.
-unpublished thesis), suggesting widespread gene flow.\\
+sagrei$, which have been shown to be highly mobile \citep{Kamath2018}, implying widespread gene flow across sites and habitats.
+Moreover, habitat-populations within each island were found to be non-monophyletic and often share identical haplotypes, based on phylogenetic analysis of mitochondrial DNA sequences (van de Schoot et al.
+unpublished thesis), suggesting gene flow between habitats may be widespread.\\
 
-Our results align with previous documented cases of persistence of dewlap color divergence despite gene flow in multiple species of anoles, sometimes in relation to environmental conditions.
+Our results align with previous documented cases of persistent dewlap color divergence despite gene flow in multiple species of anoles, sometimes in relation to environmental conditions.
 \citet{Ng2012} and \citet{Ng2016} found divergent dewlap coloration in the face of gene flow between subspecies of \textit{A.
@@
 trinitatis} on the featureless island of St Vincent \citep{Thorpe2002b}.
-Finally, divergence in body coloration, but not dewlap coloration, was also reported in \textit{A.
-conspersus} on another small island, Grand Cayman \citep{Macedonia2001}.\\
+Finally, regionally-distinct body coloration, but not dewlap coloration, is present in \textit{A.
+conspersus} on another small island, Grand Cayman, where no physical barriers to gene flow exist \citep{Macedonia2001}.\\
 
@@
 cooki} from Puerto Rico \citep{Leal2002, Leal2004}.
-Although UV reflectance was commonly involved in between-habitat divergence, we found no such patterns in \textit{A.
-sagrei}, where instead, we found the darkest dewlaps in the darkest, mesic habitat -- primary coppice forest -- on three islands, and dewlaps often differed the most between beach scrub and mangrove forest, two xeric habitats with similar, high irradiance levels \citep{Howard1950, Schoener1968}.
+Although UV reflectance was commonly involved in between-habitat divergence in \textit{A.
+cooki} and \textit{A.
+cristatellus}, we found no such patterns in \textit{A.
+sagrei}, where instead, we found the darkest dewlaps in the dark, mesic habitat -- primary coppice forest -- on three islands, and dewlaps often differed the most between beach scrub and mangrove forest, two xeric habitats with similar, high irradiance levels \citep{Howard1950, Schoener1968}.
 The inconsistent and idiosyncratic patterns we observed suggest that dewlap color variation between habitats cannot be predicted by habitat identity alone.
-Studies of Jamaican and Hispaniolan anoles similarly found between-habitat differences in dewlap coloration but no evidence for higher detectability \citep{Fleishman2009, Ng2012}.
-Habitats on different islands may also differ in other aspects than light conditions such as densities of predators or other anole species, which have been shown to affect among-island dewlap diversity \citep{Vanhooydonck2009, Baeckens2018}.
+Studies of Jamaican and Hispaniolan anoles similarly found between-habitat differences in dewlap coloration but no evidence for higher dewlap detectability in different habitats \citep{Fleishman2009, Ng2012}.
+Habitats on different islands may also differ in aspects other than light conditions, such as densities of predators or congeners, which have been shown to affect among-island dewlap diversity \citep{Vanhooydonck2009, Baeckens2018}.
 In particular, \citet{Baeckens2018} recently showed that dewlaps with spotted patterns occurred more often in \textit{A.
 sagrei} on islands with more coexisting species of anoles.
-Our data is consistent with those previous results in suggesting that adaptation to local light conditions, or at least broad habitat types, may not be a major driver of the variation in dewlap coloration in \textit{A.
+Our data are consistent with those previous results in suggesting that adaptation to local light conditions, or at least broad habitat types, may not be a major driver of the variation in dewlap coloration in \textit{A.
 sagrei}.\\
@@
 conspersus} on Grand Cayman evolved a UV-blue dewlap from an ancestral orange dewlap in 2 to 3 million years \citep{Macedonia2001}.
-We present evidence of multiple cases of potentially adaptive maintenance of habitat-associated dewlap divergence over small geographical scale in \textit{A.
+We present evidence of multiple cases of potentially adaptive maintenance of habitat-associated dewlap divergence over small geographical scales in \textit{A.
 sagrei} across the West Indies.
-While these intra-island populations do not appear to be in the process of speciation, our results suggest that the anoline dewlap has enough micro-scale, local adaptive potential to contribute of reproductive isolation, should it be recruited for assortative mating.
+While these intra-island populations do not appear to be in the process of speciation, our results suggest that the anoline dewlap has enough micro-scale, local adaptive potential to contribute to reproductive isolation, should it be recruited for assortative mating.

### 2020-07-13 ms/sections/methods.tex
almost ready to go

@@
 
-Machine learning for group comparison has become more popular in ecology and evolution in the recent years (e.g.
+Machine learning for group comparison has become more popular in ecology and evolution in recent years (e.g.
 \citealt{Pigot2020}).
@@
 For all islands where deviations from multivariate normality were detected in at least one habitat (Table \ref{suptab:multinorm}), we assessed univariate normality for each principal component (Shapiro-Wilk's test, Table \ref{suptab:normality}).
-For skewed PCs that deviated significantly from normality, we repeated the analysis using a nonparametric Kruskal-Wallis tests \citep{Hollander2013}.
+For skewed PCs that deviated significantly from normality, we repeated the analysis using a nonparametric Kruskal-Wallis test \citep{Hollander2013}.
 We found no multivariate outliers based on the Mahalanobis distance (package MVN, \citealt{Korkmaz2014}).
@@
 
-We tested for within-island spatial autocorrelation between the geographical distances among sampling sites and their euclidean distances in multivariate color space (mean PC1 to PC4 per site, Table \ref{suptab:sites}), regardless of habitat-type.
+We tested for within-island spatial autocorrelation between the geographical distances among sampling sites and their Euclidean distances in multivariate color space (mean PC1 to PC4 per site, Table \ref{suptab:sites}), regardless of habitat-type.
 Because often only a few sites were sampled per island, we could not get meaningful results from tests that use sites as units of observation, such as Moran's I test \citep{Gittleman1990}.
@@
 In this study, we were interested in the minimum spatial scale at which significant differences between habitats could be detected within islands.
-We performed multiple pairwise nonparametric Wilcoxon-Mann-Whitney's tests \citep{Hollander2013} to compare dewlap coloration between sites with different habitat-types, for each pair of habitats and each variable where significant differences were detected with our analyses of variances.
+We performed multiple pairwise nonparametric Wilcoxon-Mann-Whitney tests \citep{Hollander2013} to compare dewlap coloration between sites with different habitat-types, for each pair of habitats and each variable where significant differences were detected with our analyses of variances.
 The P-values were adjusted using a Benjamini-Hochberg correction for multiple testing \citep{Benjamini1995}.

### 2020-07-13 ms/sections/results.tex
almost ready to go

@@
 
-We tested for variation in \textit{A sagrei} dewlap coloration between populations living in three characteristic habitat types across nine islands that span the West Indian range of the brown anole (Fig \ref{fig:map}, \ref{supfig:map}).
+We tested for variation in \textit{A.
+sagrei} dewlap coloration between populations living in three characteristic habitat types across nine islands that span the West Indian range of the brown anole (Fig \ref{fig:map}, \ref{supfig:map}).
 We found that most of the variation in coloration is partitioned between islands (two-way semi-parametric MANOVA, modified ANOVA-type statistic (MATS) = 2009.6, P < 0.001, Fig.
@@
 In contrast, differences in dewlap coloration between habitats were often detected in close geographical proximity.
-For example, Bimini, Cayman Brac and Little Cayman were among the smallest islands in our study (Fig.
+For example, Bimini, Cayman Brac, and Little Cayman were among the smallest islands in our study (Fig.
 \ref{supfig:map}).
 Indeed, most significant differences in dewlap coloration involved sites that were 5-10km apart.
-Our most extreme case of significant differences occurred for PC3 between the beach scrub site and a coppice site, separated from each other by a a few hundreds of meters at most on Bimini (multiple pairwise Wilcoxon-Mann-Whitney tests, Fig.
+Our most extreme case of significant differences occurred for PC3 between the beach scrub site and a coppice site, separated from each other by a few hundreds of meters at most on Bimini (multiple pairwise Wilcoxon-Mann-Whitney tests, Fig.
 \ref{supfig:distances}).\\

### 2020-07-14 ms/sections/abstract.tex
ready to send to Jonathan

@@
 We explored variation in dewlap coloration in the most widespread species of anole, \textit{Anolis sagrei}, across three characteristic habitats spanning the Bahamas and the Cayman Islands.
-Using reflectance spectrometry as well as supervised machine learning, we found some consistent differences in spectral properties of the dewlap between habitats within small islands.
+Using reflectance spectrometry as well as supervised machine learning, we found significant differences in spectral properties of the dewlap between habitats within small islands.
 Passive divergence in dewlap phenotype associated with isolation-by-distance did not explain our results.
@@
 sagrei} populations, as one would otherwise expect differences within islands to be erased by gene flow at such small geographical scales.
-Although these habitat-specific dewlap differences vary in magnitude and direction across islands, and islands themselves differ substantially, we found a suite of consistent archipelago-wide differences between habitat types, suggesting parallel responses to similar selective pressures.
+These habitat-specific dewlap differences varied in magnitude and direction across islands, and may suggest a role of sexual selection in driving these local patterns.
 While at present, populations from these different habitats probably experience too much gene flow to follow distinct evolutionary lineages, should additional barriers arise between habitat-specific populations, the observed disruptive selection on dewlap coloration may facilitate ecological speciation.

### 2020-07-14 ms/sections/discussion.tex
ready to send to Jonathan

@@
 However, the habitat-specific variation in dewlaps was not consistent between these islands.
-Although those results are consistent with adaptation at a very local scale, other evolutionary drivers could be at work, including phenotypic plasticity, random drift, or historical contingency, including multiple colonization events.
+Although those results are consistent with adaptation at a very local scale, other evolutionary drivers could be at work, such as phenotypic plasticity, random drift, or historical contingency, including multiple colonization events.
 We reject this last explanation because all of the island populations in this study are strictly monophyletic, implying a single colonization event per island (van de Schoot, unpublished thesis; \citealt{Driessens2017, Reynolds2020}).
@@
 Such small-scale differences would be unlikely under strong gene flow \citep{Richardson2014}.
-Our study islands lack geographic barriers to the movement of $A.
-sagrei$, which have been shown to be highly mobile \citep{Kamath2018}, implying widespread gene flow across sites and habitats.
+Our study islands lack geographic barriers to the movement of \textit{A.
+sagrei}, which have been shown to be highly mobile \citep{Kamath2018}, implying widespread gene flow across sites and habitats.
 Moreover, habitat-populations within each island were found to be non-monophyletic and often share identical haplotypes, based on phylogenetic analysis of mitochondrial DNA sequences (van de Schoot et al.
@@
 sagrei} on islands with more coexisting species of anoles.
-Our data are consistent with those previous results in suggesting that adaptation to local light conditions, or at least broad habitat types, may not be a major driver of the variation in dewlap coloration in \textit{A.
+Our data are consistent with those previous results in suggesting that adaptation to local light conditions, or at least broad habitat types, may not be a major driver of the within-island variation in dewlap coloration in \textit{A.
 sagrei}.\\

### 2020-07-14 ms/sections/introduction.tex
almost ready to go
ready to send to Jonathan

@@
 between islands \citep{Vanhooydonck2009, Driessens2017} or within large islands such as Puerto Rico \citep{Leal2002, Leal2004} or Hispaniola \citep{Ng2012, Ng2016}.
-These large scales or marine barriers should reduce gene flow \citep{Ng2011, Lambert2013, Richardson2014, Ng2017}.
+These large scales and marine barriers should reduce gene flow \citep{Ng2011, Lambert2013, Richardson2014, Ng2017}.
 That said, examples do exist of divergence in dewlap coloration at smaller scales or between populations with high degrees of gene flow \citep{Thorpe2002a, Thorpe2002b, Stapley2011, Ng2016}.\\
@@
 
-Here, we analyzed the color characteristics \textit{A.
+Here, we analyzed the color characteristics of \textit{A.
 sagrei} dewlaps within nine islands in the Bahamas and Cayman Islands, combining reflectance spectrometry and supervised machine learning.

### 2020-07-14 ms/sections/methods.tex
ready to send to Jonathan

@@
 
-Significant post hoc contrasts were assessed using Tukey's Honest Significant Difference (HSD) test whenever the assumptions of normality and homogeneity of variances was met \citep{Tukey1949}, Dunnett's T3 method when only homogeneity of variances was violated but not normality \citep{Dunnett1980}, and Nemenyi's test when normality was violated \citep{Nemenyi1963}.
-All post hoc tests were performed with the R package PMCMRplus \citep{Pohlert2020}.\\
+Significant \textit{post hoc} contrasts were assessed using Tukey's Honest Significant Difference (HSD) test whenever the assumptions of normality and homogeneity of variances was met \citep{Tukey1949}, Dunnett's T3 method when only homogeneity of variances was violated but not normality \citep{Dunnett1980}, and Nemenyi's test when normality was violated \citep{Nemenyi1963}.
+All \textit{post hoc} tests were performed with the R package PMCMRplus \citep{Pohlert2020}.\\
 

### 2020-07-14 ms/sections/results.tex
ready to send to Jonathan

@@
 
-The small archipelago-wide effect of habitat-type was detected for PC1, PC2 and PC3 (mixed-effect ANOVA with island as a random effect, Table \ref{suptab:anova-pooled}), but this effect was too small for post hoc tests to find which habitats differed.
+The small archipelago-wide effect of habitat-type was detected for PC1, PC2 and PC3 (mixed-effect ANOVA with island as a random effect, Table \ref{suptab:anova-pooled}), but this effect was too small for \textit{post hoc} tests to find which habitats differed.
 Archipelago-wide differences in dewlap coloration between habitats were also detected by SVMs trained on pooled data regardless of island identity, both for PCA data and reflectance scores (Fig.
@@
 Indeed, most significant differences in dewlap coloration involved sites that were 5-10km apart.
-Our most extreme case of significant differences occurred for PC3 between the beach scrub site and a coppice site, separated from each other by a few hundreds of meters at most on Bimini (multiple pairwise Wilcoxon-Mann-Whitney tests, Fig.
+Our most extreme case of significant differences occurred for PC3 between a beach scrub site and a coppice site, separated from each other by a few hundreds of meters at most on Bimini (multiple pairwise Wilcoxon-Mann-Whitney tests, Fig.
 \ref{supfig:distances}).\\

### 2020-07-31 ms/sections/discussion.tex
wuick

@@
 However, the habitat-specific variation in dewlaps was not consistent between these islands.
-Although those results are consistent with adaptation at a very local scale, other evolutionary drivers could be at work, such as phenotypic plasticity, random drift, or historical contingency, including multiple colonization events.
+Although those results are consistent with adaptation at a very local scale, other evolutionary drivers could be at work, such as phenotypic plasticity, random drift, or multiple colonization events.
 We reject this last explanation because all of the island populations in this study are strictly monophyletic, implying a single colonization event per island (van de Schoot, unpublished thesis; \citealt{Driessens2017, Reynolds2020}).
@@
 better immune response to disease) and of indirect benefit to the offspring.
+Several studies suggest that this possibility could be the case for anoles.
 For example, \citet{Cook2013} found lower orange reflectance in dewlaps with heavily parasitized \textit{A.
@@
 
-\paragraph{A role of phenotypic plasticity is unlikely} Differences in coloration between habitat populations may not be genetically determined, and may be influenced by environmental factors such as parasite load, as mentioned above \citep{Cook2013}.
+\paragraph{A role of phenotypic plasticity is unlikely} Differences in coloration between habitat populations may not be genetically determined and instead may be influenced by environmental factors such as parasite load, as mentioned above \citep{Cook2013}.
 The yellow, orange and red coloration in anoline dewlaps are produced by pterins and carotenoids \citep{Ortiz1962, Ortiz1962a, Ortiz1963, Ortiz1966, Macedonia2000, Steffen2007, Steffen2009}.
@@
 Furthermore, developmental plasticity during the ontogeny is also unlikely because dewlap coloration develops at sexual maturity in anoles \citep{Ng2013}.
-The differences we observed could therefore be heritable.
+The differences we observed could therefore be genetically based.
 This hypothesis is further supported by \citet{Cox2017}, who found a high degree of heritability of dewlap coloration in \textit{A.

### 2020-07-31 ms/sections/introduction.tex
wuick

@@
 
-Lizards of the neotropical genus \textit{Anolis} are a model system for studying the eco-evolutionary dynamics of local adaptation and natural selection \citep{Losos2009}.
-A particularly conspicuous trait of anoles is their dewlap; an extensible flap of skin that is typically sexually dimorphic and used as a communication signal in courtship \citep{Sigmund1983, Driessens2014, Driessens2015}, competition \citep{Losos1985, Macedonia1994, Macedonia2013} as well as in predator deterrence \citep{Leal1995, Leal1997, Leal1997b}.
+Lizards of the neotropical genus \textit{Anolis} are an excellent group for studying the eco-evolutionary dynamics of local adaptation and natural selection \citep{Losos2009}.
+A particularly conspicuous trait of anoles is their dewlap; an extensible flap of skin that is typically sexually dimorphic and used as a communication signal in courtship \citep{Sigmund1983, Driessens2014, Driessens2015}, territorial displays \citep{Losos1985, Macedonia1994, Macedonia2013} as well as in predator deterrence \citep{Leal1995, Leal1997, Leal1997b}.
 Dewlap characteristics vary widely among the approximately $400$ species of the genus \citep{Nicholson2007}.
@@
 The species \textit{Anolis sagrei} is widespread across islands of the West Indies \citep{Reynolds2020}.
-It is a model organism in studies of local adaptation \citep{Losos1994, Losos1997a, Losos2001, Kolbe2012}, biological invasion \citep{Kolbe2008} and sexual selection \citep{Tokarz2002, Tokarz2005, Tokarz2006, Driessens2014, Steffen2014, Driessens2015}.
-Between-island variation in the mainly orange-red color of its dewlap was shown to be better explained by climatic variables \citep{Driessens2017} than biotic factors such as sexual selection or predation pressure \citep{Vanhooydonck2009, Baeckens2018}.
-How intra-island differences in habitat may contribute to the diversity of dewlap coloration, however, remains unexplored, and may reveal new insights into the scale of local adaptation despite gene flow.\\
+It has been the subject of study concerning local adaptation \citep{Losos1994, Losos1997a, Losos2001, Kolbe2012}, biological invasion \citep{Kolbe2008}, sexual selection \citep{Tokarz2002, Tokarz2005, Tokarz2006, Driessens2014, Steffen2014, Driessens2015} and many other topics.
+Between-island variation in the mainly orange-red color of its dewlap was shown to be better explained by climatic variables \citep{Driessens2017} than by proxies for biotic factors such as sexual selection or predation pressure \citep{Vanhooydonck2009, Baeckens2018}.
+How intra-island differences in habitat may contribute to the diversity of dewlap coloration, however, remains unexplored, and may reveal new insights into the scale of local differentiation despite gene flow.\\
 
@@
 These habitats contrast in environmental parameters including vegetation community, light irradiance, humidity and temperature \citep{Howard1950, Schoener1968}.
-Each of these island groups (Caymans and Bahamas) has been colonized independently by \textit{A.
-sagrei} from Cuba (\citealt{Driessens2017, Reynolds2020}, van de Schoot et al.
+The Cayman Islands and the Bahamas have been colonized independently by \textit{A.
+sagrei} from Cuba (\citealt{Reynolds2020}, van de Schoot et al.
 unpubl.), such that these archipelagos constitute an ideal suite of natural replicates to explore within-island dewlap diversity across multiple islands.\\
@@
 sagrei}.
-We predicted that if light conditions in the environment indeed drive color evolution, dewlaps should be most similar between beach scrub and mangrove forest, which both have high levels of light irradiance, contrary to the darker, closed-canopy coppice forest.
+We predicted that if light conditions in the environment indeed drive color evolution, dewlaps should be most similar between beach scrub and mangrove forest, which both have high levels of light irradiance, compared to the darker, closed-canopy coppice forest.
 Similar, if detectability is maximized given the local conditions, we expected darker and more contrasting dewlaps in high irradiance habitats.

### 2020-07-31 ms/sections/methods.tex
wuick

@@
 The median distance between two localities within an island was $11.18$km, with some islands being sampled at smaller or larger scales (Figure \ref{supfig:map}, Table \ref{suptab:sites}).
-$80.3$\% of all pairwise distances within islands were below $50$km.
+$80.3$\% of all pairwise distances within islands were less than $50$km.
 Additionally, there are no major barriers to dispersal (such as mountains or grassland) on any of the islands that we sampled.
@@
 
-Machine learning for group comparison has become more popular in ecology and evolution in recent years (e.g.
+Machine learning for group comparison has become more common in ecology and evolution in recent years (e.g.
 \citealt{Pigot2020}).

### 2020-07-31 ms/sections/results.tex
wuick

@@
 Differentiation in dewlap coloration occurred in multiple dimensions of color space.
-Moreover, the differences in dewlaps between habitats were not always consistent between islands, thus, we will discuss the habitat-specific variation in dewlap coloration for each island where significant differences were detected in turn (Fig.
+Moreover, the differences in dewlaps between habitats were not always consistent among islands, thus, we will discuss the habitat-specific variation in dewlap coloration for each island where significant differences were detected in turn (Fig.
 \ref{fig:anova}, Tables \ref{tab:anova}, \ref{suptab:kruskal}).

### 2020-07-31 ms/sections/introduction.tex
implementing comments

@@
 Dewlap characteristics vary widely among the approximately $400$ species of the genus \citep{Nicholson2007}.
-Interspecific variation in dewlap coloration is implicated in species recognition \citep{Rand1970b, Williams1969, Williams1977, Losos1985, Macedonia1994, Fleishman2000, Macedonia2013}, and possibly involved in speciation \citep{Lambert2013, Geneva2015, Ng2017}.\\
+Interspecific variation in dewlap coloration is implicated in species recognition \citep{Rand1970b, Williams1969, Williams1977, Losos1985, Macedonia1994, Fleishman2000, Macedonia2013}, and this function could have had a role in initiating and/or reinforcing reproductive isolation during speciation \citep{Lambert2013, Geneva2015, Ng2017}.\\
 
@@
 
-The island bank systems of the Bahamas and Cayman Islands presently, if not historically, comprise relatively small islands, with no major geographic barriers within islands limiting dispersal for this promiscuous species \citep{Kamath2018}.
+% What we did
+
+Here, we analyzed the color characteristics of \textit{A.
+sagrei} dewlaps within nine islands in the Bahamas and Cayman Islands.
+These island systems presently, if not historically, comprise relatively small islands, with no major geographic barriers within islands limiting dispersal for this promiscuous species \citep{Kamath2018}.
 These islands all share three characteristic native West Indian habitat-types -- beach scrub bush, closed-canopy primary coppice forest, and mangrove forest -- that are often spatially intermingled.
@@
 
-% What we did
-
-Here, we analyzed the color characteristics of \textit{A.
-sagrei} dewlaps within nine islands in the Bahamas and Cayman Islands, combining reflectance spectrometry and supervised machine learning.
 Our sampling design included sites in close proximity (the median distance between two sites within an island was $11.2$km).
-We tested for divergence in dewlap phenotype between habitats within islands and between islands across the range of \textit{A.
+Combining reflectance spectrometry and supervised machine learning, we tested for divergence in dewlap phenotype between habitats within islands and between islands across the range of \textit{A.
 sagrei}.
@@
 Finally, if habitat characteristics are strong determinants of dewlap color variation, similar patterns should be observed across multiple islands \citep{Losos2011}.
-We found strong support for fine-scale, within-island differences in coloration between lizards inhabiting the three habitat-types in several color space dimensions, suggesting a potentially strong effect of divergent selection.
+
+% Jonathan does not like when we sum up the results at the end of the intro
+
+% We found strong support for fine-scale, within-island differences in coloration between lizards inhabiting the three habitat-types in several color space dimensions, suggesting a potentially strong effect of divergent selection.
 However, the divergence patterns we observed did not match our \textit{a priori} predictions and were inconsistent between islands.

### 2020-07-31 ms/sections/methods.tex
implementing comments

@@
 We sampled 466 male \textit{Anolis sagrei} from seven islands in the Bahamas Archipelago -- Abaco, North Andros, South Andros, South Bimini, Eleuthera, Long Island, Ragged Island -- and two in the Cayman Islands -- Cayman Brac and Little Cayman (Figure \ref{fig:map}).
-These islands and island banks were chosen to span the breadth of the West Indian range of \textit{Anolis sagrei}, because they have highly similar habitat types,  and because the \textit{A.
+These islands were chosen to span the breadth of the West Indian range of \textit{Anolis sagrei}, because they have highly similar habitat types,  and because the \textit{A.
 sagrei} on each island group are derived from ancient and distinct colonization events from Cuba (i.e.
@@
 Our sampling design enabled us to test for differences between habitats at a coarse and fine geographical scale.
-The median distance between two localities within an island was $11.18$km, with some islands being sampled at smaller or larger scales (Figure \ref{supfig:map}, Table \ref{suptab:sites}).
+The median distance between two localities within an island was $\sim 11$km, with some islands being sampled at smaller or larger scales (Figure \ref{supfig:map}, Table \ref{suptab:sites}).
 $80.3$\% of all pairwise distances within islands were less than $50$km.

### 2020-07-31 ms/sections/results.tex
implementing comments

@@
 \ref{supfig:reflectance}, explained variance $\eta^2 = 44.3$\%, MANOVA approximation).
-Nonetheless, we did find evidence for differences in dewlap coloration between habitat-types, and those were mostly island-specific (habitat-by-island interaction term, MATS = 384.4, P < 0.001, explained variance $\eta^2 = 11.4$\%), leaving a small but significant portion of the variation explained by an archipelago-wide habitat effect (MATS = 42.5, P = 0.001, $\eta^2 = 4.8$\%).\\
+Nonetheless, we did find evidence for differences in dewlap coloration between habitat-types, and those were mostly island-specific (habitat-by-island interaction term, MATS = 384.4, P < 0.001, explained variance $\eta^2 = 11.4$\%), with a small but significant portion of the variation explained by an archipelago-wide habitat effect (MATS = 42.5, P = 0.001, $\eta^2 = 4.8$\%).\\
 

### 2020-08-02 ms/main.tex
working with tables
second review
done implemented the feedback left by Jonathan in the ms

@@
     \textsc{Rapha\"{e}l Scherrer$^{1,3}$ \thanks{Corresponding author: r.scherrer@rug.nl}, Colin M. Donihue$^{1,4}$, }\\
-	\textsc{R. Graham Reynolds$^2$, Jonathan B. Losos$^{1,4}$ and Anthony J. Geneva$^1$} \\[1ex]
+	\textsc{R. Graham Reynolds$^2$, Jonathan B. Losos$^{1,4}$ and Anthony J. Geneva$^{1,5}$} \\[1ex]
 	%\thanks{A thank you} \\[1ex] % Your name
 	\normalsize $^1$ Department of Organismic and Evolutionary Biology and Museum of Comparative Zoology \\ \normalsize Harvard University, Cambridge, MA, USA \\
-	\normalsize $^2$ Department of Biology, University of North Carolina Asheville, Asheville, NC, USA\\ % Your institutionhttps://www.overleaf.com/project/5a0705e5f4573c1ace38dbdf
+	\normalsize $^2$ Department of Biology, University of North Carolina Asheville, Asheville, NC, USA\\ 
 	\normalsize $^3$ Current address: Groningen Institute for Evolutionary Life Sciences,\\
@@
 	\normalsize $^4$ Current address: Department of Biology, Washington University, St. Louis, MO, USA\\
+	\normalsize $^5$ Current address: Department of Biology, Center for Computational and Integrative Biology,\\ \normalsize Rutgers University--Camden, Camden, NJ, USA
 }
@@
 \section*{Introduction}
-
-\begin{table}
-	\input{../analyses/01-counts/table_counts.tex}
-\end{table}
 

### 2020-08-02 ms/sections/discussion.tex
done implemented the feedback left by Jonathan in the ms

@@
 
-Besides, although this process is well-established in some systems such as African cichlids (e.g.
-\citealt{Seehausen1997}), previous work suggests that dewlap coloration in \textit{Anolis} may act as a mating cue according to a "good genes" model of sexual selection \citep{Andersson1994}, rather than a Fisherian process.
+Alternatively, dewlap coloration in \textit{Anolis} may act as a mating cue according to a "good genes" model of sexual selection, rather than a Fisherian process.
 In the good genes model, the cue under sexual selection is an indicator of individual quality (e.g.
-better immune response to disease) and of indirect benefit to the offspring.
+better immune response to disease) and of indirect benefit to the offspring \citep{Andersson1994}.
 Several studies suggest that this possibility could be the case for anoles.
@@
 This hypothesis is further supported by \citet{Cox2017}, who found a high degree of heritability of dewlap coloration in \textit{A.
-sagrei}.
-Although most studies used one or two-generation common garden experiments and thus could not rule out transgenerational plastic effects \citep{Tariel2020}, dewlap coloration generally seems to not be a plastic trait.
-This further reinforces an adaptive explanation, where dewlap color could be under differential natural and/or sexual selection in these different habitats.\\
+sagrei}.\\ % playing it safer here by not saying it applies to all anoles, but also not mentioning the transgenerational effects anymore
+
+% For these reasons, we consider implausible that dewlap coloration is a plastic trait in our study species.\\
+
+% Too broad a statemtent to make
+% Although most studies used one or two-generation common garden experiments and thus could not rule out transgenerational plastic effects \citep{Tariel2020}, dewlap coloration generally seems to not be a plastic trait.
+
+This further reinforces an adaptive explanation, where dewlap color could be under differential natural and/or sexual selection in these different habitats.
 
@@
 Evidence suggests that dewlap coloration could take this role in anoles \citep{Ng2011, Lambert2013, Geneva2015, Ng2017}, or at least that it is frequently involved in species recognition \citep{Williams1969, Williams1977, Losos1985, Macedonia1994, Fleishman2000, Macedonia2013, Ingram2016, Baeckens2018}.
-Although this signal is not detected at the phylogenetic scale of the whole genus \citep{Nicholson2007, Harrison2012, Ingram2016}, sexual signals are often evolutionarily very labile \citep{Kraaijeveld2011}, and the anole dewlap in particular is capable of rapid macroevolution; for example, \textit{A.
+Although a correlation between dewlap coloration and reproductive isolation is not detected at the phylogenetic scale of the whole genus \citep{Nicholson2007, Harrison2012, Ingram2016}, sexual signals are often evolutionarily very labile \citep{Kraaijeveld2011}, and the anole dewlap in particular is capable of rapid macroevolution throughout its macroevolutionary history \citep{Nicholson2007}.
+For example, \textit{A.
 conspersus} on Grand Cayman evolved a UV-blue dewlap from an ancestral orange dewlap in 2 to 3 million years \citep{Macedonia2001}.

### 2020-08-02 ms/sections/results.tex
done implemented the feedback left by Jonathan in the ms

@@
 Sensitivity analyses on these machines suggest a relatively small role of long wavelengths (red reflectance) in driving this pattern (Fig.
-\ref{supfig:importance-svm-refl-pooled}), but did not reveal strong differences between the PCs in relative importance (Fig.
+\ref{supfig:importance-svm-refl-pooled}) -- \textbf{be more explicit about how the colors vary among habitats} --, but did not reveal strong differences between the PCs in relative importance (Fig.
 \ref{supfig:importance-svm-pca-pooled}).
@@
 Differentiation in dewlap coloration occurred in multiple dimensions of color space.
-Moreover, the differences in dewlaps between habitats were not always consistent among islands, thus, we will discuss the habitat-specific variation in dewlap coloration for each island where significant differences were detected in turn (Fig.
+Moreover, the differences in dewlaps between habitats generally were not consistent among islands, thus, we will discuss the habitat-specific variation in dewlap coloration for each island where significant differences were detected in turn (Fig.
 \ref{fig:anova}, Tables \ref{tab:anova}, \ref{suptab:kruskal}).

### 2020-08-02 ms/sections/abstract.tex
massaging the discussion to put less emphasis on adaptation, as well as the abstract

@@
 Dewlap color in \textit{Anolis} lizards can be highly variable between populations in relation to both biotic and abiotic adaptive drivers, albeit at relatively large geographical scales.
+At smaller scales, gene flow is likely to erase phenotypic differentiation between localities.
 Here, we investigated local adaptation of the dewlap across habitat-types at a small spatial scale, as this may give an indication of how conditions for the early stages of speciation may be met.
 We explored variation in dewlap coloration in the most widespread species of anole, \textit{Anolis sagrei}, across three characteristic habitats spanning the Bahamas and the Cayman Islands.
-Using reflectance spectrometry as well as supervised machine learning, we found significant differences in spectral properties of the dewlap between habitats within small islands.
+Using reflectance spectrometry as well as supervised machine learning, we found significant differences in spectral properties of the dewlap between habitats within small islands, sometimes over very short distances.
 Passive divergence in dewlap phenotype associated with isolation-by-distance did not explain our results.
-Instead, the observed patterns in dewlap coloration are more consistent with an adaptive explanation in these \textit{A.
-sagrei} populations, as one would otherwise expect differences within islands to be erased by gene flow at such small geographical scales.
-These habitat-specific dewlap differences varied in magnitude and direction across islands, and may suggest a role of sexual selection in driving these local patterns.
-While at present, populations from these different habitats probably experience too much gene flow to follow distinct evolutionary lineages, should additional barriers arise between habitat-specific populations, the observed disruptive selection on dewlap coloration may facilitate ecological speciation.
+On the other hand, these habitat-specific dewlap differences varied in magnitude and direction across islands, and thus our primary test for adaptation -- parallel responses across islands -- was falsified.
+We suggest, however, that selection could be involved and present several ways this might have happened, including sexual selection.
+(This one may be too much of a leap) While at present, populations from these different habitats probably experience too much gene flow to follow distinct evolutionary lineages, should additional barriers arise between habitat-specific populations, the observed disruptive selection on dewlap coloration may facilitate speciation.

### 2020-08-02 ms/sections/discussion.tex
massaging the discussion to put less emphasis on adaptation, as well as the abstract

@@
 % There are inconsistent differences that are not explained by distance
+
+\textbf{Jonathan's take: too much push in favor of selection, without really having much evidence.
+Simple models of genetic drift are ruled out and we don’t think it is plasticity.
+However, our primary test for adaptation—looking for parallel responses across islands—also was falsified.
+We suggest, however, that selection could be involved, and suggest several ways this might have happened.}
 
@@
 However, the habitat-specific variation in dewlaps was not consistent between these islands.
-Although those results are consistent with adaptation at a very local scale, other evolutionary drivers could be at work, such as phenotypic plasticity, random drift, or multiple colonization events.
+Although those results are consistent with selection acting at a very local scale, other evolutionary drivers could be at work, such as phenotypic plasticity, random drift, or multiple colonization events.
 We reject this last explanation because all of the island populations in this study are strictly monophyletic, implying a single colonization event per island (van de Schoot, unpublished thesis; \citealt{Driessens2017, Reynolds2020}).
@@
 Moreover, habitat-populations within each island were found to be non-monophyletic and often share identical haplotypes, based on phylogenetic analysis of mitochondrial DNA sequences (van de Schoot et al.
-unpublished thesis), suggesting gene flow between habitats may be widespread.\\
-
+unpublished thesis), suggesting gene flow between habitats may be widespread.
 Our results align with previous documented cases of persistent dewlap color divergence despite gene flow in multiple species of anoles, sometimes in relation to environmental conditions.
@@
 
-\paragraph{Dewlap coloration could be locally adapted} Although phenotypic divergence at small spatial scales may be an indicator of local adaptation, the inconsistency of the between-habitat divergence patterns we observed across islands implies that neutral processes cannot be definitely ruled out \citep{Losos2011}.
+% \paragraph{No conclusive evidence for local adaptation of dewlap coloration} Although phenotypic divergence at small spatial scales may be an indicator of local adaptation, the inconsistency of the between-habitat divergence patterns we observed across islands implies that neutral processes cannot be definitely ruled out \citep{Losos2011}.
 Conversely, the absence of parallel divergence across islands does not rule out local adaptation either \citep{Losos2011}.
@@
 
+\paragraph{No conclusive evidence for local adaptation of dewlap coloration} One of the most informative tests for adaptation is the convergence of differentiation patterns across replicate islands or localities \citep{Losos2009, Losos2011}.
 Previous studies have described convergent patterns of dewlap color evolution in similar environments across islands and species \citep{Thorpe2002a, Thorpe2002b}.
@@
 sagrei}, where instead, we found the darkest dewlaps in the dark, mesic habitat -- primary coppice forest -- on three islands, and dewlaps often differed the most between beach scrub and mangrove forest, two xeric habitats with similar, high irradiance levels \citep{Howard1950, Schoener1968}.
-The inconsistent and idiosyncratic patterns we observed suggest that dewlap color variation between habitats cannot be predicted by habitat identity alone.
+The inconsistent and idiosyncratic patterns we observed suggest that dewlap color variation between habitats cannot be predicted by habitat identity alone, and the test of convergence across islands is rejected.
 Studies of Jamaican and Hispaniolan anoles similarly found between-habitat differences in dewlap coloration but no evidence for higher dewlap detectability in different habitats \citep{Fleishman2009, Ng2012}.
@@
 sagrei} on islands with more coexisting species of anoles.
-Our data are consistent with those previous results in suggesting that adaptation to local light conditions, or at least broad habitat types, may not be a major driver of the within-island variation in dewlap coloration in \textit{A.
-sagrei}.\\
+Our data are consistent with those previous results in suggesting that adaptation to local light conditions, or at least broad habitat types, is not a major driver of the within-island variation in dewlap coloration in \textit{A.
+sagrei}.
 
-% The evidence for sexual selection
+\paragraph{Sexual selection could be at play} Selection, however, needs not necessarily be linked to habitat type, and may take the form of arbitrary, "Fisherian" sexual selection, where female preferences differ between localities for reasons unrelated to the environment, driving the divergent evolution of male ornaments \citep{Andersson1994}.
+This process could account for the idiosyncratic patterns of within-island divergence we report, where initial differences in female preferences could have arisen for nonselective reasons (e.g.
+plasticity or random drift).
+Substantial levels of promiscuity in \textit{A.
+sagrei} suggest ample opportunity for female mate choice \citep{Kamath2018}, and are in line with this scenario.
+However, \citet{Baeckens2018} found no link between \textit{A.
+sagrei} dewlap coloration and size dimorphism (a proxy for sexual selection) in an among-island study of the same archipelagos.
+Another form of sexual selection is the "good genes" model, where the cue under sexual selection is an indicator of individual quality (e.g.
+better immune response to disease) and of indirect benefit to the offspring \citep{Andersson1994}.
+Several studies suggest that this possibility could be the case for anoles.
+For example, \citet{Cook2013} found lower orange reflectance in dewlaps with heavily parasitized \textit{A.
+brevirostris}, suggesting a trade-off in carotenoid use between the immune response and pigment deposition.
+\citet{Steffen2014} found that lower UV and orange-red reflectance predict contest-winning success between \textit{A.
+sagrei} males, while \citet{Driessens2015} further found that more yellow and red dewlaps (relative to UV) predict better body condition, and that higher yellow and UV reflectance at the margin of the dewlap predict higher hematocrit (the concentration of red blood cells), indicating a better health.
+Other aspects of the dewlap than color have also been found to be indicators of individual quality, such as dewlap size \citep{Vanhooydonck2005, Vanhooydonck2009}, but not dewlap display frequency \citep{Tokarz2002, Tokarz2005, Driessens2015}.
+If dewlap coloration is indeed an indicator of male quality in \textit{A.
+sagrei}, however, this would explain the observed patterns only if cues associated with good genes differ between sites, which is unlikely as it implies an association of male fitness with different dewlap coloration in different habitats, the evolution of which may be counteracted by gene flow (citation to say that if gene flow counters the divergence of one trait, then it is even worse for the association of two traits).
+Basically, one would expect good genes-sexual selection to favor the same coloration across the archipelago, and would therefore not result in local differences between habitats.
 
-\paragraph{Sexual selection could be at play} Sexual selection is an important component of an adaptive, sensory drive hypothesis of dewlap color divergence across environmental conditions, because in this scenario habitat-dependent selection is executed by the signal recipients \citep{Endler1988}, such as mating partners \citep{Driessens2014}.
-An alternative yet non-mutually exclusive explanation for the observed idiosyncratic patterns of within-island divergence is that of "Fisherian" sexual selection \citep{Andersson1994}.
-Under the Fisherian model of sexual selection, arbitrary female preferences (i.e.
+% An alternative yet non-mutually exclusive explanation for the observed idiosyncratic patterns of within-island divergence is that of "Fisherian" sexual selection \citep{Andersson1994}.
+
+% Under the Fisherian model of sexual selection, arbitrary female preferences (i.e.
 independent of the environment) for certain male ornaments may drive divergent evolution of male signals, such as dewlap coloration, if female preferences differ between localities for other reasons than environment-dependent perception abilities (a situation that could have arisen e.g.
@@
 
-Alternatively, dewlap coloration in \textit{Anolis} may act as a mating cue according to a "good genes" model of sexual selection, rather than a Fisherian process.
+% Alternatively, dewlap coloration in \textit{Anolis} may act as a mating cue according to a "good genes" model of sexual selection, rather than a Fisherian process.
 In the good genes model, the cue under sexual selection is an indicator of individual quality (e.g.
@@
 This hypothesis is further supported by \citet{Cox2017}, who found a high degree of heritability of dewlap coloration in \textit{A.
-sagrei}.\\ % playing it safer here by not saying it applies to all anoles, but also not mentioning the transgenerational effects anymore
+sagrei}.
+Response to parasite load is definitely an example of plasticity!\\ % playing it safer here by not saying it applies to all anoles, but also not mentioning the transgenerational effects anymore
 
@@
 
-This further reinforces an adaptive explanation, where dewlap color could be under differential natural and/or sexual selection in these different habitats.
+% This further reinforces an adaptive explanation, where dewlap color could be under differential natural and/or sexual selection in these different habitats.
 
@@
 conspersus} on Grand Cayman evolved a UV-blue dewlap from an ancestral orange dewlap in 2 to 3 million years \citep{Macedonia2001}.
-We present evidence of multiple cases of potentially adaptive maintenance of habitat-associated dewlap divergence over small geographical scales in \textit{A.
+We present evidence of multiple cases of within-island dewlap divergence over small geographical scales in \textit{A.
 sagrei} across the West Indies.

### 2020-08-02 ms/sections/discussion.tex
eod

@@
 
-\textbf{Jonathan's take: too much push in favor of selection, without really having much evidence.
+%\textbf{Jonathan's take: too much push in favor of selection, without really having much evidence.
 Simple models of genetic drift are ruled out and we don’t think it is plasticity.
@@
 
-\paragraph{No conclusive evidence for local adaptation of dewlap coloration} One of the most informative tests for adaptation is the convergence of differentiation patterns across replicate islands or localities \citep{Losos2009, Losos2011}.
+\paragraph{No conclusive evidence for adaptation of dewlap coloration to habitat type} One of the most informative tests for adaptation is the convergence of differentiation patterns across replicate islands or localities \citep{Losos2009, Losos2011}.
 Previous studies have described convergent patterns of dewlap color evolution in similar environments across islands and species \citep{Thorpe2002a, Thorpe2002b}.
-Others have suggested that dewlap coloration may have evolved to be maximally detectable under local light conditions imposed by the environment, primarily through UV contrast (i.e.
+However, the inconsistent and idiosyncratic patterns we observed reject the test of convergence across islands, suggesting that dewlap color variation between habitats cannot be predicted by habitat identity alone in \textit{A.
+sagrei}.
+Other studies have proposed that dewlap coloration may have evolved to be maximally detectable under local light conditions imposed by the environment, primarily through UV contrast (i.e.
 UV-brighter dewlaps in UV-dark, mesic habitats and UV-darker dewlaps in UV-bright, xeric habitats), in \textit{A.
@@
 sagrei}, where instead, we found the darkest dewlaps in the dark, mesic habitat -- primary coppice forest -- on three islands, and dewlaps often differed the most between beach scrub and mangrove forest, two xeric habitats with similar, high irradiance levels \citep{Howard1950, Schoener1968}.
-The inconsistent and idiosyncratic patterns we observed suggest that dewlap color variation between habitats cannot be predicted by habitat identity alone, and the test of convergence across islands is rejected.
 Studies of Jamaican and Hispaniolan anoles similarly found between-habitat differences in dewlap coloration but no evidence for higher dewlap detectability in different habitats \citep{Fleishman2009, Ng2012}.
+Our data are consistent with those previous results in suggesting that adaptation to local light conditions, or at least broad habitat types, is not a major driver of the within-island variation in dewlap coloration in \textit{A.
+sagrei}.
 Habitats on different islands may also differ in aspects other than light conditions, such as densities of predators or congeners, which have been shown to affect among-island dewlap diversity \citep{Vanhooydonck2009, Baeckens2018}.
@@
 sagrei} on islands with more coexisting species of anoles.
-Our data are consistent with those previous results in suggesting that adaptation to local light conditions, or at least broad habitat types, is not a major driver of the within-island variation in dewlap coloration in \textit{A.
-sagrei}.
+Therefore, if local adaptation occurs, it is more likely to involve components of the environment that do not encompass our broad habitat categories.
 
@@
 Other aspects of the dewlap than color have also been found to be indicators of individual quality, such as dewlap size \citep{Vanhooydonck2005, Vanhooydonck2009}, but not dewlap display frequency \citep{Tokarz2002, Tokarz2005, Driessens2015}.
-If dewlap coloration is indeed an indicator of male quality in \textit{A.
-sagrei}, however, this would explain the observed patterns only if cues associated with good genes differ between sites, which is unlikely as it implies an association of male fitness with different dewlap coloration in different habitats, the evolution of which may be counteracted by gene flow (citation to say that if gene flow counters the divergence of one trait, then it is even worse for the association of two traits).
-Basically, one would expect good genes-sexual selection to favor the same coloration across the archipelago, and would therefore not result in local differences between habitats.
+However, one would expect good genes sexual selection to favor the same coloration across the archipelago, and therefore this mechanism is unlikely to have produced idiosyncratic patterns of within-island divergence.
 
@@
 sagrei}.
-Response to parasite load is definitely an example of plasticity!\\ % playing it safer here by not saying it applies to all anoles, but also not mentioning the transgenerational effects anymore
+% playing it safer here by not saying it applies to all anoles, but also not mentioning the transgenerational effects anymore
 
@@
 Evidence suggests that dewlap coloration could take this role in anoles \citep{Ng2011, Lambert2013, Geneva2015, Ng2017}, or at least that it is frequently involved in species recognition \citep{Williams1969, Williams1977, Losos1985, Macedonia1994, Fleishman2000, Macedonia2013, Ingram2016, Baeckens2018}.
-Although a correlation between dewlap coloration and reproductive isolation is not detected at the phylogenetic scale of the whole genus \citep{Nicholson2007, Harrison2012, Ingram2016}, sexual signals are often evolutionarily very labile \citep{Kraaijeveld2011}, and the anole dewlap in particular is capable of rapid macroevolution throughout its macroevolutionary history \citep{Nicholson2007}.
+Although a correlation between dewlap coloration and reproductive isolation is not detected at the phylogenetic scale of the whole genus \citep{Nicholson2007, Harrison2012, Ingram2016}, sexual signals are often evolutionarily very labile \citep{Kraaijeveld2011}, and the anole dewlap in particular has been capable of rapid macroevolution throughout its phylogenetic history \citep{Nicholson2007}.
 For example, \textit{A.

### 2020-08-08 ms/main.tex
jonathans comments implemented

@@
 
-\title{Dewlap color variation in \textit{Anolis sagrei} is maintained between habitats within islands of the West Indies}
+\title{Dewlap color variation in \textit{Anolis sagrei} is maintained among habitats within islands of the West Indies}
 \author{
@@
 	\normalsize $^3$ Current address: Groningen Institute for Evolutionary Life Sciences,\\
-	\normalsize Groningen, The Netherlands\\
+	\normalsize University of Groningen, Groningen, The Netherlands\\
 	\normalsize $^4$ Current address: Department of Biology, Washington University, St. Louis, MO, USA\\
@@
 
-\textbf{Keywords} --- \textit{Anolis}, reflectance, local adaptation, sexual signal, supervised machine learning
+\textbf{Keywords} --- reflectance, adaptation, sexual signal, machine learning, polymorphism
 

### 2020-08-08 ms/sections/abstract.tex
jonathans comments implemented

@@
 At smaller scales, gene flow is likely to erase phenotypic differentiation between localities.
-Here, we investigated local adaptation of the dewlap across habitat-types at a small spatial scale, as this may give an indication of how conditions for the early stages of speciation may be met.
+Here, we investigated differentiation of dewlap coloration among habitat-types at a small spatial scale, within multiple islands of the West Indies, as this may give new insights into the local scale at which adaptation is possible.
 We explored variation in dewlap coloration in the most widespread species of anole, \textit{Anolis sagrei}, across three characteristic habitats spanning the Bahamas and the Cayman Islands.
@@
 On the other hand, these habitat-specific dewlap differences varied in magnitude and direction across islands, and thus our primary test for adaptation -- parallel responses across islands -- was falsified.
-We suggest, however, that selection could be involved and present several ways this might have happened, including sexual selection.
-(This one may be too much of a leap) While at present, populations from these different habitats probably experience too much gene flow to follow distinct evolutionary lineages, should additional barriers arise between habitat-specific populations, the observed disruptive selection on dewlap coloration may facilitate speciation.
+We suggest, however, that selection could be involved in several ways, including sexual selection.
+Our results shed new light on the scale at which signal color polymorphism can be maintained in the presence of gene flow, and the relative role of local adaptation and other processes in driving these patterns.

### 2020-08-08 ms/sections/acknowledgments.tex
jonathans comments implemented

@@
 Glor, Alberto R.
-Puente-Rol\'{o}n, Kevin Aviles-Rodriguez, Kristin Winchell, Jason Fredette and Melissa Kemp for assistance in the field and Max Lambert and James Stroud for helpful discussions.
+Puente-Rol\'{o}n, Kevin Aviles-Rodriguez, Kristin Winchell, Jason Fredette and Melissa Kemp for assistance in the field and Pratik Gupte, Max Lambert and James Stroud for helpful discussions.
 Funding for this work was provided by the Templeton Foundation (to JBL), NSF DEB~\#1927194 (to JBL and AJG), NSF DEB~\#1500761 (to AJG), NSF DBI~\#1609284 (to CMD), and a Harvard Museum of Comparative Zoology Putnam Expedition Grant (to RGR).

### 2020-08-08 ms/sections/discussion.tex
jonathans comments implemented

@@
 
-%\textbf{Jonathan's take: too much push in favor of selection, without really having much evidence.
+%firef\textbf{Jonathan's take: too much push in favor of selection, without really having much evidence.
 Simple models of genetic drift are ruled out and we don’t think it is plasticity.
@@
 Finally, regionally-distinct body coloration, but not dewlap coloration, is present in \textit{A.
-conspersus} on another small island, Grand Cayman, where no physical barriers to gene flow exist \citep{Macedonia2001}.\\
+conspersus} on another small island, Grand Cayman, where no physical barriers to gene flow exist \citep{Macedonia2001}.
 
@@
 conspersus} on Grand Cayman evolved a UV-blue dewlap from an ancestral orange dewlap in 2 to 3 million years \citep{Macedonia2001}.
-We present evidence of multiple cases of within-island dewlap divergence over small geographical scales in \textit{A.
+We present evidence of multiple cases of within-island dewlap color divergence over small geographical scales in \textit{A.
 sagrei} across the West Indies.
-While these intra-island populations do not appear to be in the process of speciation, our results suggest that the anoline dewlap has enough micro-scale, local adaptive potential to contribute to reproductive isolation, should it be recruited for assortative mating.
+While these intra-island populations do not appear to be in the process of speciation, our results suggest that the anoline dewlap could have enough micro-scale, local adaptive potential to contribute to reproductive isolation, should it be recruited for assortative mating.

### 2020-08-08 ms/sections/results.tex
new figure on why did archipelago wide machines detect anything

@@
 \ref{supfig:classif-svm-pca-pooled}, \ref{supfig:classif-svm-refl-pooled}).
-This pattern seemed to be driven by mangrove lizards being correctly reassigned more often than predicted by chance.
-Sensitivity analyses on these machines suggest a relatively small role of long wavelengths (red reflectance) in driving this pattern (Fig.
-\ref{supfig:importance-svm-refl-pooled}) -- \textbf{be more explicit about how the colors vary among habitats} --, but did not reveal strong differences between the PCs in relative importance (Fig.
+This seemed to be driven primarily by mangrove lizards being correctly reassigned more often than predicted by chance.
+Sensitivity analyses on these machines suggest that wavelengths from a large range, between 300 and 600nm, overall played a more important role in successful identification than wavelengths above 600nm (Fig.
+\ref{supfig:importance-svm-refl-pooled}), and that PC3 and PC4 were more important than PC2 (Fig.
 \ref{supfig:importance-svm-pca-pooled}).
-Archipelago-wide differences were not detected by LDA classifiers at all (Fig.
-\ref{supfig:classif-lda-pca-pooled}, \ref{supfig:classif-lda-refl-pooled}).\\ %So, the differences could be in something other than means.\\
+This suggests that archipelago-wide differences may involve fine deviations from the main axes of variation (such as PC2) in multiple wavelengths, and possibly nonlinear combinations of wavelengths, that only the SVMs, not the LDAs, could pick up (Fig.
+\ref{supfig:classif-lda-pca-pooled}, \ref{supfig:classif-lda-refl-pooled}).
+This pattern was weak, with machine accuracy scores narrowly distributed around about $50\%$, which is suggestive of only small deviations and a large degree of overlap in color space (Fig.
+\ref{supfig:pooled} and \ref{supfig:reflectance-pooled}).\\
 

### 2020-08-08 analyses/02-reflectance/PlotReflectance.R

@@ -38,3 +38,35 @@ p <- ggplot(smr, aes(x = wl, color = habitat)) +
 p
 
 ggsave("figure_reflectance.png", height = 6, width = 6, dpi = 300)
+
+#####################
+
+# At the scale of the whole archipelago now
+
+# Reorganize into a long table
+smr <- data %>%
+  gather_("wl", "reflectance", paste0("wl", 300:700)) %>%
+  mutate(wl = as.numeric(gsub("wl", "", wl)))
+
+# Prepare ribbons
+smr <- smr %>% group_by(wl, habitat)
+smr <- smr %>%
+  summarize(
+    mid = quantile(reflectance, 0.05),
+    lower = median(reflectance),
+    upper = quantile(reflectance, 0.95)
+  )
+
+# Plot the spectra
+p <- ggplot(smr, aes(x = wl, color = habitat)) +
+  geom_ribbon(aes(ymin = lower, ymax = upper, fill = habitat), alpha = 0.2) +
+  #geom_line(aes(y = mid)) +
+  xlab("Wavelength (nm)") +
+  ylab("Reflectance (%)") +
+  theme_bw() +
+  labs(fill = "Habitat", color = "Habitat") +
+  scale_fill_manual(values = colors) +
+  scale_color_manual(values = colors)
+p
+
+ggsave("figure_reflectance_pooled.png", height = 2, width = 4, dpi = 300)

### 2020-08-08 ms/fetch_content.sh

@@ -21,6 +21,7 @@ cp ../analyses/10-distances/figure_distances2.png suppfigures
 cp ../analyses/10-distances/figure_distances.png suppfigures
 cp ../analyses/06-pooled/figure_pooled.png suppfigures
 cp ../analyses/02-reflectance/figure_reflectance.png suppfigures
+cp ../analyses/02-reflectance/figure_reflectance_pooled.png suppfigures
 cp ../analyses/04-machine-learning/plots/importance_lda_pca.png suppfigures
 cp ../analyses/04-machine-learning/plots/importance_lda_pca_pooled.png suppfigures
 cp ../analyses/04-machine-learning/plots/importance_lda_refl.png suppfigures

### 2020-10-28 ms/sections/abstract.tex
backup

@@
 Animal signals evolve in an ecological context.
-Moreover, locally adapting animal sexual signals can be especially important for initiating or reinforcing reproductive isolation during the early stages of speciation.
-Dewlap color in \textit{Anolis} lizards can be highly variable between populations in relation to both biotic and abiotic adaptive drivers, albeit at relatively large geographical scales.
-At smaller scales, gene flow is likely to erase phenotypic differentiation between localities.
+Locally adapting animal sexual signals can be especially important for initiating or reinforcing reproductive isolation during the early stages of speciation.
+Previous studies have demonstrated that dewlap color in \textit{Anolis} lizards can be highly variable between populations in relation to both biotic and abiotic adaptive drivers at relatively large geographical scales.
 Here, we investigated differentiation of dewlap coloration among habitat-types at a small spatial scale, within multiple islands of the West Indies, as this may give new insights into the local scale at which adaptation is possible.
@@
 Passive divergence in dewlap phenotype associated with isolation-by-distance did not explain our results.
-On the other hand, these habitat-specific dewlap differences varied in magnitude and direction across islands, and thus our primary test for adaptation -- parallel responses across islands -- was falsified.
+On the other hand, these habitat-specific dewlap differences varied in magnitude and direction across islands, and thus our primary test for adaptation -- parallel responses across islands -- was not supported.
 We suggest, however, that selection could be involved in several ways, including sexual selection.

### 2020-10-28 ms/sections/acknowledgments.tex
backup

@@
-Collection permission was granted by the Bahamas Environment, Science and Technology Commission, the Bahamas National Trust, and the Cayman Islands Department of the Environment.
+Collection permission was granted by the Bahamas Environment, Science and Technology Commission, the Bahamas National Trust, the Bahamas Ministry of Agriculture, and the Cayman Islands Department of the Environment.
 The authors thank Sofia Prado-Irwin, Pavitra Muralidhar, Nicholas Herrmann, Richard E.

### 2020-10-28 ms/sections/discussion.tex
backup

@@
-% There are inconsistent differences that are not explained by distance
+\paragraph{Dewlap coloration differs between habitats, but the differentiation is inconsistent with adaptive convergence} Dewlaps are highly conspicuous in anoles and used as communication signals in courtship \citep{Sigmund1983, Driessens2014, Driessens2015} and territorial displays \citep{Losos1985, Macedonia1994, Macedonia2013}.
+On five of the nine islands we sampled in the West Indies, we found that \textit{A.
+sagrei} male dewlap coloration significantly varied between individuals living in three different habitat types  (beach scrub bush, primary coppice forest and mangrove forest), even though those habitats were in close proximity to each other and the \textit{sagrei} populations are continuously distributed between the habitats.\\
 
-%firef\textbf{Jonathan's take: too much push in favor of selection, without really having much evidence.
-Simple models of genetic drift are ruled out and we don’t think it is plasticity.
-However, our primary test for adaptation—looking for parallel responses across islands—also was falsified.
-We suggest, however, that selection could be involved, and suggest several ways this might have happened.}
+One compelling indicator of adaptation by natural selection is the repetition of patterns of differentiation across localities \citep{Losos2009, Losos2011}.
+In the context of this study, had we found consistent patterns of dewlap differentiation between habitats, we would have concluded that there was likely an adaptive basis for this pattern.
+However, the inconsistent and idiosyncratic patterns of dewlap variation we observed contradict this prediction.
+What then explains the observed differences among the populations?\\
 
-\paragraph{Dewlap coloration differs between habitat-types} We found that male dewlap coloration in \textit{A.
-sagrei} significantly varied between fine-scale habitat-types (beach scrub bush, primary coppice forest and mangrove forest) on five islands of the West Indies: Abaco, Bimini, Cayman Brac, Little Cayman and Long Island.
-However, the habitat-specific variation in dewlaps was not consistent between these islands.
-Although those results are consistent with selection acting at a very local scale, other evolutionary drivers could be at work, such as phenotypic plasticity, random drift, or multiple colonization events.
-We reject this last explanation because all of the island populations in this study are strictly monophyletic, implying a single colonization event per island (van de Schoot, unpublished thesis; \citealt{Driessens2017, Reynolds2020}).
+One explanation is that this pattern is the result of multiple island colonizations of phenotypically different individuals to these habitats.
+However, we reject this possibility because all of the island populations in this study are strictly monophyletic, reflecting a single colonization event per island (van de Schoot, unpublished thesis; \citealt{Driessens2017, Reynolds2020}).
+Other drivers could be at work, including genetic drift, phenotypic plasticity, and divergent natural selection.
+Of these, we find divergent natural selection the most likely explanation, suggesting that future research with additional granularity on the drivers of dewlap coloration is needed.
 
-\paragraph{A role of neutral drift is unlikely} Differences in organismal traits between environments are not necessarily proof of adaptation or selection, and genetic drift may result in patterns similar to local adaptation \citep{Miles2019}.
-Nevertheless, two lines of evidence from our data suggest that this scenario may be implausible.
-First, we found little evidence for a role of phenotypic isolation-by-distance (spatial autocorrelation) in explaining the differences we report.
-We did detect a significant signal of isolation-by-distance on Eleuthera, but there were no differences in dewlap coloration between habitats on this island.
-Second, we detected differences between habitats at relatively small spatial scales, most of the time between sites 5-10km apart, sometimes a few hundred meters away, on Bimini for example.
-Such small-scale differences would be unlikely under strong gene flow \citep{Richardson2014}.
-Our study islands lack geographic barriers to the movement of \textit{A.
-sagrei}, which have been shown to be highly mobile \citep{Kamath2018}, implying widespread gene flow across sites and habitats.
-Moreover, habitat-populations within each island were found to be non-monophyletic and often share identical haplotypes, based on phylogenetic analysis of mitochondrial DNA sequences (van de Schoot et al.
-unpublished thesis), suggesting gene flow between habitats may be widespread.
-Our results align with previous documented cases of persistent dewlap color divergence despite gene flow in multiple species of anoles, sometimes in relation to environmental conditions.
-\citet{Ng2012} and \citet{Ng2016} found divergent dewlap coloration in the face of gene flow between subspecies of \textit{A.
-distichus} across Hispaniola, and proposed this as a mechanism of reproductive isolation in the early stages of speciation \citep{Ng2011, Lambert2013, Ng2017}.
+\paragraph{A role of neutral drift is unlikely} Genetic drift may lead to patterns of trait divergence similar to local adaptation \citep{Miles2019}.
+We think this scenario is implausible, however, as we found significant differences in dewlap color over distances too small for genetic drift to have realistically counteracted the effect of homogenizing gene flow \citep{Richardson2014}.
+On Bimini, for example, we detected differences between sites only a few hundred meters away and \textit{A.
+sagrei}, have previously been shown to be highly mobile \citep{Kamath2018}.
+Moreover, habitat-populations within each island were found to be non-monophyletic and often share identical haplotypes, based on phylogenetic analysis of mitochondrial DNA (van de Schoot et al.
+unpublished thesis).
+This is further evidence of gene flow between habitats.\\
+
+It is possible that complex drift scenarios other than simple distance-mediated, diffusion-like processes may be at play \citep{Miles2019}.
+For example,  we did detect a significant signal of isolation-by-distance on Eleuthera, without finding differences in dewlap coloration between habitats, suggesting a possible role of drift on this island.
+Nonetheless, by and large our results align with previously documented cases of highly localized  dewlap color divergence despite gene flow in multiple species of anoles.
+\citet{Ng2012} and \citet{Ng2016} found divergent dewlap coloration despite gene flow between subspecies of \textit{A.
+distichus} on Hispaniola, and proposed this as a mechanism of reproductive isolation in the early stages of speciation \citep{Ng2011, Lambert2013, Ng2017}.
 \citet{Stapley2011} found that dewlap color polymorphism was maintained in the absence of genetic structure between populations of \textit{A.
@@
 
-% \paragraph{No conclusive evidence for local adaptation of dewlap coloration} Although phenotypic divergence at small spatial scales may be an indicator of local adaptation, the inconsistency of the between-habitat divergence patterns we observed across islands implies that neutral processes cannot be definitely ruled out \citep{Losos2011}.
-Conversely, the absence of parallel divergence across islands does not rule out local adaptation either \citep{Losos2011}.
-If selection has an effect, it may be dependent on aspects of the environment that are not encompassed by our broad habitat-type classification into coastal dry scrub, primary coppice forest and mangrove forest, and dewlap coloration may be influenced by a mosaic of interacting local environmental factors, which need not be the same across islands.\\
-
-\paragraph{No conclusive evidence for adaptation of dewlap coloration to habitat type} One of the most informative tests for adaptation is the convergence of differentiation patterns across replicate islands or localities \citep{Losos2009, Losos2011}.
-Previous studies have described convergent patterns of dewlap color evolution in similar environments across islands and species \citep{Thorpe2002a, Thorpe2002b}.
-However, the inconsistent and idiosyncratic patterns we observed reject the test of convergence across islands, suggesting that dewlap color variation between habitats cannot be predicted by habitat identity alone in \textit{A.
-sagrei}.
-Other studies have proposed that dewlap coloration may have evolved to be maximally detectable under local light conditions imposed by the environment, primarily through UV contrast (i.e.
-UV-brighter dewlaps in UV-dark, mesic habitats and UV-darker dewlaps in UV-bright, xeric habitats), in \textit{A.
-cristatellus} and \textit{A.
-cooki} from Puerto Rico \citep{Leal2002, Leal2004}.
-Although UV reflectance was commonly involved in between-habitat divergence in \textit{A.
-cooki} and \textit{A.
-cristatellus}, we found no such patterns in \textit{A.
-sagrei}, where instead, we found the darkest dewlaps in the dark, mesic habitat -- primary coppice forest -- on three islands, and dewlaps often differed the most between beach scrub and mangrove forest, two xeric habitats with similar, high irradiance levels \citep{Howard1950, Schoener1968}.
-Studies of Jamaican and Hispaniolan anoles similarly found between-habitat differences in dewlap coloration but no evidence for higher dewlap detectability in different habitats \citep{Fleishman2009, Ng2012}.
-Our data are consistent with those previous results in suggesting that adaptation to local light conditions, or at least broad habitat types, is not a major driver of the within-island variation in dewlap coloration in \textit{A.
-sagrei}.
-Habitats on different islands may also differ in aspects other than light conditions, such as densities of predators or congeners, which have been shown to affect among-island dewlap diversity \citep{Vanhooydonck2009, Baeckens2018}.
-In particular, \citet{Baeckens2018} recently showed that dewlaps with spotted patterns occurred more often in \textit{A.
-sagrei} on islands with more coexisting species of anoles.
-Therefore, if local adaptation occurs, it is more likely to involve components of the environment that do not encompass our broad habitat categories.
-
-\paragraph{Sexual selection could be at play} Selection, however, needs not necessarily be linked to habitat type, and may take the form of arbitrary, "Fisherian" sexual selection, where female preferences differ between localities for reasons unrelated to the environment, driving the divergent evolution of male ornaments \citep{Andersson1994}.
-This process could account for the idiosyncratic patterns of within-island divergence we report, where initial differences in female preferences could have arisen for nonselective reasons (e.g.
-plasticity or random drift).
-Substantial levels of promiscuity in \textit{A.
-sagrei} suggest ample opportunity for female mate choice \citep{Kamath2018}, and are in line with this scenario.
-However, \citet{Baeckens2018} found no link between \textit{A.
-sagrei} dewlap coloration and size dimorphism (a proxy for sexual selection) in an among-island study of the same archipelagos.
-Another form of sexual selection is the "good genes" model, where the cue under sexual selection is an indicator of individual quality (e.g.
-better immune response to disease) and of indirect benefit to the offspring \citep{Andersson1994}.
-Several studies suggest that this possibility could be the case for anoles.
-For example, \citet{Cook2013} found lower orange reflectance in dewlaps with heavily parasitized \textit{A.
-brevirostris}, suggesting a trade-off in carotenoid use between the immune response and pigment deposition.
-\citet{Steffen2014} found that lower UV and orange-red reflectance predict contest-winning success between \textit{A.
-sagrei} males, while \citet{Driessens2015} further found that more yellow and red dewlaps (relative to UV) predict better body condition, and that higher yellow and UV reflectance at the margin of the dewlap predict higher hematocrit (the concentration of red blood cells), indicating a better health.
-Other aspects of the dewlap than color have also been found to be indicators of individual quality, such as dewlap size \citep{Vanhooydonck2005, Vanhooydonck2009}, but not dewlap display frequency \citep{Tokarz2002, Tokarz2005, Driessens2015}.
-However, one would expect good genes sexual selection to favor the same coloration across the archipelago, and therefore this mechanism is unlikely to have produced idiosyncratic patterns of within-island divergence.
-
-% An alternative yet non-mutually exclusive explanation for the observed idiosyncratic patterns of within-island divergence is that of "Fisherian" sexual selection \citep{Andersson1994}.
-
-% Under the Fisherian model of sexual selection, arbitrary female preferences (i.e.
-independent of the environment) for certain male ornaments may drive divergent evolution of male signals, such as dewlap coloration, if female preferences differ between localities for other reasons than environment-dependent perception abilities (a situation that could have arisen e.g.
-in an initial phase of genetic drift).
-Substantial levels of promiscuity in \textit{A.
-sagrei} suggest ample opportunity for female mate choice \citep{Kamath2018}, and are in line with this explanation.
-Therefore, arbitrary, habitat-independent female preferences could further explain the inconsistent patterns of divergence across islands.
-In contrast with this, however, \citet{Baeckens2018} found no link between \textit{A.
-sagrei} dewlap coloration and size dimorphism (a proxy for sexual selection) in an among-island study of the same archipelagos.\\
-
-% Alternatively, dewlap coloration in \textit{Anolis} may act as a mating cue according to a "good genes" model of sexual selection, rather than a Fisherian process.
-In the good genes model, the cue under sexual selection is an indicator of individual quality (e.g.
-better immune response to disease) and of indirect benefit to the offspring \citep{Andersson1994}.
-Several studies suggest that this possibility could be the case for anoles.
-For example, \citet{Cook2013} found lower orange reflectance in dewlaps with heavily parasitized \textit{A.
-brevirostris}, suggesting a trade-off in carotenoid use between the immune response and pigment deposition.
-\citet{Steffen2014} found that lower UV and orange-red reflectance predict contest-winning success between \textit{A.
-sagrei} males, while \citet{Driessens2015} further found that more yellow and red dewlaps (relative to UV) predict better body condition, and that higher yellow and UV reflectance at the margin of the dewlap predict higher hematocrit (the concentration of red blood cells), indicating a better health.
-Other aspects of the dewlap than color have also been found to be indicators of individual quality, such as dewlap size \citep{Vanhooydonck2005, Vanhooydonck2009}, but not dewlap display frequency \citep{Tokarz2002, Tokarz2005, Driessens2015}.\\
-
-\paragraph{A role of phenotypic plasticity is unlikely} Differences in coloration between habitat populations may not be genetically determined and instead may be influenced by environmental factors such as parasite load, as mentioned above \citep{Cook2013}.
+\paragraph{A role of phenotypic plasticity is unlikely} Differences in coloration between habitat populations may not be genetically determined and instead may be influenced by environmental factors.
 The yellow, orange and red coloration in anoline dewlaps are produced by pterins and carotenoids \citep{Ortiz1962, Ortiz1962a, Ortiz1963, Ortiz1966, Macedonia2000, Steffen2007, Steffen2009}.
@@
 distichus} \citep{Ng2013}, which also has an orange-based dewlap.
-This makes a plastic response to differences in diet across habitats unlikely.
-Furthermore, developmental plasticity during the ontogeny is also unlikely because dewlap coloration develops at sexual maturity in anoles \citep{Ng2013}.
-The differences we observed could therefore be genetically based.
+(Dewlap color has also been shown to change in relation to parasite load \citep{Cook2013}, though parasites were not investigated in this study.
+-- to remove) These laboratory results  suggest a plastic response in dewlap color to differences in diet across habitats unlikely, suggesting the differences we observed could therefore be genetically based.
 This hypothesis is further supported by \citet{Cox2017}, who found a high degree of heritability of dewlap coloration in \textit{A.
 sagrei}.
-% playing it safer here by not saying it applies to all anoles, but also not mentioning the transgenerational effects anymore
 
-% For these reasons, we consider implausible that dewlap coloration is a plastic trait in our study species.\\
+\paragraph{Sensory Drive - maybe drop?} One hypothesis that a relationship would exist between dewlap color and habitat stems from the idea that the communication signals evolve adaptively in response to light environment (the sensory drive hypothesis, \citealt{Endler1988, Endler1992, Endler1998}).
+However, we find this explanation unlikely because the differences we observed were both inconsistent among islands and inconsistent with predictions of the sensory drive hypothesis.
+Previous studies have proposed that dewlap coloration may have evolved to be maximally detectable under local light conditions, primarily through UV contrast (i.e.
+UV-brighter dewlaps in UV-dark, mesic habitats and UV-darker dewlaps in UV-bright, xeric habitats), in \textit{ A.
+cristatellus} and \textit{A.
+cooki} from Puerto Rico \citep{Leal2002, Leal2004}.
+On the contrary, we found no apparent habitat-dependent maximization of UV-contrast, or just any contrast in \textit{ A.
+sagrei}.
+Instead, we found for example the darkest dewlaps in the dark, mesic habitat -- primary coppice forest -- on three islands, and dewlaps often differed the most between beach scrub and mangrove forest, two xeric habitats with similar, high irradiance levels \citep{Howard1950, Schoener1968}.
+Studies of Jamaican and Hispaniolan anoles similarly found between-habitat differences in dewlap coloration but no evidence for higher dewlap detectability in different habitats \citep{Fleishman2009, Ng2012}.
+Our data are consistent with those previous results in suggesting that adaptation to local light conditions, or at least broad habitat types, is not a major driver of the within-island variation in dewlap coloration in \textit{A.
+sagrei}.
 
-% Too broad a statemtent to make
-% Although most studies used one or two-generation common garden experiments and thus could not rule out transgenerational plastic effects \citep{Tariel2020}, dewlap coloration generally seems to not be a plastic trait.
+\paragraph{Fisherian sexual selection - maybe drop?} Selection, however, needs not necessarily be linked to habitat type, and may take the form of arbitrary, "Fisherian" sexual selection, where female preferences differ between localities for reasons unrelated to the environment, driving the divergent evolution of male ornaments \citep{Lande1981, Andersson1994, Higashi1999}.
+This process could account for the idiosyncratic patterns of within-island divergence we report, where initial differences in female preferences could have arisen for nonselective reasons (e.g.
+plasticity or random drift).
+Substantial levels of promiscuity in \textit{A.
+sagrei} suggest ample opportunity for female mate choice \citep{Kamath2018}, and are in line with this scenario.
+However, \citep{Baeckens2018} found no link between \textit{A.
+sagrei} dewlap coloration and size dimorphism (a proxy for sexual selection) in an among-island study of the same archipelagos.
 
-% This further reinforces an adaptive explanation, where dewlap color could be under differential natural and/or sexual selection in these different habitats.
+\paragraph{Natural selection may proceed differently across islands and habitats} A further potential explanation for the inconsistencies in dewlap coloration among habitats between islands is that the observed differences in dewlap color are the result of divergent natural selection.
+If that is the case, however, either the habitats we have identified are not as similar as we thought, such that the same habitat type on different islands favors different colors, or, alternatively, the selective pressures causing dewlap divergence are related to some environmental factor other than habitat type that we did not measure.
+The habitats on different islands may, for example, differ in densities of predators or congeners, which have been shown to affect among-island dewlap diversity \cite{Vanhooydonck2009, Baeckens2018}.
+In particular, \citet{Baeckens2018} recently showed that dewlaps with spotted patterns occurred more often in \textit{A.
+sagrei} on islands with more coexisting species of anoles.
+Therefore, if local adaptation occurs, it may be more likely to involve components of the environment that do not encompass our broad habitat categories leading to island-specific differences between habitats rather than parallel differences expected if habitat identity was the primary selective pressure driving divergence.\\
 
-% Conclusion, and a word about speciation
+One last possibility to explain idiosyncratic patterns of between-habitat differences we observe is that, if island populations are different in color on average, they may be experiencing different portions of a within-island fitness landscape whose shape changes depending on the starting phenotype.
+In this study, we indeed found that dewlap coloration differed much more between than within islands.
+So, a population may not respond to selection in the same manner depending on its starting conditions.
+This is true not only in terms of phenotype (average dewlap color differs among islands) but also genetic composition (e.g.
+independent colonizations may have important genetic founder effects, \citealt{Reynolds2020}), and so the selective pressures applying to populations with different starting phenotypes may drive these divergent outcomes.
+For example, \citet{Fleishman2020} found that yellow stimuli are less detectable than red stimuli in high-light environments while both colors are readily detectable in low-light environments.
+Consequently, the environmental pressures applying to yellower dewlaps may differ than those for redder dewlaps, as red dewlaps should be always equally or more detectable than yellow dewlaps across light environments, whereas yellow dewlaps could experience more directional selection towards more red when in a high-light environments.
 
-\paragraph{Implications in the context of speciation} Local adaptation can be a precursor to ecological speciation, a process that may have given rise to the adaptive radiation of \textit{Anolis} lizards \citep{Harmon2003, Gavrilets2009}.
-Ecologically-mediated divergence of a sexual signal may be a potent path to the evolution of reproductive isolation through divergent sexual selection \citep{Reynolds2007, Servedio2011}.
-Evidence suggests that dewlap coloration could take this role in anoles \citep{Ng2011, Lambert2013, Geneva2015, Ng2017}, or at least that it is frequently involved in species recognition \citep{Williams1969, Williams1977, Losos1985, Macedonia1994, Fleishman2000, Macedonia2013, Ingram2016, Baeckens2018}.
-Although a correlation between dewlap coloration and reproductive isolation is not detected at the phylogenetic scale of the whole genus \citep{Nicholson2007, Harrison2012, Ingram2016}, sexual signals are often evolutionarily very labile \citep{Kraaijeveld2011}, and the anole dewlap in particular has been capable of rapid macroevolution throughout its phylogenetic history \citep{Nicholson2007}.
-For example, \textit{A.
-conspersus} on Grand Cayman evolved a UV-blue dewlap from an ancestral orange dewlap in 2 to 3 million years \citep{Macedonia2001}.
-We present evidence of multiple cases of within-island dewlap color divergence over small geographical scales in \textit{A.
-sagrei} across the West Indies.
-While these intra-island populations do not appear to be in the process of speciation, our results suggest that the anoline dewlap could have enough micro-scale, local adaptive potential to contribute to reproductive isolation, should it be recruited for assortative mating.
+\paragraph{Conclusion} We identified patterns of dewlap color differentiation in populations of \textit{A.
+sagrei} from different habitat-types on multiple small islands of the West Indies.
+However, our main hypothesis to explain our findings -- local adaptation and sensory drive -- received little support from our multiple-island dataset.
+We also found other mechanisms such as drift or plasticity unlikely candidates to explain the observed patterns.
+This suggests that combinations of multiple factors interacting in subtle ways, and not necessarily in parallel across islands, may have contributed to present-day dewlap color diversity.

### 2020-10-28 ms/sections/introduction.tex
backup

@@
 The staggering diversity of animal communication signals has long been of interest to evolutionary biologists.
-Animals use chemical, mechanical, electromagnetic, and visual signals to communicate in a wide variety of contexts, including, for example, competition for mates, species recognition, aposematism, and cooperation \citep{Bradbury2011}.
-A primary evolutionary factor shaping communication signals is the sensory system and behavior of their recipient(s) (the sensory drive hypothesis; \citealt{Endler1988,Endler1992,Endler1998}).
+Animals use chemical, mechanical, electromagnetic, and visual signals to communicate in a wide variety of contexts, including, competition for mates, species recognition, aposematism, and cooperation \citep{Bradbury2011}.
+A primary evolutionary factor shaping communication signals is the sensory system and behavior of their recipients (the sensory drive hypothesis; \citealt{Endler1988,Endler1992,Endler1998}).
 Over the past decades, scientists have established that signals evolve in an ecological context and are dependent on environmental conditions \citep{Endler1992,Endler1993,Endler1993a}.
@@
 Population genetics theory suggests that gene flow may counteract local adaptation between localities and prevent divergence altogether, especially at small spatial scales, because of the inflow of maladapted alleles or because of the breaking of linkage between coevolving loci \citep{Felsenstein1976, Garcia-Ramos1997, Dieckmann1999, Lenormand2002, Hendry2007}.
-This has been confirmed empirically in systems such as stick-insects \citep{Nosil2004} and sticklebacks \citep{Hendry2007a}.
+This genetic homogenization has been confirmed empirically in systems such as stick-insects \citep{Nosil2004} and sticklebacks \citep{Hendry2007a}.
 Yet, examples of microgeographic adaptation, i.e.
-adaptation at smaller scales than the range of dispersal, exist, highlighting a high potential of some organisms to respond to selection in the face of gene flow (see \citealt{Richardson2014} and references therein).
+adaptation at smaller scales than the range of dispersal, exist, highlighting the potential of some organisms to respond to selection in the face of gene flow (see \citealt{Richardson2014} and references therein).
 Examples include small scale adaptation in fragmented areas in Australian fruit flies \citep{Willi2012}, or local adaptation to predation pressure in North American salamanders \citep{Richardson2013}.
@@
 Lizards of the neotropical genus \textit{Anolis} are an excellent group for studying the eco-evolutionary dynamics of local adaptation and natural selection \citep{Losos2009}.
-A particularly conspicuous trait of anoles is their dewlap; an extensible flap of skin that is typically sexually dimorphic and used as a communication signal in courtship \citep{Sigmund1983, Driessens2014, Driessens2015}, territorial displays \citep{Losos1985, Macedonia1994, Macedonia2013} as well as in predator deterrence \citep{Leal1995, Leal1997, Leal1997b}.
+A particularly conspicuous trait of anoles is their dewlap; an extensible flap of skin that is typically sexually dimorphic and used as a communication signal in courtship \citep{Sigmund1983, Driessens2014, Driessens2015} and territorial displays \citep{Losos1985, Macedonia1994, Macedonia2013} as well as in predator deterrence \citep{Leal1995, Leal1997, Leal1997b}.
 Dewlap characteristics vary widely among the approximately $400$ species of the genus \citep{Nicholson2007}.
@@
 Some studies suggest that those differences may be adaptive, and that dewlaps may have evolved to maximize detectability given local light conditions \citep{Fleishman2001, Leal2002, Leal2004}.
-Other studies testing this hypothesis, however, found no pattern \citep{Fleishman2009, Ng2012, Macedonia2014a}.\\
-
-% So what? Introducing Anolis sagrei as a model species
-
-% I am re-citing the same papers here, is that okay?
-% This is the room for sagrei in particular
+Although this claim is further supported by recent findings that dewlap colors are perceived differently under different levels of shading \citep{Fleishman2020}, other studies found conflicting patterns of between-habitat variation that did not support the sensory drive hypothesis \citep{Fleishman2009, Ng2012, Macedonia2014a}.\\
 
@@
 
-Our sampling design included sites in close proximity (the median distance between two sites within an island was $11.2$km).
+Our sampling design included sites in close proximity; the median distance between two sites within an island was $11.2$km.
 Combining reflectance spectrometry and supervised machine learning, we tested for divergence in dewlap phenotype between habitats within islands and between islands across the range of \textit{A.

### 2020-10-28 ms/sections/methods.tex
backup

@@
 We sampled 466 male \textit{Anolis sagrei} from seven islands in the Bahamas Archipelago -- Abaco, North Andros, South Andros, South Bimini, Eleuthera, Long Island, Ragged Island -- and two in the Cayman Islands -- Cayman Brac and Little Cayman (Figure \ref{fig:map}).
-These islands were chosen to span the breadth of the West Indian range of \textit{Anolis sagrei}, because they have highly similar habitat types,  and because the \textit{A.
+These islands were chosen to span the breadth of the West Indian range of \textit{A.
+sagrei}, because they have highly similar habitat types,  and because the \textit{A.
 sagrei} on each island group are derived from ancient and distinct colonization events from Cuba (i.e.
@@
 Three habitats were sampled on each island based on characterizations by \citet{Howard1950} and \citet{Schoener1968}.
-Each habitat is clearly distinguishable by their dominant vegetation type --- xeric coastal scrub (open, relatively dry habitat consisting of low vegetation or isolated trees), primary coppice forest (closed-canopy forest) and mangrove forest (wet coastal habitat with trees growing in brackish water and high light penetration).
+Each habitat is clearly distinguishable by its dominant vegetation type --- xeric coastal scrub (open, relatively dry habitat consisting of low vegetation or isolated trees), primary coppice forest (closed-canopy forest) and mangrove forest (wet coastal habitat with trees growing in brackish water and high light penetration).
 Sample sizes are given in Table \ref{suptab:counts}.
 Our sampling design enabled us to test for differences between habitats at a coarse and fine geographical scale.
-The median distance between two localities within an island was $\sim 11$km, with some islands being sampled at smaller or larger scales (Figure \ref{supfig:map}, Table \ref{suptab:sites}).
-$80.3$\% of all pairwise distances within islands were less than $50$km.
+The median distance between two localities within an island was $\sim 11$km, with some islands being sampled at smaller or larger scales (Figure \ref{supfig:map}, Table \ref{suptab:sites}), and $80.3$\% of all pairwise distances within islands were less than $50$km.
 Additionally, there are no major barriers to dispersal (such as mountains or grassland) on any of the islands that we sampled.
@@
 
-All analyses in this study were performed in R 3.6.1 \citep{RCoreTeam2019}.
+All analyses in this study were performed in R 3.6.1 \citep{RCoreTeam2019} and the source code can be found at https://github.com/rscherrer/dewlap, presently private.
+We used our newly developed R package nmgc (Nested Multivariate Group Comparisons; https://github.com/rscherrer/nmgc), to perform all analyses.
+This package provides a streamlined and powerful interface to perform group comparison analyses in hierarchical and/or nested by combining both classical statistical tools (e.g.
+MANOVA, ANOVAs, Kruskal-Wallis) and machine learning tools (e.g.
+LDA, SVMs) into complete analysis pipelines.
+The package also facilitates PCA-transformations, assumption testing and automatic choice of suitable posthoc tests, along with full cross-validation, binomial testing of accuracy scores and independent replicate trials for machine learning procedures.
 
@@
 
-In addition to within-island PCA, we performed a PCA on pooled data from the whole archipelago.
+In addition to within-island PCA, we performed a PCA on pooled data from all islands.
 The first four principal components explained 91.3\% of the variance (Table \ref{suptab:pcavariances}).
@@
 
-Because of the aforementioned violations of the MANOVA assumptions, and to reduce the chances of false discovery, we conducted multivariate group comparisons using support vector machines (SVMs), a model-free, powerful nonparametric supervised machine learning technique.\\
+Because of the aforementioned violations of MANOVA assumptions, and to reduce the chances of false discovery, we conducted multivariate group comparisons using support vector machines (SVMs), a model-free, powerful nonparametric supervised machine learning technique.\\
 

### 2020-10-28 analyses/04-machine-learning/MachineLearning.R

@@ -42,9 +42,9 @@ blank <- ggplot() + theme_void()
 
 p <- plot_classif(res, facets = "island")
 plot_grid(p, plot_grid(blank, confplot, blank, nrow = 3), ncol = 2, rel_widths = c(2, 1))
-ggsave("classif_svm_pca.png", height = 6, width = 9, dpi = 300)
+ggsave("analyses/04-machine-learning/plots/classif_svm_pca.png", height = 6, width = 9, dpi = 300)
 
-fname <- "analyses/04-machine learning/table_classif_%s"
+fname <- "analyses/04-machine-learning/table_classif_%s"
 colnames <- c("Island", "Accuracy", "$N$", "$p_{\\mbox{test}}$", "$n_{\\mbox{test}}$", "$P$", "")
 
 t1 <- res$mean
@@ -61,7 +61,7 @@ res$imp %>%
   facet_wrap(. ~ island) +
   geom_hline(yintercept = 0.25, lty = 2)
 
-ggsave("importance_svm_pca.png", height = 4, width = 5, dpi = 300)
+ggsave("analyses/04-machine-learning/plots/importance_svm_pca.png", height = 4, width = 5, dpi = 300)
 
 # LDA-classification on PCs
 
@@ -73,7 +73,7 @@ res <- classify(
 
 p <- plot_classif(res, facets = "island", fill = "coral")
 plot_grid(p, plot_grid(blank, confplot, blank, nrow = 3), ncol = 2, rel_widths = c(2, 1))
-ggsave("classif_lda_pca.png", height = 6, width = 9, dpi = 300)
+ggsave("analyses/04-machine-learning/plots/classif_lda_pca.png", height = 6, width = 9, dpi = 300)
 
 t2 <- res$mean
 t2_fname <- sprintf(fname, "lda_pca")
@@ -89,7 +89,7 @@ res$imp %>%
   facet_wrap(. ~ island) +
   geom_hline(yintercept = 0.25, lty = 2)
 
-ggsave("importance_lda_pca.png", height = 4, width = 5, dpi = 300)
+ggsave("analyses/04-machine-learning/plots/importance_lda_pca.png", height = 4, width = 5, dpi = 300)
 
 # Reflectance data
 
@@ -104,7 +104,7 @@ res <- classify(
 
 p <- plot_classif(res, facets = "island")
 plot_grid(p, plot_grid(blank, confplot, blank, nrow = 3), ncol = 2, rel_widths = c(2, 1))
[…]

### 2020-10-28 ms/fetch_content.sh

@@ -21,7 +21,6 @@ cp ../analyses/10-distances/figure_distances2.png suppfigures
 cp ../analyses/10-distances/figure_distances.png suppfigures
 cp ../analyses/06-pooled/figure_pooled.png suppfigures
 cp ../analyses/02-reflectance/figure_reflectance.png suppfigures
-cp ../analyses/02-reflectance/figure_reflectance_pooled.png suppfigures
 cp ../analyses/04-machine-learning/plots/importance_lda_pca.png suppfigures
 cp ../analyses/04-machine-learning/plots/importance_lda_pca_pooled.png suppfigures
 cp ../analyses/04-machine-learning/plots/importance_lda_refl.png suppfigures

### 2020-11-01 ms/sections/introduction.tex
reran the analyses with corrected nmgc classify function

@@
 Previous studies investigating variation in anole dewlaps compared populations at relatively large geographical scales, e.g.
-between islands \citep{Vanhooydonck2009, Driessens2017} or within large islands such as Puerto Rico \citep{Leal2002, Leal2004} or Hispaniola \citep{Ng2012, Ng2016}.
+between islands \citep{Vanhooydonck2009, Driessens2017} or within large islands such as Puerto Rico \citep{Leal2004} or Hispaniola \citep{Ng2012, Ng2016}.
 These large scales and marine barriers should reduce gene flow \citep{Ng2011, Lambert2013, Richardson2014, Ng2017}.
@@
 The species \textit{Anolis sagrei} is widespread across islands of the West Indies \citep{Reynolds2020}.
-It has been the subject of study concerning local adaptation \citep{Losos1994, Losos1997a, Losos2001, Kolbe2012}, biological invasion \citep{Kolbe2008}, sexual selection \citep{Tokarz2002, Tokarz2005, Tokarz2006, Driessens2014, Steffen2014, Driessens2015} and many other topics.
+It has been the subject of numerous studies concerning local adaptation \citep{Losos1994, Losos1997a, Losos2001, Kolbe2012}, biological invasion \citep{Kolbe2008}, sexual selection \citep{Tokarz2002, Tokarz2005, Tokarz2006, Driessens2014, Steffen2014, Driessens2015} and many other topics.
 Between-island variation in the mainly orange-red color of its dewlap was shown to be better explained by climatic variables \citep{Driessens2017} than by proxies for biotic factors such as sexual selection or predation pressure \citep{Vanhooydonck2009, Baeckens2018}.

### 2020-11-01 analyses/04-machine-learning/MachineLearning.R

@@ -174,7 +174,7 @@ confplot <- plot_classif(res, type = "confusion") +
   labs(x = "True habitat", y = "Predicted habitat", fill = "Frequency") +
   theme(axis.text.x = element_text(angle = 60, hjust = 1))
 plot_grid(p, plot_grid(blank, confplot, blank, nrow = 3, rel_heights = c(1, 3, 1)), ncol = 2, rel_widths = c(1.1, 1))
-ggsave("classif_svm_pca_pooled.png", height = 3, width = 6, dpi = 300)
+ggsave("analyses/04-machine-learning/plots/classif_svm_pca_pooled.png", height = 3, width = 6, dpi = 300)
 
 t5 <- res$mean
 t5_fname <- sprintf(fname, "svm_pca_pooled")
@@ -205,7 +205,7 @@ confplot <- plot_classif(res, type = "confusion") +
   labs(x = "True habitat", y = "Predicted habitat", fill = "Frequency") +
   theme(axis.text.x = element_text(angle = 60, hjust = 1))
 plot_grid(p, plot_grid(blank, confplot, blank, nrow = 3, rel_heights = c(1, 3, 1)), ncol = 2, rel_widths = c(1.1, 1))
-ggsave("classif_lda_pca_pooled.png", height = 3, width = 6, dpi = 300)
+ggsave("analyses/04-machine-learning/plots/classif_lda_pca_pooled.png", height = 3, width = 6, dpi = 300)
 
 t6 <- res$mean
 t6_fname <- sprintf(fname, "lda_pca_pooled")
@@ -239,7 +239,7 @@ confplot <- plot_classif(res, type = "confusion") +
   labs(x = "True habitat", y = "Predicted habitat", fill = "Frequency") +
   theme(axis.text.x = element_text(angle = 60, hjust = 1))
 plot_grid(p, plot_grid(blank, confplot, blank, nrow = 3, rel_heights = c(1, 3, 1)), ncol = 2, rel_widths = c(1.1, 1))
-ggsave("classif_svm_refl_pooled.png", height = 3, width = 6, dpi = 300)
+ggsave("analyses/04-machine-learning/plots/classif_svm_refl_pooled.png", height = 3, width = 6, dpi = 300)
 
 t7 <- res$mean
 t7_fname <- sprintf(fname, "svm_refl_pooled")

### 2021-04-10 ms/sections/discussion2.tex
many things happened

Two main insights follow from our results. First, excluding North Andros where the follow-up univariate analyses were not significant, we detected significant differences in dewlap coloration between habitats within seven out of the nine islands investigated, suggesting a putatively high potential for local differentiation of dewlap coloration in \textit{Anolis sagrei}. Second, we found differences in coloration along different dimensions of color space, and in different directions, on different islands.\\

Detectable differences in dewlap color between habitat-populations are surprising, as habitats were often in close geographical proximity to each other (as close as a few hundred meters on Bimini and most of the time within ten kilometers). Indeed, given that (1) the populations were continuously distributed between the habitats, (2) different habitat-populations were not monophyletic with respect to mitochondrial haplotypes (van de Schoot, unpublished thesis), and (3) \textit{A. sagrei} have been shown to be a highly mobile species within these islands \citep{Kamath2018}, we would have expected more homogeneous distributions of color phenotypes within islands due to extensive gene flow, with fewer differences between populations, especially those in close proximity.\\

Several scenarios could account for these findings. One explanation is an adaptive one. Indeed, populations living in different habitats could be phenotypically adapted to their local environmental conditions. Given that the brightly colored dewlap of \textit{A. sagrei} is used as a communication signal, its color may be a target for selection if the transmission or perception of the signal differs from one habitat to another, for example because of differences in ambient light, according to the sensory-drive hypothesis \citep{Endler1988, Endler1992, Endler1998}. The sensory-drive hypothesis has been tested multiple times for dewlap coloration in \textit{Anolis} lizards, with mixed results. Some authors found support for it \citep{Leal2002, Leal2004}, while others did find differences in dewlap coloration between habitats, but those were inconsistent with the sensory-drive hypothesis \citep{Fleishman2009, Ng2012}.\\

If our results were an example of sensory drive, we would have expected to see consistent differences between habitat-populations across islands (a pattern that would have been a compelling indicator of adaptation at all, \citealt{Losos2011}). This is because environmental conditions that may be relevant to color signal detectability such as light, temperature, moisture and vegetation, are consistent within the three main and clearly distinct habitat-types found across the sampled islands, i.e. beach scrub, primary coppice and mangroves \citep{Howard1950, Schoener1968}. Moreover, the patterns of divergence expected under a sensory drive scenario should be consistent with increased detectability given the local light conditions, such as the high contrasts with background vegetation found in the UV-range by \citet{Leal2002} and \citet{Leal2004}.\\

Instead, we found differences in the way dewlap color differs between habitats across islands. While short-wavelengths (UV reflectance) were often involved in color differences, they were not involved on all islands where significant differences were detected. On some islands, other or additional variables differed, such as brightness, red reflectance or the reflectance at the ends of the spectrum visible to \textit{Anolis} lizards (UV and red, \citealt{Lazareva2012}) relative to intermediate wavelengths (blue-to-yellow). Similar portions of the spectrum were sometimes involved in opposite directions on different islands, such as on Abaco and Cayman Brac, where mangrove lizards had a higher UV-reflectance than coastal lizards on the former, but a lower UV-reflectance on the latter. Under a sensory-drive scenario, we would have expected the same variables to be consistently divergent between habitats, or at least in a consistent direction.\\

Not only consistent patterns across islands would have been a good clue for a sensory-drive explanation, but in particular consistent differences between habitats that are most different in their local conditions regarding the ecological function of the dewlap, such as ambient light. For example, if ambient light is an important factor shaping dewlap coloration, we would expect mangrove and beach scrub lizards, both inhabiting areas with high light penetration, to harbor more similar dewlaps, and to differ significantly from lizards from the coppice habitat, where irradiance is low. Overall, the observed heterogeneity of divergence patterns across islands provides no support to a sensory-drive explanation.\\

Phenotypic plasticity could be another cause for dewlap color variation between habitats, where different conditions would favor different phenotypes in different habitats, without genetic changes. Indeed, the yellow, orange and red colors in anoline dewlaps are produced by pterins and carotenoids \citep{Ortiz1962, Ortiz1962a, Ortiz1963, Ortiz1966, Macedonia2000, Steffen2007, Steffen2009}. Animals can be synthesize pterins from nucleotides, but lack the ability to synthesize carotenoids \citep{Goodwin1984, Hill2002, Hill2006}. Different food qualities across sites within islands could therefore potentially cause detectable differences in coloration. Alternatively, more subtle effects on dewlap color could arise from developmental plasticity and depend, e.g. on differences in egg-rearing conditions. However, more data are needed to test these hypotheses, and although some work has shown plastic responses of dewlap color in response to parasites in \textit{A. sagrei} \citep{Cook2013}, we find it unlikely to account for the widespread habitat differences we found. Besides, studies testing the effect of carotenoid deprivation \citep{Steffen2010, Ng2013} and heritability \citep{Cox2017} of dewlap coloration in \textit{A. sagrei} and another species with a carotenoid-based dewlap, \textit{A. distichus}, found little support for phenotypic and developmental plasticity in dewlap coloration.\\

Genetic drift is another process that can account for differences in phenotype between localities, especially in small populations. One way this could proceed is through isolation-by-distance, where more distant populations accumulate more differences through time because of the reduced effect of gene flow at larger geographical scales relative to the dispersal range of the species \citep{Rousset2004}. Here, we only found a significant correlation between phenotypic and geographical distances on Eleuthera to support this scenario. On all the other islands, in contrast, populations from closer sites were not phenotypically more similar, which argues against isolation-by-distance. That said, there were often few sampling sites per island in our study, whose locations were not uniformly chosen within the islands, and so the true extent of isolation-by-distance may be difficult to test. Other, less trivial forms of drift may be at play than isolation-by-distance, but nevertheless, we did find significant differences in color phenotype at relatively small spatial scales, sometimes in neighboring habitats, on islands where gene flow is probably highly pervasive, as suggested by high rates of encounter between males and females \citep{Kamath2018}, making the divergence of habitat-populations by drift in relative genetic isolation an unlikely scenario.\\

A number of alternative explanations remain. First, there could sexual selection for different dewlap colors in different locations. Indeed, although the sensory-drive hypothesis may include a sexual selection aspect, e.g. if the optimal male phenotype in a given habitat is a function of female perception, sexual preferences may also be arbitrary and independent of the habitat \citep{Andersson1994}, and so could differ across habitats and islands. However, one previous study has found no link between dewlap coloration and body size dimorphism in \textit{A. sagrei} (a proxy for the intensity of sexual selection) in an among-island comparison \citep{Baeckens2018}, and although within-island data are lacking, this scenario may have little plausibility.\\

Alternatively, selective pressures may be different in similar habitats from one island to another, because of other environmental variables not accounted for by the habitat-type classification we used \citep{Howard1950, Schoener1968}. The islands we sampled indeed exhibit variation in some climatic variables but also in densities of predators and anole congeners, which have all been shown to correlate with variation in \textit{A. sagrei} mean dewlap coloration among islands \citep{Vanhooydonck2009, Baeckens2018}.\\

Finally, different island-populations could also respond differently to similar selective pressures, resulting in various between-habitat divergence patterns across islands. Several factors could account for this. For example, the founder populations of each island, which we know colonized the islands independently (van de Schoot, unpublished thesis; \citealt{Driessens2017, Reynolds2020}), could have exhibited different dewlap colors at the time of colonization, as may be suggested by the larger differences we observed between than within islands. In turn, different initial phenotypes could have led to different ways in which populations would have diverged between habitats. Moreover, the different founding populations may have also consisted in different subsets of the standing genetic variation of their Cuban ancestor due to potential bottlenecks \citep{Reynolds2020}, which may have constrained the way they would later respond to the local selective pressures of their respective islands.\\

Altogether, our results show that dewlap color of \textit{A. sagrei} commonly varies between habitat-types, even in close geographical proximity, within islands of the West Indies, and that coloration differs in different ways from one island to another. We discussed several non-mutually exclusive mechanisms that could explain these observations, but more data are needed to thoroughly test each of these. Nevertheless, heterogeneous patterns of divergence across islands do not support an adaptive sensory-drive scenario, and our we propose that within-island dewlap color variation may be underlain by a more subtle mosaic of factors.

% What is missing?
% It reads quite complete, although I haven't talked about all the examples where dewlaps were found to vary between localities in the literature --- at least that would be a dilute a bit the large number of speculations in the intro.
% Also it would be nice to echo the concept of adaptation at small spatial scales talked about in the intro
% Maybe let it rest and read it all first to see what is missing
% How did closely related island-populations diverge between habitats relative to each other? e.g. Cayman Islands?

### 2021-04-10 ms/sections/figures.tex
tables and figures look good
implemented colin's comments and refactored the discussion to put more emphasis on neutral processes
many things happened

@@
-% Map of the West Indies
 \begin{figure}[H]
 	\centering
-	\includegraphics[width=0.8\textwidth]{../maps/map.pdf}
-	\caption{Map of the West Indies with sampled islands highlighted in black.}
-	\label{fig:map}
+	\includegraphics[width=\textwidth]{figures/Abaco.png}
+	\caption{Comparison of dewlap coloration across habitats on Abaco.
+(A) Confusion matrix showing the proportion of lizards from each (true) habitat reassigned to each (predicted) habitat by the random forests, based on the first four within-island principal components and averaged across replicates.
+Each column sums to one.
+(B) One-dimensional sensitivity analysis showing the relative importance (mean decrease in accuracy) of the various wavelengths in random forest classification of the whole spectrum.
+(C) Reflectance profiles of all the dewlaps on the island.
+(D) Within-island principal component scores across habitats.
+Bars indicate significant contrasts.
+*, $P < 0.05$; **, $P < 0.01$; ***, $P < 0.001$.
+(E) How reflectance profiles map onto the within-island principal components.
+(F) Map of the island with the sampling sites colored by habitat.
+(G) Geographical distance between sites where significant differences were detected in within-island principal component scores (Wilcoxon test, Benjamini-Hochberg correction, $P < 0.05$), including only pairs of sites whose habitats were involved in between-habitat dewlap differences.}
+	\label{fig:Abaco}
 \end{figure}
-
-% Boxplots and analyses of variance
-\begin{sidewaysfigure}
-	\centering
-	\includegraphics[width=18cm]{../analyses/07-ANOVA/figure_anova.png}
-	\caption{Dewlap color variation between habitat-types across the most significant islands.
-(A) Mapping of reflectance at various wavelengths onto the principal components (loadings from the PCA rotation matrix).
-(B) Distribution of PC scores between habitats along the first four PCs on each island where significant between-habitat differences were detected using SVMs.
-P-values are reported for univariate ANOVA (or Kruskal-Wallis tests when applicable, see Methods).
-Post hoc significant differences at a 0.05 error rate are indicated with horizontal bars.
-*, P < 0.05.}
-	\label{fig:anova}
-\end{sidewaysfigure}
-
-% Classification accuracy of SVMs
-\begin{figure}[H]
-    \centering
-	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_svm_pca"}
-	\caption{SVM classification accuracy across islands based on principal component data.
-Histograms show accuracy distributions over 100 replicates for each five cross-validation bins per island.
-The dashed line is the density of a corresponding null binomial distribution, which would be expected under random guessing (testing sets with 20\% of the observations for each island and success probability of $1/3$).
-Inset plots show the corresponding average confusion matrices and represent the proportion of lizards from each habitat (columns) reassigned in each other habitat (rows), with an interpretation guide in the right panel.
-Binomial test P-values indicate deviations of the mean classification accuracy to the null distribution.
-*, P < 0.05.}
-	\label{fig:classif-svm-pca}
-\end{figure}

### 2021-04-10 ms/sections/methods2.tex
many things happened

% Citation for classifiers as a good group comparison tool
% Yang and Yang on LDA being performed on PC

\subsection*{Data collection}

We sampled 466 male \textit{Anolis sagrei} from seven islands in the Bahamas Archipelago -- Abaco, North Andros, South Andros, South Bimini, Eleuthera, Long Island, Ragged Island -- and two in the Cayman Islands -- Cayman Brac and Little Cayman (Figure \ref{fig:maps}A). These islands were chosen to span the breadth of the West Indian range of \textit{A. sagrei}, because they have highly similar habitat types,  and because the \textit{A. sagrei} on each island group are derived from ancient and distinct colonization events from Cuba (i.e. relatively evolutionarily independent, \citealt{Reynolds2020}). Three habitats were sampled on each island based on characterizations by \citet{Howard1950} and \citet{Schoener1968}. Each habitat is clearly distinguishable by its dominant vegetation type --- xeric coastal scrub (open, relatively dry habitat consisting of low vegetation or isolated trees), primary coppice forest (closed-canopy forest) and mangrove forest (wet coastal habitat with trees growing in brackish water and high light penetration). Sample sizes are given in Table \ref{tab:counts}. Our sampling design enabled us to test for differences between habitats at a coarse and fine geographical scale. The median distance between two localities within an island was $\sim 11$km, with some islands being sampled at smaller or larger scales (Figure \ref{fig:maps}B), and $80.3$\% of all pairwise distances within islands were less than $50$km. Additionally, there are no major barriers to dispersal (such as mountains or grassland) on any of the islands that we sampled.

\subsection*{Reflectance measurements}

We measured reflectance between 300 and 700nm wavelength, a range that encompasses the colors visible to most lizards and vertebrates in general \citep{Lazareva2012}. Measurements were taken with an Ocean Optics USB4000 spectrometer, a pulsed Xenon light source (PX-2, Ocean Optics, Largo, FL, USA) and a reflectance probe protected by a black anodized aluminum sheath. Measurements were taken with a 45-degree inclination to prevent specular reflection \citep{Endler1990}. The device was regularly standardized with a Spectralon white standard (Labsphere, North Sutton, NH, USA). Reflectance was measured at the center of the dewlap. Reflectance curves were smoothed using the R package pavo \citep{Maia2013} as well as with custom R functions, down to one reflectance value at each nanometer in wavelength from 300 to 700nm.

\subsection*{Analysis}

We tested for detectable differences in dewlap coloration between populations from different habitats across islands by following an analytic pipeline in several steps. First, we used multivariate analyses of variance to assess the relative contributions of islands, habitats and habitat-by-island interactions on the partitioning of variation in color space. Second, and provided that habitat-by-island interactions were found, we investigated habitat-differences in dewlap color for each island separately using machine learning classification. Third, for each island where multivariate differences were detected using our machine learning pipeline, we used univariate analyses of variance to decompose the signal among the different dimensions of color space. Fourth, for each significant between-habitat variation found in univariate analyses, we used posthoc tests to determine which habitats were involved. Last, to get insights into the spatial scale of phenotypic variation, for each significant contrast between two habitats detected along a given dimension on a given island we performed multiple pairwise Wilcoxon tests to assess differences in dewlap coloration among the sites involved in that significant contrast, and recorded the geographical distance between sites that were found significant. In parallel, we tested a possible effect of isolation-by-distance, an alternative cause of phenotypic divergence between localities based on diffusion approximation and dispersal distance, irrespective of habitat-types. We did so using a permutation test to assess the significance of the correlation between geographical distances and phenotypic distances among sites within each island.\\

All analyses in this study were performed in R 3.6.1 \citep{RCoreTeam2019} and the source code can be found at https://github.com/rscherrer/dewlap, presently private.

\subsubsection*{Dimensionality reduction}

Because neighboring wavelengths are highly collinear and redundant in reflectance, we reduced the dimensionality of the data using principal component analysis (PCA), as per \citet{Cuthill1999} and \citet{Leal2002}. We performed PCA on data from all islands combined, as well as on each island separately and systematically retained the first four principal components (PC), which together always explained more than $88.8\%$ of the variance across islands (Table \ref{tab:pcavariances}). PCs need not represent the same wavelengths across islands because they are fitted on different datasets. Nevertheless, PC1 was highly collinear with brightness for all islands (Figure \ref{fig:brightness}), while the other PCs captured the chromatic variation (i.e. irrespective of brightness) in dewlap color.

\subsubsection*{Among-island variance partitioning}

We performed a two-way nonparametric multivariate analysis of variance (PERMANOVA, \citealt{Anderson2001}, R package vegan, \citealt{Oksanen2019}) to identify differences in coloration between islands, habitats and habitats within islands, using principal components fitted on data from all islands together. We used a nonparametric test because although no multivariate outliers were detected based on the Mahalanobis distance, the assumption of multivariate normality was violated in several habitats on several islands (Henze-Zirkler's test, \citealt{Henze1990}, R package MVN, \citealt{Korkmaz2014}, $P < 0.05$, Table \ref{tab:multinorm}).

\subsubsection*{Within-island machine learning}

We performed a machine learning classification analysis on the first four principal components within each island separately, using random forests \citep{Breiman2001}. Random forests are a versatile, intuitive and powerful algorithm commonly used in machine learning, using decision trees to predict the labels of particular observations, based on their multivariate coordinates, i.e. variables, passed through a series of successive decision nodes, each examining a given variable of any given observation \citep{James2013}. The prediction for each observation is an aggregate over a large number of decision trees, each being trained on a subset of observations sampled with replacement from the dataset, and each allowed to examine only a subset of the variables. This allows the random forest to overcome the individual errors of all trees in the predictions it makes.\\

To detect differences in dewlap coloration between habitats, we measured the success of random forests in reassigning individual lizards to their correct habitat of origin, based solely on their principal component scores. In machine learning, this so-called cross-validation procedure is typically done in two steps \citep{James2013}. First, a random forest is trained in recognizing features of dewlap coloration most associated with the different habitats, by being presented multiple observations, making predictions about them and updating its own decision rules based on whether the prediction deviates from the truth. Then, once trained, the patterns that the random forest has learned to recognize are tested by presenting new, previously unseen observations to the random forest, and measuring the proportion of correct predictions. This proportion, or success score, can then be statistically assessed against random guessing using a binomial test.\\

The cross-validation procedure requires that the data be split into a training set and a testing set. To remove any bias due to the set that is being sampled for training, it is common practice to use k-fold cross-validation \citep{James2013}, where the data is split into $k$ random bins and $k$ independent machines are trained, each taking one of the bins as a testing set and the rest for training, and where classification success is measured by summing all correct classifications from the $k$ machines.\\

Here, we used a k-fold cross-validation procedure with $k = 5$, where each training set consisted of 80\% of the data and the machine was tested on the remaining 20\%. Each training set was conditioned on containing at least 5 lizards from each of the three habitats. We also down-sampled the training set to the sample size of the least represented habitat, to ensure that the different habitats were equally represented. To further remove any bias due to the specific random split into the different bins, we replicated each k-fold cross-validation five times. We then averaged the five resulting confusion matrices across replicates, where each confusion matrix shows the number of lizards from each habitat reassigned into each habitat. For each island, we then used the average proportion of correctly reassigned lizards (i.e. the proportion of observations on the diagonal of the average confusion matrix) as an estimate of classification success. This score was tested against random guessing by comparing it to a binomial distribution with number of trials being the number of lizards on that island and success probability $1/3$, representing the rate of successful classification by chance when three habitats are involved.\\

We used the machine learning fitting functions in the R package rminer \citep{Cortez2020}, which call random forest routines from the randomForest package (\citealt{Liaw2002}, implementation from the original random forest algorithm, \citealp{Breiman2001}). For each random forest, we optimized the number of trees in the forest and the number of variables examined by each tree using the grid hyperparameter search procedure implemented in rminer, to choose between two numbers of trees (500 or 1,000) and four numbers of principal components examined per tree (1 to 4), using rminer's ordered holdout validation method with $2/3$ of the data used for training.\\

We validated the results of our analysis by using two other widely used machine learning classification methods: linear discriminant analysis and support vector machines \citep{Cristianini2000, James2013}, both accessible in rminer \citep{Cortez2020}.\\

To know which wavelengths were most used to assign data points to each habitat, we trained another set of random forests, this time directly on reflectance data (taken every 5nm from 300 to 700nm) instead of principal components. We recorded the relative importance of each wavelength for each habitat, as measured by the mean decrease in accuracy during wavelength permutation, implemented in the randomForest package \citep{Liaw2002}.

\subsubsection*{Univariate analyses}

For each island where significant differences in dewlap coloration were detected between habitats, we used multiple univariate analyses of variance (ANOVA) to identify possible principal components underlying the observed differences. We constructed our ANOVA models in two steps, as per \citet{Zuur2009}. In a first step, we accounted for heterogeneity of variances across groups by systematically comparing the goodness-of-fit of an ANOVA model estimated with ordinary least squares (OLS) with that of a model estimated with generalized least squares (GLS), which allowed one estimate of residual variance per habitat (using the R package nlme, \citealt{Pinheiro2000, Pinheiro2020}). Both models were fitted with restricted maximum likelihood (REML). Goodness-of-fit was estimated using Akaike's Information Criterion corrected for small sample sizes (AICc, R package MuMIn, \citealt{Barton2019}), and the estimation method yielding the lowest AICc was retained. In a second step, we re-fitted the retained model with maximum likelihood (ML) to test for the effect of habitat-type using likelihood ratio tests (LRT) between a model including a habitat-term and a null model lacking the habitat-term.\\

We evaluated the normality of the standardized residuals (residuals divided by their standard error, which can differ among habitats in a GLS model) of each fitted ANOVA model using Shapiro-Wilk's test, with P-values adjusted for multiple testing using the Benjamini-Hochberg correction \citep{Benjamini1995}. In cases where significant deviations from normality were detected ($P_{adj} < 0.05$, Table \ref{tab:normality}) we performed Kruskal-Wallis's nonparametric test to back up the ANOVA results.\\

To know which habitat-populations were different from which in dewlap coloration, we performed different post-hoc multiple comparison tests (all implemented in the PMCMRplus package, \citealp{Pohlert2020}), depending on which assumptions were met. In cases where normality and homoskedasticity were met (i.e. OLS-ANOVA was the best fit), we used Tukey's honest significant difference test. When normality was met but not homoskedasticity (i.e. GLS-ANOVA was the best fit), we used Dunnett's T3 test. Finally, whenever we used Kruskal-Wallis's test because the ANOVA residuals were not normally distributed, we used Nemenyi's test for post-hoc comparisons.

\subsubsection*{Spatial autocorrelation}

We tested for within-island spatial autocorrelation between the geographical distances among sampling sites and their Euclidean distances in multivariate color space (mean PC1 to PC4 per site, Table \ref{tab:sites}), regardless of habitat-type. Because often only a few sites were sampled per island, we could not get meaningful results from tests that use sites as units of observation, such as Moran's I test \citep{Gittleman1990}. Instead, we designed a permutation test where we randomly reshuffled individual lizards across sites within islands 1,000 times each, and systematically recalculated Pearson's correlation coefficient between geographic distances (computed as geodesic distances in the R package geosphere; \citealt{Hijmans2019}) and phenotypic distances. We used the resulting null distributions of correlation coefficients to assess the significance of the observed spatial autocorrelation for each island.

\subsubsection*{Site differences}

In this study, we were interested in the minimum spatial scale at which significant differences between habitats could be detected within islands. We performed multiple pairwise nonparametric Wilcoxon-Mann-Whitney tests \citep{Hollander2013} to compare dewlap coloration between sites with different habitat-types, for each pair of habitats and each variable where significant differences were detected with our analyses of variances. The P-values were adjusted using a Benjamini-Hochberg correction for multiple testing \citep{Benjamini1995}.

### 2021-04-10 ms/sections/results2.tex
many things happened

% Little but detectable differences on five islands

We tested for variation in \textit{A. sagrei} dewlap coloration between populations living in three characteristic habitat types across nine islands that span the West Indian range of the brown anole (coastal scrub, primary coppice and mangroves). We found that most of the variation in coloration was partitioned between islands (two-way PERMANOVA, approx. $F(\text{df} = 8) = 43.7$, $P = 0.0001$, Fig. \ref{fig:reflectance}, explained variance $R^2 = 40.9$\%). Nonetheless, we did find evidence for differences in dewlap coloration between habitat-types, and those were mostly island-specific (habitat-by-island interaction term, approx. $F(16) = 3.53$, $P = 0.0001$, $R^2 = 6.6$\%), with a small but significant portion of the variation explained by an archipelago-wide habitat effect (approx. $F(2) = 4.7$, $P = 0.0001$, $R^2 = 1.1$\%).\\

We subsequently tested for differences in dewlap coloration between habitat-populations within each island, using within-island principal component scores, that is, computed from the data specific to each island (to maximize the variation captured for each island). Our within-island random forest classification analyses revealed detectable differences in dewlap coloration on seven out of the nine islands in our sample: Abaco, Bimini, Cayman Brac, Eleuthera, Little Cayman, Long Island and South Andros. On these islands, our classifiers could reassign individual dewlaps to their correct habitat more often than expected by chance (Table \ref{tab:randomforests}). We obtained nearly identical results using support vector machines (Table \ref{tab:ksvms}) and linear discriminant analysis (Table \ref{tab:ldas}) for classification. We did not find evidence of spatial autocorrelation in dewlap coloration between the sites within islands, except for Eleuthera (Table \ref{tab:autocorrelation}). We now describe the specific differences detected on each island.\\

On Abaco, we detected the strongest differences between the coastal and mangrove habitats, while dewlaps from the coppice habitat were more difficult to classify (Fig. \ref{fig:Abaco}A). Importance analysis revealed that coastal and mangrove lizards mostly differed in reflectance in the UV-end of the spectrum (below 400nm, Fig. \ref{fig:Abaco}B), where mangrove dewlaps had higher UV reflectance relative to coastal lizards, and coppice lizards had an intermediate UV reflectance between the two other habitats (Fig. \ref{fig:Abaco}C). Consistent with this, our analyses of variance detected significantly lower PC2 scores in mangrove lizards than in the two other habitats, representing a higher UV-reflectance relative to red (Fig. \ref{fig:Abaco}D, E, Table \ref{tab:anova}). Coastal lizards also scored lower on PC3, indicating less curvature of the reflectance profile and relatively higher reflectance at intermediate wavelengths (blue-to-yellow) than at the ends of the range (Fig. \ref{fig:Abaco}D, E). Differences were detected between sites both at large ($\sim$ 100km) and short ($<$ 1km) distances (Fig. \ref{fig:Abaco}F, G).\\

On Bimini, the random forests mostly correctly classified lizards from the coastal habitat (Fig. \ref{fig:Bimini}A), with a relatively flat importance profile suggesting that brightness was used instead of a particular wavelength (Fig. \ref{fig:Bimini}B). Indeed, some coastal dewlaps were substantially brighter than the rest (Fig. \ref{fig:Bimini}C), a pattern that was captured by our analysis of variance along PC1 (i.e. brightness, Fig. \ref{fig:Bimini}D, E, Table \ref{tab:anova}). Coastal lizards were also characterized by elevated red reflectance relative to UV (as represented by PC2, (Fig. \ref{fig:Bimini}D, E)), and coastal and mangrove lizards were characterized by a more even distribution of the reflectance along the spectrum (as represented by PC3, (Fig. \ref{fig:Bimini}D, E)), in contrast to coppice lizards which harbored a stronger curvature at intermediate wavelengths (Fig. \ref{fig:Bimini}D, E). On this island, the coastal and coppice habitats were separated by a few hundred meters, making this contrast the smallest geographical scale at which differences in coloration were found in our study (Fig. \ref{fig:Bimini}F, G).\\

On Cayman Brac, all three habitats could be fairly well discriminated against each other (Fig. \ref{fig:CaymanBrac}A), with UV reflectance appearing to be again an important variable (Fig. \ref{fig:CaymanBrac}B). Coastal and mangrove lizards were the best differentiated habitats (Fig. \ref{fig:CaymanBrac}A). At a distance between 2 and 3km (Fig. \ref{fig:CaymanBrac}F, G), dewlaps in the coastal habitat reflected more red light (as represented by PC2, Fig. \ref{fig:CaymanBrac}D, E) and more UV (as represented by PC3) than in the mangrove habitat. Coppice lizards were characterized by darker dewlaps than the rest (represented by PC1), and a higher UV reflectance than mangrove dewlaps (PC3, Fig. \ref{fig:CaymanBrac}D, E, Table \ref{tab:anova}).\\

Eleuthera was the only island where we detected significant spatial autocorrelation (Table \ref{tab:autocorrelation}), that is, sites that were closer geographically tended to have populations of lizards with more similar dewlap colors. The strongest identified differences were between coastal and mangrove lizards (Fig. \ref{fig:Eleuthera}A), where coastal lizards had higher levels of red reflectance and mangrove lizards higher levels of UV reflectance (as represented by PC2, Fig. \ref{fig:Eleuthera}D, E, Table \ref{tab:anova}).\\

Little Cayman was also characterized by a good discrimination of all three habitats, particularly of mangrove lizards (Fig. \ref{fig:LittleCayman}A). These lizards differed in short wavelengths from the rest (Fig. \ref{fig:LittleCayman}B), with significantly lower UV reflectance (as represented by PC2, Fig. \ref{fig:LittleCayman}D, E, Table \ref{tab:anova}). Coastal lizards were characterized by brighter dewlaps than coppice lizards (PC1), and also more convex curves, i.e. slightly higher UV and red reflectance (as represented by higher PC3 scores), than lizards from the other two habitats (Fig. \ref{fig:LittleCayman}D, E, Table \ref{tab:anova}).\\

On Long Island the three habitats were well discriminated too, with the largest differences between coastal and mangrove lizards (Fig. \ref{fig:LongIsland}A). Coastal lizards had more curved reflectance profiles than in either of the two other habitats, with higher levels of UV and red reflectance relative to intermediate wavelengths (PC3, Fig. \ref{fig:LongIsland}D, E, Table \ref{tab:anova}). Coastal lizards also differed from mangrove lizards along PC4 (Fig. \ref{fig:LongIsland}D), which represents a rather small portion of the variance not already explained by the first three principal components, and is therefore difficult to interpret (Fig. \ref{fig:LongIsland}E). Coppice lizards were significantly darker than mangrove and coastal lizards (PC1, Fig. \ref{fig:LongIsland}D, E, Table \ref{tab:anova}).\\

On North Andros, although the random forest classification was significant ($P = 0.0216$, Table \ref{tab:randomforests}) and the average confusion matrix indicated that lizards from beach scrub were particularly well predicted (Fig. \ref{fig:NorthAndros}A), no significant univariate differences were detected along any of the four PCs (Fig. \ref{fig:NorthAndros}D, Table \ref{tab:anova}). Importance analysis of full-spectrum random forests showed higher importance scores near the UV-end of the spectrum in discriminating beach scrub dewlaps from the rest (Fig. \ref{fig:NorthAndros}B). Besides, reflectance curves of beach scrub dewlaps appeared more similar to each other in the UV range than dewlaps from other habitats (Fig. \ref{fig:NorthAndros}C), suggesting that the machines may have used this low within-habitat variance, as opposed to between-habitat differences in means, to correctly classify beach scrub lizards. A small sample size on this islands may also have contributed to a lack of power in detecting univariate differences using analyses of variance (Table \ref{tab:counts}).\\

On South Andros classification was not as successful as on other islands, but coastal and coppice dewlaps could be discriminated better than expected by chance (Fig. \ref{fig:SouthAndros}A). Coppice lizards had more curved reflectance profiles than coastal lizards (PC3), and lizards from both habitats differed along PC4, which is again more difficult to interpret (Fig. \ref{fig:SouthAndros}D, E, Table \ref{tab:anova}). Coastal lizards also differed from mangrove lizards in PC4 (Fig. \ref{fig:SouthAndros}D, E, Table \ref{tab:anova}).\\

Classification success was not significantly better than expected by chance on Ragged Island (Table \ref{tab:randomforests}) where nearly no habitat could be differentiated from any other based on reflectance.

### 2021-04-10 ms/sections/suppfigures.tex
new figure on why did archipelago wide machines detect anything
many things happened

--- 2020-05-24 ms/sections/suppfigures.tex
@@
-% Map of each island
 \begin{figure}[H]
 	\centering
-	\includegraphics[width=\textwidth]{../maps/detailed_map.pdf}
-	\caption{Map of the sampling sites and corresponding habitats across nine islands of the West Indies.}
-	\label{supfig:map}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/03-PCA/figure_brightness"}
-	\caption{Correlation between dewlap brightness (as measured by the mean reflectance from 300 to 700nm in wavelength) and PC1 score for each island.
-Pearson's squared correlation coefficients are reported.
-*, P < 0.05.}
-	\label{supfig:brightness}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.5\textwidth]{"../analyses/03-PCA/figure_brightness_pooled"}
-	\caption{Correlation between dewlap brightness (as measured by the mean reflectance from 300 to 700nm in wavelength) and PC1 score across the whole archipelago.
-Pearson's squared correlation coefficient is reported.
-*, P < 0.05.}
-	\label{supfig:brightness_pooled}
-\end{figure}
-
-% Reflectance curves
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/02-reflectance/figure_reflectance"}
-	\caption{5-95th percentile range of lizard dewlap reflectance values (in \% of incoming light) across wavelengths for each island and each habitat.}
-	\label{supfig:reflectance}
+	\includegraphics[width=\textwidth]{figures/all_maps.png}
+	\caption{Maps of the islands. (A) Map of the West Indies with sampled islands highlighted in black. (B) Sampling sites within islands colored after their respective habitat-types.}
+	\label{fig:maps}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/classif_svm_pca_pooled"}
-	\caption{Archipelago-wide SVM classification accuracy based on principal component data.
-Machines were trained on individual dewlaps regardless of island identity.
-The histogram shows the accuracy distribution over 100 replicates for each five cross-validation bins.
-The legend is the same as in Figure \ref{fig:classif-svm-pca}.}
-	\label{supfig:classif-svm-pca-pooled}
+	\includegraphics[width=\textwidth]{figures/Bimini.png}
+	\caption{Comparison of dewlap coloration across habitats on Bimini. Legend is as per Figure \ref{fig:Abaco}.}
+	\label{fig:Bimini}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/classif_svm_refl_pooled"}
-	\caption{Archipelago-wide SVM classification accuracy based on reflectance data at 50nm-intervals in wavelength (see Methods).
-Machines were trained on individual dewlaps regardless of island identity.
-The histogram shows the accuracy distribution over 100 replicates for each five cross-validation bins.
-The legend is the same as in Figure \ref{fig:classif-svm-pca}.}
-	\label{supfig:classif-svm-refl-pooled}
+	\includegraphics[width=\textwidth]{figures/CaymanBrac.png}
+	\caption{Comparison of dewlap coloration across habitats on Cayman Brac. Legend is as per Figure \ref{fig:Abaco}.}
+	\label{fig:CaymanBrac}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.4\textwidth]{"../analyses/04-machine learning/plots/importance_svm_pca_pooled"}
-	\caption{Sensitivity analyses of the different input variables in the archipelago-wide SVM classification on principal component data (Figure \ref{supfig:classif-svm-pca-pooled}), with relative importance computed for every machine.}
-	\label{supfig:importance-svm-pca-pooled}
+	\includegraphics[width=\textwidth]{figures/Eleuthera.png}
+	\caption{Comparison of dewlap coloration across habitats on Eleuthera. Legend is as per Figure \ref{fig:Abaco}.}
+	\label{fig:Eleuthera}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{"../analyses/04-machine learning/plots/importance_svm_refl_pooled"}
-	\caption{Sensitivity analyses of the different input variables in the archipelago-wide SVM classification on reflectance data at 50nm-intervals in wavelength (Figure \ref{supfig:classif-svm-refl-pooled}), with relative importance computed for every machine.}
-	\label{supfig:importance-svm-refl-pooled}
+	\includegraphics[width=\textwidth]{figures/LittleCayman.png}
+	\caption{Comparison of dewlap coloration across habitats on Little Cayman. Legend is as per Figure \ref{fig:Abaco}.}
+	\label{fig:LittleCayman}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/classif_lda_pca_pooled"}
-	\caption{Archipelago-wide LDA classification accuracy based on principal component data.
-Machines were trained on individual dewlaps regardless of island identity.
-The histogram shows the accuracy distribution over 100 replicates for each five cross-validation bins.
-The legend is the same as in Figure \ref{fig:classif-svm-pca}.}
-	\label{supfig:classif-lda-pca-pooled}
+	\includegraphics[width=\textwidth]{figures/LongIsland.png}
+	\caption{Comparison of dewlap coloration across habitats on Long Island. Legend is as per Figure \ref{fig:Abaco}.}
+	\label{fig:LongIsland}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/classif_lda_refl_pooled"}
-	\caption{Archipelago-wide LDA classification accuracy based on reflectance data at 50nm-intervals in wavelength (see Methods).
-Machines were trained on individual dewlaps regardless of island identity.
-The histogram shows the accuracy distribution over 100 replicates for each five cross-validation bins.
-The legend is the same as in Figure \ref{fig:classif-svm-pca}.}
-	\label{supfig:classif-lda-refl-pooled}
+	\includegraphics[width=\textwidth]{figures/NorthAndros.png}
+	\caption{Comparison of dewlap coloration across habitats on North Andros. Legend is as per Figure \ref{fig:Abaco}.}
+	\label{fig:NorthAndros}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_lda_pca"}
-	\caption{LDA classification accuracy across islands based on principal component data.
-Histograms show accuracy distributions over 100 replicates for each five cross-validation bins per island.
-The legend is the same as in Figure \ref{fig:classif-svm-pca}.}
-	\label{supfig:classif-lda-pca}
+	\includegraphics[width=\textwidth]{figures/SouthAndros.png}
+	\caption{Comparison of dewlap coloration across habitats on South Andros. Legend is as per Figure \ref{fig:Abaco}.}
+	\label{fig:SouthAndros}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_svm_refl"}
-	\caption{SVM classification accuracy across islands based on reflectance data at 50nm-intervals in wavelength (see Methods).
-Histograms show accuracy distributions over 100 replicates for each five cross-validation bins per island.
-The legend is the same as in Figure \ref{fig:classif-svm-pca}.}
-	\label{supfig:classif-svm-refl}
+	\includegraphics[width=0.8\textwidth]{figures/brightness.png}
+	\caption{PC1 captures brightness across all islands.
+(A) Correlation between dewlap brightness (as measured by the mean reflectance from 300 to 700nm in wavelength) and PC1 score across the whole archipelago.
+(B) Correlation between brightness and PC1 for each island.
+Pearson's correlation coefficients are reported.
+***, $P < 0.001$.}
+	\label{fig:brightness}
 \end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\textwidth]{"../analyses/04-machine learning/plots/classif_lda_refl"}
-	\caption{LDA classification accuracy across islands based on reflectance data at 50nm-intervals in wavelength (see Methods).
-Histograms show accuracy distributions over 100 replicates for each five cross-validation bins per island.
-The legend is the same as in Figure \ref{fig:classif-svm-pca}.}
-	\label{supfig:classif-lda-refl}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/importance_svm_pca"}
-	\caption{Sensitivity analyses of the different input variables in the within-island SVM classification on principal component data (Figure \ref{supfig:classif-svm-pca}), with relative importance computed for every machine.}
-	\label{supfig:importance-svm-pca}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/importance_lda_pca"}
-	\caption{Sensitivity analyses of the different input variables in the within-island LDA classification on principal component data (Figure \ref{supfig:classif-lda-pca}), with relative importance computed for every machine.}
-	\label{supfig:importance-lda-pca}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/importance_svm_pca"}
-	\caption{Sensitivity analyses of the different input variables in the archipelago-wide SVM classification on reflectance at 50nm-intervals in wavelength (Figure \ref{supfig:classif-svm-refl}), with relative importance computed for every machine.}
-	\label{supfig:importance-svm-refl}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{"../analyses/04-machine learning/plots/importance_lda_pca"}
-	\caption{Sensitivity analyses of the different input variables in the archipelago-wide LDA classification on reflectance at 50nm-intervals in wavelength (Figure \ref{supfig:classif-lda-refl}), with relative importance computed for every machine.}
-	\label{supfig:importance-lda-refl}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\textwidth]{"../analyses/10-distances/figure_distances2"}
-	\caption{Spatial scale of between-habitat differences in dewlap coloration.
-For each variable and each pair of habitats where significant differences were detected (Figure \ref{fig:anova}), we performed multiple post hoc pairwise comparisons between the sites involved (Figure \ref{supfig:map}, Table \ref{suptab:sites}), using nonparametric Wilcoxon-Mann-Whitney tests.
-Here we report, for each pair of habitats, the distances between sites that significantly differed in dewlap coloration at an error rate of 0.05 (P-values corrected with the Benjamini-Hochberg procedure for multiple testing).}
-	\label{supfig:distances}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.7\textwidth]{"../analyses/09-parallelism/figure_contrasts"}
-	\caption{Test of parallel divergence between islands.
-Differences in habitat-means, or contrasts, are shown for two pairs of habitats for each principal component on each island, rescaled so the standard deviation of the means along each principal component is one.
-The contrasts represent the patterns of between-habtiat variation on each island, for a given principal component.
-The absence of clustering of islands by variable indicates that islands differ in their between-habitat divergence patterns.
-This is confirmed by a non-significant MANOVA test of the between versus within-variable variance in contrasts.}
-	\label{supfig:contrasts}
-\end{figure}

### 2021-04-10 ms/sections/tables.tex
tables and figures look good
nearly there
done?
implemented colin's comments and refactored the discussion to put more emphasis on neutral processes
many things happened

@@
-\begin{table}
+\begin{table}[H]
+	\caption{Random forest classification results.
+For each island are shown the sample size ($N$) and the proportion of correctly reassigned observations (or success score).
+P-values were computed using a binomial test and assess the significance of the observed success score relative to the score expected under random guessing.
+*, $P < 0.05$; **, $P < 0.01$, ***, $P < 0.001$.}
+	\centering
+	\input{"./tables/randomforests.tex"}
+	\label{tab:randomforests}
+\end{table}
+
+\begin{table}[H]
+	\caption{Test of spatial autocorrelation.
+For each island are shown the correlation (Pearson's $\rho$) between the matrix of phenotypic distances between populations from each site and the matrix of geographic distances between sites, where phenotypic distances are Euclidean distances between the mean phenotypes of each site in the multivariate space consisting of the first four within-island principal components.
+P-values assess the significance of the observed correlation against the correlation expected if lizards were randomly permuted among sites (1,000 permutations).
+*, $P < 0.05$; **, $P < 0.01$, ***, $P < 0.001$.}
+	\centering
+	\input{"./tables/autocorrelation.tex"}
+	\label{tab:autocorrelation}
+\end{table}
+
+\begin{sidewaystable}
+	\caption{Significance of habitat differences in dewlap coloration, using ANOVA for all islands where significant multivariate differences in dewlap coloration were detected by random forests.
+Model, best-fitting model (either OLS or GLS).
+AICc, corrected AIC score of the best-fitting model.
+$\Delta$AICc, difference in AICc between the best-fitting model and the OLS-model.
+AICcw, AICc weight.
+Log-lik., log-likelihood.
+$\chi^2$, likelihood ratio.
+df, degrees of freedom.
+*, $P < 0.05$; **, $P < 0.01$, ***, $P < 0.001$.}
+	\centering
+	\input{"./tables/anova.tex"}
 	\label{tab:anova}
-	\caption{}
-	\centering
-	\input{"../analyses/07-ANOVA/table_anova.tex"}
-\end{table}
+\end{sidewaystable}

### 2021-04-10 R/classifyX.R

@@ -1,0 +1,293 @@
+#' Classification analysis --- testing versions
+#'
+#' This function is an old version of `classify`, used only to compare different
+#' ways of computing P-values in their tendency to yield false positives.
+#' We concluded that comparing a distribution of observed accuracy scores to
+#' a null distribution was prone to false discovery because we have total
+#' control over the number of replicates, i.e. the number of points in the
+#' vector of observed scores.
+#'
+#' @export
+
+classifyX <- function(
+  data,
+  variables,
+  grouping,
+  nesting = NULL,
+  method = "SVM",
+  k = 5,
+  nrep = 1,
+  nperm = 0,
+  minsize = 5,
+  seed = NULL,
+  importance = FALSE,
+  getmachine = FALSE,
+  verbose = TRUE,
+  pb = TRUE,
+  digest = TRUE,
+  topcomp = NULL,
+  center = TRUE,
+  scale = TRUE
+) {
+
+  # Compute principal components if needed
+  if (!is.null(topcomp)) {
+    data <- data %>%
+      cbind(
+        npcomp(
+          data, topcomp, center, scale, nesting, combine = TRUE,
+          reduce = variables
+        )$x
+      )
+  }
+
+  # Random seed
+  if (!is.null(seed)) set.seed(seed)
+
+  # Proportion of observations in the testing set
+  ptesting <- 1 / k
+
+  assertthat::assert_that(nrow(data) > k)
+  assertthat::assert_that(floor(ptesting * nrow(data)) > 0)
+
+  # Define the possible labels
+  labels <- unique(data[, grouping])
+
+  nested <- TRUE
+
+  # Artificial single nesting level if no nesting supplied
+  if (is.null(nesting)) {
+    nested <- FALSE
+    data$nesting <- factor(1)
+    nesting <- "nesting"
+  }
+
+  # Split the data into each nesting level
+  data <- data %>% split(f = .[, nesting])
+
+  # Decide on a looping function depending on whether we want a progress bar
+  if (!verbose) pb <- FALSE
+  thislapply1 <- thislapply2 <- lapply
+  if (pb) {
+    if (nested) {
+      thislapply1 <- pbapply::pblapply
+    } else {
+      thislapply2 <- pbapply::pblapply
+    }
+  }
+
+  # For each nesting level...
+  results <- thislapply1(data, function(data) {
+
+    # For each replicate (and randomized) analysis...
+    thislapply2(seq(nrep + nperm), function(i) {
+
[…]

### 2021-04-10 R/csv2latex.R

@@ -1,0 +1,15 @@
+rm(list = ls())
+
+library(readr)
+library(knitr)
+
+filename <- "results/group_comparisons/machine_learning/PC/lda/summary.csv"
+
+tab <- read_csv(filename)
+
+# Process significance column
+colnames(tab)[ncol(tab)] <- ""
+tab[[ncol(tab)]][is.na(tab[[ncol(tab)]])] <- ""
+
+# Make a latex table
+kable(tab, "latex", booktabs = TRUE, linesep = "", escape = FALSE)

### 2021-04-10 analyses.R

@@ -1,0 +1,1980 @@
+## ---------------------------
+##
+## Script name: analyses.R
+##
+## Purpose of script: This script contains all the analyses of this study.
+##
+## Author: Raphael Scherrer
+##
+## Date Created: 2021-02-06
+##
+## Copyright (c) Raphael Scherrer, 2021
+##
+## Email:
+## r.scherrer@rug.nl
+## raphael.scherrer@evobio.eu
+## raph.rjfs@hotmail.fr
+##
+## ---------------------------
+
+# Notation used for naming datasets in this script:
+# D = the original dataset
+# X = placeholder within a function
+# P = temporary data frame for plotting or containing plots
+
+rm(list = ls())
+
+library(nmgc)
+library(tidyverse)
+library(ggridges)
+library(ggiraphExtra)
+library(broom)
+library(heplots)
+library(MVN)
+library(assertthat)
+library(rminer)
+library(geosphere)
+library(MuMIn)
+library(nlme)
+library(PMCMRplus)
+library(RColorBrewer)
+library(pairwiseAdonis)
+library(scales)
+library(vegan)
+
+#### 1. Set-up ####
+
+# Set plotting theme
+theme_set(theme_classic())
+
+# Graphical parameters
+island_colors <- brewer.pal(8, "Set2")
+island_colors <- c(island_colors, "darkgreen")
+habitat_colors <- c("goldenrod", "forestgreen", "mediumseagreen")
+
+# Global settings
+pc_names <- paste0("PC", 1:4)
+wpc_names <- paste0("w", pc_names)
+wl <- 300:700
+wl_names <- paste0("wl", wl)
+
+# Whether to load machines already fitted to whole spectrum data
+# (instead of rerunning the very lengthy classification)
+read_fitted <- TRUE
+
+#### 2. Read reflectance data ####
+
+# Read the data
+D <- read_csv("data/reflectance.csv")
+
+# Useful things to keep in mind
+island_names <- unique(D$island)
+habitat_names <- unique(D$habitat)
+
+# Save sample sizes
+with(D, table(island, habitat)) %>%
+  as_tibble() %>%
+  pivot_wider(names_from = habitat, values_from = n) %>%
+  write_csv("metadata/counts.csv")
+
+#### 2.1. Plot reflectance profiles ####
+
+# Prepare to plot reflectance
+P <- D %>%
[…]

### 2021-04-10 analyses/04-machine-learning/MachineLearning.R

@@ -9,39 +9,97 @@ data <- read.csv("data/reflectance.csv", header = TRUE)
 
 #### Analysis within each island ####
 
-# Principal components
+# Setup principal components
+pcomps <- paste0("PC", 1:4)
+wlengths <- paste0("wl", 300:700)
+somewl <- paste0("wl", seq(350, 650, 50))
 
-variables <- paste0("PC", 1:4)
+# Prepare a few ways to perform our analysis
+methods <- c("SVM", "LDA") # type of machine to fit
+datasets <- c("PC", "WL") # type of data to use
 
-# SVM-classification on PCs
+# For each method and each dataset
+for (i in seq_along(methods)) {
+  for (j in seq_along(datasets)) {
 
-res <- classify(
-  data, variables, grouping = "habitat", nesting = "island",
-  to_pcomp = paste0("wl", 300:700), digest = TRUE, test = TRUE, k = 5,
-  nrep = 100, seed = 24, method = "SVM", importance = TRUE
-)
+    # Set up
+    if (datasets[j] == "PC") {
+      variables <- pcomps
+      to_pcomp <- wlengths
+    } else {
+      variables <- somewl
+      to_pcomp <- NULL
+    }
 
-# Make sure that this script works with the new output of the classify function
-# before changing the package further
-classify2(
-  data, variables, grouping = "habitat", nesting = "island",
-  to_pcomp = paste0("wl", 300:700), k = 5,
-  nrep = 10, seed = 24, method = "SVM", importance = TRUE
-)
+    # Perform the analysis
+    res <- classify(
+      data, variables, grouping = "habitat", nesting = "island",
+      topcomp = to_pcomp, k = 5, nrep = 100, seed = 24, method = methods[i],
+      importance = TRUE, nperm = 1000
+    )
 
-# Can we try classify2 on made-up data?
-data2 <- data %>%
+    # Rename for practical purposes
+    res[[1]] <- rename(res[[1]], island = "nesting")
+
+    # Prepare file names for saving
+    outputname <- sprintf("results/classif_%s_%s.rds", methods[i], datasets[j])
+    tablename <- str_replace(outputname, ".rds", ".csv")
+    confplotname <- str_replace(outputname, ".rds", ".png")
+    confplotname <- str_replace(confplotname, "classif", "confplot")
+    accuplotname <- str_replace(confplotname, "confplot", "accuplot")
+
+    # Save
[…]

### 2021-04-10 classify3.R

@@ -1,0 +1,296 @@
+#' Classification analysis
+#'
+#' Perform a replicated classification analysis of a multivariate dataset into categorical labels using machine learning tools and k-fold cross validation
+#'
+#' @param data A data frame
+#' @param variables The variables used to classify
+#' @param grouping Name of the grouping variable (the labels)
+#' @param nesting Optional nesting variable, if the analysis must be conducted separately on different subsets of the data
+#' @param method The data mining model used. Currently supports "SVM" and "LDA".
+#' @param k Number of bins for the k-fold cross-validation procedure
+#' @param nrep Number of replicate analyses (i.e. number of k-fold cross validations)
+#' @param minsize Minimum size required per group for a training data set
+#' @param seed Optional random seed to reset at the beginning
+#' @param importance Whether to perform sensitivity analysis on the input (takes a while)
+#' @param return_machine Whether to return the machines (takes space)
+#' @param verbose Whether to display messages
+#' @param pb Whether to display progress bars
+#' @param to_pcomp Variable to perform PCA on
+#' @param center Center the PCA
+#' @param scale Scale the PCA
+#'
+#' @return A list of two tibbles.
+#'
+#' The first tibble, `full`, contains information about each fitted machine in the analysis, including its nesting level `nesting`, replicate number `repl` and cross-validation bin `kbin` but also classification testing outputs such as its confusion matrix `confmat`, its prediction `accuracy` and the number of observations in the testing set `ntested`. If `return_machine` is TRUE, an extra column is added, that is a list of the actual fitted machines. If `importance` is TRUE, extra columns are added, named after the input variables and containing their respective relative importance scores (these are computed using the `rminer::Importance` function for one-dimensional sensitivity analysis).
+#'
[…]

### 2021-04-10 classify4.R

@@ -1,0 +1,296 @@
+#' Classification analysis
+#'
+#' Perform a replicated classification analysis of a multivariate dataset into categorical labels using machine learning tools and k-fold cross validation
+#'
+#' @param data A data frame
+#' @param variables The variables used to classify
+#' @param grouping Name of the grouping variable (the labels)
+#' @param nesting Optional nesting variable, if the analysis must be conducted separately on different subsets of the data
+#' @param method The data mining model used. Currently supports "SVM" and "LDA".
+#' @param k Number of bins for the k-fold cross-validation procedure
+#' @param nrep Number of replicate analyses (i.e. number of k-fold cross validations)
+#' @param minsize Minimum size required per group for a training data set
+#' @param seed Optional random seed to reset at the beginning
+#' @param importance Whether to perform sensitivity analysis on the input (takes a while)
+#' @param return_machine Whether to return the machines (takes space)
+#' @param verbose Whether to display messages
+#' @param pb Whether to display progress bars
+#' @param to_pcomp Variable to perform PCA on
+#' @param center Center the PCA
+#' @param scale Scale the PCA
+#'
+#' @return A list of two tibbles.
+#'
+#' The first tibble, `full`, contains information about each fitted machine in the analysis, including its nesting level `nesting`, replicate number `repl` and cross-validation bin `kbin` but also classification testing outputs such as its confusion matrix `confmat`, its prediction `accuracy` and the number of observations in the testing set `ntested`. If `return_machine` is TRUE, an extra column is added, that is a list of the actual fitted machines. If `importance` is TRUE, extra columns are added, named after the input variables and containing their respective relative importance scores (these are computed using the `rminer::Importance` function for one-dimensional sensitivity analysis).
+#'
[…]

### 2021-04-10 mutate_signif.R

@@ -1,0 +1,26 @@
+# Function to add a significance column with asterisk symbols to a tibble
+# (We are going to need that)
+# Use on a tibble that contains P-values from test results in a column
+# Arguments:
+# D: the data frame
+# col: the name of the column containing the P-values
+# levels: the significance levels for the three labels *, ** and ***, in that
+# order
+# dropname: whether to replace the name of the extra column by a blank
+
+mutate_signif <- function(
+  D, col = "pvalue", levels = c(0.05, 0.01, 0.001), dropname = TRUE
+) {
+
+  D <- D %>%
+    mutate(
+      .signif = ifelse(get(col) < levels[1], "*", ""),
+      .signif = ifelse(get(col) < levels[2], "**", .signif),
+      .signif = ifelse(get(col) < levels[3], "***", .signif)
+    )
+
+  if (dropname) D <- D %>% rename(" " = ".signif")
+
+  return(D)
+
+}

### 2021-04-10 tables.R

@@ -1,0 +1,227 @@
+## ---------------------------
+##
+## Script name: tables.R
+##
+## Purpose of script:
+##
+## This script loads relevant tables from the results and
+## turns them into nice-looking LaTeX tables
+##
+## Author: Raphael Scherrer
+##
+## Date Created: 2021-04-05
+##
+## Copyright (c) Raphael Scherrer, 2021
+##
+## Email:
+## r.scherrer@rug.nl
+## raphael.scherrer@evobio.eu
+## raph.rjfs@hotmail.fr
+##
+## ---------------------------
+
+rm(list = ls())
+
+library(tidyverse)
+library(knitr)
+library(kableExtra)
+
+# Function to add a significance column with asterisk symbols to a tibble
+mutate_signif <- function(
+  D, col = "pvalue", levels = c(0.05, 0.01, 0.001), dropname = TRUE
+) {
+
+  D <- D %>%
+    mutate(
+      .signif = ifelse(get(col) < levels[1], "*", ""),
+      .signif = ifelse(get(col) < levels[2], "**", .signif),
+      .signif = ifelse(get(col) < levels[3], "***", .signif)
+    )
+
+  if (dropname) D <- D %>% rename(" " = ".signif")
+
+  return(D)
+
+}
+
+# Round and format P-values
+round_pvalue <- function(pvalue, digits = 4) {
+
+  threshold <- 1 / 10^digits
+  template <- paste0("%.", sprintf("%sf", digits))
+  label <- sprintf(paste("<", template), threshold, digits)
+  rounded <- sprintf(template, round(pvalue, digits), digits)
+  ifelse(pvalue < threshold, label, rounded)
+
+}
+
+# Random forest results
+read_csv("results/group_comparisons/machine_learning/PC/randomforest/summary.csv") %>%
+  mutate_signif() %>%
+  mutate(pvalue = round_pvalue(pvalue, digits = 4)) %>%
+  kable(
+    "latex",
+    col.names = c("Island", "$N$", "Score", "$P$", ""),
+    digits = c(0, 0, 3, 0, 0),
+    align = "lrrrl",
+    booktabs = TRUE,
+    linesep = "",
+    escape = FALSE
+  ) %>%
+  cat(file = "ms/tables/randomforests.tex")
[…]

### 2021-04-10 ms/sections/suppfigures.tex
added explained variance of for all islands to the explained variance table

@@
 	\caption{PC1 captures brightness across all islands.
-(A) Correlation between dewlap brightness (as measured by the mean reflectance from 300 to 700nm in wavelength) and PC1 score across the whole archipelago.
-(B) Correlation between brightness and PC1 for each island.
+(A) Correlation between dewlap brightness (as measured by the mean reflectance from 300 to 700nm in wavelength) and PC1 score across all islands.
+(B) Correlation between brightness and within-island PC1, for each island.
 Pearson's correlation coefficients are reported.

### 2021-04-10 analyses.R

@@ -602,14 +602,30 @@ with(P, walk2(data, island, function(X, island) {
 #### 6.4. Explained variance ####
 
 # Make a table with the variance explained by each PC on each island
-map(wPCA, ~ with(.x, sdev[1:4] / sum(sdev))) %>%
+X <- map(wPCA, ~ with(.x, sdev[1:4] / sum(sdev))) %>%
   do.call("rbind", .) %>%
   as_tibble(.name_repair = ) %>%
   mutate(expvar = rowSums(.)) %>%
   mutate(island = island_names) %>%
   setNames(c(pc_names, "expvar", "island")) %>%
-  select(island, expvar, pc_names) %>%
-  write_csv("results/pc_expvars/pc_expvars.csv")
+  select(island, expvar, pc_names)
+
+# Record explained variance for the global PCA on all islands
+pc_expvar <- with(PCA, sdev[1:4] / sum(sdev))
+names(pc_expvar) <- pc_names
+
+# Prepare extra row for all islands
+new_row <- c(
+  island = "All islands",
+  expvar = sum(pc_expvar),
+  as.list(pc_expvar)
+)
+
+# Append extra row
+X <- X %>% add_row(!!!new_row)
+
+# Save the table
+write_csv(X, "results/pc_expvars/pc_expvars.csv")
 
 #### 6.5. Mapping of the PC onto the reflectance curves ####
 

### 2021-04-10 tables.R

@@ -26,6 +26,8 @@ library(tidyverse)
 library(knitr)
 library(kableExtra)
 
+#### 0. Accessory functions ####
+
 # Function to add a significance column with asterisk symbols to a tibble
 mutate_signif <- function(
   D, col = "pvalue", levels = c(0.05, 0.01, 0.001), dropname = TRUE
@@ -55,6 +57,8 @@ round_pvalue <- function(pvalue, digits = 4) {
 
 }
 
+#### 1. Random forest table ####
+
 # Random forest results
 read_csv("results/group_comparisons/machine_learning/PC/randomforest/summary.csv") %>%
   mutate_signif() %>%
@@ -69,6 +73,8 @@ read_csv("results/group_comparisons/machine_learning/PC/randomforest/summary.csv
     escape = FALSE
   ) %>%
   cat(file = "ms/tables/randomforests.tex")
+
+#### 2. LDA table ####
 
 # LDA results
 read_csv("results/group_comparisons/machine_learning/PC/lda/summary.csv") %>%
@@ -87,6 +93,8 @@ read_csv("results/group_comparisons/machine_learning/PC/lda/summary.csv") %>%
   ) %>%
   cat(file = "ms/tables/ldas.tex")
 
+#### 3. SVM table ####
+
 # SVM results
 read_csv("results/group_comparisons/machine_learning/PC/ksvm/summary.csv") %>%
   mutate_signif() %>%
@@ -103,6 +111,8 @@ read_csv("results/group_comparisons/machine_learning/PC/ksvm/summary.csv") %>%
     escape = FALSE
   ) %>%
   cat(file = "ms/tables/ksvms.tex")
+
+#### 4. ANOVA table ####
 
 # ANOVA table
 read_csv("results/group_comparisons/anovas.csv") %>%
@@ -122,6 +132,8 @@ read_csv("results/group_comparisons/anovas.csv") %>%
   ) %>%
   cat(file = "ms/tables/anova.tex")
 
+#### 5. Spatial autocorrelation table ####
+
 # Spatial autocorrelation
 read_csv("results/spatial_correlation/spatial_correlation.csv") %>%
   mutate_signif() %>%
@@ -139,6 +151,8 @@ read_csv("results/spatial_correlation/spatial_correlation.csv") %>%
   ) %>%
   cat(file = "ms/tables/autocorrelation.tex")
 
+#### 6. Sample size table ####
+
[…]

### 2021-04-10 ms/sections/results2.tex
revised results but latex broken

@@
 We tested for variation in \textit{A.
-sagrei} dewlap coloration between populations living in three characteristic habitat types across nine islands that span the West Indian range of the brown anole (coastal scrub, primary coppice and mangroves).
-We found that most of the variation in coloration was partitioned between islands (two-way PERMANOVA, approx.
-$F(\text{df} = 8) = 43.7$, $P = 0.0001$, Fig.
-\ref{fig:reflectance}, explained variance $R^2 = 40.9$\%).
-Nonetheless, we did find evidence for differences in dewlap coloration between habitat-types, and those were mostly island-specific (habitat-by-island interaction term, approx.
-$F(16) = 3.53$, $P = 0.0001$, $R^2 = 6.6$\%), with a small but significant portion of the variation explained by an archipelago-wide habitat effect (approx.
-$F(2) = 4.7$, $P = 0.0001$, $R^2 = 1.1$\%).\\
+sagrei} dewlap coloration between populations living in three characteristic habitat-types across nine islands that span the West Indian range of the brown anole (coastal scrub, primary coppice and mangroves).
+We found that most of the variation in coloration was partitioned between islands (two-way PERMANOVA, $F(df = 8) = 43.7$, $P = 0.0001$}, explained variance $R^2 = 40.9$\%).
+Nonetheless, we did find evidence for differences in dewlap coloration between habitat-types, and those were mostly island-specific (habitat-by-island interaction term, $F(16) = 3.53$, $P = 0.0001$, $R^2 = 6.6$\%), with a significant portion of the variation explained by an habitat effect across all islands, but this effect was very small ($F(2) = 4.7$, $P = 0.0001$, $R^2 = 1.1$\%).\\
 
-We subsequently tested for differences in dewlap coloration between habitat-populations within each island, using within-island principal component scores, that is, computed from the data specific to each island (to maximize the variation captured for each island).
-Our within-island random forest classification analyses revealed detectable differences in dewlap coloration on seven out of the nine islands in our sample: Abaco, Bimini, Cayman Brac, Eleuthera, Little Cayman, Long Island and South Andros.
+We subsequently tested for differences in dewlap coloration between habitat-populations within each island, using within-island principal component scores, that is, from components computed from the data specific to each island (to maximize the variation captured for each island, see Methods).
+Our within-island random forest classification analyses revealed detectable differences in dewlap coloration on seven out of the nine islands in our sample: Abaco, Bimini, Cayman Brac, Eleuthera, Little Cayman, Long Island, North Andros and South Andros.
 On these islands, our classifiers could reassign individual dewlaps to their correct habitat more often than expected by chance (Table \ref{tab:randomforests}).
-We obtained nearly identical results using support vector machines (Table \ref{tab:ksvms}) and linear discriminant analysis (Table \ref{tab:ldas}) for classification.
+We obtained similar results using other machine learning approaches such as support vector machines (Table \ref{tab:ksvms}) and linear discriminant analysis (Table \ref{tab:ldas}), except that these methods did not detect significant differences on Eleuthera and North Andros.
 We did not find evidence of spatial autocorrelation in dewlap coloration between the sites within islands, except for Eleuthera (Table \ref{tab:autocorrelation}).
@@
 
-On Abaco, we detected the strongest differences between the coastal and mangrove habitats, while dewlaps from the coppice habitat were more difficult to classify (Fig.
+On Abaco, dewlaps from the mangrove habitat were the best discriminated, while dewlaps from the coastal habitat were often mistaken for dewlaps from the coppice habitat (Fig.
 \ref{fig:Abaco}A).
@@
 
-On Bimini, the random forests mostly correctly classified lizards from the coastal habitat (Fig.
-\ref{fig:Bimini}A), with a relatively flat importance profile suggesting that brightness was used instead of a particular wavelength (Fig.
+On Bimini, the random forests mostly correctly classified lizards from the coppice and mangrove habitats while often misclassifying lizards from the coastal habitat (Fig.
+\ref{fig:Bimini}A).
+Relatively flat importance profiles for coastal lizards suggested that brightness was used instead of a particular wavelength to identify some of the coastal dewlaps (Fig.
 \ref{fig:Bimini}B).
@@
 \ref{fig:Bimini}D, E, Table \ref{tab:anova}).
-Coastal lizards were also characterized by elevated red reflectance relative to UV (as represented by PC2, (Fig.
+The random forests also used UV reflectance to discriminate between coppice and mangrove dewlaps (Fig.
+\ref{fig:Bimini}B), which could reflect the significant difference we detected along PC3 between these two habitats (Fig.
+\ref{fig:Bimini}D, Table \ref{tab:anova}).
+Coastal lizards were characterized by elevated red reflectance relative to UV (as represented by PC2, (Fig.
 \ref{fig:Bimini}D, E)), and coastal and mangrove lizards were characterized by a more even distribution of the reflectance along the spectrum (as represented by PC3, (Fig.
@@
 
-On Cayman Brac, all three habitats could be fairly well discriminated against each other (Fig.
-\ref{fig:CaymanBrac}A), with UV reflectance appearing to be again an important variable (Fig.
+On Cayman Brac, all three habitats could be well discriminated against each other (Fig.
+\ref{fig:CaymanBrac}A), with UV reflectance appearing to be an important variable differentiating coastal and mangrove dewlaps (Fig.
 \ref{fig:CaymanBrac}B).
-Coastal and mangrove lizards were the best differentiated habitats (Fig.
-\ref{fig:CaymanBrac}A).
+In contrast, coppice dewlaps had a relatively flat importance profile, suggesting that brightness made them more distinct rather than any particular wavelength (Fig.
+\ref{fig:CaymanBrac}B).
+Consistent with this, coppice dewlaps were significantly different from all other dewlaps along PC1 (Fig.
+\ref{fig:CaymanBrac}D, Table \ref{tab:anova}).
 At a distance between 2 and 3km (Fig.
@@
 \ref{fig:CaymanBrac}D, E) and more UV (as represented by PC3) than in the mangrove habitat.
-Coppice lizards were characterized by darker dewlaps than the rest (represented by PC1), and a higher UV reflectance than mangrove dewlaps (PC3, Fig.
+Coppice lizards were also characterized by a higher UV reflectance than mangrove dewlaps (PC3, Fig.
 \ref{fig:CaymanBrac}D, E, Table \ref{tab:anova}).\\
@@
 Eleuthera was the only island where we detected significant spatial autocorrelation (Table \ref{tab:autocorrelation}), that is, sites that were closer geographically tended to have populations of lizards with more similar dewlap colors.
-The strongest identified differences were between coastal and mangrove lizards (Fig.
-\ref{fig:Eleuthera}A), where coastal lizards had higher levels of red reflectance and mangrove lizards higher levels of UV reflectance (as represented by PC2, Fig.
+Although random forests detected between-habitat differences in dewlap color, other approaches did not (Tables \ref{tab:ldas} and \ref{tab:ksvms}), suggesting that the differences may be small.
+Consistent with this, the only significant univariate difference detected was for PC2 between coastal and mangrove lizards, where coastal lizards had higher levels of red reflectance and mangrove lizards higher levels of UV reflectance (Fig.
 \ref{fig:Eleuthera}D, E, Table \ref{tab:anova}).\\
 
-Little Cayman was also characterized by a good discrimination of all three habitats, particularly of mangrove lizards (Fig.
+Little Cayman was characterized by a better discrimination of mangrove lizards from the rest than between coastal and coppice lizards (Fig.
 \ref{fig:LittleCayman}A).
-These lizards differed in short wavelengths from the rest (Fig.
+Mangrove dewlaps were possibly most distinct with respect to their reflectance in short wavelengths (Fig.
 \ref{fig:LittleCayman}B), with significantly lower UV reflectance (as represented by PC2, Fig.
@@
 
-On Long Island the three habitats were well discriminated too, with the largest differences between coastal and mangrove lizards (Fig.
+On Long Island the three habitats were relatively well discriminated (Fig.
 \ref{fig:LongIsland}A).
+Importance profiles indicated that short wavelengths were used to discriminate between coastal and mangrove lizards (Fig.
+\ref{fig:LongIsland}B).
 Coastal lizards had more curved reflectance profiles than in either of the two other habitats, with higher levels of UV and red reflectance relative to intermediate wavelengths (PC3, Fig.
@@
 Coastal lizards also differed from mangrove lizards along PC4 (Fig.
-\ref{fig:LongIsland}D), which represents a rather small portion of the variance not already explained by the first three principal components, and is therefore difficult to interpret (Fig.
+\ref{fig:LongIsland}D), which represented a rather small portion of the variance not already explained by the first three principal components, and is therefore difficult to interpret (Fig.
 \ref{fig:LongIsland}E).
@@
 
-On South Andros classification was not as successful as on other islands, but coastal and coppice dewlaps could be discriminated better than expected by chance (Fig.
-\ref{fig:SouthAndros}A).
+On South Andros coastal and coppice dewlaps could be discriminated better against each other than with mangrove dewlaps (Fig.
+\ref{fig:SouthAndros}A), with importance profiles supported UV-reflectance as a predictor of coppice lizards (Fig.
+\ref{fig:SouthAndros}B).
 Coppice lizards had more curved reflectance profiles than coastal lizards (PC3), and lizards from both habitats differed along PC4, which is again more difficult to interpret (Fig.

### 2021-04-25 ms/sections/results2.tex
fixed latex

@@
 sagrei} dewlap coloration between populations living in three characteristic habitat-types across nine islands that span the West Indian range of the brown anole (coastal scrub, primary coppice and mangroves).
-We found that most of the variation in coloration was partitioned between islands (two-way PERMANOVA, $F(df = 8) = 43.7$, $P = 0.0001$}, explained variance $R^2 = 40.9$\%).
+We found that most of the variation in coloration was partitioned between islands (two-way PERMANOVA, $F(df = 8) = 43.7$, $P = 0.0001$, explained variance $R^2 = 40.9$\%).
 Nonetheless, we did find evidence for differences in dewlap coloration between habitat-types, and those were mostly island-specific (habitat-by-island interaction term, $F(16) = 3.53$, $P = 0.0001$, $R^2 = 6.6$\%), with a significant portion of the variation explained by an habitat effect across all islands, but this effect was very small ($F(2) = 4.7$, $P = 0.0001$, $R^2 = 1.1$\%).\\

### 2021-04-25 ms/sections/discussion2.tex
replaced coastal with beach scrub in main text

@@
 On some islands, other or additional variables differed, such as brightness, red reflectance or the reflectance at the ends of the spectrum visible to \textit{Anolis} lizards (UV and red, \citealt{Lazareva2012}) relative to intermediate wavelengths (blue-to-yellow).
-Similar portions of the spectrum were sometimes involved in opposite directions on different islands, such as on Abaco and Cayman Brac, where mangrove lizards had a higher UV-reflectance than coastal lizards on the former, but a lower UV-reflectance on the latter.
+Similar portions of the spectrum were sometimes involved in opposite directions on different islands, such as on Abaco and Cayman Brac, where mangrove lizards had a higher UV-reflectance than beach scrub lizards on the former, but a lower UV-reflectance on the latter.
 Under a sensory-drive scenario, we would have expected the same variables to be consistently divergent between habitats, or at least in a consistent direction.\\

### 2021-04-25 ms/sections/methods2.tex
replaced coastal with beach scrub in main text

@@
 Three habitats were sampled on each island based on characterizations by \citet{Howard1950} and \citet{Schoener1968}.
-Each habitat is clearly distinguishable by its dominant vegetation type --- xeric coastal scrub (open, relatively dry habitat consisting of low vegetation or isolated trees), primary coppice forest (closed-canopy forest) and mangrove forest (wet coastal habitat with trees growing in brackish water and high light penetration).
+Each habitat is clearly distinguishable by its dominant vegetation type --- xeric beach scrub (open, relatively dry habitat consisting of low vegetation or isolated trees), primary coppice forest (closed-canopy forest) and mangrove forest (wet coastal habitat with trees growing in brackish water and high light penetration).
 Sample sizes are given in Table \ref{tab:counts}.

### 2021-04-25 ms/sections/results2.tex
replaced coastal with beach scrub in main text

@@
 We tested for variation in \textit{A.
-sagrei} dewlap coloration between populations living in three characteristic habitat-types across nine islands that span the West Indian range of the brown anole (coastal scrub, primary coppice and mangroves).
+sagrei} dewlap coloration between populations living in three characteristic habitat-types across nine islands that span the West Indian range of the brown anole (beach scrub, primary coppice and mangroves).
 We found that most of the variation in coloration was partitioned between islands (two-way PERMANOVA, $F(df = 8) = 43.7$, $P = 0.0001$, explained variance $R^2 = 40.9$\%).
@@
 
-On Abaco, dewlaps from the mangrove habitat were the best discriminated, while dewlaps from the coastal habitat were often mistaken for dewlaps from the coppice habitat (Fig.
+On Abaco, dewlaps from the mangrove habitat were the best discriminated, while dewlaps from the beach scrub habitat were often mistaken for dewlaps from the coppice habitat (Fig.
 \ref{fig:Abaco}A).
-Importance analysis revealed that coastal and mangrove lizards mostly differed in reflectance in the UV-end of the spectrum (below 400nm, Fig.
-\ref{fig:Abaco}B), where mangrove dewlaps had higher UV reflectance relative to coastal lizards, and coppice lizards had an intermediate UV reflectance between the two other habitats (Fig.
+Importance analysis revealed that beach scrub and mangrove lizards mostly differed in reflectance in the UV-end of the spectrum (below 400nm, Fig.
+\ref{fig:Abaco}B), where mangrove dewlaps had higher UV reflectance relative to beach scrub lizards, and coppice lizards had an intermediate UV reflectance between the two other habitats (Fig.
 \ref{fig:Abaco}C).
@@
 \ref{fig:Abaco}D, E, Table \ref{tab:anova}).
-Coastal lizards also scored lower on PC3, indicating less curvature of the reflectance profile and relatively higher reflectance at intermediate wavelengths (blue-to-yellow) than at the ends of the range (Fig.
+Beach scrub lizards also scored lower on PC3, indicating less curvature of the reflectance profile and relatively higher reflectance at intermediate wavelengths (blue-to-yellow) than at the ends of the range (Fig.
 \ref{fig:Abaco}D, E).
@@
 
-On Bimini, the random forests mostly correctly classified lizards from the coppice and mangrove habitats while often misclassifying lizards from the coastal habitat (Fig.
+On Bimini, the random forests mostly correctly classified lizards from the coppice and mangrove habitats while often misclassifying lizards from the beach scrub habitat (Fig.
 \ref{fig:Bimini}A).
-Relatively flat importance profiles for coastal lizards suggested that brightness was used instead of a particular wavelength to identify some of the coastal dewlaps (Fig.
+Relatively flat importance profiles for beach scrub lizards suggested that brightness was used instead of a particular wavelength to identify some of the beach scrub dewlaps (Fig.
 \ref{fig:Bimini}B).
-Indeed, some coastal dewlaps were substantially brighter than the rest (Fig.
+Indeed, some beach scrub dewlaps were substantially brighter than the rest (Fig.
 \ref{fig:Bimini}C), a pattern that was captured by our analysis of variance along PC1 (i.e.
@@
 \ref{fig:Bimini}D, Table \ref{tab:anova}).
-Coastal lizards were characterized by elevated red reflectance relative to UV (as represented by PC2, (Fig.
-\ref{fig:Bimini}D, E)), and coastal and mangrove lizards were characterized by a more even distribution of the reflectance along the spectrum (as represented by PC3, (Fig.
+Beach scrub lizards were characterized by elevated red reflectance relative to UV (as represented by PC2, (Fig.
+\ref{fig:Bimini}D, E)), and beach scrub and mangrove lizards were characterized by a more even distribution of the reflectance along the spectrum (as represented by PC3, (Fig.
 \ref{fig:Bimini}D, E)), in contrast to coppice lizards which harbored a stronger curvature at intermediate wavelengths (Fig.
 \ref{fig:Bimini}D, E).
-On this island, the coastal and coppice habitats were separated by a few hundred meters, making this contrast the smallest geographical scale at which differences in coloration were found in our study (Fig.
+On this island, the beach scrub and coppice habitats were separated by a few hundred meters, making this contrast the smallest geographical scale at which differences in coloration were found in our study (Fig.
 \ref{fig:Bimini}F, G).\\
@@
 On Cayman Brac, all three habitats could be well discriminated against each other (Fig.
-\ref{fig:CaymanBrac}A), with UV reflectance appearing to be an important variable differentiating coastal and mangrove dewlaps (Fig.
+\ref{fig:CaymanBrac}A), with UV reflectance appearing to be an important variable differentiating beach scrub and mangrove dewlaps (Fig.
 \ref{fig:CaymanBrac}B).
@@
 At a distance between 2 and 3km (Fig.
-\ref{fig:CaymanBrac}F, G), dewlaps in the coastal habitat reflected more red light (as represented by PC2, Fig.
+\ref{fig:CaymanBrac}F, G), dewlaps in the beach scrub habitat reflected more red light (as represented by PC2, Fig.
 \ref{fig:CaymanBrac}D, E) and more UV (as represented by PC3) than in the mangrove habitat.
@@
 Although random forests detected between-habitat differences in dewlap color, other approaches did not (Tables \ref{tab:ldas} and \ref{tab:ksvms}), suggesting that the differences may be small.
-Consistent with this, the only significant univariate difference detected was for PC2 between coastal and mangrove lizards, where coastal lizards had higher levels of red reflectance and mangrove lizards higher levels of UV reflectance (Fig.
+Consistent with this, the only significant univariate difference detected was for PC2 between beach scrub and mangrove lizards, where beach scrub lizards had higher levels of red reflectance and mangrove lizards higher levels of UV reflectance (Fig.
 \ref{fig:Eleuthera}D, E, Table \ref{tab:anova}).\\
 
-Little Cayman was characterized by a better discrimination of mangrove lizards from the rest than between coastal and coppice lizards (Fig.
+Little Cayman was characterized by a better discrimination of mangrove lizards from the rest than between beach scrub and coppice lizards (Fig.
 \ref{fig:LittleCayman}A).
@@
 \ref{fig:LittleCayman}D, E, Table \ref{tab:anova}).
-Coastal lizards were characterized by brighter dewlaps than coppice lizards (PC1), and also more convex curves, i.e.
+Beach scrub lizards were characterized by brighter dewlaps than coppice lizards (PC1), and also more convex curves, i.e.
 slightly higher UV and red reflectance (as represented by higher PC3 scores), than lizards from the other two habitats (Fig.
@@
 \ref{fig:LongIsland}A).
-Importance profiles indicated that short wavelengths were used to discriminate between coastal and mangrove lizards (Fig.
+Importance profiles indicated that short wavelengths were used to discriminate between beach scrub and mangrove lizards (Fig.
 \ref{fig:LongIsland}B).
-Coastal lizards had more curved reflectance profiles than in either of the two other habitats, with higher levels of UV and red reflectance relative to intermediate wavelengths (PC3, Fig.
+Beach scrub lizards had more curved reflectance profiles than in either of the two other habitats, with higher levels of UV and red reflectance relative to intermediate wavelengths (PC3, Fig.
 \ref{fig:LongIsland}D, E, Table \ref{tab:anova}).
-Coastal lizards also differed from mangrove lizards along PC4 (Fig.
+Beach scrub lizards also differed from mangrove lizards along PC4 (Fig.
 \ref{fig:LongIsland}D), which represented a rather small portion of the variance not already explained by the first three principal components, and is therefore difficult to interpret (Fig.
 \ref{fig:LongIsland}E).
-Coppice lizards were significantly darker than mangrove and coastal lizards (PC1, Fig.
+Coppice lizards were significantly darker than mangrove and beach scrub lizards (PC1, Fig.
 \ref{fig:LongIsland}D, E, Table \ref{tab:anova}).\\
@@
 
-On South Andros coastal and coppice dewlaps could be discriminated better against each other than with mangrove dewlaps (Fig.
+On South Andros beach scrub and coppice dewlaps could be discriminated better against each other than with mangrove dewlaps (Fig.
 \ref{fig:SouthAndros}A), with importance profiles supported UV-reflectance as a predictor of coppice lizards (Fig.
 \ref{fig:SouthAndros}B).
-Coppice lizards had more curved reflectance profiles than coastal lizards (PC3), and lizards from both habitats differed along PC4, which is again more difficult to interpret (Fig.
+Coppice lizards had more curved reflectance profiles than beach scrub lizards (PC3), and lizards from both habitats differed along PC4, which is again more difficult to interpret (Fig.
 \ref{fig:SouthAndros}D, E, Table \ref{tab:anova}).
-Coastal lizards also differed from mangrove lizards in PC4 (Fig.
+Beach scrub lizards also differed from mangrove lizards in PC4 (Fig.
 \ref{fig:SouthAndros}D, E, Table \ref{tab:anova}).\\

### 2021-04-25 ms/sections/discussion2.tex
few last changes to discussion

@@
 Indeed, the yellow, orange and red colors in anoline dewlaps are produced by pterins and carotenoids \citep{Ortiz1962, Ortiz1962a, Ortiz1963, Ortiz1966, Macedonia2000, Steffen2007, Steffen2009}.
-Animals can be synthesize pterins from nucleotides, but lack the ability to synthesize carotenoids \citep{Goodwin1984, Hill2002, Hill2006}.
+Animals can synthesize pterins from nucleotides, but lack the ability to synthesize carotenoids \citep{Goodwin1984, Hill2002, Hill2006}.
 Different food qualities across sites within islands could therefore potentially cause detectable differences in coloration.
@@
 if the optimal male phenotype in a given habitat is a function of female perception, sexual preferences may also be arbitrary and independent of the habitat \citep{Andersson1994}, and so could differ across habitats and islands.
-However, one previous study has found no link between dewlap coloration and body size dimorphism in \textit{A.
-sagrei} (a proxy for the intensity of sexual selection) in an among-island comparison \citep{Baeckens2018}, and although within-island data are lacking, this scenario may have little plausibility.\\
+Although one previous study has found no link between dewlap coloration and body size dimorphism in \textit{A.
+sagrei} (a proxy for the intensity of sexual selection) in an among-island comparison \citep{Baeckens2018}, within-island data are lacking to test the plausibility of this scenario.\\
 

### 2021-05-23 ms/sections/abstract.tex
addressed most immediate comments

@@
 We explored variation in dewlap coloration in the most widespread species of anole, \textit{Anolis sagrei}, across three characteristic habitats spanning the Bahamas and the Cayman Islands.
-Using reflectance spectrometry as well as supervised machine learning, we found significant differences in spectral properties of the dewlap between habitats within small islands, sometimes over very short distances.
+Using reflectance spectrometry paired with supervised machine learning, we found significant differences in spectral properties of the dewlap between habitats within small islands, sometimes over very short distances.
 Passive divergence in dewlap phenotype associated with isolation-by-distance did not explain our results.
 On the other hand, these habitat-specific dewlap differences varied in magnitude and direction across islands, and thus our primary test for adaptation -- parallel responses across islands -- was not supported.
-We suggest, however, that selection could be involved in several ways, including sexual selection.
+We suggest that selection could be involved in several ways, including sexual selection.
 Our results shed new light on the scale at which signal color polymorphism can be maintained in the presence of gene flow, and the relative role of local adaptation and other processes in driving these patterns.

### 2021-05-23 ms/sections/acknowledgments.tex
addressed most immediate comments

@@
 Glor, Alberto R.
-Puente-Rol\'{o}n, Kevin Aviles-Rodriguez, Kristin Winchell, Jason Fredette and Melissa Kemp for assistance in the field and Pratik Gupte, Max Lambert and James Stroud for helpful discussions.
+Puente-Rol\'{o}n, Kevin Aviles-Rodriguez, Kristin Winchell, Jason Fredette, Quyhn Quach, Wendy Jesse, Inbar Maayan, Alexis Harrison and Melissa Kemp for assistance in the field and Pratik Gupte, Max Lambert and James Stroud for helpful discussions.
 Funding for this work was provided by the Templeton Foundation (to JBL), NSF DEB~\#1927194 (to JBL and AJG), NSF DEB~\#1500761 (to AJG), NSF DBI~\#1609284 (to CMD), and a Harvard Museum of Comparative Zoology Putnam Expedition Grant (to RGR).

### 2021-05-23 ms/sections/discussion2.tex
addressed most immediate comments

@@
 Detectable differences in dewlap color between habitat-populations are surprising, as habitats were often in close geographical proximity to each other (as close as a few hundred meters on Bimini and most of the time within ten kilometers).
-Indeed, given that (1) the populations were continuously distributed between the habitats, (2) different habitat-populations were not monophyletic with respect to mitochondrial haplotypes (van de Schoot, unpublished thesis), and (3) \textit{A.
-sagrei} have been shown to be a highly mobile species within these islands \citep{Kamath2018}, we would have expected more homogeneous distributions of color phenotypes within islands due to extensive gene flow, with fewer differences between populations, especially those in close proximity.\\
+Indeed, given that (1) the populations were continuously distributed between the habitats, (2) different habitat-populations were not monophyletic with respect to mitochondrial haplotypes (\citealt{vandeSchoot2016} unpubl.), and (3) \textit{A.
+sagrei} have been shown to be a highly mobile species \citep{Kamath2018}, we would have expected more homogeneous distributions of color phenotypes within islands due to extensive gene flow, with fewer differences between populations, especially those in close proximity.\\
 
@@
 
-Not only consistent patterns across islands would have been a good clue for a sensory-drive explanation, but in particular consistent differences between habitats that are most different in their local conditions regarding the ecological function of the dewlap, such as ambient light.
+Consistent patterns across islands would have been a good clue for a sensory-drive explanation, but in particular consistent differences between habitats that are most different in their local conditions regarding the ecological function of the dewlap, such as ambient light.
 For example, if ambient light is an important factor shaping dewlap coloration, we would expect mangrove and beach scrub lizards, both inhabiting areas with high light penetration, to harbor more similar dewlaps, and to differ significantly from lizards from the coppice habitat, where irradiance is low.
@@
 A number of alternative explanations remain.
-First, there could sexual selection for different dewlap colors in different locations.
+First, there could be sexual selection for different dewlap colors in different locations.
 Indeed, although the sensory-drive hypothesis may include a sexual selection aspect, e.g.

### 2021-05-23 ms/sections/introduction.tex
addressed most immediate comments

@@
 Animals use chemical, mechanical, electromagnetic, and visual signals to communicate in a wide variety of contexts, including, competition for mates, species recognition, aposematism, and cooperation \citep{Bradbury2011}.
-A primary evolutionary factor shaping communication signals is the sensory system and behavior of their recipients (the sensory drive hypothesis; \citealt{Endler1988,Endler1992,Endler1998}).
+A primary evolutionary factor shaping communication signals is the sensory system and behavior of recipients (the sensory drive hypothesis; \citealt{Endler1988,Endler1992,Endler1998}).
 Over the past decades, scientists have established that signals evolve in an ecological context and are dependent on environmental conditions \citep{Endler1992,Endler1993,Endler1993a}.
@@
 Population genetics theory suggests that gene flow may counteract local adaptation between localities and prevent divergence altogether, especially at small spatial scales, because of the inflow of maladapted alleles or because of the breaking of linkage between coevolving loci \citep{Felsenstein1976, Garcia-Ramos1997, Dieckmann1999, Lenormand2002, Hendry2007}.
-This genetic homogenization has been confirmed empirically in systems such as stick-insects \citep{Nosil2004} and sticklebacks \citep{Hendry2007a}.
+This genetic homogenization has been confirmed empirically in systems such as stick-insects \citep{Nosil2004} and stickleback \citep{Hendry2007a}.
 Yet, examples of microgeographic adaptation, i.e.
@@
 Lizards of the neotropical genus \textit{Anolis} are an excellent group for studying the eco-evolutionary dynamics of local adaptation and natural selection \citep{Losos2009}.
-A particularly conspicuous trait of anoles is their dewlap; an extensible flap of skin that is typically sexually dimorphic and used as a communication signal in courtship \citep{Sigmund1983, Driessens2014, Driessens2015} and territorial displays \citep{Losos1985, Macedonia1994, Macedonia2013} as well as in predator deterrence \citep{Leal1995, Leal1997, Leal1997b}.
+A particularly conspicuous trait of anoles is their dewlap; an extensible flap of skin that is typically sexually dimorphic and used as a communication signal in courtship \citep{Sigmund1983, Driessens2014, Driessens2015} and territorial displays \citep{Losos1985, Macedonia1994, Macedonia2013} as well as in predator deterrence \citep{Leal1995, Leal1997, Leal1997a}.
 Dewlap characteristics vary widely among the approximately $400$ species of the genus \citep{Nicholson2007}.
-Interspecific variation in dewlap coloration is implicated in species recognition \citep{Rand1970b, Williams1969, Williams1977, Losos1985, Macedonia1994, Fleishman2000, Macedonia2013}, and this function could have had a role in initiating and/or reinforcing reproductive isolation during speciation \citep{Lambert2013, Geneva2015, Ng2017}.\\
+Interspecific variation in dewlap coloration is implicated in species recognition \citep{Rand1970, Williams1969, Williams1977, Losos1985, Macedonia1994, Fleishman2000, Macedonia2013}, and this function could have had a role in initiating and/or reinforcing reproductive isolation during speciation \citep{Lambert2013, Geneva2015, Ng2017}.\\
 
-Within species, studies have shown a link between variation in dewlap coloration and differences in habitats or climatic conditions \citep{Macedonia2001, Leal2002, Thorpe2002a, Thorpe2002b, Leal2004, Vanhooydonck2009, Ng2012, Ng2013, Ng2016, Vanhooydonck2009, Driessens2017}.
+Within species, studies have shown a link between variation in dewlap coloration and differences in habitats or climatic conditions \citep{Macedonia2001, Leal2002, Thorpe2002, Thorpe2002a, Leal2004, Vanhooydonck2009, Ng2012, Ng2013, Ng2016, Vanhooydonck2009, Driessens2017}.
 Some studies suggest that those differences may be adaptive, and that dewlaps may have evolved to maximize detectability given local light conditions \citep{Fleishman2001, Leal2002, Leal2004}.
-Although this claim is further supported by recent findings that dewlap colors are perceived differently under different levels of shading \citep{Fleishman2020}, other studies found conflicting patterns of between-habitat variation that did not support the sensory drive hypothesis \citep{Fleishman2009, Ng2012, Macedonia2014a}.\\
+Although this claim is further supported by recent findings that dewlap colors are perceived differently under different levels of shading \citep{Fleishman2020}, other studies found conflicting patterns of between-habitat variation that did not support the sensory drive hypothesis \citep{Fleishman2009, Ng2012, Macedonia2014}.\\
 
@@
 These large scales and marine barriers should reduce gene flow \citep{Ng2011, Lambert2013, Richardson2014, Ng2017}.
-That said, examples do exist of divergence in dewlap coloration at smaller scales or between populations with high degrees of gene flow \citep{Thorpe2002a, Thorpe2002b, Stapley2011, Ng2016}.\\
+That said, examples do exist of divergence in dewlap coloration at smaller scales or between populations with high degrees of gene flow \citep{Thorpe2002, Thorpe2002a, Stapley2011, Ng2016}.\\
 
 The species \textit{Anolis sagrei} is widespread across islands of the West Indies \citep{Reynolds2020}.
-It has been the subject of numerous studies concerning local adaptation \citep{Losos1994, Losos1997a, Losos2001, Kolbe2012}, biological invasion \citep{Kolbe2008}, sexual selection \citep{Tokarz2002, Tokarz2005, Tokarz2006, Driessens2014, Steffen2014, Driessens2015} and many other topics.
+It has been the subject of numerous studies concerning local adaptation \citep{Losos1994, Losos1997a, Losos2001, Kolbe2012}, biological invasion \citep{Kolbe2008}, sexual selection \citep{Tokarz2002, Tokarz2005, Tokarz2006, Driessens2014, Steffen2014, Driessens2015} among many other topics.
 Between-island variation in the mainly orange-red color of its dewlap was shown to be better explained by climatic variables \citep{Driessens2017} than by proxies for biotic factors such as sexual selection or predation pressure \citep{Vanhooydonck2009, Baeckens2018}.
@@
 These island systems presently, if not historically, comprise relatively small islands, with no major geographic barriers within islands limiting dispersal for this promiscuous species \citep{Kamath2018}.
-These islands all share three characteristic native West Indian habitat-types -- beach scrub bush, closed-canopy primary coppice forest, and mangrove forest -- that are often spatially intermingled.
-These habitats contrast in environmental parameters including vegetation community, light irradiance, humidity and temperature \citep{Howard1950, Schoener1968}.
+These islands all share three characteristic native West Indian small-island habitat-types -- beach scrub bush, closed-canopy primary coppice forest, and mangrove forest -- that are often spatially intermingled.
+These habitats contrast in environmental parameters including vegetation community, light irradiance, humidity, and temperature \citep{Howard1950, Schoener1968}.
 The Cayman Islands and the Bahamas have been colonized independently by \textit{A.
-sagrei} from Cuba (\citealt{Reynolds2020}, van de Schoot et al.
-unpubl.), such that these archipelagos constitute an ideal suite of natural replicates to explore within-island dewlap diversity across multiple islands.\\
+sagrei} from Cuba (\citealt{vandeSchoot2016} unpublished thesis; \citealt{Reynolds2020}), such that these archipelagos constitute an ideal suite of natural replicates to explore within-island dewlap diversity across multiple islands.\\
 

### 2021-05-23 ms/sections/methods2.tex
addressed most immediate comments

@@
 We performed a machine learning classification analysis on the first four principal components within each island separately, using random forests \citep{Breiman2001}.
-Random forests are a versatile, intuitive and powerful algorithm commonly used in machine learning, using decision trees to predict the labels of particular observations, based on their multivariate coordinates, i.e.
-variables, passed through a series of successive decision nodes, each examining a given variable of any given observation \citep{James2013}.
-The prediction for each observation is an aggregate over a large number of decision trees, each being trained on a subset of observations sampled with replacement from the dataset, and each allowed to examine only a subset of the variables.
+Random forests are a versatile, intuitive, and powerful algorithm commonly used in machine learning, using decision trees to predict the labels of particular observations based on their multivariate coordinates.
+These coordinates, or variables, are passed through a series of successive decision nodes, each examining a given variable of any given observation \citep{James2013}.
+The prediction for each observation is an aggregate over a large number of decision trees, each tree being trained on a subset of observations sampled with replacement from the dataset, and each tree being allowed to examine only a subset of the variables.
 This allows the random forest to overcome the individual errors of all trees in the predictions it makes.\\
@@
 In machine learning, this so-called cross-validation procedure is typically done in two steps \citep{James2013}.
-First, a random forest is trained in recognizing features of dewlap coloration most associated with the different habitats, by being presented multiple observations, making predictions about them and updating its own decision rules based on whether the prediction deviates from the truth.
+First, a random forest is trained in recognizing features of dewlap coloration most associated with the different habitats, by being presented with multiple observations, making predictions about them, and updating its own decision rules based on whether the prediction deviates from the truth.
 Then, once trained, the patterns that the random forest has learned to recognize are tested by presenting new, previously unseen observations to the random forest, and measuring the proportion of correct predictions.
@@
 The cross-validation procedure requires that the data be split into a training set and a testing set.
-To remove any bias due to the set that is being sampled for training, it is common practice to use k-fold cross-validation \citep{James2013}, where the data is split into $k$ random bins and $k$ independent machines are trained, each taking one of the bins as a testing set and the rest for training, and where classification success is measured by summing all correct classifications from the $k$ machines.\\
+To remove any bias due to the set that is being sampled for training, it is common practice to use k-fold cross-validation \citep{James2013}, where the data are split into $k$ random bins and $k$ independent machines are trained, each taking one of the bins as a testing set and the rest for training, and where classification success is measured by summing all correct classifications from the $k$ machines.\\
 
@@
 
-We used the machine learning fitting functions in the R package rminer \citep{Cortez2020}, which call random forest routines from the randomForest package (\citealt{Liaw2002}, implementation from the original random forest algorithm, \citealp{Breiman2001}).
+We used the machine learning fitting functions in the R package rminer \citep{Cortez2020}, which calls random forest routines from the randomForest package (\citealt{Liaw2002}, implementation from the original random forest algorithm, \citealp{Breiman2001}).
 For each random forest, we optimized the number of trees in the forest and the number of variables examined by each tree using the grid hyperparameter search procedure implemented in rminer, to choose between two numbers of trees (500 or 1,000) and four numbers of principal components examined per tree (1 to 4), using rminer's ordered holdout validation method with $2/3$ of the data used for training.\\
@@
 In this study, we were interested in the minimum spatial scale at which significant differences between habitats could be detected within islands.
-We performed multiple pairwise nonparametric Wilcoxon-Mann-Whitney tests \citep{Hollander2013} to compare dewlap coloration between sites with different habitat-types, for each pair of habitats and each variable where significant differences were detected with our analyses of variances.
+We performed multiple pairwise nonparametric Wilcoxon-Mann-Whitney tests \citep{Hollander2013} to compare dewlap coloration between sites with different habitat-types, for each pair of habitats and each variable where significant differences were detected with our analyses of variance.
 The P-values were adjusted using a Benjamini-Hochberg correction for multiple testing \citep{Benjamini1995}.

### 2021-05-23 ms/sections/results2.tex
addressed most immediate comments

@@
 We found that most of the variation in coloration was partitioned between islands (two-way PERMANOVA, $F(df = 8) = 43.7$, $P = 0.0001$, explained variance $R^2 = 40.9$\%).
-Nonetheless, we did find evidence for differences in dewlap coloration between habitat-types, and those were mostly island-specific (habitat-by-island interaction term, $F(16) = 3.53$, $P = 0.0001$, $R^2 = 6.6$\%), with a significant portion of the variation explained by an habitat effect across all islands, but this effect was very small ($F(2) = 4.7$, $P = 0.0001$, $R^2 = 1.1$\%).\\
+Nonetheless, we did find evidence for differences in dewlap coloration between habitat-types, and those were mostly island-specific (habitat-by-island interaction term, $F(16) = 3.53$, $P = 0.0001$, $R^2 = 6.6$\%), with a significant portion of the variation explained by an habitat effect across all islands, but this effect was relatively small ($F(2) = 4.7$, $P = 0.0001$, $R^2 = 1.1$\%).\\
 
-We subsequently tested for differences in dewlap coloration between habitat-populations within each island, using within-island principal component scores, that is, from components computed from the data specific to each island (to maximize the variation captured for each island, see Methods).
-Our within-island random forest classification analyses revealed detectable differences in dewlap coloration on seven out of the nine islands in our sample: Abaco, Bimini, Cayman Brac, Eleuthera, Little Cayman, Long Island, North Andros and South Andros.
-On these islands, our classifiers could reassign individual dewlaps to their correct habitat more often than expected by chance (Table \ref{tab:randomforests}).
+We subsequently tested for differences in dewlap coloration between habitat-populations within each island, using within-island principal component scores (to maximize the variation captured for each island, see Methods).
+Our within-island random forest classification analyses revealed detectable differences in dewlap coloration on eight out of the nine islands in our sample: Abaco, Bimini, Cayman Brac, Eleuthera, Little Cayman, Long Island, North Andros and South Andros.
+The accuracy of Random forest classification exceeded random expectation more often than expected by chance for all these islands (Table \ref{tab:randomforests}).
+Accuracy was as high as 73\% for Cayman Brac.
 We obtained similar results using other machine learning approaches such as support vector machines (Table \ref{tab:ksvms}) and linear discriminant analysis (Table \ref{tab:ldas}), except that these methods did not detect significant differences on Eleuthera and North Andros.
@@
 We now describe the specific differences detected on each island.\\
+
+The accuracy of Random forest classification exceeded random expectation more often than expected by chance for all islands. Accuracy was as high as 73% for Cayman Brac.
 
@@
 \ref{fig:LittleCayman}A).
-Mangrove dewlaps were possibly most distinct with respect to their reflectance in short wavelengths (Fig.
+Mangrove dewlaps were most distinct with respect to their reflectance in short wavelengths (Fig.
 \ref{fig:LittleCayman}B), with significantly lower UV reflectance (as represented by PC2, Fig.
@@
 On South Andros beach scrub and coppice dewlaps could be discriminated better against each other than with mangrove dewlaps (Fig.
-\ref{fig:SouthAndros}A), with importance profiles supported UV-reflectance as a predictor of coppice lizards (Fig.
+\ref{fig:SouthAndros}A), with importance profiles supporting UV-reflectance as a predictor of coppice lizards (Fig.
 \ref{fig:SouthAndros}B).

### 2021-05-29 ms/sections/figures.tex
island maps and island figures refurbished

@@
 	\centering
-	\includegraphics[width=\textwidth]{figures/Abaco.png}
+	\includegraphics[width=\textwidth]{figures/Abaco_2.2.png}
 	\caption{Comparison of dewlap coloration across habitats on Abaco.
-(A) Confusion matrix showing the proportion of lizards from each (true) habitat reassigned to each (predicted) habitat by the random forests, based on the first four within-island principal components and averaged across replicates.
+(A) Map of the island with the sampling sites colored by habitat.
+(B) Reflectance profiles of all the dewlaps on the island.
+(C) How reflectance profiles map onto the within-island principal components.
+(D) Confusion matrix showing the proportion of lizards from each (true) habitat reassigned to each (predicted) habitat by the random forests, based on the first four within-island principal components and averaged across replicates.
 Each column sums to one.
-(B) One-dimensional sensitivity analysis showing the relative importance (mean decrease in accuracy) of the various wavelengths in random forest classification of the whole spectrum.
-(C) Reflectance profiles of all the dewlaps on the island.
-(D) Within-island principal component scores across habitats.
+(E) Within-island principal component scores across habitats.
 Bars indicate significant contrasts.
-*, $P < 0.05$; **, $P < 0.01$; ***, $P < 0.001$.
-(E) How reflectance profiles map onto the within-island principal components.
-(F) Map of the island with the sampling sites colored by habitat.
-(G) Geographical distance between sites where significant differences were detected in within-island principal component scores (Wilcoxon test, Benjamini-Hochberg correction, $P < 0.05$), including only pairs of sites whose habitats were involved in between-habitat dewlap differences.}
+*, $P < 0.05$; **, $P < 0.01$; ***, $P < 0.001$.}
 	\label{fig:Abaco}

### 2021-05-29 ms/sections/suppfigures.tex
island maps and island figures refurbished

@@
 	\centering
-	\includegraphics[width=\textwidth]{figures/all_maps.png}
+	\includegraphics[width=\textwidth]{figures/island_maps_white_2.0.png}
 	\caption{Maps of the islands. (A) Map of the West Indies with sampled islands highlighted in black. (B) Sampling sites within islands colored after their respective habitat-types.}
@@
 	\centering
-	\includegraphics[width=\textwidth]{figures/Bimini.png}
-	\caption{Comparison of dewlap coloration across habitats on Bimini. Legend is as per Figure \ref{fig:Abaco}.}
+	\includegraphics[width=\textwidth]{figures/Abaco_supplement_1.0b.png}
+	\caption{Comparison of dewlap coloration across habitats on Abaco, with extended results.
+(A--E) Legend as per Figure \ref{fig:Abaco}.
+(F) One-dimensional sensitivity analysis showing the relative importance (mean decrease in accuracy) of the various wavelengths in random forest classification of the whole spectrum.
+(G) Geographical distance between sites where significant differences were detected in within-island principal component scores (Wilcoxon test, Benjamini-Hochberg correction, $P < 0.05$), including only pairs of sites whose habitats were involved in between-habitat dewlap differences.}
+	\label{fig:Abaco_supplement}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{figures/Bimini_supplement.png}
+	\caption{Comparison of dewlap coloration across habitats on Bimini. Legend is as per Figure \ref{fig:Abaco_supplement}.}
 	\label{fig:Bimini}
@@
 	\centering
-	\includegraphics[width=\textwidth]{figures/CaymanBrac.png}
-	\caption{Comparison of dewlap coloration across habitats on Cayman Brac. Legend is as per Figure \ref{fig:Abaco}.}
+	\includegraphics[width=\textwidth]{figures/CaymanBrac_supplement.png}
+	\caption{Comparison of dewlap coloration across habitats on Cayman Brac. Legend is as per Figure \ref{fig:Abaco_supplement}.}
 	\label{fig:CaymanBrac}
@@
 	\centering
-	\includegraphics[width=\textwidth]{figures/Eleuthera.png}
-	\caption{Comparison of dewlap coloration across habitats on Eleuthera. Legend is as per Figure \ref{fig:Abaco}.}
+	\includegraphics[width=\textwidth]{figures/Eleuthera_supplement.png}
+	\caption{Comparison of dewlap coloration across habitats on Eleuthera. Legend is as per Figure \ref{fig:Abaco_supplement}.}
 	\label{fig:Eleuthera}
@@
 	\centering
-	\includegraphics[width=\textwidth]{figures/LittleCayman.png}
-	\caption{Comparison of dewlap coloration across habitats on Little Cayman. Legend is as per Figure \ref{fig:Abaco}.}
+	\includegraphics[width=\textwidth]{figures/LittleCayman_supplement.png}
+	\caption{Comparison of dewlap coloration across habitats on Little Cayman. Legend is as per Figure \ref{fig:Abaco_supplement}.}
 	\label{fig:LittleCayman}
@@
 	\centering
-	\includegraphics[width=\textwidth]{figures/LongIsland.png}
-	\caption{Comparison of dewlap coloration across habitats on Long Island. Legend is as per Figure \ref{fig:Abaco}.}
+	\includegraphics[width=\textwidth]{figures/LongIsland_supplement.png}
+	\caption{Comparison of dewlap coloration across habitats on Long Island. Legend is as per Figure \ref{fig:Abaco_supplement}.}
 	\label{fig:LongIsland}
@@
 	\centering
-	\includegraphics[width=\textwidth]{figures/NorthAndros.png}
-	\caption{Comparison of dewlap coloration across habitats on North Andros. Legend is as per Figure \ref{fig:Abaco}.}
+	\includegraphics[width=\textwidth]{figures/NorthAndros_supplement.png}
+	\caption{Comparison of dewlap coloration across habitats on North Andros. Legend is as per Figure \ref{fig:Abaco_supplement}, but without panel G.}
 	\label{fig:NorthAndros}
@@
 	\centering
-	\includegraphics[width=\textwidth]{figures/SouthAndros.png}
-	\caption{Comparison of dewlap coloration across habitats on South Andros. Legend is as per Figure \ref{fig:Abaco}.}
+	\includegraphics[width=\textwidth]{figures/RaggedIsland_supplement.png}
+	\caption{Comparison of dewlap coloration across habitats on Ragged Island. Legend is as per Figure \ref{fig:Abaco_supplement}, but without panel G.}
+	\label{fig:RaggedIsland}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{figures/SouthAndros_supplement.png}
+	\caption{Comparison of dewlap coloration across habitats on South Andros. Legend is as per Figure \ref{fig:Abaco_supplement}.}
 	\label{fig:SouthAndros}

### 2021-05-29 ms/sections/figures.tex
added Colin's figure

@@
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\textwidth]{figures/Figure1_CMD-01_cropped.png}
+	\caption{Overview of our study design, including a map of the Bahamas and the Cayman Islands, on which are indicated the nine islands we sampled, two representatives of our study species \textit{Anolis sagrei} with their dewlaps deployed, and the three types of habitats we considered on each island.}
+	\label{fig:overview}
+\end{figure}
+
 \begin{figure}[H]

### 2021-06-10 ms/sections/methods2.tex
ms 3.1

@@
 
-We measured reflectance between 300 and 700nm wavelength, a range that encompasses the colors visible to most lizards and vertebrates in general \citep{Lazareva2012}.
+We measured reflectance between 300nm and 700nm wavelength, a range from ultraviolet to red that encompasses the colors visible to most lizards and vertebrates in general \citep{Lazareva2012}.
 Measurements were taken with an Ocean Optics USB4000 spectrometer, a pulsed Xenon light source (PX-2, Ocean Optics, Largo, FL, USA) and a reflectance probe protected by a black anodized aluminum sheath.

### 2021-06-10 ms/sections/results2.tex
ms 3.1

@@
 Our within-island random forest classification analyses revealed detectable differences in dewlap coloration on eight out of the nine islands in our sample: Abaco, Bimini, Cayman Brac, Eleuthera, Little Cayman, Long Island, North Andros and South Andros.
-The accuracy of Random forest classification exceeded random expectation more often than expected by chance for all these islands (Table \ref{tab:randomforests}).
+The accuracy of random forest classification exceeded random expectation more often than expected by chance for all these islands (Table \ref{tab:randomforests}).
 Accuracy was as high as 73\% for Cayman Brac.
@@
 
-The accuracy of Random forest classification exceeded random expectation more often than expected by chance for all islands. Accuracy was as high as 73% for Cayman Brac.
-
 On Abaco, dewlaps from the mangrove habitat were the best discriminated, while dewlaps from the beach scrub habitat were often mistaken for dewlaps from the coppice habitat (Fig.
-\ref{fig:Abaco}A).
-Importance analysis revealed that beach scrub and mangrove lizards mostly differed in reflectance in the UV-end of the spectrum (below 400nm, Fig.
-\ref{fig:Abaco}B), where mangrove dewlaps had higher UV reflectance relative to beach scrub lizards, and coppice lizards had an intermediate UV reflectance between the two other habitats (Fig.
+\ref{fig:Abaco}D).
+Importance analysis revealed that beach scrub and mangrove lizards mostly differed in reflectance in the ultraviolet (UV) end of the spectrum (below 400nm, Fig.
+\ref{fig:Abaco_supplement}F), where mangrove dewlaps had higher UV reflectance relative to beach scrub lizards, and coppice lizards had an intermediate UV reflectance between the two other habitats (Fig.
+\ref{fig:Abaco}B).
+Consistent with this, our analyses of variance detected significantly lower PC2 scores in mangrove lizards than in the two other habitats (Fig.
+\ref{fig:Abaco}E, Table \ref{tab:anova}), representing a higher UV-reflectance relative to red (Fig.
 \ref{fig:Abaco}C).
-Consistent with this, our analyses of variance detected significantly lower PC2 scores in mangrove lizards than in the two other habitats, representing a higher UV-reflectance relative to red (Fig.
-\ref{fig:Abaco}D, E, Table \ref{tab:anova}).
-Beach scrub lizards also scored lower on PC3, indicating less curvature of the reflectance profile and relatively higher reflectance at intermediate wavelengths (blue-to-yellow) than at the ends of the range (Fig.
-\ref{fig:Abaco}D, E).
+Beach scrub lizards also scored lower on PC3 (Fig.
+\ref{fig:Abaco}E, Table \ref{tab:anova}), indicating less curvature of the reflectance profile and relatively higher reflectance at intermediate wavelengths (blue-to-yellow) than at the ends of the range (Fig.
+\ref{fig:Abaco}C).
 Differences were detected between sites both at large ($\sim$ 100km) and short ($<$ 1km) distances (Fig.
-\ref{fig:Abaco}F, G).\\
+\ref{fig:Abaco_supplement}G).\\
 
 On Bimini, the random forests mostly correctly classified lizards from the coppice and mangrove habitats while often misclassifying lizards from the beach scrub habitat (Fig.
-\ref{fig:Bimini}A).
+\ref{fig:Bimini}D).
 Relatively flat importance profiles for beach scrub lizards suggested that brightness was used instead of a particular wavelength to identify some of the beach scrub dewlaps (Fig.
-\ref{fig:Bimini}B).
+\ref{fig:Bimini}F).
 Indeed, some beach scrub dewlaps were substantially brighter than the rest (Fig.
-\ref{fig:Bimini}C), a pattern that was captured by our analysis of variance along PC1 (i.e.
+\ref{fig:Bimini}B, C), a pattern that was captured by our analysis of variance along PC1 (i.e.
 brightness, Fig.
-\ref{fig:Bimini}D, E, Table \ref{tab:anova}).
+\ref{fig:Bimini}C, E, Table \ref{tab:anova}).
 The random forests also used UV reflectance to discriminate between coppice and mangrove dewlaps (Fig.
-\ref{fig:Bimini}B), which could reflect the significant difference we detected along PC3 between these two habitats (Fig.
-\ref{fig:Bimini}D, Table \ref{tab:anova}).
+\ref{fig:Bimini}F), which could reflect the significant difference we detected along PC3 between these two habitats (Fig.
+\ref{fig:Bimini}C, E, Table \ref{tab:anova}).
 Beach scrub lizards were characterized by elevated red reflectance relative to UV (as represented by PC2, (Fig.
-\ref{fig:Bimini}D, E)), and beach scrub and mangrove lizards were characterized by a more even distribution of the reflectance along the spectrum (as represented by PC3, (Fig.
-\ref{fig:Bimini}D, E)), in contrast to coppice lizards which harbored a stronger curvature at intermediate wavelengths (Fig.
-\ref{fig:Bimini}D, E).
+\ref{fig:Bimini}C, E)), and beach scrub and mangrove lizards were characterized by a more even distribution of the reflectance along the spectrum (as represented by PC3, (Fig.
+\ref{fig:Bimini}C, E)), in contrast to coppice lizards which harbored a stronger curvature at intermediate wavelengths (Fig.
+\ref{fig:Bimini}C, E).
 On this island, the beach scrub and coppice habitats were separated by a few hundred meters, making this contrast the smallest geographical scale at which differences in coloration were found in our study (Fig.
-\ref{fig:Bimini}F, G).\\
+\ref{fig:Bimini}G).\\
 
 On Cayman Brac, all three habitats could be well discriminated against each other (Fig.
-\ref{fig:CaymanBrac}A), with UV reflectance appearing to be an important variable differentiating beach scrub and mangrove dewlaps (Fig.
-\ref{fig:CaymanBrac}B).
+\ref{fig:CaymanBrac}D), with UV reflectance appearing to be an important variable differentiating beach scrub and mangrove dewlaps (Fig.
+\ref{fig:CaymanBrac}F).
 In contrast, coppice dewlaps had a relatively flat importance profile, suggesting that brightness made them more distinct rather than any particular wavelength (Fig.
-\ref{fig:CaymanBrac}B).
+\ref{fig:CaymanBrac}F).
 Consistent with this, coppice dewlaps were significantly different from all other dewlaps along PC1 (Fig.
-\ref{fig:CaymanBrac}D, Table \ref{tab:anova}).
+\ref{fig:CaymanBrac}E, Table \ref{tab:anova}).
 At a distance between 2 and 3km (Fig.
-\ref{fig:CaymanBrac}F, G), dewlaps in the beach scrub habitat reflected more red light (as represented by PC2, Fig.
-\ref{fig:CaymanBrac}D, E) and more UV (as represented by PC3) than in the mangrove habitat.
+\ref{fig:CaymanBrac}G), dewlaps in the beach scrub habitat reflected more red light (as represented by PC2, Fig.
+\ref{fig:CaymanBrac}C, E) and more UV (as represented by PC3, Fig.
+\ref{fig:CaymanBrac}C, E) than in the mangrove habitat.
 Coppice lizards were also characterized by a higher UV reflectance than mangrove dewlaps (PC3, Fig.
-\ref{fig:CaymanBrac}D, E, Table \ref{tab:anova}).\\
+\ref{fig:CaymanBrac}C, E, Table \ref{tab:anova}).\\
 
@@
 Consistent with this, the only significant univariate difference detected was for PC2 between beach scrub and mangrove lizards, where beach scrub lizards had higher levels of red reflectance and mangrove lizards higher levels of UV reflectance (Fig.
-\ref{fig:Eleuthera}D, E, Table \ref{tab:anova}).\\
+\ref{fig:Eleuthera}C, E, Table \ref{tab:anova}).\\
 
 Little Cayman was characterized by a better discrimination of mangrove lizards from the rest than between beach scrub and coppice lizards (Fig.
-\ref{fig:LittleCayman}A).
+\ref{fig:LittleCayman}D).
 Mangrove dewlaps were most distinct with respect to their reflectance in short wavelengths (Fig.
-\ref{fig:LittleCayman}B), with significantly lower UV reflectance (as represented by PC2, Fig.
-\ref{fig:LittleCayman}D, E, Table \ref{tab:anova}).
+\ref{fig:LittleCayman}F), with significantly lower UV reflectance (as represented by PC2, Fig.
+\ref{fig:LittleCayman}C, E, Table \ref{tab:anova}).
 Beach scrub lizards were characterized by brighter dewlaps than coppice lizards (PC1), and also more convex curves, i.e.
 slightly higher UV and red reflectance (as represented by higher PC3 scores), than lizards from the other two habitats (Fig.
-\ref{fig:LittleCayman}D, E, Table \ref{tab:anova}).\\
+\ref{fig:LittleCayman}C, E, Table \ref{tab:anova}).\\
 
 On Long Island the three habitats were relatively well discriminated (Fig.
-\ref{fig:LongIsland}A).
+\ref{fig:LongIsland}D).
 Importance profiles indicated that short wavelengths were used to discriminate between beach scrub and mangrove lizards (Fig.
-\ref{fig:LongIsland}B).
+\ref{fig:LongIsland}F).
 Beach scrub lizards had more curved reflectance profiles than in either of the two other habitats, with higher levels of UV and red reflectance relative to intermediate wavelengths (PC3, Fig.
-\ref{fig:LongIsland}D, E, Table \ref{tab:anova}).
+\ref{fig:LongIsland}C, E, Table \ref{tab:anova}).
 Beach scrub lizards also differed from mangrove lizards along PC4 (Fig.
-\ref{fig:LongIsland}D), which represented a rather small portion of the variance not already explained by the first three principal components, and is therefore difficult to interpret (Fig.
-\ref{fig:LongIsland}E).
+\ref{fig:LongIsland}E), which represented a rather small portion of the variance not already explained by the first three principal components, and is therefore difficult to interpret (Fig.
+\ref{fig:LongIsland}C).
 Coppice lizards were significantly darker than mangrove and beach scrub lizards (PC1, Fig.
-\ref{fig:LongIsland}D, E, Table \ref{tab:anova}).\\
+\ref{fig:LongIsland}C, E, Table \ref{tab:anova}).\\
 
 On North Andros, although the random forest classification was significant ($P = 0.0216$, Table \ref{tab:randomforests}) and the average confusion matrix indicated that lizards from beach scrub were particularly well predicted (Fig.
-\ref{fig:NorthAndros}A), no significant univariate differences were detected along any of the four PCs (Fig.
-\ref{fig:NorthAndros}D, Table \ref{tab:anova}).
+\ref{fig:NorthAndros}D), no significant univariate differences were detected along any of the four PCs (Fig.
+\ref{fig:NorthAndros}E, Table \ref{tab:anova}).
 Importance analysis of full-spectrum random forests showed higher importance scores near the UV-end of the spectrum in discriminating beach scrub dewlaps from the rest (Fig.
-\ref{fig:NorthAndros}B).
+\ref{fig:NorthAndros}F).
 Besides, reflectance curves of beach scrub dewlaps appeared more similar to each other in the UV range than dewlaps from other habitats (Fig.
-\ref{fig:NorthAndros}C), suggesting that the machines may have used this low within-habitat variance, as opposed to between-habitat differences in means, to correctly classify beach scrub lizards.
+\ref{fig:NorthAndros}B), suggesting that the machines may have used this low within-habitat variance, as opposed to between-habitat differences in means, to correctly classify beach scrub lizards.
 A small sample size on this islands may also have contributed to a lack of power in detecting univariate differences using analyses of variance (Table \ref{tab:counts}).\\
@@
 On South Andros beach scrub and coppice dewlaps could be discriminated better against each other than with mangrove dewlaps (Fig.
-\ref{fig:SouthAndros}A), with importance profiles supporting UV-reflectance as a predictor of coppice lizards (Fig.
-\ref{fig:SouthAndros}B).
+\ref{fig:SouthAndros}D), with importance profiles supporting UV-reflectance as a predictor of coppice lizards (Fig.
+\ref{fig:SouthAndros}F).
 Coppice lizards had more curved reflectance profiles than beach scrub lizards (PC3), and lizards from both habitats differed along PC4, which is again more difficult to interpret (Fig.
-\ref{fig:SouthAndros}D, E, Table \ref{tab:anova}).
+\ref{fig:SouthAndros}C, E, Table \ref{tab:anova}).
 Beach scrub lizards also differed from mangrove lizards in PC4 (Fig.
-\ref{fig:SouthAndros}D, E, Table \ref{tab:anova}).\\
+\ref{fig:SouthAndros}C, E, Table \ref{tab:anova}).\\
 

### 2021-06-10 ms/run_pandoc.sh
mail setn to jonathan

@@ -1,5 +1,5 @@
 # Run this script to convert LaTeX files into ODT and DOCX documents
 # Requires pandoc and pandoc-crossref being installed
 
-pandoc main.tex --filter=pandoc-crossref --bibliography=library.bib --csl=journal-of-evolutionary-biology.csl -o scherrer2020.odt
-pandoc main.tex --filter=pandoc-crossref --bibliography=library.bib --csl=journal-of-evolutionary-biology.csl -o scherrer2020.docx
+pandoc main.tex --filter=pandoc-crossref --bibliography=library.bib --csl=journal-of-evolutionary-biology.csl -o scherrer2021_v3.1.odt
+pandoc main.tex --filter=pandoc-crossref --bibliography=library.bib --csl=journal-of-evolutionary-biology.csl -o scherrer2021_v3.1.docx

### 2021-07-11 ms/sections/abstract.tex
ms 3.1
first rofirst round of resolving JBL's comments

@@
 Previous studies have demonstrated that dewlap color in \textit{Anolis} lizards can be highly variable between populations in relation to both biotic and abiotic adaptive drivers at relatively large geographical scales.
-Here, we investigated differentiation of dewlap coloration among habitat-types at a small spatial scale, within multiple islands of the West Indies, as this may give new insights into the local scale at which adaptation is possible.
-We explored variation in dewlap coloration in the most widespread species of anole, \textit{Anolis sagrei}, across three characteristic habitats spanning the Bahamas and the Cayman Islands.
+Here, we investigated differentiation of dewlap coloration among habitat-types at a small spatial scale, within multiple islands of the West Indies, to test the hypothesis that similar local adaptive processes occur over smaller spatial scales.
+We explored variation in dewlap coloration in the most widespread species of anole, \textit{Anolis sagrei}, across three characteristic habitats spanning the Bahamas and the Cayman Islands, namely beach scrub, primary coppice forest and mangrove forest.
 Using reflectance spectrometry paired with supervised machine learning, we found significant differences in spectral properties of the dewlap between habitats within small islands, sometimes over very short distances.
@@
 On the other hand, these habitat-specific dewlap differences varied in magnitude and direction across islands, and thus our primary test for adaptation -- parallel responses across islands -- was not supported.
-We suggest that selection could be involved in several ways, including sexual selection.
+We suggest that neutral processes or selection could be involved in several ways, including sexual selection.
 Our results shed new light on the scale at which signal color polymorphism can be maintained in the presence of gene flow, and the relative role of local adaptation and other processes in driving these patterns.

### 2021-07-11 ms/sections/discussion2.tex
first rofirst round of resolving JBL's comments

@@
 Two main insights follow from our results.
-First, excluding North Andros where the follow-up univariate analyses were not significant, we detected significant differences in dewlap coloration between habitats within seven out of the nine islands investigated, suggesting a putatively high potential for local differentiation of dewlap coloration in \textit{Anolis sagrei}.
+First, we detected significant differences in dewlap coloration between habitats within seven out of the nine islands investigated (excluding North Andros where the follow-up univariate analyses were not significant), suggesting a putatively high potential for local differentiation of dewlap coloration in \textit{Anolis sagrei}.
 Second, we found differences in coloration along different dimensions of color space, and in different directions, on different islands.\\
 
-Detectable differences in dewlap color between habitat-populations are surprising, as habitats were often in close geographical proximity to each other (as close as a few hundred meters on Bimini and most of the time within ten kilometers).
-Indeed, given that (1) the populations were continuously distributed between the habitats, (2) different habitat-populations were not monophyletic with respect to mitochondrial haplotypes (\citealt{vandeSchoot2016} unpubl.), and (3) \textit{A.
-sagrei} have been shown to be a highly mobile species \citep{Kamath2018}, we would have expected more homogeneous distributions of color phenotypes within islands due to extensive gene flow, with fewer differences between populations, especially those in close proximity.\\
+Detectable differences in dewlap color between populations are surprising, as habitats were often in close geographical proximity to each other (as close as a few hundred meters on Bimini and most of the time within ten kilometers).
+Indeed, given that (1) the populations were continuously distributed between the habitats, (2) populations from different habitats were not monophyletic with respect to mitochondrial haplotypes (\citealt{vandeSchoot2016} unpubl.), and (3) \textit{A.
+sagrei} has been shown to be a highly mobile species \citep{Kamath2018}, we would have expected more homogeneous distributions of color phenotypes within islands due to extensive gene flow, with fewer differences between populations, especially those in close proximity.\\
 
 Several scenarios could account for these findings.
-One explanation is an adaptive one.
-Indeed, populations living in different habitats could be phenotypically adapted to their local environmental conditions.
+One explanation is an adaptive one: populations living in different habitats could be phenotypically adapted to their local environmental conditions.
 Given that the brightly colored dewlap of \textit{A.
@@
 
-If our results were an example of sensory drive, we would have expected to see consistent differences between habitat-populations across islands (a pattern that would have been a compelling indicator of adaptation at all, \citealt{Losos2011}).
+If our results were an example of sensory drive, we would have expected to see consistent differences between habitat-populations across islands (a pattern that would have been a compelling indicator of adaptation, \citealt{Losos2011}).
 This is because environmental conditions that may be relevant to color signal detectability such as light, temperature, moisture and vegetation, are consistent within the three main and clearly distinct habitat-types found across the sampled islands, i.e.
@@
 
-Phenotypic plasticity could be another cause for dewlap color variation between habitats, where different conditions would favor different phenotypes in different habitats, without genetic changes.
-Indeed, the yellow, orange and red colors in anoline dewlaps are produced by pterins and carotenoids \citep{Ortiz1962, Ortiz1962a, Ortiz1963, Ortiz1966, Macedonia2000, Steffen2007, Steffen2009}.
+Phenotypic plasticity could be another cause for dewlap color variation between habitats, where different environmental conditions produce different phenotypes in different habitats.
+The yellow, orange and red colors in anoline dewlaps are produced by pterins and carotenoids \citep{Ortiz1962, Ortiz1962a, Ortiz1963, Ortiz1966, Macedonia2000, Steffen2007, Steffen2009}.
 Animals can synthesize pterins from nucleotides, but lack the ability to synthesize carotenoids \citep{Goodwin1984, Hill2002, Hill2006}.
 Different food qualities across sites within islands could therefore potentially cause detectable differences in coloration.
-Alternatively, more subtle effects on dewlap color could arise from developmental plasticity and depend, e.g.
-on differences in egg-rearing conditions.
-However, more data are needed to test these hypotheses, and although some work has shown plastic responses of dewlap color in response to parasites in \textit{A.
-sagrei} \citep{Cook2013}, we find it unlikely to account for the widespread habitat differences we found.
+Alternatively, more subtle effects on dewlap color could arise from developmental plasticity and depend, e.g., on differences in egg-rearing conditions.
+However, more data are needed to test these hypotheses, and although \citet{Cook2013} have shown plastic responses of dewlap color in response to parasites, we find it unlikely to account for the widespread habitat differences we found.
 Besides, studies testing the effect of carotenoid deprivation \citep{Steffen2010, Ng2013} and heritability \citep{Cox2017} of dewlap coloration in \textit{A.
@@
 That said, there were often few sampling sites per island in our study, whose locations were not uniformly chosen within the islands, and so the true extent of isolation-by-distance may be difficult to test.
-Other, less trivial forms of drift may be at play than isolation-by-distance, but nevertheless, we did find significant differences in color phenotype at relatively small spatial scales, sometimes in neighboring habitats, on islands where gene flow is probably highly pervasive, as suggested by high rates of encounter between males and females \citep{Kamath2018}, making the divergence of habitat-populations by drift in relative genetic isolation an unlikely scenario.\\
+Other forms of drift may be at play than isolation-by-distance, but nevertheless, we did find significant differences in color phenotype at relatively small spatial scales, sometimes in neighboring habitats, on islands where gene flow is probably highly pervasive, as suggested by high rates of encounter between males and females \citep{Kamath2018}, making the divergence of habitat-populations by drift in relative genetic isolation an unlikely scenario.\\
 
@@
 First, there could be sexual selection for different dewlap colors in different locations.
-Indeed, although the sensory-drive hypothesis may include a sexual selection aspect, e.g.
-if the optimal male phenotype in a given habitat is a function of female perception, sexual preferences may also be arbitrary and independent of the habitat \citep{Andersson1994}, and so could differ across habitats and islands.
+Indeed, although the sensory-drive hypothesis may include a sexual selection aspect, e.g., if the optimal male phenotype in a given habitat is a function of female perception, sexual preferences may also be arbitrary and independent of the habitat \citep{Andersson1994}, and so could differ across habitats and islands.
 Although one previous study has found no link between dewlap coloration and body size dimorphism in \textit{A.
@@
 Alternatively, selective pressures may be different in similar habitats from one island to another, because of other environmental variables not accounted for by the habitat-type classification we used \citep{Howard1950, Schoener1968}.
-The islands we sampled indeed exhibit variation in some climatic variables but also in densities of predators and anole congeners, which have all been shown to correlate with variation in \textit{A.
+The islands we sampled exhibit variation in some climatic variables, but also in densities of predators and anole congeners, which have all been shown to correlate with variation in \textit{A.
 sagrei} mean dewlap coloration among islands \citep{Vanhooydonck2009, Baeckens2018}.\\
@@
 Several factors could account for this.
-For example, the founder populations of each island, which we know colonized the islands independently (van de Schoot, unpublished thesis; \citealt{Driessens2017, Reynolds2020}), could have exhibited different dewlap colors at the time of colonization, as may be suggested by the larger differences we observed between than within islands.
+For example, the founder populations of each island, which colonized the islands independently (van de Schoot, unpublished thesis; \citealt{Driessens2017, Reynolds2020}), could have exhibited different dewlap colors at the time of colonization, as may be suggested by the larger differences we observed between than within islands.
 In turn, different initial phenotypes could have led to different ways in which populations would have diverged between habitats.
@@
 Altogether, our results show that dewlap color of \textit{A.
-sagrei} commonly varies between habitat-types, even in close geographical proximity, within islands of the West Indies, and that coloration differs in different ways from one island to another.
-We discussed several non-mutually exclusive mechanisms that could explain these observations, but more data are needed to thoroughly test each of these.
-Nevertheless, heterogeneous patterns of divergence across islands do not support an adaptive sensory-drive scenario, and our we propose that within-island dewlap color variation may be underlain by a more subtle mosaic of factors.
+sagrei} commonly varies between habitat types, even in close geographical proximity, within islands of the West Indies, and that coloration differs in different ways from one island to another.
+We discussed several non-mutually exclusive mechanisms that could explain these observations, but more data are needed to thoroughly test these hypotheses.
+Nevertheless, heterogeneous patterns of divergence across islands do not support an adaptive sensory-drive scenario, and we propose that within-island dewlap color variation may be underlain by a more subtle mosaic of factors.
 

### 2021-07-11 ms/sections/introduction.tex
first rofirst round of resolving JBL's comments

@@
 Over the past decades, scientists have established that signals evolve in an ecological context and are dependent on environmental conditions \citep{Endler1992,Endler1993,Endler1993a}.
-Just as different habitats may favor different combinations of eco-morphological traits to maximize performance and fitness \citep{Arnold1983}, they may also shape different forms of a signal, so as to maximize its transmission and detection (e.g.
+Just as different habitats may favor different combinations of ecomorphological traits to maximize performance and fitness \citep{Arnold1983}, they may also shape different forms of a signal, so as to maximize its transmission and detection (e.g.
 \citealt{Seehausen1997}), or reduce its detection by unintended recipients such as predators \citep{Endler1984,Endler1990,Endler1991,Halfwerk2014}.
@@
 Population genetics theory suggests that gene flow may counteract local adaptation between localities and prevent divergence altogether, especially at small spatial scales, because of the inflow of maladapted alleles or because of the breaking of linkage between coevolving loci \citep{Felsenstein1976, Garcia-Ramos1997, Dieckmann1999, Lenormand2002, Hendry2007}.
-This genetic homogenization has been confirmed empirically in systems such as stick-insects \citep{Nosil2004} and stickleback \citep{Hendry2007a}.
+This genetic homogenization has been confirmed empirically in systems such as stick insects \citep{Nosil2004} and stickleback \citep{Hendry2007a}.
 Yet, examples of microgeographic adaptation, i.e.
 adaptation at smaller scales than the range of dispersal, exist, highlighting the potential of some organisms to respond to selection in the face of gene flow (see \citealt{Richardson2014} and references therein).
-Examples include small scale adaptation in fragmented areas in Australian fruit flies \citep{Willi2012}, or local adaptation to predation pressure in North American salamanders \citep{Richardson2013}.
-Therefore, despite evidence that local adaptation may be particularly difficult at small spatial scales where gene flow tends to cause adjoining populations to remain genetically homogeneous, the potential adaptive response of species traits, in particular communication signals, to localized differences in habitats remains relatively unknown \citep{Richardson2014}.\\
-
-% Anolis lizards and their dewlap
-
+Examples include small scale adaptation in fragmented areas in Australian fruit flies \citep{Willi2012}, and local adaptation to predation pressure in North American salamanders \citep{Richardson2013}.
+Therefore, despite evidence that local adaptation may be particularly difficult at small spatial scales where gene flow tends to cause adjoining populations to remain genetically homogeneous, the potential adaptive response of species traits, in particular communication signals, to localized differences in habitats remains relatively unknown \citep{Richardson2014}.
 Lizards of the neotropical genus \textit{Anolis} are an excellent group for studying the eco-evolutionary dynamics of local adaptation and natural selection \citep{Losos2009}.
-A particularly conspicuous trait of anoles is their dewlap; an extensible flap of skin that is typically sexually dimorphic and used as a communication signal in courtship \citep{Sigmund1983, Driessens2014, Driessens2015} and territorial displays \citep{Losos1985, Macedonia1994, Macedonia2013} as well as in predator deterrence \citep{Leal1995, Leal1997, Leal1997a}.
+A particularly conspicuous trait of anoles is their dewlap, an extensible flap of skin that is typically sexually dimorphic and used as a communication signal in courtship \citep{Sigmund1983, Driessens2014, Driessens2015} and territorial displays \citep{Losos1985, Macedonia1994, Macedonia2013} as well as in predator deterrence \citep{Leal1995, Leal1997, Leal1997a}.
 Dewlap characteristics vary widely among the approximately $400$ species of the genus \citep{Nicholson2007}.
-Interspecific variation in dewlap coloration is implicated in species recognition \citep{Rand1970, Williams1969, Williams1977, Losos1985, Macedonia1994, Fleishman2000, Macedonia2013}, and this function could have had a role in initiating and/or reinforcing reproductive isolation during speciation \citep{Lambert2013, Geneva2015, Ng2017}.\\
+Interspecific variation in dewlap coloration is implicated in species recognition \citep{Rand1970, Williams1969, Williams1977, Losos1985, Macedonia1994, Fleishman2000, Macedonia2013}, and this function could have had a role in initiating or reinforcing reproductive isolation during speciation \citep{Lambert2013, Geneva2015, Ng2017}.\\
 
 Within species, studies have shown a link between variation in dewlap coloration and differences in habitats or climatic conditions \citep{Macedonia2001, Leal2002, Thorpe2002, Thorpe2002a, Leal2004, Vanhooydonck2009, Ng2012, Ng2013, Ng2016, Vanhooydonck2009, Driessens2017}.
-Some studies suggest that those differences may be adaptive, and that dewlaps may have evolved to maximize detectability given local light conditions \citep{Fleishman2001, Leal2002, Leal2004}.
+Some studies suggest that those differences may be adaptive and that dewlaps may have evolved to maximize detectability given local light conditions \citep{Fleishman2001, Leal2002, Leal2004}.
 Although this claim is further supported by recent findings that dewlap colors are perceived differently under different levels of shading \citep{Fleishman2020}, other studies found conflicting patterns of between-habitat variation that did not support the sensory drive hypothesis \citep{Fleishman2009, Ng2012, Macedonia2014}.\\
@@
 The species \textit{Anolis sagrei} is widespread across islands of the West Indies \citep{Reynolds2020}.
-It has been the subject of numerous studies concerning local adaptation \citep{Losos1994, Losos1997a, Losos2001, Kolbe2012}, biological invasion \citep{Kolbe2008}, sexual selection \citep{Tokarz2002, Tokarz2005, Tokarz2006, Driessens2014, Steffen2014, Driessens2015} among many other topics.
+It has been the subject of numerous studies concerning local adaptation \citep{Losos1994, Losos1997a, Losos2001, Kolbe2012}, biological invasion \citep{Kolbe2008}, and sexual selection \citep{Tokarz2002, Tokarz2005, Tokarz2006, Driessens2014, Steffen2014, Driessens2015} among many other topics.
 Between-island variation in the mainly orange-red color of its dewlap was shown to be better explained by climatic variables \citep{Driessens2017} than by proxies for biotic factors such as sexual selection or predation pressure \citep{Vanhooydonck2009, Baeckens2018}.
@@
 Our sampling design included sites in close proximity; the median distance between two sites within an island was $11.2$km.
-Combining reflectance spectrometry and supervised machine learning, we tested for divergence in dewlap phenotype between habitats within islands and between islands across the range of \textit{A.
+Combining reflectance spectrometry and supervised machine learning, we tested for divergence in dewlap phenotype between habitats within islands and between islands across part of the range of \textit{A.
 sagrei}.

### 2021-07-11 ms/sections/methods2.tex
first rofirst round of resolving JBL's comments

@@
 
-We sampled 466 male \textit{Anolis sagrei} from seven islands in the Bahamas Archipelago -- Abaco, North Andros, South Andros, South Bimini, Eleuthera, Long Island, Ragged Island -- and two in the Cayman Islands -- Cayman Brac and Little Cayman (Figure \ref{fig:maps}A).
+We sampled 466 male \textit{A.
+sagrei} from seven islands in the Bahamas Archipelago -- Abaco, North Andros, South Andros, South Bimini, Eleuthera, Long Island, and Ragged Island -- and two in the Cayman Islands -- Cayman Brac and Little Cayman (Figure \ref{fig:maps}A).
 These islands were chosen to span the breadth of the West Indian range of \textit{A.
@@
 Our sampling design enabled us to test for differences between habitats at a coarse and fine geographical scale.
-The median distance between two localities within an island was $\sim 11$km, with some islands being sampled at smaller or larger scales (Figure \ref{fig:maps}B), and $80.3$\% of all pairwise distances within islands were less than $50$km.
+The median distance between two localities within an island was $\sim 11$km (Figure \ref{fig:maps}B), and $80.3$\% of all pairwise distances within islands were less than $50$km.
 Additionally, there are no major barriers to dispersal (such as mountains or grassland) on any of the islands that we sampled.
@@
 Third, for each island where multivariate differences were detected using our machine learning pipeline, we used univariate analyses of variance to decompose the signal among the different dimensions of color space.
-Fourth, for each significant between-habitat variation found in univariate analyses, we used posthoc tests to determine which habitats were involved.
-Last, to get insights into the spatial scale of phenotypic variation, for each significant contrast between two habitats detected along a given dimension on a given island we performed multiple pairwise Wilcoxon tests to assess differences in dewlap coloration among the sites involved in that significant contrast, and recorded the geographical distance between sites that were found significant.
-In parallel, we tested a possible effect of isolation-by-distance, an alternative cause of phenotypic divergence between localities based on diffusion approximation and dispersal distance, irrespective of habitat-types.
+Fourth, for each significant between-habitat variation found in univariate analyses, we used post-hoc tests to determine which habitats were responsible for the differences.
+Last, to get insights into the spatial scale of phenotypic variation, for each significant contrast between two habitats detected along a given dimension on a given island, we performed multiple pairwise Wilcoxon tests to assess differences in dewlap coloration among the sites involved in that significant contrast, and recorded the geographical distance between sites that were found significant.
+In parallel, we tested a possible effect of isolation-by-distance, an alternative cause of phenotypic divergence between localities, based on diffusion approximation and dispersal distance, irrespective of habitat-types.
 We did so using a permutation test to assess the significance of the correlation between geographical distances and phenotypic distances among sites within each island.\\
@@
 Here, we used a k-fold cross-validation procedure with $k = 5$, where each training set consisted of 80\% of the data and the machine was tested on the remaining 20\%.
-Each training set was conditioned on containing at least 5 lizards from each of the three habitats.
+Each training set was conditioned on containing at least five lizards from each of the three habitats.
 We also down-sampled the training set to the sample size of the least represented habitat, to ensure that the different habitats were equally represented.
@@
 
-We used the machine learning fitting functions in the R package rminer \citep{Cortez2020}, which calls random forest routines from the randomForest package (\citealt{Liaw2002}, implementation from the original random forest algorithm, \citealp{Breiman2001}).
+We used the machine learning fitting functions in the R package rminer \citep{Cortez2020}, which calls random forest routines from the randomForest package (\citealt{Liaw2002}, implementation from the original random forest algorithm \citealp{Breiman2001}).
 For each random forest, we optimized the number of trees in the forest and the number of variables examined by each tree using the grid hyperparameter search procedure implemented in rminer, to choose between two numbers of trees (500 or 1,000) and four numbers of principal components examined per tree (1 to 4), using rminer's ordered holdout validation method with $2/3$ of the data used for training.\\
@@
 To know which habitat-populations were different from which in dewlap coloration, we performed different post-hoc multiple comparison tests (all implemented in the PMCMRplus package, \citealp{Pohlert2020}), depending on which assumptions were met.
-In cases where normality and homoskedasticity were met (i.e.
+In cases where normality and homoscedasticity were met (i.e.
 OLS-ANOVA was the best fit), we used Tukey's honest significant difference test.
-When normality was met but not homoskedasticity (i.e.
+When normality was met but not homoscedasticity (i.e.
 GLS-ANOVA was the best fit), we used Dunnett's T3 test.

### 2021-07-26 ms/sections/appendix.tex
v3.2

--- 2021-06-10 ms/sections/results2.tex
@@
-% Little but detectable differences on five islands
-
-We tested for variation in \textit{A.
-sagrei} dewlap coloration between populations living in three characteristic habitat-types across nine islands that span the West Indian range of the brown anole (beach scrub, primary coppice and mangroves).
-We found that most of the variation in coloration was partitioned between islands (two-way PERMANOVA, $F(df = 8) = 43.7$, $P = 0.0001$, explained variance $R^2 = 40.9$\%).
-Nonetheless, we did find evidence for differences in dewlap coloration between habitat-types, and those were mostly island-specific (habitat-by-island interaction term, $F(16) = 3.53$, $P = 0.0001$, $R^2 = 6.6$\%), with a significant portion of the variation explained by an habitat effect across all islands, but this effect was relatively small ($F(2) = 4.7$, $P = 0.0001$, $R^2 = 1.1$\%).\\
-
-We subsequently tested for differences in dewlap coloration between habitat-populations within each island, using within-island principal component scores (to maximize the variation captured for each island, see Methods).
-Our within-island random forest classification analyses revealed detectable differences in dewlap coloration on eight out of the nine islands in our sample: Abaco, Bimini, Cayman Brac, Eleuthera, Little Cayman, Long Island, North Andros and South Andros.
-The accuracy of random forest classification exceeded random expectation more often than expected by chance for all these islands (Table \ref{tab:randomforests}).
-Accuracy was as high as 73\% for Cayman Brac.
-We obtained similar results using other machine learning approaches such as support vector machines (Table \ref{tab:ksvms}) and linear discriminant analysis (Table \ref{tab:ldas}), except that these methods did not detect significant differences on Eleuthera and North Andros.
-We did not find evidence of spatial autocorrelation in dewlap coloration between the sites within islands, except for Eleuthera (Table \ref{tab:autocorrelation}).
-We now describe the specific differences detected on each island.\\
+Here we describe more precisely the patterns identified on each island.\\
 
@@
 Differences were detected between sites both at large ($\sim$ 100km) and short ($<$ 1km) distances (Fig.
-\ref{fig:Abaco_supplement}G).\\
+\ref{fig:Abaco_supplement}G).
+Abaco was the only island where we detected significant spatial autocorrelation (Table \ref{tab:autocorrelation}), that is, sites that were closer geographically tended to have populations of lizards with more similar dewlap colors.\\
 
@@
 
-Eleuthera was the only island where we detected significant spatial autocorrelation (Table \ref{tab:autocorrelation}), that is, sites that were closer geographically tended to have populations of lizards with more similar dewlap colors.
-Although random forests detected between-habitat differences in dewlap color, other approaches did not (Tables \ref{tab:ldas} and \ref{tab:ksvms}), suggesting that the differences may be small.
+On Eleuthera, although random forests detected between-habitat differences in dewlap color, other approaches did not (Tables \ref{tab:ldas} and \ref{tab:ksvms}), suggesting that the differences may be small.
 Consistent with this, the only significant univariate difference detected was for PC2 between beach scrub and mangrove lizards, where beach scrub lizards had higher levels of red reflectance and mangrove lizards higher levels of UV reflectance (Fig.

### 2021-07-26 ms/sections/discussion2.tex
v3.2

@@
 Detectable differences in dewlap color between populations are surprising, as habitats were often in close geographical proximity to each other (as close as a few hundred meters on Bimini and most of the time within ten kilometers).
-Indeed, given that (1) the populations were continuously distributed between the habitats, (2) populations from different habitats were not monophyletic with respect to mitochondrial haplotypes (\citealt{vandeSchoot2016} unpubl.), and (3) \textit{A.
+Indeed, given that (1) the populations were continuously distributed between the habitats, (2) populations from different habitats were not monophyletic with respect to mitochondrial haplotypes (\citealt{vandeSchoot2016} unpublished thesis), and (3) \textit{A.
 sagrei} has been shown to be a highly mobile species \citep{Kamath2018}, we would have expected more homogeneous distributions of color phenotypes within islands due to extensive gene flow, with fewer differences between populations, especially those in close proximity.\\
@@
 
-If our results were an example of sensory drive, we would have expected to see consistent differences between habitat-populations across islands (a pattern that would have been a compelling indicator of adaptation, \citealt{Losos2011}).
-This is because environmental conditions that may be relevant to color signal detectability such as light, temperature, moisture and vegetation, are consistent within the three main and clearly distinct habitat-types found across the sampled islands, i.e.
-beach scrub, primary coppice and mangroves \citep{Howard1950, Schoener1968}.
-Moreover, the patterns of divergence expected under a sensory drive scenario should be consistent with increased detectability given the local light conditions, such as the high contrasts with background vegetation found in the UV-range by \citet{Leal2002} and \citet{Leal2004}.\\
-
-Instead, we found differences in the way dewlap color differs between habitats across islands.
+If our results were an example of sensory drive, we would have expected to see consistent differences between populations from different habitats across islands, given the apparent environmental consistency each of the three habitat types across the islands we sampled.
+In particular, we would have expected divergence in line with increased detectability given local light conditions, such as the high contrasts with background vegetation found in the UV range in \citet{Leal2002} and \citet{Leal2004}.
+We might also have expected mangrove and beach scrub lizards, both inhabiting areas with high light penetration, to harbor more similar dewlaps, and to differ significantly form lizards from the coppice habitat, where irradiance is low.
+Instead, we found inconsistencies in the way dewlap color differed between habitats across islands.
 While short-wavelengths (UV reflectance) were often involved in color differences, they were not involved on all islands where significant differences were detected.
@@
 Similar portions of the spectrum were sometimes involved in opposite directions on different islands, such as on Abaco and Cayman Brac, where mangrove lizards had a higher UV-reflectance than beach scrub lizards on the former, but a lower UV-reflectance on the latter.
-Under a sensory-drive scenario, we would have expected the same variables to be consistently divergent between habitats, or at least in a consistent direction.\\
-
-Consistent patterns across islands would have been a good clue for a sensory-drive explanation, but in particular consistent differences between habitats that are most different in their local conditions regarding the ecological function of the dewlap, such as ambient light.
-For example, if ambient light is an important factor shaping dewlap coloration, we would expect mangrove and beach scrub lizards, both inhabiting areas with high light penetration, to harbor more similar dewlaps, and to differ significantly from lizards from the coppice habitat, where irradiance is low.
 Overall, the observed heterogeneity of divergence patterns across islands provides no support to a sensory-drive explanation.\\
 
-Phenotypic plasticity could be another cause for dewlap color variation between habitats, where different environmental conditions produce different phenotypes in different habitats.
-The yellow, orange and red colors in anoline dewlaps are produced by pterins and carotenoids \citep{Ortiz1962, Ortiz1962a, Ortiz1963, Ortiz1966, Macedonia2000, Steffen2007, Steffen2009}.
-Animals can synthesize pterins from nucleotides, but lack the ability to synthesize carotenoids \citep{Goodwin1984, Hill2002, Hill2006}.
-Different food qualities across sites within islands could therefore potentially cause detectable differences in coloration.
-Alternatively, more subtle effects on dewlap color could arise from developmental plasticity and depend, e.g., on differences in egg-rearing conditions.
-However, more data are needed to test these hypotheses, and although \citet{Cook2013} have shown plastic responses of dewlap color in response to parasites, we find it unlikely to account for the widespread habitat differences we found.
-Besides, studies testing the effect of carotenoid deprivation \citep{Steffen2010, Ng2013} and heritability \citep{Cox2017} of dewlap coloration in \textit{A.
-sagrei} and another species with a carotenoid-based dewlap, \textit{A.
-distichus}, found little support for phenotypic and developmental plasticity in dewlap coloration.\\
+% Besides, pterins (the other orange pigment in dewlaps) can be synthesized \citep{Ortiz1962, Ortiz1962a, Ortiz1963, Ortiz1966, Macedonia2000, Steffen2007, Steffen2009}
 
-Genetic drift is another process that can account for differences in phenotype between localities, especially in small populations.
-One way this could proceed is through isolation-by-distance, where more distant populations accumulate more differences through time because of the reduced effect of gene flow at larger geographical scales relative to the dispersal range of the species \citep{Rousset2004}.
-Here, we only found a significant correlation between phenotypic and geographical distances on Eleuthera to support this scenario.
-On all the other islands, in contrast, populations from closer sites were not phenotypically more similar, which argues against isolation-by-distance.
-That said, there were often few sampling sites per island in our study, whose locations were not uniformly chosen within the islands, and so the true extent of isolation-by-distance may be difficult to test.
-Other forms of drift may be at play than isolation-by-distance, but nevertheless, we did find significant differences in color phenotype at relatively small spatial scales, sometimes in neighboring habitats, on islands where gene flow is probably highly pervasive, as suggested by high rates of encounter between males and females \citep{Kamath2018}, making the divergence of habitat-populations by drift in relative genetic isolation an unlikely scenario.\\
+It is presently not known if the reported differences in coloration have a genetic basis.
+Yet, we find it unlikely that these differences arose through phenotypic plasticity, as although the carotenoids that partly make up the red and orange colors of anole dewlaps must be found in the diet \citep{Goodwin1984, Hill2002, Hill2006}, studies testing the effect of carotenoid deprivation \citep{Steffen2010, Ng2013} and heritability \citep{Cox2017} of dewlap coloration in \textit{A.
+sagrei} and \textit{A.
+distichus} (another species with a carotenoid-based dewlap), found little support for phenotypic and developmental plasticity in dewlap coloration.
+One exception is a study demonstrating that lizards heavily parasitized by skin mites had duller dewlaps \citep{Cook2013}, but we found no sign of that in our study.\\
 
-A number of alternative explanations remain.
-First, there could be sexual selection for different dewlap colors in different locations.
-Indeed, although the sensory-drive hypothesis may include a sexual selection aspect, e.g., if the optimal male phenotype in a given habitat is a function of female perception, sexual preferences may also be arbitrary and independent of the habitat \citep{Andersson1994}, and so could differ across habitats and islands.
-Although one previous study has found no link between dewlap coloration and body size dimorphism in \textit{A.
-sagrei} (a proxy for the intensity of sexual selection) in an among-island comparison \citep{Baeckens2018}, within-island data are lacking to test the plausibility of this scenario.\\
+We found no evidence for a role of genetic drift in explaining the observed patterns either.
+First, \textit{A.
+sagrei} was distributed across the islands continuously, usually at relatively high population densities, and not in small and isolated populations where drift might be expected to have a strong effect.
+Second, we found no evidence of isolation-by-distance except on Abaco, which was sampled at the largest geographical scale, with sites nearly a hundred kilometers apart from each other.
+Hence, while isolation-by-distance may explain long-range differences on this island, most of the differences among habitats across islands, which were often in close proximity (less than 10km), are unlikely to be explained by genetic drift.\\
 
-Alternatively, selective pressures may be different in similar habitats from one island to another, because of other environmental variables not accounted for by the habitat-type classification we used \citep{Howard1950, Schoener1968}.
-The islands we sampled exhibit variation in some climatic variables, but also in densities of predators and anole congeners, which have all been shown to correlate with variation in \textit{A.
-sagrei} mean dewlap coloration among islands \citep{Vanhooydonck2009, Baeckens2018}.\\
-
-Finally, different island-populations could also respond differently to similar selective pressures, resulting in various between-habitat divergence patterns across islands.
-Several factors could account for this.
-For example, the founder populations of each island, which colonized the islands independently (van de Schoot, unpublished thesis; \citealt{Driessens2017, Reynolds2020}), could have exhibited different dewlap colors at the time of colonization, as may be suggested by the larger differences we observed between than within islands.
-In turn, different initial phenotypes could have led to different ways in which populations would have diverged between habitats.
-Moreover, the different founding populations may have also consisted in different subsets of the standing genetic variation of their Cuban ancestor due to potential bottlenecks \citep{Reynolds2020}, which may have constrained the way they would later respond to the local selective pressures of their respective islands.\\
+In this study, we found larger differences among than within islands, a pattern already reported and linked to climatic conditions \citep{Vanhooydonck2009} and to densities of predators and of anole congeners \citep{Baeckens2018}.
+Differences among habitats within islands, however, are still difficult to account for.
+Remaining hypotheses may include, for example, runaway sexual selection arbitrarily operating in different directions across islands \citep{Andersson1994}, but no evidence so far suggests that dewlap is a target of mate choice in anoles \citep{Tokarz2002, Tokarz2005, Nicholson2007}.
+Another hypothesis is that the different genetic constitutions of different islands, perhaps resulting from founder effects (the islands have been colonized independently from Cuba, \citealt{vandeSchoot2016} unpubl.; \citealt{Driessens2017, Reynolds2020}), may have predisposed populations to adapt differently to similarly selective circumstances.
+Either way, new data would be needed to test these hypotheses.\\
 
 Altogether, our results show that dewlap color of \textit{A.
-sagrei} commonly varies between habitat types, even in close geographical proximity, within islands of the West Indies, and that coloration differs in different ways from one island to another.
-We discussed several non-mutually exclusive mechanisms that could explain these observations, but more data are needed to thoroughly test these hypotheses.
+sagrei} commonly varies between habitat types, even in close geographical proximity, within islands of the West Indies.
+Besides, coloration differs in different ways across similar habitats from one island to another.
+We discussed several non-mutually exclusive mechanisms that could explain these observations, but more data are needed to investigate these scenarios.
 Nevertheless, heterogeneous patterns of divergence across islands do not support an adaptive sensory-drive scenario, and we propose that within-island dewlap color variation may be underlain by a more subtle mosaic of factors.

### 2021-07-26 ms/sections/introduction.tex
v3.2

@@
 sagrei} dewlaps within nine islands in the Bahamas and Cayman Islands.
-These island systems presently, if not historically, comprise relatively small islands, with no major geographic barriers within islands limiting dispersal for this promiscuous species \citep{Kamath2018}.
+These island systems presently, if not historically, comprise relatively small islands, with no major geographic barriers within islands limiting dispersal for this species \citep{Kamath2018}.
 These islands all share three characteristic native West Indian small-island habitat-types -- beach scrub bush, closed-canopy primary coppice forest, and mangrove forest -- that are often spatially intermingled.
@@
 Similar, if detectability is maximized given the local conditions, we expected darker and more contrasting dewlaps in high irradiance habitats.
-Finally, if habitat characteristics are strong determinants of dewlap color variation, similar patterns should be observed across multiple islands \citep{Losos2011}.
+Finally, if habitat characteristics are strong determinants of dewlap color variation, similar patterns should be observed across multiple islands \citep{Harvey1991, Losos2011}.
 

### 2021-07-26 ms/sections/methods2.tex
v3.2

@@
 Three habitats were sampled on each island based on characterizations by \citet{Howard1950} and \citet{Schoener1968}.
-Each habitat is clearly distinguishable by its dominant vegetation type --- xeric beach scrub (open, relatively dry habitat consisting of low vegetation or isolated trees), primary coppice forest (closed-canopy forest) and mangrove forest (wet coastal habitat with trees growing in brackish water and high light penetration).
+Each habitat is clearly distinguishable by its dominant vegetation type --- xeric beach scrub (open, relatively dry habitat consisting of low vegetation or isolated trees), primary coppice forest (closed-canopy forest) and mangrove forest (wet coastal habitat with trees growing in brackish water and high light penetration, although lizards were sampled in dry soil areas).
 Sample sizes are given in Table \ref{tab:counts}.
@@
 Last, to get insights into the spatial scale of phenotypic variation, for each significant contrast between two habitats detected along a given dimension on a given island, we performed multiple pairwise Wilcoxon tests to assess differences in dewlap coloration among the sites involved in that significant contrast, and recorded the geographical distance between sites that were found significant.
-In parallel, we tested a possible effect of isolation-by-distance, an alternative cause of phenotypic divergence between localities, based on diffusion approximation and dispersal distance, irrespective of habitat-types.
+In parallel, we tested a possible effect of isolation-by-distance, an alternative cause of phenotypic divergence between localities, based on diffusion approximation and dispersal distance, irrespective of habitat types.
 We did so using a permutation test to assess the significance of the correlation between geographical distances and phenotypic distances among sites within each island.\\
@@
 Goodness-of-fit was estimated using Akaike's Information Criterion corrected for small sample sizes (AICc, R package MuMIn, \citealt{Barton2019}), and the estimation method yielding the lowest AICc was retained.
-In a second step, we re-fitted the retained model with maximum likelihood (ML) to test for the effect of habitat-type using likelihood ratio tests (LRT) between a model including a habitat-term and a null model lacking the habitat-term.\\
+In a second step, we re-fitted the retained model with maximum likelihood (ML) to test for the effect of habitat type using likelihood ratio tests (LRT) between a model including a habitat-term and a null model lacking the habitat-term.\\
 
@@
 
-We tested for within-island spatial autocorrelation between the geographical distances among sampling sites and their Euclidean distances in multivariate color space (mean PC1 to PC4 per site, Table \ref{tab:sites}), regardless of habitat-type.
-Because often only a few sites were sampled per island, we could not get meaningful results from tests that use sites as units of observation, such as Moran's I test \citep{Gittleman1990}.
-Instead, we designed a permutation test where we randomly reshuffled individual lizards across sites within islands 1,000 times each, and systematically recalculated Pearson's correlation coefficient between geographic distances (computed as geodesic distances in the R package geosphere; \citealt{Hijmans2019}) and phenotypic distances.
-We used the resulting null distributions of correlation coefficients to assess the significance of the observed spatial autocorrelation for each island.
+We tested for within-island spatial autocorrelation between the geographical distances among sampling sites and their Euclidean distances in multivariate color space (mean PC1 to PC4 per site, Table \ref{tab:sites}), regardless of habitat type.
+For this, we performed Mantel's test (\citealt{Legendre2012}, R package vegan; \citealt{Oksanen2019}) on each island, using 999 permutations and geographical distances computed as geodesic distances from latitude and longitude data (R package geosphere, \citealt{Hijmans2019}).\\
 
@@
 In this study, we were interested in the minimum spatial scale at which significant differences between habitats could be detected within islands.
-We performed multiple pairwise nonparametric Wilcoxon-Mann-Whitney tests \citep{Hollander2013} to compare dewlap coloration between sites with different habitat-types, for each pair of habitats and each variable where significant differences were detected with our analyses of variance.
+We performed multiple pairwise nonparametric Wilcoxon-Mann-Whitney tests \citep{Hollander2013} to compare dewlap coloration between sites with different habitat types, for each pair of habitats and each variable where significant differences were detected with our analyses of variance.
 The P-values were adjusted using a Benjamini-Hochberg correction for multiple testing \citep{Benjamini1995}.

### 2021-07-26 ms/sections/results2.tex
v3.2

@@
 We tested for variation in \textit{A.
-sagrei} dewlap coloration between populations living in three characteristic habitat-types across nine islands that span the West Indian range of the brown anole (beach scrub, primary coppice and mangroves).
+sagrei} dewlap coloration between populations living in three characteristic habitat types across nine islands that span the West Indian range of the brown anole (beach scrub, primary coppice and mangroves).
 We found that most of the variation in coloration was partitioned between islands (two-way PERMANOVA, $F(df = 8) = 43.7$, $P = 0.0001$, explained variance $R^2 = 40.9$\%).
-Nonetheless, we did find evidence for differences in dewlap coloration between habitat-types, and those were mostly island-specific (habitat-by-island interaction term, $F(16) = 3.53$, $P = 0.0001$, $R^2 = 6.6$\%), with a significant portion of the variation explained by an habitat effect across all islands, but this effect was relatively small ($F(2) = 4.7$, $P = 0.0001$, $R^2 = 1.1$\%).\\
+Nonetheless, we did find evidence for differences in dewlap coloration between habitat types, and those were mostly island-specific (habitat-by-island interaction term, $F(16) = 3.53$, $P = 0.0001$, $R^2 = 6.6$\%), with a significant portion of the variation explained by an habitat effect across all islands, but this effect was relatively small ($F(2) = 4.7$, $P = 0.0001$, $R^2 = 1.1$\%).\\
 
@@
 We obtained similar results using other machine learning approaches such as support vector machines (Table \ref{tab:ksvms}) and linear discriminant analysis (Table \ref{tab:ldas}), except that these methods did not detect significant differences on Eleuthera and North Andros.
-We did not find evidence of spatial autocorrelation in dewlap coloration between the sites within islands, except for Eleuthera (Table \ref{tab:autocorrelation}).
-We now describe the specific differences detected on each island.\\
+We describe in details the specific differences detected on each island in the Appendix, and instead focus here on the general patterns emerging from our data.\\
 
-On Abaco, dewlaps from the mangrove habitat were the best discriminated, while dewlaps from the beach scrub habitat were often mistaken for dewlaps from the coppice habitat (Fig.
-\ref{fig:Abaco}D).
-Importance analysis revealed that beach scrub and mangrove lizards mostly differed in reflectance in the ultraviolet (UV) end of the spectrum (below 400nm, Fig.
-\ref{fig:Abaco_supplement}F), where mangrove dewlaps had higher UV reflectance relative to beach scrub lizards, and coppice lizards had an intermediate UV reflectance between the two other habitats (Fig.
-\ref{fig:Abaco}B).
-Consistent with this, our analyses of variance detected significantly lower PC2 scores in mangrove lizards than in the two other habitats (Fig.
-\ref{fig:Abaco}E, Table \ref{tab:anova}), representing a higher UV-reflectance relative to red (Fig.
-\ref{fig:Abaco}C).
-Beach scrub lizards also scored lower on PC3 (Fig.
-\ref{fig:Abaco}E, Table \ref{tab:anova}), indicating less curvature of the reflectance profile and relatively higher reflectance at intermediate wavelengths (blue-to-yellow) than at the ends of the range (Fig.
-\ref{fig:Abaco}C).
-Differences were detected between sites both at large ($\sim$ 100km) and short ($<$ 1km) distances (Fig.
-\ref{fig:Abaco_supplement}G).\\
+Overall, we found significant differences in dewlap coloration between populations that were often in close geographical proximity.
+On Bimini, notably, we found a significant difference between dewlaps from beach scrub and primary coppice forest, at a distance of a few hundred meters, making this contrast the smallest geographical scale at which differences in coloration were found in our study (Fig.
+\ref{fig:Bimini}).
+We also detected significant differences in dewlap coloration at distances below one kilometer on Abaco (Fig.
+\ref{fig:Abaco_supplement}G), and at distances between one and ten kilometers on Bimini (Fig.
+\ref{fig:Bimini}G), Cayman Brac (Fig.
+\ref{fig:CaymanBrac}G), Little Cayman (Fig.
+\ref{fig:LittleCayman}G), Long Island (Fig.
+\ref{fig:LongIsland}G) and South Andros (Fig.
+\ref{fig:SouthAndros}G).\\
 
-On Bimini, the random forests mostly correctly classified lizards from the coppice and mangrove habitats while often misclassifying lizards from the beach scrub habitat (Fig.
+We found evidence of spatial autocorrelation in dewlap coloration between the sites within islands for Abaco (Table \ref{tab:autocorrelation}), suggesting that populations from closer sites tend to have more similar dewlaps on this island than expected by chance.
+Abaco was the island we sampled at the largest scale, with some sites nearly a hundred kilometers away from each other (Fig.
+\ref{fig:Abaco}A).
+That said, some sites were also in close proximity, and significant differences in coloration were detected between habitats sometimes less than a kilometer away (Fig.
+\ref{fig:Abaco_supplement}G), suggesting that differences in dewlap coloration between distant sites may be partly attributable to isolation-by-distance, but this may not necessarily be the case for sites in close proximity.
+We did not find evidence for spatial autocorrelation on other islands than Abaco (Table \ref{tab:autocorrelation}).\\
+
+A striking feature of our data was inconsistencies in between-habitat differences among islands, in terms of which habitats differ from which, which dimensions of coloration were involved, and in which direction.
+For example, while on Cayman Brac the random forests could well distinguish between all three habitats (Fig.
+\ref{fig:CaymanBrac}D), on Abaco dewlaps from beach scrub and primary coppice were often mistaken, and on Bimini beach scrub dewlaps were more often classified into primary coppice or mangrove than into beach srub (Fig.
 \ref{fig:Bimini}D).
-Relatively flat importance profiles for beach scrub lizards suggested that brightness was used instead of a particular wavelength to identify some of the beach scrub dewlaps (Fig.
-\ref{fig:Bimini}F).
-Indeed, some beach scrub dewlaps were substantially brighter than the rest (Fig.
-\ref{fig:Bimini}B, C), a pattern that was captured by our analysis of variance along PC1 (i.e.
-brightness, Fig.
-\ref{fig:Bimini}C, E, Table \ref{tab:anova}).
-The random forests also used UV reflectance to discriminate between coppice and mangrove dewlaps (Fig.
-\ref{fig:Bimini}F), which could reflect the significant difference we detected along PC3 between these two habitats (Fig.
-\ref{fig:Bimini}C, E, Table \ref{tab:anova}).
-Beach scrub lizards were characterized by elevated red reflectance relative to UV (as represented by PC2, (Fig.
-\ref{fig:Bimini}C, E)), and beach scrub and mangrove lizards were characterized by a more even distribution of the reflectance along the spectrum (as represented by PC3, (Fig.
-\ref{fig:Bimini}C, E)), in contrast to coppice lizards which harbored a stronger curvature at intermediate wavelengths (Fig.
-\ref{fig:Bimini}C, E).
-On this island, the beach scrub and coppice habitats were separated by a few hundred meters, making this contrast the smallest geographical scale at which differences in coloration were found in our study (Fig.
-\ref{fig:Bimini}G).\\
-
-On Cayman Brac, all three habitats could be well discriminated against each other (Fig.
-\ref{fig:CaymanBrac}D), with UV reflectance appearing to be an important variable differentiating beach scrub and mangrove dewlaps (Fig.
-\ref{fig:CaymanBrac}F).
-In contrast, coppice dewlaps had a relatively flat importance profile, suggesting that brightness made them more distinct rather than any particular wavelength (Fig.
-\ref{fig:CaymanBrac}F).
-Consistent with this, coppice dewlaps were significantly different from all other dewlaps along PC1 (Fig.
-\ref{fig:CaymanBrac}E, Table \ref{tab:anova}).
-At a distance between 2 and 3km (Fig.
-\ref{fig:CaymanBrac}G), dewlaps in the beach scrub habitat reflected more red light (as represented by PC2, Fig.
-\ref{fig:CaymanBrac}C, E) and more UV (as represented by PC3, Fig.
-\ref{fig:CaymanBrac}C, E) than in the mangrove habitat.
-Coppice lizards were also characterized by a higher UV reflectance than mangrove dewlaps (PC3, Fig.
-\ref{fig:CaymanBrac}C, E, Table \ref{tab:anova}).\\
-
-Eleuthera was the only island where we detected significant spatial autocorrelation (Table \ref{tab:autocorrelation}), that is, sites that were closer geographically tended to have populations of lizards with more similar dewlap colors.
-Although random forests detected between-habitat differences in dewlap color, other approaches did not (Tables \ref{tab:ldas} and \ref{tab:ksvms}), suggesting that the differences may be small.
-Consistent with this, the only significant univariate difference detected was for PC2 between beach scrub and mangrove lizards, where beach scrub lizards had higher levels of red reflectance and mangrove lizards higher levels of UV reflectance (Fig.
-\ref{fig:Eleuthera}C, E, Table \ref{tab:anova}).\\
-
-Little Cayman was characterized by a better discrimination of mangrove lizards from the rest than between beach scrub and coppice lizards (Fig.
-\ref{fig:LittleCayman}D).
-Mangrove dewlaps were most distinct with respect to their reflectance in short wavelengths (Fig.
-\ref{fig:LittleCayman}F), with significantly lower UV reflectance (as represented by PC2, Fig.
-\ref{fig:LittleCayman}C, E, Table \ref{tab:anova}).
-Beach scrub lizards were characterized by brighter dewlaps than coppice lizards (PC1), and also more convex curves, i.e.
-slightly higher UV and red reflectance (as represented by higher PC3 scores), than lizards from the other two habitats (Fig.
-\ref{fig:LittleCayman}C, E, Table \ref{tab:anova}).\\
-
-On Long Island the three habitats were relatively well discriminated (Fig.
-\ref{fig:LongIsland}D).
-Importance profiles indicated that short wavelengths were used to discriminate between beach scrub and mangrove lizards (Fig.
-\ref{fig:LongIsland}F).
-Beach scrub lizards had more curved reflectance profiles than in either of the two other habitats, with higher levels of UV and red reflectance relative to intermediate wavelengths (PC3, Fig.
-\ref{fig:LongIsland}C, E, Table \ref{tab:anova}).
-Beach scrub lizards also differed from mangrove lizards along PC4 (Fig.
-\ref{fig:LongIsland}E), which represented a rather small portion of the variance not already explained by the first three principal components, and is therefore difficult to interpret (Fig.
-\ref{fig:LongIsland}C).
-Coppice lizards were significantly darker than mangrove and beach scrub lizards (PC1, Fig.
-\ref{fig:LongIsland}C, E, Table \ref{tab:anova}).\\
-
-On North Andros, although the random forest classification was significant ($P = 0.0216$, Table \ref{tab:randomforests}) and the average confusion matrix indicated that lizards from beach scrub were particularly well predicted (Fig.
-\ref{fig:NorthAndros}D), no significant univariate differences were detected along any of the four PCs (Fig.
-\ref{fig:NorthAndros}E, Table \ref{tab:anova}).
-Importance analysis of full-spectrum random forests showed higher importance scores near the UV-end of the spectrum in discriminating beach scrub dewlaps from the rest (Fig.
-\ref{fig:NorthAndros}F).
-Besides, reflectance curves of beach scrub dewlaps appeared more similar to each other in the UV range than dewlaps from other habitats (Fig.
-\ref{fig:NorthAndros}B), suggesting that the machines may have used this low within-habitat variance, as opposed to between-habitat differences in means, to correctly classify beach scrub lizards.
-A small sample size on this islands may also have contributed to a lack of power in detecting univariate differences using analyses of variance (Table \ref{tab:counts}).\\
-
-On South Andros beach scrub and coppice dewlaps could be discriminated better against each other than with mangrove dewlaps (Fig.
-\ref{fig:SouthAndros}D), with importance profiles supporting UV-reflectance as a predictor of coppice lizards (Fig.
-\ref{fig:SouthAndros}F).
-Coppice lizards had more curved reflectance profiles than beach scrub lizards (PC3), and lizards from both habitats differed along PC4, which is again more difficult to interpret (Fig.
-\ref{fig:SouthAndros}C, E, Table \ref{tab:anova}).
-Beach scrub lizards also differed from mangrove lizards in PC4 (Fig.
-\ref{fig:SouthAndros}C, E, Table \ref{tab:anova}).\\
-
-Classification success was not significantly better than expected by chance on Ragged Island (Table \ref{tab:randomforests}) where nearly no habitat could be differentiated from any other based on reflectance.
+In terms of variable importance, for multiple islands the random forests used information in the UV range to discriminate between at least some habitats, particularly on Abaco (Fig.
+\ref{fig:Abaco_supplement}F), Bimini (Fig.
+\ref{fig:Bimini}F), Cayman Brac (Fig.
+\ref{fig:CaymanBrac}F), Little Cayman (Fig.
+\ref{fig:LittleCayman}F) and Long Island (Fig.
+\ref{fig:LongIsland}F), but differences in UV reflectance involved different habitats and were in different directions among these islands.\\

### 2021-07-26 ms/sections/tables.tex
v3.2

@@
 \begin{table}[H]
-	\caption{Test of spatial autocorrelation.
+	\caption{Mantel's test of spatial autocorrelation.
 For each island are shown the correlation (Pearson's $\rho$) between the matrix of phenotypic distances between populations from each site and the matrix of geographic distances between sites, where phenotypic distances are Euclidean distances between the mean phenotypes of each site in the multivariate space consisting of the first four within-island principal components.
-P-values assess the significance of the observed correlation against the correlation expected if lizards were randomly permuted among sites (1,000 permutations).
-*, $P < 0.05$; **, $P < 0.01$, ***, $P < 0.001$.}
+P-values assess the significance of the observed correlation against the correlation expected if population means were randomly permuted among sites (999 permutations).
+*, $P < 0.05$.}
 	\centering
-	\input{"./tables/autocorrelation.tex"}
+	\input{"./tables/mantel.tex"}
 	\label{tab:autocorrelation}

### 2021-07-26 analyses.R

@@ -24,7 +24,7 @@
 
 rm(list = ls())
 
-library(nmgc)
+#library(nmgc)
 library(tidyverse)
 library(ggridges)
 library(ggiraphExtra)
@@ -41,6 +41,7 @@ library(RColorBrewer)
 library(pairwiseAdonis)
 library(scales)
 library(vegan)
+library(ape)
 
 #### 1. Set-up ####
 
@@ -1170,7 +1171,7 @@ islands <- c(
   "Long Island", "North Andros", "South Andros"
 )
 
-#### 8.3.4. Run on principal components ####
+#### 8.3.4. Run on reflectance data ####
 
 set.seed(35)
 
@@ -1311,6 +1312,8 @@ D %>%
   summarize(across(wpc_names, mean)) %>%
   write_csv("metadata/sites.csv")
 
+#### 9.1. Permutation test ####
+
 set.seed(1994)
 
 # Permutation test
@@ -1380,6 +1383,64 @@ D %>%
   ) %>%
   select(robs, pvalue) %>%
   write_csv("results/spatial_correlation/spatial_correlation.csv")
+
+#### 9.2. Mantel's test ####
+
+set.seed(99)
+
+D %>%
+  group_by(island) %>%
+  nest() %>%
+  mutate(
+
+    # Get a vector of observed and permuted correlation coefficients
+    mantel_test = map(data, function(X) {
+
+      # Compute mean phenotypes per site
+      S <- X %>%
+        group_by(longitude, latitude) %>%
+        nest() %>%
+        mutate(
+
+          data = map(data, function(data) {
+
+            tibble(
+              variable = wpc_names,
+              mean = colMeans(data[wpc_names])
+            )
+
+          })
+
+        ) %>%
+        unnest(cols = c(data)) %>%
+        pivot_wider(names_from = "variable", values_from = "mean")
+
+      # Matrix of geographic distances between sites
+      G <- as.dist(
+        distm(
+          S[c("longitude", "latitude")], fun = distGeo
+        )
+      )
+
+      # Matrix of euclidean distances between site means in phenotype space
+      M <- dist(S[wpc_names])
+
+      # Mantel's test of correlation between these two matrices
+      test_result <- mantel(G, M)
+      test_stat <- test_result$statistic
+      p_value <- test_result$signif
+
+      return(list(test_stat, p_value))
+
+    }),
+
+    # Extract the test statistics
[…]

### 2021-07-26 ms/run_pandoc.sh

@@ -1,5 +1,5 @@
 # Run this script to convert LaTeX files into ODT and DOCX documents
 # Requires pandoc and pandoc-crossref being installed
 
-pandoc main.tex --filter=pandoc-crossref --bibliography=library.bib --csl=journal-of-evolutionary-biology.csl -o scherrer2021_v3.1.odt
-pandoc main.tex --filter=pandoc-crossref --bibliography=library.bib --csl=journal-of-evolutionary-biology.csl -o scherrer2021_v3.1.docx
+pandoc main.tex --filter=pandoc-crossref --bibliography=library.bib --csl=journal-of-evolutionary-biology.csl -o scherrer2021_v3.2.odt
+pandoc main.tex --filter=pandoc-crossref --bibliography=library.bib --csl=journal-of-evolutionary-biology.csl -o scherrer2021_v3.2.docx

### 2021-07-26 tables.R

@@ -132,9 +132,26 @@ read_csv("results/group_comparisons/anovas.csv") %>%
   ) %>%
   cat(file = "ms/tables/anova.tex")
 
-#### 5. Spatial autocorrelation table ####
+#### 5. Spatial autocorrelation tables ####
 
-# Spatial autocorrelation
+# Mantel's test of spatial autocorrelation
+read_csv("results/spatial_correlation/mantel_test.csv") %>%
+  mutate_signif() %>%
+  mutate(pvalue = round_pvalue(pvalue, digits = 3)) %>%
+  kable(
+    "latex",
+    col.names = c(
+      "Island", "$\\rho$", "$P$", ""
+    ),
+    digits = c(0, 3, 0, 0),
+    align = "lrrl",
+    booktabs = TRUE,
+    linesep = "",
+    escape = FALSE
+  ) %>%
+  cat(file = "ms/tables/mantel.tex")
+
+# Spatial autocorrelation (custom permutation test)
 read_csv("results/spatial_correlation/spatial_correlation.csv") %>%
   mutate_signif() %>%
   mutate(pvalue = round_pvalue(pvalue, digits = 3)) %>%

### 2021-09-06 ms/main_title.tex
stuff

--- 2020-08-08 ms/main.tex
@@
 \documentclass{article}
+
+% General rule for P-values
+% * < 0.05
+% ** < 0.01
+% *** < 0.001
+% Write "< 0.0001" or the value up to the fourth decimal
 
@@
 \author{
-    \textsc{Rapha\"{e}l Scherrer$^{1,3}$ \thanks{Corresponding author: r.scherrer@rug.nl}, Colin M. Donihue$^{1,4}$, }\\
-	\textsc{R. Graham Reynolds$^2$, Jonathan B. Losos$^{1,4}$ and Anthony J. Geneva$^{1,5}$} \\[1ex]
-	%\thanks{A thank you} \\[1ex] % Your name
+	\textsc{Rapha\"{e}l Scherrer}$^{1,3}$ \thanks{Corresponding author} (raphael.scherrer@evobio.eu, ORCID 0000-0002-1447-7630),\\
+    \textsc{Colin M. Donihue}$^{1,4}$ (colindonihue@gmail.com, 0000-0003-1096-8536),\\
+	\textsc{R. Graham Reynolds}$^2$ (greynold@unca.edu, 0000-0002-0056-8304),\\ 
+    \textsc{Jonathan B. Losos}$^{1,4}$ (losos@wustl.edu, 0000-0003-4712-8258) and\\
+	\textsc{Anthony J. Geneva}$^{1,5}$ (anthony.geneva@gmail.com, 0000-0001-8253-6527)\\[1ex]
 	\normalsize $^1$ Department of Organismic and Evolutionary Biology and Museum of Comparative Zoology \\ \normalsize Harvard University, Cambridge, MA, USA \\
@@
 
-\begin{abstract}
-    \input{sections/abstract.tex}
-\end{abstract}
-
-\textbf{Keywords} --- reflectance, adaptation, sexual signal, machine learning, polymorphism
-
-\section*{Introduction}
-
-\input{sections/introduction.tex}
-
-\section*{Methods}
-
-\input{sections/methods.tex}
-
-\section*{Results}
-
-\input{sections/results.tex}
-
-\section*{Discussion}
-
-\input{sections/discussion.tex}
-
-\subsection*{Acknowledgements}
-
-\input{sections/acknowledgments.tex}
-
-\pagebreak
-
-\section*{Figures}
-
-\input{sections/figures.tex}
-
-\pagebreak
-
-\section*{Tables}
-
-\input{sections/tables.tex}
-
-\pagebreak
-
-\bibliographystyle{apalike}
-\bibliography{library.bib}
-
-\beginsupplement
-
-%\section*{Supplementary Figures}
-
-%\input{sections/suppfigures.tex}
-
-\pagebreak
-
-\section*{Supplementary Figures}
-
-\input{sections/suppfigures}
-
-\pagebreak
-
-\section*{Supplementary Tables}
-
-\input{sections/supptables.tex}
+	\textbf{Running title} --- Dewlap coloration in \textit{Anolis sagrei}.
 

### 2021-09-06 ms/sections/discussion2.tex
stuff

@@
 Indeed, given that (1) the populations were continuously distributed between the habitats, (2) populations from different habitats were not monophyletic with respect to mitochondrial haplotypes (\citealt{vandeSchoot2016} unpublished thesis), and (3) \textit{A.
-sagrei} has been shown to be a highly mobile species \citep{Kamath2018}, we would have expected more homogeneous distributions of color phenotypes within islands due to extensive gene flow, with fewer differences between populations, especially those in close proximity.\\
+sagrei} is highly mobile \citep{Kamath2018}, we would have expected more homogeneous distributions of color phenotypes within islands due to gene flow, with fewer differences between populations, especially those in close proximity.\\
 
@@
 The sensory-drive hypothesis has been tested multiple times for dewlap coloration in \textit{Anolis} lizards, with mixed results.
-Some authors found support for it \citep{Leal2002, Leal2004}, while others did find differences in dewlap coloration between habitats, but those were inconsistent with the sensory-drive hypothesis \citep{Fleishman2009, Ng2012}.\\
+Some authors found support for it \citep{Leal2002, Leal2004}, while others found differences in dewlap coloration between habitats inconsistent with the sensory-drive hypothesis \citep{Fleishman2009, Ng2012}.\\
 
@@
 In particular, we would have expected divergence in line with increased detectability given local light conditions, such as the high contrasts with background vegetation found in the UV range in \citet{Leal2002} and \citet{Leal2004}.
-We might also have expected mangrove and beach scrub lizards, both inhabiting areas with high light penetration, to harbor more similar dewlaps, and to differ significantly form lizards from the coppice habitat, where irradiance is low.
+We might also have expected mangrove and beach scrub lizards, both inhabiting areas with high light penetration, to have more similar dewlaps, and to differ significantly from lizards from the coppice habitat, where irradiance is low.
 Instead, we found inconsistencies in the way dewlap color differed between habitats across islands.
@@
 First, \textit{A.
-sagrei} was distributed across the islands continuously, usually at relatively high population densities, and not in small and isolated populations where drift might be expected to have a strong effect.
+sagrei} was distributed across the islands continuously, usually at relatively high population densities, rather than in small and isolated populations where drift might be expected to have a strong effect.
 Second, we found no evidence of isolation-by-distance except on Abaco, which was sampled at the largest geographical scale, with sites nearly a hundred kilometers apart from each other.
-Hence, while isolation-by-distance may explain long-range differences on this island, most of the differences among habitats across islands, which were often in close proximity (less than 10km), are unlikely to be explained by genetic drift.\\
+Hence, while isolation-by-distance may explain long-range differences on this island, most of the differences among habitats across the rest of the sampling region are unlikely to be explained by genetic drift, as habitats were often in close proximity (less than 10km).\\
 
@@
 Remaining hypotheses may include, for example, runaway sexual selection arbitrarily operating in different directions across islands \citep{Andersson1994}, but no evidence so far suggests that dewlap is a target of mate choice in anoles \citep{Tokarz2002, Tokarz2005, Nicholson2007}.
-Another hypothesis is that the different genetic constitutions of different islands, perhaps resulting from founder effects (the islands have been colonized independently from Cuba, \citealt{vandeSchoot2016} unpubl.; \citealt{Driessens2017, Reynolds2020}), may have predisposed populations to adapt differently to similarly selective circumstances.
+Another hypothesis is that the different genetic constitutions of different islands, perhaps resulting from founder effects (the islands have been colonized independently, \citealt{vandeSchoot2016} unpubl.; \citealt{Driessens2017, Reynolds2020}), may have predisposed populations to adapt differently to similarly selective circumstances.
 Either way, new data would be needed to test these hypotheses.\\
@@
 sagrei} commonly varies between habitat types, even in close geographical proximity, within islands of the West Indies.
-Besides, coloration differs in different ways across similar habitats from one island to another.
-We discussed several non-mutually exclusive mechanisms that could explain these observations, but more data are needed to investigate these scenarios.
+However, coloration differs in different ways across similar habitats from one island to another.
+We discussed several non-mutually exclusive mechanisms that could explain these observations.
 Nevertheless, heterogeneous patterns of divergence across islands do not support an adaptive sensory-drive scenario, and we propose that within-island dewlap color variation may be underlain by a more subtle mosaic of factors.

### 2021-09-06 ms/sections/introduction.tex
stuff

@@
 
-The species \textit{Anolis sagrei} is widespread across islands of the West Indies \citep{Reynolds2020}.
+\textit{Anolis sagrei} is widespread across islands of the West Indies \citep{Reynolds2020}.
 It has been the subject of numerous studies concerning local adaptation \citep{Losos1994, Losos1997a, Losos2001, Kolbe2012}, biological invasion \citep{Kolbe2008}, and sexual selection \citep{Tokarz2002, Tokarz2005, Tokarz2006, Driessens2014, Steffen2014, Driessens2015} among many other topics.
@@
 We predicted that if light conditions in the environment indeed drive color evolution, dewlaps should be most similar between beach scrub and mangrove forest, which both have high levels of light irradiance, compared to the darker, closed-canopy coppice forest.
-Similar, if detectability is maximized given the local conditions, we expected darker and more contrasting dewlaps in high irradiance habitats.
+If detectability is maximized given the local conditions, we expected darker and more contrasting dewlaps in high irradiance habitats.
 Finally, if habitat characteristics are strong determinants of dewlap color variation, similar patterns should be observed across multiple islands \citep{Harvey1991, Losos2011}.

### 2021-09-06 ms/sections/results2.tex
stuff

@@
 We obtained similar results using other machine learning approaches such as support vector machines (Table \ref{tab:ksvms}) and linear discriminant analysis (Table \ref{tab:ldas}), except that these methods did not detect significant differences on Eleuthera and North Andros.
-We describe in details the specific differences detected on each island in the Appendix, and instead focus here on the general patterns emerging from our data.\\
+We describe in details the specific differences detected on each island in the Appendix, and focus here on the general patterns emerging from our data.\\
 

### 2021-09-06 ms-jeb/sections/abstract.tex
almost ready

--- 2021-07-11 ms/sections/abstract.tex
@@
 Previous studies have demonstrated that dewlap color in \textit{Anolis} lizards can be highly variable between populations in relation to both biotic and abiotic adaptive drivers at relatively large geographical scales.
-Here, we investigated differentiation of dewlap coloration among habitat-types at a small spatial scale, within multiple islands of the West Indies, to test the hypothesis that similar local adaptive processes occur over smaller spatial scales.
+Here, we investigated differentiation of dewlap coloration among habitat types at a small spatial scale, within multiple islands of the West Indies, to test the hypothesis that similar local adaptive processes occur over smaller spatial scales.
 We explored variation in dewlap coloration in the most widespread species of anole, \textit{Anolis sagrei}, across three characteristic habitats spanning the Bahamas and the Cayman Islands, namely beach scrub, primary coppice forest and mangrove forest.

### 2021-09-06 ms/main_title.tex
stuff
almost ready

@@
 \documentclass{article}
-
-% General rule for P-values
-% * < 0.05
-% ** < 0.01
-% *** < 0.001
-% Write "< 0.0001" or the value up to the fourth decimal
 
@@
 \usepackage[T1]{fontenc}
-\usepackage{verbatim} % multiline comments
 
@@
 
-%% Useful packages
-\usepackage{amsmath}
-\usepackage{float}
-\usepackage{rotating}
-\usepackage{booktabs}
-%\restylefloat{figure} % for force placing figures
-%\floatplacement{figure}{H}
-%\floatplacement{table}{H}
-%\floatplacement{longtable}{H}
-%\usepackage[markers,nolists]{endfloat}
-\usepackage{graphicx}
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
-%\usepackage{apalike}
-\usepackage[round]{natbib}
-\usepackage[switch]{lineno} % line numbers
-\usepackage{longtable}
-\usepackage{multirow}
-% Supp mat
-\newcommand{\beginsupplement}{%
-	\setcounter{table}{0}
-	\renewcommand{\thetable}{S\arabic{table}}%
-	\setcounter{figure}{0}
-	\renewcommand{\thefigure}{S\arabic{figure}}%
-}
+\title{Dewlap color variation in \textit{Anolis sagrei} is maintained among habitats within islands of the West Indies}
 
-\title{Dewlap color variation in \textit{Anolis sagrei} is maintained among habitats within islands of the West Indies}
 \author{
+
 	\textsc{Rapha\"{e}l Scherrer}$^{1,3}$ \thanks{Corresponding author} (raphael.scherrer@evobio.eu, ORCID 0000-0002-1447-7630),\\
@@
 	\textsc{Anthony J. Geneva}$^{1,5}$ (anthony.geneva@gmail.com, 0000-0001-8253-6527)\\[1ex]
+
 	\normalsize $^1$ Department of Organismic and Evolutionary Biology and Museum of Comparative Zoology \\ \normalsize Harvard University, Cambridge, MA, USA \\
@@
 	\normalsize $^5$ Current address: Department of Biology, Center for Computational and Integrative Biology,\\ \normalsize Rutgers University--Camden, Camden, NJ, USA
+
 }
+
 \date{} % Leave empty to omit a date
@@
 \begin{document}
-
-	\linenumbers
 

### 2021-09-06 R/classifyX.R

@@ -1,293 +1,0 @@
-#' Classification analysis --- testing versions
-#'
-#' This function is an old version of `classify`, used only to compare different
-#' ways of computing P-values in their tendency to yield false positives.
-#' We concluded that comparing a distribution of observed accuracy scores to
-#' a null distribution was prone to false discovery because we have total
-#' control over the number of replicates, i.e. the number of points in the
-#' vector of observed scores.
-#'
-#' @export
-
-classifyX <- function(
-  data,
-  variables,
-  grouping,
-  nesting = NULL,
-  method = "SVM",
-  k = 5,
-  nrep = 1,
-  nperm = 0,
-  minsize = 5,
-  seed = NULL,
-  importance = FALSE,
-  getmachine = FALSE,
-  verbose = TRUE,
-  pb = TRUE,
-  digest = TRUE,
-  topcomp = NULL,
-  center = TRUE,
-  scale = TRUE
-) {
-
-  # Compute principal components if needed
-  if (!is.null(topcomp)) {
-    data <- data %>%
-      cbind(
-        npcomp(
-          data, topcomp, center, scale, nesting, combine = TRUE,
-          reduce = variables
-        )$x
-      )
-  }
-
-  # Random seed
-  if (!is.null(seed)) set.seed(seed)
-
-  # Proportion of observations in the testing set
-  ptesting <- 1 / k
-
-  assertthat::assert_that(nrow(data) > k)
-  assertthat::assert_that(floor(ptesting * nrow(data)) > 0)
-
-  # Define the possible labels
-  labels <- unique(data[, grouping])
-
-  nested <- TRUE
-
-  # Artificial single nesting level if no nesting supplied
-  if (is.null(nesting)) {
-    nested <- FALSE
-    data$nesting <- factor(1)
-    nesting <- "nesting"
-  }
-
-  # Split the data into each nesting level
-  data <- data %>% split(f = .[, nesting])
-
-  # Decide on a looping function depending on whether we want a progress bar
-  if (!verbose) pb <- FALSE
-  thislapply1 <- thislapply2 <- lapply
-  if (pb) {
-    if (nested) {
-      thislapply1 <- pbapply::pblapply
-    } else {
-      thislapply2 <- pbapply::pblapply
-    }
-  }
-
-  # For each nesting level...
-  results <- thislapply1(data, function(data) {
-
-    # For each replicate (and randomized) analysis...
-    thislapply2(seq(nrep + nperm), function(i) {
-
[…]

### 2021-09-06 R/csv2latex.R

@@ -1,15 +1,0 @@
-rm(list = ls())
-
-library(readr)
-library(knitr)
-
-filename <- "results/group_comparisons/machine_learning/PC/lda/summary.csv"
-
-tab <- read_csv(filename)
-
-# Process significance column
-colnames(tab)[ncol(tab)] <- ""
-tab[[ncol(tab)]][is.na(tab[[ncol(tab)]])] <- ""
-
-# Make a latex table
-kable(tab, "latex", booktabs = TRUE, linesep = "", escape = FALSE)

### 2021-09-06 R/save_table.R

@@ -1,26 +1,0 @@
-#' Save tables for this study
-#'
-#' @param tab The table
-#' @param nm File name (without the extension)
-#' @param ... Extra parameters for `knitr::kable`
-#'
-#' @return Nothing. Just saves.
-#'
-#' @export
-
-save_table <- function(tab, nm, ...) {
-
-  library(knitr)
-
-  # Save table as CSV
-  write.csv(tab, paste0(nm, ".csv"), row.names = FALSE)
-
-  # Save LaTeX table
-  tab <- knitr::kable(
-    tab, "latex", booktabs = TRUE, linesep = "", escape = FALSE, ...
-  )
-  texfile <- file(paste0(nm, ".tex"))
-  writeLines(tab, texfile)
-  close(texfile)
-
-}

### 2021-09-06 analyses.R

@@ -1,2057 +1,0 @@
-## ---------------------------
-##
-## Script name: analyses.R
-##
-## Purpose of script: This script contains all the analyses of this study.
-##
-## Author: Raphael Scherrer
-##
-## Date Created: 2021-02-06
-##
-## Copyright (c) Raphael Scherrer, 2021
-##
-## Email:
-## r.scherrer@rug.nl
-## raphael.scherrer@evobio.eu
-## raph.rjfs@hotmail.fr
-##
-## ---------------------------
-
-# Notation used for naming datasets in this script:
-# D = the original dataset
-# X = placeholder within a function
-# P = temporary data frame for plotting or containing plots
-
-rm(list = ls())
-
-#library(nmgc)
-library(tidyverse)
-library(ggridges)
-library(ggiraphExtra)
-library(broom)
-library(heplots)
-library(MVN)
-library(assertthat)
-library(rminer)
-library(geosphere)
-library(MuMIn)
-library(nlme)
-library(PMCMRplus)
-library(RColorBrewer)
-library(pairwiseAdonis)
-library(scales)
-library(vegan)
-library(ape)
-
-#### 1. Set-up ####
-
-# Set plotting theme
-theme_set(theme_classic())
-
-# Graphical parameters
-island_colors <- brewer.pal(8, "Set2")
-island_colors <- c(island_colors, "darkgreen")
-habitat_colors <- c("goldenrod", "forestgreen", "mediumseagreen")
-
-# Global settings
-pc_names <- paste0("PC", 1:4)
-wpc_names <- paste0("w", pc_names)
-wl <- 300:700
-wl_names <- paste0("wl", wl)
-
-# Whether to load machines already fitted to whole spectrum data
-# (instead of rerunning the very lengthy classification)
-read_fitted <- TRUE
-
-#### 2. Read reflectance data ####
-
-# Read the data
-D <- read_csv("data/reflectance.csv")
-
-# Useful things to keep in mind
-island_names <- unique(D$island)
-habitat_names <- unique(D$habitat)
-
-# Save sample sizes
-with(D, table(island, habitat)) %>%
-  as_tibble() %>%
-  pivot_wider(names_from = habitat, values_from = n) %>%
-  write_csv("metadata/counts.csv")
-
-#### 2.1. Plot reflectance profiles ####
-
-# Prepare to plot reflectance
-P <- D %>%
[…]

### 2021-09-06 analyses/01-counts/Counts.R

@@ -1,10 +1,0 @@
-# This script was used to generate the table of counts across islands and habitats
-
-rm(list = ls())
-
-library(dewlap)
-
-data <- read.csv("data/reflectance.csv", header = TRUE)
-tab <- table(data$island, data$habitat)
-
-save_table(tab, "analyses/01-counts/table_counts")

### 2021-09-06 analyses/02-reflectance/PlotReflectance.R

@@ -1,72 +1,0 @@
-# Plot reflectance spectra
-
-rm(list = ls())
-
-library(tidyverse)
-
-colors <- c(coastal = "darkgoldenrod1", coppice = "forestgreen", mangrove = "mediumseagreen")
-
-# Read the data
-data <- read.csv("data/reflectance.csv", header = TRUE)
-
-# Reorganize into a long table
-smr <- data %>%
-  gather_("wl", "reflectance", paste0("wl", 300:700)) %>%
-  mutate(wl = as.numeric(gsub("wl", "", wl)))
-
-# Prepare ribbons
-smr <- smr %>% group_by(wl, habitat, island)
-smr <- smr %>%
-  summarize(
-    mid = quantile(reflectance, 0.05),
-    lower = median(reflectance),
-    upper = quantile(reflectance, 0.95)
-  )
-
-# Plot the spectra
-p <- ggplot(smr, aes(x = wl, color = habitat)) +
-  geom_ribbon(aes(ymin = lower, ymax = upper, fill = habitat), alpha = 0.2) +
-  #geom_line(aes(y = mid)) +
-  facet_wrap(. ~ island) +
-  xlab("Wavelength (nm)") +
-  ylab("Reflectance (%)") +
-  theme_bw() +
-  labs(fill = "Habitat", color = "Habitat") +
-  theme(legend.position = "bottom") +
-  scale_fill_manual(values = colors) +
-  scale_color_manual(values = colors)
-p
-
-ggsave("figure_reflectance.png", height = 6, width = 6, dpi = 300)
-
-#####################
-
-# At the scale of the whole archipelago now
-
-# Reorganize into a long table
-smr <- data %>%
-  gather_("wl", "reflectance", paste0("wl", 300:700)) %>%
-  mutate(wl = as.numeric(gsub("wl", "", wl)))
-
-# Prepare ribbons
-smr <- smr %>% group_by(wl, habitat)
-smr <- smr %>%
-  summarize(
-    mid = quantile(reflectance, 0.05),
-    lower = median(reflectance),
-    upper = quantile(reflectance, 0.95)
-  )
-
-# Plot the spectra
-p <- ggplot(smr, aes(x = wl, color = habitat)) +
-  geom_ribbon(aes(ymin = lower, ymax = upper, fill = habitat), alpha = 0.2) +
-  #geom_line(aes(y = mid)) +
-  xlab("Wavelength (nm)") +
-  ylab("Reflectance (%)") +
[…]

### 2021-09-06 analyses/03-PCA/PCA.R

@@ -1,112 +1,0 @@
-rm(list = ls())
-
-# Exploration of the principal component analysis
-
-library(dewlap)
-library(nmgc)
-library(tidyverse)
-
-# Here we are checking first the correlation between PC1 and brightness
-# and second, the variance explained by each PC on each island
-
-data <- read.csv("data/reflectance.csv", header = TRUE)
-
-# Perform PCA on each island
-pca <- npcomp(data, paste0("wl", 300:700), nesting = "island", combine = TRUE, reduce = 1:4)
-
-# And on the whole archipelago
-pca_pooled <- npcomp(data, paste0("wl", 300:700))
-
-####  1. Correlation between brightness and PC1 ####
-
-# Calculate brightness (mean reflectance)
-data$brightness <- data %>% select_at(paste0("wl", 300:700)) %>% apply(1, mean)
-
-# Attach PC1
-data <- data %>% select_at(c("island", "habitat", "brightness"))
-data$PC1 <- pca$x$PC1
-data$PC1pooled <- pca_pooled$x[, 1]
-data <- data %>%
-  gather_("variable", "PC1", c("PC1", "PC1pooled")) %>%
-  mutate(island = ifelse(variable == "PC1pooled", "Archipelago", as.character(island))) %>%
-  dplyr::select(-variable)
-
-# Measure correlation
-res_cor <- data %>%
-  split(f = .[["island"]]) %>%
-  map_dfr(
-    ~ cor.test(.x$PC1, .x$brightness) %>%
-      .[c("estimate", "p.value")] %>%
-      data.frame %>%
-      rename(r2 = "estimate", pvalue = "p.value") %>%
-      mutate(r2 = r2^2),
-    .id = "island"
-  )
-
-# Reorder
-res_cor <- res_cor %>%
-  filter(island != "Archipelago") %>%
-  rbind(., res_cor %>% filter(island == "Archipelago"))
-
-# Save table
-t1 <- res_cor
-t1 <- t1 %>%
-  add_signif() %>%
-  mutate(pvalue = ifelse(pvalue < 0.0001, "< 0.0001", pvalue))
-col.names <- c("Island", "$r^2$", "$P$", "")
-save_table(
-  t1, "analyses/03-PCA/table_brightness", digits = c(0, 3, 0),
-  col.names = col.names, align = "lrrl"
-)
-
-
[…]

### 2021-09-06 analyses/04-machine-learning/MachineLearning.R

@@ -1,369 +1,0 @@
-rm(list = ls())
-
-library(dewlap)
-library(nmgc)
-library(tidyverse)
-library(cowplot)
-
-data <- read.csv("data/reflectance.csv", header = TRUE)
-
-#### Analysis within each island ####
-
-# Setup principal components
-pcomps <- paste0("PC", 1:4)
-wlengths <- paste0("wl", 300:700)
-somewl <- paste0("wl", seq(350, 650, 50))
-
-# Prepare a few ways to perform our analysis
-methods <- c("SVM", "LDA") # type of machine to fit
-datasets <- c("PC", "WL") # type of data to use
-
-# For each method and each dataset
-for (i in seq_along(methods)) {
-  for (j in seq_along(datasets)) {
-
-    # Set up
-    if (datasets[j] == "PC") {
-      variables <- pcomps
-      to_pcomp <- wlengths
-    } else {
-      variables <- somewl
-      to_pcomp <- NULL
-    }
-
-    # Perform the analysis
-    res <- classify(
-      data, variables, grouping = "habitat", nesting = "island",
-      topcomp = to_pcomp, k = 5, nrep = 100, seed = 24, method = methods[i],
-      importance = TRUE, nperm = 1000
-    )
-
-    # Rename for practical purposes
-    res[[1]] <- rename(res[[1]], island = "nesting")
-
-    # Prepare file names for saving
-    outputname <- sprintf("results/classif_%s_%s.rds", methods[i], datasets[j])
-    tablename <- str_replace(outputname, ".rds", ".csv")
-    confplotname <- str_replace(outputname, ".rds", ".png")
-    confplotname <- str_replace(confplotname, "classif", "confplot")
-    accuplotname <- str_replace(confplotname, "confplot", "accuplot")
-
-    # Save
-    saveRDS(res, outputname)
-    write_csv(res[[1]][, 1:6], tablename)
-    ggsave(confplotname, res[[2]], width = 4.5, height = 4)
-    ggsave(accuplotname, res[[3]], width = 7.5, height = 4)
-
-  }
-}
-
-# Compare results across analyses
-all_res <- map(methods, function(method) {
-  map(datasets, function(dataset) {
-
[…]

### 2021-09-06 analyses/05-assumptions/Assumptions.R

@@ -1,98 +1,0 @@
-rm(list = ls())
-
-# Here we test for the assumptions of (M)ANOVA across different levels
-
-library(nmgc)
-library(dewlap)
-
-data <- read.csv("data/reflectance.csv", header = TRUE)
-wl <- paste0("wl", 300:700)
-variables <- paste0("PC", 1:4)
-
-#### 1. Pooled data -- across the whole archipelago ####
-
-data <- cbind(data, data.frame(npcomp(data, wl)$x)[, variables])
-
-# 1.1. Multivariate normality
-t1 <- test_multinorm(data, variables, grouping = "habitat")[, -1]
-
-# 1.2. Heterogeneity of covariance matrices
-t2 <- test_covariance(data, variables, grouping = "habitat")[, -1]
-
-# 1.3. Univariate normality
-t3 <- test_multinorm(data, variables, grouping = "habitat", univariate = TRUE)[, -1]
-
-# 1.4. Heterogeneity of variances
-t4 <- test_covariance(data, variables, grouping = "habitat", univariate = TRUE)[, -1]
-
-t1_names <- c("Habitat", "$HZ$", "$P$", "")
-t2_names <- c("$\\chi^2$", "df", "$P$", "")
-t3_names <- c("Habitat", "Variable", "$W$", "$P$", "")
-t4_names <- c("Variable", "$K^2$", "df", "$P$", "")
-
-fname <- "analyses/05-assumptions/table_%s"
-t1_fname <- sprintf(fname, "multinorm_pooled")
-t2_fname <- sprintf(fname, "covariance_pooled")
-t3_fname <- sprintf(fname, "normality_pooled")
-t4_fname <- sprintf(fname, "variance_pooled")
-
-save_table(t1, t1_fname, digits = c(0, 2, 0, 0), col.names = t1_names)
-save_table(t2, t2_fname, digits = c(1, 0, 4, 0), col.names = t2_names)
-save_table(t3, t3_fname, digits = c(0, 0, 3, 4, 0), col.names = t3_names)
-save_table(t4, t4_fname, digits = c(0, 2, 0, 4), col.names = t4_names)
-
-#### 2. Within islands ####
-
-# 2.1. Multivariate normality
-t5 <- test_multinorm(data, variables, grouping = "habitat", nesting = "island")
-
-# 2.2. Homogeneity of covariance matrices
-t6 <- test_covariance(data, variables, grouping = "habitat", nesting = "island")
-
[…]

### 2021-09-06 analyses/06-pooled/Pooled.R

@@ -1,170 +1,0 @@
-rm(list = ls())
-
-# Here we explore dewlap coloration between habitats across the whole archipelago
-
-library(dewlap)
-library(nmgc)
-library(GGally)
-library(MANOVA.RM)
-library(heplots)
-library(cowplot)
-
-data <- read.csv("data/reflectance.csv", header = TRUE)
-wl <- paste0("wl", 300:700)
-variables <- paste0("PC", 1:4)
-
-data <- cbind(data, data.frame(npcomp(data, wl)$x)[, variables])
-
-#### 1. Eyeball ####
-
-# We plot the color variables across habitats at the scale of the whole archipelago
-
-colors <- c(coastal = "darkgoldenrod1", coppice = "forestgreen", mangrove = "mediumseagreen")
-alpha <- 0.2
-
-lowerplot <- function(data, mapping, ...) {
-  ggplot(data = data, mapping = mapping) +
-    geom_point(..., alpha = alpha) +
-    stat_ellipse() +
-    scale_color_manual(values = colors)
-}
-
-diagplot <- function(data, mapping, ...) {
-  ggplot(data = data, mapping = mapping) +
-    geom_density(..., alpha = alpha) +
-    scale_fill_manual(values = colors) +
-    scale_color_manual(values = colors)
-}
-
-diagplot2 <- function(data, mapping, ...) {
-  ggplot(data = data, mapping = mapping) +
-    geom_boxplot(..., alpha = alpha) +
-    scale_color_manual(values = colors) +
-    coord_flip()
-}
-
-p <- ggpairs(
-  data, columns = variables, mapping = aes(color = habitat),
-  upper = NULL, lower = list(continuous = lowerplot), diag = list(continuous = diagplot2)
-) +
-  theme_bw()
-
-p
-
-# There seems to be no difference between habitats.
-# To confirm this we run a two-way MANOVA with islands and habitats as effects
-
-#### 2. 2-way MANOVA ####
-
-# 2.1. Parametric MANOVA
-
-res_manova <- manova(cbind(PC1, PC2, PC3, PC4) ~ island * habitat, data = data)
-res_manova %>% summary(test = "Pillai")
-
-# This test suggests that all effects are significant, including an archipelago-wide
-# habitat-effect
-# How big is this effect?
-
-# 2.2. Effect sizes
-
-etasq(res_manova)
-
[…]

### 2021-09-06 analyses/07-ANOVA/ANOVA.R

@@ -1,258 +1,0 @@
-# This pipeline performs the following tests
-# 1. MANOVA and tests of MANOVA assumptions
-# 2. Multiple ANOVAs accounting for heteroskedasticity and posthoc tests
-# 3. Plotting of the results
-# 4. Identification of deviations from normality
-# 5. Repeat tests with Kruskal-Wallis when such deviations occur
-
-rm(list = ls())
-
-library(nmgc)
-library(tidyverse)
-library(cowplot)
-library(knitr)
-library(dewlap)
-
-data <- read.csv("data/reflectance.csv", header = TRUE)
-variables <- paste0("PC", 1:4)
-wl <- paste0("wl", 300:700)
-
-#### 1. Multiple one-way MANOVAs ####
-
-# Are there multivariate differences between habitats on each island?
-
-# 1.1. Multiple parametric MANOVAs
-
-res_manova <- nanova(
-  data, variables, grouping = "habitat", nesting = "island", to_pcomp = wl,
-  univariate = FALSE, test = "Pillai"
-)
-
-res_manova$res # MANOVA table
-
-# It looks like on some islands, yes
-# But we know that multivariate normality and homogeneity of covariance matrices
-# are violated on multiple islands
-read.csv("analyses/05-assumptions/table_multinorm.csv", header = TRUE)
-read.csv("analyses/05-assumptions/table_covariance.csv", header = TRUE)
-
-# 1.2. Multiple semi-paramatric MANOVAs
-
-res_smanova <- nanova(
-  data, variables, grouping = "habitat", nesting = "island", to_pcomp = wl,
-  univariate = FALSE, parametric = FALSE, seed = 42, iter = 1000
-)
-
-t1 <- res_smanova$res
-t1_fname <- "analyses/07-ANOVA/table_smanova"
-t1_names <- c("Island", "MATS", "$P$", "")
-save_table(t1, t1_fname, digits = c(0, 1, 4, 0), col.names = t1_names)
-
-#### 2. Multiple ANOVAs on significant islands ####
-
-# Keep the most significant islands only (based on machine learning)
-sislands <- c("Abaco", "Bimini", "Cayman Brac", "Little Cayman", "Long Island")
-data <- data %>% filter(island %in% sislands) %>% droplevels
-
[…]

### 2021-09-06 analyses/08-spatial-correlation/SpatialAutocorrelation.R

@@ -1,28 +1,0 @@
-# Spatial autocorrelation by permutation test
-
-rm(list = ls())
-
-library(dewlap)
-library(tidyverse)
-library(nmgc)
-
-data <- read.csv("data/reflectance.csv", header = TRUE)
-
-# Spatial autocorrelation test
-res_cortest <- nspcortest(
-  data, variables = paste0("PC", 1:4), nesting = "island", nperm = 1000, seed = 42,
-  to_pcomp = paste0("wl", 300:400), keep = "habitat"
-)
-
-# Save table
-t1 <- res_cortest$res
-t1 <- t1 %>% add_signif()
-t1_fname <- "analyses/08-spatial correlation/table_spatial"
-t1_names <- c("Island", "$r$", "$P$", "$N$", "")
-save_table(t1, t1_fname, digits = c(0, 3, 3, 0), col.names = t1_names)
-
-# Table with multivariate means per site
-t2 <- res_cortest$sites
-t2_fname <- "analyses/08-spatial correlation/table_sites"
-t2_names <- c("Island", "Longitude", "Latitude", "Habitat", "PC1", "PC2", "PC3", "PC4")
-save_table(t2, t2_fname, digits = c(0, 4, 4, 0, 4, 4, 4, 4), col.names = t2_names)

### 2021-09-06 analyses/09-parallelism/Parallelism.R

@@ -1,85 +1,0 @@
-rm(list = ls())
-
-# Analysis of parallel divergence across islands
-
-library(dewlap)
-library(tidyverse)
-library(nmgc)
-library(RColorBrewer)
-
-# Are differences in coloration similar on different islands?
-# To test this, perform a PCA on the whole archipelago
-# Then, measure and plot the pairwise differences between the habitats
-# in color space, along each PC
-# Re-scale these contrasts if needed, because some islands may be more
-# variable than others
-# If islands undergo the same divergence process, then they should cluster
-# together in this contrast space
-
-data <- read.csv("data/reflectance.csv", header = TRUE)
-variables <- paste0("PC", 1:4)
-wl <- paste0("wl", 300:700)
-
-# Perform PCA at the scale of the whole archipelago
-data <- data <- cbind(data, data.frame(npcomp(data, wl)$x)[, variables])
-
-# Compute contrast scores across islands and variables
-newdata <- data %>%
-  group_by(island, habitat) %>%
-  summarize_at(variables, mean) %>%
-  gather_("variable", "score", variables) %>%
-  spread_("habitat", "score") %>%
-  mutate(
-    coa_cop = coastal - coppice,
-    coa_man = coastal - mangrove,
-    cop_man = coppice - mangrove
-  ) %>%
-  ungroup() %>%
-  group_by(variable) %>%
-  mutate_at(
-    c("coa_cop", "coa_man", "cop_man"),
-    function(x) scale(x, center = FALSE, scale = TRUE)
-  ) %>%
-  filter(island %in% c("Abaco", "Bimini", "Cayman Brac", "Little Cayman", "Long Island")) %>%
-  ungroup()
-
-newdata <- newdata %>% mutate(
-  x = coa_cop,
-  y = cop_man
-)
-
-newdata <- newdata %>% rename(vbl = "variable")
-
-p <- ggplot(newdata, aes(x = x, y = y, fill = vbl, shape = island)) +
-  geom_point(size = 5) +
-  xlab("Coastal versus coppice") +
-  ylab("Coppice versus mangrove") +
-  theme_bw() +
-  labs(shape = "Island", fill = "Variable") +
-  scale_fill_manual(values = rev(brewer.pal(length(variables), "YlOrRd"))) +
-  scale_shape_manual(values = 21:25) +
[…]

### 2021-09-06 analyses/10-distances/Distances.R

@@ -1,147 +1,0 @@
-rm(list = ls())
-
-# Here we explore the distances at which differences between sites are detected
-
-library(dewlap)
-library(tidyverse)
-library(nmgc)
-library(RColorBrewer)
-
-data <- read.csv("data/reflectance.csv", header = TRUE)
-wl <- paste0("wl", 300:700)
-variables <- paste0("PC", 1:4)
-data <- cbind(data, npcomp(data, wl, nesting = "island", reduce = 1:4)$x)
-data <- data[, -grep("island", colnames(data))[2]]
-
-# Kepp only the most significant islands
-keep <- c("Abaco", "Bimini", "Cayman Brac", "Little Cayman", "Long Island")
-data <- data %>% filter(island %in% keep)
-data <- data %>% select(island, habitat, site, longitude, latitude, PC1, PC2, PC3, PC4)
-
-# Rearrange the data
-data <- data %>% gather_("variable", "score", variables)
-
-# Function to compare sites within an island for a given variable
-compare_sites <- function(data) {
-
-  data <- data %>%
-    group_by(site, habitat, longitude, latitude) %>%
-    nest()
-
-  ii <- 1:(nrow(data) - 1)
-  names(ii) <- ii
-  out <- map_dfr(ii, function(i) {
-    jj <- (i + 1):(nrow(data))
-    names(jj) <- jj
-    map_dfr(jj, function(j) {
-      res <- wilcox.test(data$data[[i]]$score, data$data[[j]]$score)
-      data.frame(U = res$statistic, pvalue = res$p.value)
-    }, .id = "j")
-  }, .id = "i")
-
-  out <- out %>% mutate_at(c("i", "j"), as.numeric)
-  out <- out %>% mutate(
-    habitat_i = map_chr(i, ~ as.character(data$habitat[.x])),
-    habitat_j = map_chr(j, ~ as.character(data$habitat[.x]))
-  )
-  out <- out %>% filter(habitat_i != habitat_j)
-  out <- out %>% mutate(
-    lon_i = map_dbl(i, ~ data$longitude[.x]),
-    lat_i = map_dbl(i, ~ data$latitude[.x]),
-    lon_j = map_dbl(j, ~ data$longitude[.x]),
-    lat_j = map_dbl(j, ~ data$latitude[.x])
-  )
-  out <- out %>% mutate(
[…]

### 2021-09-06 classify3.R

@@ -1,296 +1,0 @@
-#' Classification analysis
-#'
-#' Perform a replicated classification analysis of a multivariate dataset into categorical labels using machine learning tools and k-fold cross validation
-#'
-#' @param data A data frame
-#' @param variables The variables used to classify
-#' @param grouping Name of the grouping variable (the labels)
-#' @param nesting Optional nesting variable, if the analysis must be conducted separately on different subsets of the data
-#' @param method The data mining model used. Currently supports "SVM" and "LDA".
-#' @param k Number of bins for the k-fold cross-validation procedure
-#' @param nrep Number of replicate analyses (i.e. number of k-fold cross validations)
-#' @param minsize Minimum size required per group for a training data set
-#' @param seed Optional random seed to reset at the beginning
-#' @param importance Whether to perform sensitivity analysis on the input (takes a while)
-#' @param return_machine Whether to return the machines (takes space)
-#' @param verbose Whether to display messages
-#' @param pb Whether to display progress bars
-#' @param to_pcomp Variable to perform PCA on
-#' @param center Center the PCA
-#' @param scale Scale the PCA
-#'
-#' @return A list of two tibbles.
-#'
-#' The first tibble, `full`, contains information about each fitted machine in the analysis, including its nesting level `nesting`, replicate number `repl` and cross-validation bin `kbin` but also classification testing outputs such as its confusion matrix `confmat`, its prediction `accuracy` and the number of observations in the testing set `ntested`. If `return_machine` is TRUE, an extra column is added, that is a list of the actual fitted machines. If `importance` is TRUE, extra columns are added, named after the input variables and containing their respective relative importance scores (these are computed using the `rminer::Importance` function for one-dimensional sensitivity analysis).
-#'
[…]

### 2021-09-06 classify4.R

@@ -1,296 +1,0 @@
-#' Classification analysis
-#'
-#' Perform a replicated classification analysis of a multivariate dataset into categorical labels using machine learning tools and k-fold cross validation
-#'
-#' @param data A data frame
-#' @param variables The variables used to classify
-#' @param grouping Name of the grouping variable (the labels)
-#' @param nesting Optional nesting variable, if the analysis must be conducted separately on different subsets of the data
-#' @param method The data mining model used. Currently supports "SVM" and "LDA".
-#' @param k Number of bins for the k-fold cross-validation procedure
-#' @param nrep Number of replicate analyses (i.e. number of k-fold cross validations)
-#' @param minsize Minimum size required per group for a training data set
-#' @param seed Optional random seed to reset at the beginning
-#' @param importance Whether to perform sensitivity analysis on the input (takes a while)
-#' @param return_machine Whether to return the machines (takes space)
-#' @param verbose Whether to display messages
-#' @param pb Whether to display progress bars
-#' @param to_pcomp Variable to perform PCA on
-#' @param center Center the PCA
-#' @param scale Scale the PCA
-#'
-#' @return A list of two tibbles.
-#'
-#' The first tibble, `full`, contains information about each fitted machine in the analysis, including its nesting level `nesting`, replicate number `repl` and cross-validation bin `kbin` but also classification testing outputs such as its confusion matrix `confmat`, its prediction `accuracy` and the number of observations in the testing set `ntested`. If `return_machine` is TRUE, an extra column is added, that is a list of the actual fitted machines. If `importance` is TRUE, extra columns are added, named after the input variables and containing their respective relative importance scores (these are computed using the `rminer::Importance` function for one-dimensional sensitivity analysis).
-#'
[…]

### 2021-09-06 ms/fetch_content.sh

@@ -1,47 +1,0 @@
-# Run this script to copy selected files from the analysis folders to the manuscript folder
-
-# Fetch figures
-cp ../analyses/04-machine-learning/plots/classif_svm_pca.png figures
-cp ../analyses/07-ANOVA/figure_anova.png figures
-cp ../maps/map.pdf figures
-
-# Fetch supplementary figures
-cp ../analyses/04-machine-learning/plots/classif_lda_pca.png suppfigures
-cp ../analyses/04-machine-learning/plots/classif_lda_pca_pooled.png suppfigures
-cp ../analyses/04-machine-learning/plots/classif_lda_refl.png suppfigures
-cp ../analyses/04-machine-learning/plots/classif_lda_refl_pooled.png suppfigures
-cp ../analyses/04-machine-learning/plots/classif_svm_pca_pooled.png suppfigures
-cp ../analyses/04-machine-learning/plots/classif_svm_refl.png suppfigures
-cp ../analyses/04-machine-learning/plots/classif_svm_refl_pooled.png suppfigures
-cp ../maps/detailed_map.pdf suppfigures
-cp ../analyses/03-PCA/figure_brightness.png suppfigures
-cp ../analyses/03-PCA/figure_brightness_pooled.png suppfigures
-cp ../analyses/09-parallelism/figure_contrasts.png suppfigures
-cp ../analyses/10-distances/figure_distances2.png suppfigures
-cp ../analyses/10-distances/figure_distances.png suppfigures
-cp ../analyses/06-pooled/figure_pooled.png suppfigures
-cp ../analyses/02-reflectance/figure_reflectance.png suppfigures
-cp ../analyses/04-machine-learning/plots/importance_lda_pca.png suppfigures
-cp ../analyses/04-machine-learning/plots/importance_lda_pca_pooled.png suppfigures
-cp ../analyses/04-machine-learning/plots/importance_lda_refl.png suppfigures
-cp ../analyses/04-machine-learning/plots/importance_lda_refl_pooled.png suppfigures
-cp ../analyses/04-machine-learning/plots/importance_svm_pca.png suppfigures
-cp ../analyses/04-machine-learning/plots/importance_svm_pca_pooled.png suppfigures
-cp ../analyses/04-machine-learning/plots/importance_svm_refl.png suppfigures
-cp ../analyses/04-machine-learning/plots/importance_svm_refl_pooled.png suppfigures
-
-# Fetch tables
[…]

### 2021-09-06 ms/run_pandoc.sh

@@ -1,5 +1,5 @@
 # Run this script to convert LaTeX files into ODT and DOCX documents
 # Requires pandoc and pandoc-crossref being installed
 
-pandoc main.tex --filter=pandoc-crossref --bibliography=library.bib --csl=journal-of-evolutionary-biology.csl -o scherrer2021_v3.2.odt
-pandoc main.tex --filter=pandoc-crossref --bibliography=library.bib --csl=journal-of-evolutionary-biology.csl -o scherrer2021_v3.2.docx
+pandoc main.tex --filter=pandoc-crossref --bibliography=library.bib --csl=journal-of-evolutionary-biology.csl -o scherrer2021.odt
+pandoc main.tex --filter=pandoc-crossref --bibliography=library.bib --csl=journal-of-evolutionary-biology.csl -o scherrer2021.docx

### 2021-09-06 mutate_signif.R

@@ -1,26 +1,0 @@
-# Function to add a significance column with asterisk symbols to a tibble
-# (We are going to need that)
-# Use on a tibble that contains P-values from test results in a column
-# Arguments:
-# D: the data frame
-# col: the name of the column containing the P-values
-# levels: the significance levels for the three labels *, ** and ***, in that
-# order
-# dropname: whether to replace the name of the extra column by a blank
-
-mutate_signif <- function(
-  D, col = "pvalue", levels = c(0.05, 0.01, 0.001), dropname = TRUE
-) {
-
-  D <- D %>%
-    mutate(
-      .signif = ifelse(get(col) < levels[1], "*", ""),
-      .signif = ifelse(get(col) < levels[2], "**", .signif),
-      .signif = ifelse(get(col) < levels[3], "***", .signif)
-    )
-
-  if (dropname) D <- D %>% rename(" " = ".signif")
-
-  return(D)
-
-}

### 2021-09-06 scripts/01_run_analyses.R

@@ -1,0 +1,2054 @@
+## ---------------------------
+##
+## Script name: analyses.R
+##
+## Purpose of script: This script contains all the analyses of this study.
+##
+## Author: Raphael Scherrer
+##
+## Date Created: 2021-02-06
+##
+## Copyright (c) Raphael Scherrer, 2021
+##
+## Email:
+## r.scherrer@rug.nl
+## raphael.scherrer@evobio.eu
+## raph.rjfs@hotmail.fr
+##
+## ---------------------------
+
+# Notation used for naming datasets in this script:
+# D = the original dataset
+# X = placeholder within a function
+# P = temporary data frame for plotting or containing plots
+
+rm(list = ls())
+
+library(tidyverse)
+library(ggridges) # ridge plots
+library(broom) # tidy statistical tables
+library(heplots) # Box M test
+library(MVN) # multivariate normality
+library(assertthat) # assertions within functions
+library(rminer) # machine learning
+library(geosphere) # geographical distances
+library(MuMIn) # AICc
+library(nlme) # univariate ANOVAs
+library(PMCMRplus) # posthoc tests
+library(RColorBrewer) # color palette
+library(pairwiseAdonis) # posthoc PERMANOVA
+library(scales) # color palette
+library(vegan) # PERMANOVA
+library(ape) # Mantel test
+
+#### 1. Set-up ####
+
+# Set plotting theme
+theme_set(theme_classic())
+
+# Graphical parameters
+island_colors <- brewer.pal(8, "Set2")
+island_colors <- c(island_colors, "darkgreen")
+habitat_colors <- c("goldenrod", "forestgreen", "mediumseagreen")
+
+# Global settings
+pc_names <- paste0("PC", 1:4)
+wpc_names <- paste0("w", pc_names)
+wl <- 300:700
+wl_names <- paste0("wl", wl)
+
+# Whether to load machines already fitted to whole spectrum data
+# (instead of rerunning the very lengthy classification)
+read_fitted <- TRUE
+
+#### 2. Read reflectance data ####
+
+# Read the data
+D <- read_csv("data/reflectance.csv")
+
+# Useful things to keep in mind
+island_names <- unique(D$island)
+habitat_names <- unique(D$habitat)
+
+# Save sample sizes
+with(D, table(island, habitat)) %>%
[…]

### 2021-09-07 ms-jeb/main_cover.tex
almost there

\documentclass[12pt]{letter}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

\begin{document}

	\thispagestyle{empty}

	\begin{flushright}
		Rapha\"{e}l Scherrer\\
		Groningen Institute for Evolutionary Life Sciences\\
		Nijenborgh 7, 9747AG\\
		Groningen, the Netherlands\\
		raphael.scherrer@evobio.eu\\[5ex]
	\end{flushright}

	\textbf{Submission of a manuscript to \textit{Journal of Evolutionary Biology}}\\

	Dear Editor,\\

	I am writing to express our wish to submit our manuscript ``Dewlap color variation in \textit{Anolis sagrei} is maintained among habitats within islands of the West Indies'' for publication as an original research article in \textit{Journal of Evolutionary Biology}.

	Combining reflectance spectrometry and machine learning, we investigated the patterns of color variation in the dewlap of one species of \textit{Anolis} lizard across habitats spanning multiple islands in the West Indies, testing the idea that differences in coloration may be linked to environmental conditions and habitat-types. We found that coloration indeed differed between different habitats, sometimes in close proximity. However, these differences were inconsistent across islands, suggesting that the diversity of this communication signal is maintained between habitats within islands, but that adaptation to habitat-type is unlikely to explain this pattern.

	Our work builds on previous studies published in \textit{J. Evol. Biol.}, such as Vanhooydonck et al. (2009, DOI 10.1111/j.1420-9101.2008.01643.x) and Driessens et al. (2017, DOI 10.1111/jeb.13144) who investigated similar questions in the same study species, albeit by comparing populations from different islands and not habitats within islands. By doing so, we hope that our study will constitute a logical and complementary follow-up to theirs for the readership of \textit{J. Evol. Biol.}.

	This manuscript has not been published before and is not currently under consideration by another journal. All authors have agreed to submit this manuscript to \textit{J. Evol. Biol.}, and declare no conflicts of interest.\\

	Yours sincerely,\\

	Rapha\"{e}l Scherrer

\end{document}

### 2021-09-07 ms/main_title.tex
submitted

@@
     \textsc{Jonathan B. Losos}$^{1,4}$ (losos@wustl.edu, 0000-0003-4712-8258) and\\
-	\textsc{Anthony J. Geneva}$^{1,5}$ (anthony.geneva@gmail.com, 0000-0001-8253-6527)\\[1ex]
+	\textsc{Anthony J. Geneva}$^{1,5}$ (anthony.geneva@rutgers.edu, 0000-0001-8253-6527)\\[1ex]
 
@@
 
-	\textbf{Running title} --- Dewlap coloration in \textit{Anolis sagrei}.
+	\section*{Running title}
+
+	Dewlap coloration in \textit{Anolis sagrei}.
+
+	\section*{Acknowledgements}
+
+	\input{sections/acknowledgments.tex}
+
+	\section*{Author Contributions}
+
+	CMD, RGR, AJG and JBL designed the study.
+CMD, RGR and AJG organized the fieldwork and collected the data.
+RS performed the analyses and wrote the manuscript, under the supervision of CMD, RGR and AJG.
+All authors contributed to revisions of the manuscript.
+JBL provided the resources.\\
+
+	The authors have no conflict of interest to declare.
 

### 2021-12-10 ms/response_JEB_20211024.txt
answered reviewer comments

EDITOR. One other more minor point is that lack of evidence for spatial autocorrelation is not the same as evidence for a lack of spatial autocorrelation (indeed there is a reasonably strong, albeit not quite "significant" spatial correlation on Eleuthera), so some of the conclusions about a lack of isolation by distance and genetic drift should be toned down slightly.

(R) We changed the whole discussion paragraph on the effect of drift accordingly (new citations were added there), and also a sentence in the results to point out that Eleuthera had almost significant spatial correlation. We also toned down that statement in the abstract. Re-read this paragraph later on (it could benefit from info about the cruising range of an anolis). Note that the citations have been updated, with new ones accompanying the requested changes by the reviewers.

EDITOR. How similar are the habitats between islands or could differences in habitat between islands explain some of the variation between islands?

(R) We do not have these data unfortunately (do we not? ask Colin).

REVIEWER1. Do you know whether such colour variation nicely described by your analyses, is actually detectable by conspecific, congeners and/or predators, the “agents” of selection?  For instance, do you think all dewlaps will be similarly seen under similar light/background conditions? Or whether observers can tell apart dewlaps coming from different habitats? In the first case, it is likely that the non-directional variation that you describe is not under selection, as to observers’ eyes, all dewlap colours look similar. In the second case, the conclusions will be similar to the ones you get with your current approach. This question can be tested applying vision modelling using your reflectance profiles, and an irradiance profile of at least one of the habitats that you studied. Perhaps this has been already done in other studies. If such is the case, you can clearly mention it.

(R) We do not know how detectable the reported differences in coloration are. Indeed, to know this we would need to do some visual modelling, and for this we would also need some habitat irradiance data, which we do not have unfortunately. Visual modelling has been performed in Anolis lizards, but to the best of our knowledge it has not been done on our study populations, and so we do not know if the lizards can detect the differences in coloration of conspecifics from different habitats. Maybe a discussion paragraph about that?

REVIEWER1. Among the differences that can be found on the three studied habitats, authors mostly talk about differences in light incidence. Are there any differences also in background colour? Predator community? Anole community? If so, how will they contribute or not to the obtained results?

(R) Background color definitely varies between habitats, which can be seen with the naked eye from photos, but we did not quantify this precisely (we do not have irradiance profiles for the habitats telling us which wavelengths are different). Anole and predator communities have been shown to vary between islands (at least in density of the main species, which are overall present across all islands), but we do not know whether they also vary between the habitats within islands.

REVIEWER1. How big is the home range of an Anolis? Sorry of I missed it. This information will help understanding the importance of your experimental design

(R) Ambika's paper reports wider territories than previously thought in A. sagrei: 225m2 for males (which is 15 by 15m), and 36m2 for females (6 by 6m), with with evidence for multiple matings in both sexes. These estimates were measured on a campus in Florida and may not represent the real territory size on our study islands, however. Nevertheless, although the species is still classified as territorial, its ability for movement and promiscuity, its continuous distribution at high densities throughout the islands and the polyphyly of individuals from different habitats within islands suggests that long-range gene flow is highly plausible. (Maybe the guys will know something about the cruising range of an anolis...)

REVIEWER1. In line 91, what do you mean by climatic variables? Light incidence?

(R) We have now precised that these climatic variables include precipitations and solar incidence, among others, and those were interpreted as causing differences in types of habitat, with different light environments, among islands.

REVIEWER1. In line 413, perhaps is worth it including a short definition of runaway sexual selection.

(R) This is now clarified.

REVIEWER2. Of course, also the size and use of the dewlap is essential during signalling (see e.g. papers by Lailvaux, Irschick, Vanhooydonck, Driessens, Husak, Baeckens, to name a few). While I understand this is not the goal of the study, I think it’s still important to invest a couple of sentences or a short paragraph on how other characteristics of the dewlap (aside from colouration) are important too, and that those characteristics (possibly interaction with colouration) might vary within- and among-islands as well.

(R) We have modified the second-to-last paragraph to mention those other facets of anole display behavior, and added new references to back them up.

### 2021-12-10 ms/sections/abstract.tex
answered reviewer comments

@@
 Previous studies have demonstrated that dewlap color in \textit{Anolis} lizards can be highly variable between populations in relation to both biotic and abiotic adaptive drivers at relatively large geographical scales.
-Here, we investigated differentiation of dewlap coloration among habitat-types at a small spatial scale, within multiple islands of the West Indies, to test the hypothesis that similar local adaptive processes occur over smaller spatial scales.
+Here, we investigated differentiation of dewlap coloration among habitat types at a small spatial scale, within multiple islands of the West Indies, to test the hypothesis that similar local adaptive processes occur over smaller spatial scales.
 We explored variation in dewlap coloration in the most widespread species of anole, \textit{Anolis sagrei}, across three characteristic habitats spanning the Bahamas and the Cayman Islands, namely beach scrub, primary coppice forest and mangrove forest.
 Using reflectance spectrometry paired with supervised machine learning, we found significant differences in spectral properties of the dewlap between habitats within small islands, sometimes over very short distances.
-Passive divergence in dewlap phenotype associated with isolation-by-distance did not explain our results.
+Passive divergence in dewlap phenotype associated with isolation-by-distance did not \textcolor{olive}{seem to} explain our results.
 On the other hand, these habitat-specific dewlap differences varied in magnitude and direction across islands, and thus our primary test for adaptation -- parallel responses across islands -- was not supported.

### 2021-12-10 ms/sections/discussion.tex
answered reviewer comments

@@
-\paragraph{Dewlap coloration differs between habitats, but the differentiation is inconsistent with adaptive convergence} Dewlaps are highly conspicuous in anoles and used as communication signals in courtship \citep{Sigmund1983, Driessens2014, Driessens2015} and territorial displays \citep{Losos1985, Macedonia1994, Macedonia2013}.
-On five of the nine islands we sampled in the West Indies, we found that \textit{A.
-sagrei} male dewlap coloration significantly varied between individuals living in three different habitat types  (beach scrub bush, primary coppice forest and mangrove forest), even though those habitats were in close proximity to each other and the \textit{sagrei} populations are continuously distributed between the habitats.\\
+Two main insights follow from our results.
+First, we detected significant differences in dewlap coloration between habitats within seven out of the nine islands investigated (excluding North Andros where the follow-up univariate analyses were not significant), suggesting a putatively high potential for local differentiation of dewlap coloration in \textit{Anolis sagrei}.
+Second, we found differences in coloration along different dimensions of color space, and in different directions, on different islands.\\
 
-One compelling indicator of adaptation by natural selection is the repetition of patterns of differentiation across localities \citep{Losos2009, Losos2011}.
-In the context of this study, had we found consistent patterns of dewlap differentiation between habitats, we would have concluded that there was likely an adaptive basis for this pattern.
-However, the inconsistent and idiosyncratic patterns of dewlap variation we observed contradict this prediction.
-What then explains the observed differences among the populations?\\
+Detectable differences in dewlap color between populations are surprising, as habitats were often in close geographical proximity to each other (as close as a few hundred meters on Bimini and most of the time within ten kilometers).
+Indeed, given that (1) the populations were continuously distributed between the habitats, (2) populations from different habitats were not monophyletic with respect to mitochondrial haplotypes (\citealt{vandeSchoot2016} unpublished thesis), and (3) \textit{A.
+sagrei} is highly mobile \citep{Kamath2018}, we would have expected more homogeneous distributions of color phenotypes within islands due to gene flow, with fewer differences between populations, especially those in close proximity.\\
 
-One explanation is that this pattern is the result of multiple island colonizations of phenotypically different individuals to these habitats.
-However, we reject this possibility because all of the island populations in this study are strictly monophyletic, reflecting a single colonization event per island (van de Schoot, unpublished thesis; \citealt{Driessens2017, Reynolds2020}).
-Other drivers could be at work, including genetic drift, phenotypic plasticity, and divergent natural selection.
-Of these, we find divergent natural selection the most likely explanation, suggesting that future research with additional granularity on the drivers of dewlap coloration is needed.
+Several scenarios could account for these findings.
+One explanation is an adaptive one: populations living in different habitats could be phenotypically adapted to their local environmental conditions.
+Given that the brightly colored dewlap of \textit{A.
+sagrei} is used as a communication signal, its color may be a target for selection if the transmission or perception of the signal differs from one habitat to another, for example because of differences in ambient light, according to the sensory-drive hypothesis \citep{Endler1988, Endler1992, Endler1998}.
+The sensory-drive hypothesis has been tested multiple times for dewlap coloration in \textit{Anolis} lizards, with mixed results.
+Some authors found support for it \citep{Leal2002, Leal2004}, while others found differences in dewlap coloration between habitats inconsistent with the sensory-drive hypothesis \citep{Fleishman2009, Ng2012}.\\
 
-\paragraph{A role of neutral drift is unlikely} Genetic drift may lead to patterns of trait divergence similar to local adaptation \citep{Miles2019}.
-We think this scenario is implausible, however, as we found significant differences in dewlap color over distances too small for genetic drift to have realistically counteracted the effect of homogenizing gene flow \citep{Richardson2014}.
-On Bimini, for example, we detected differences between sites only a few hundred meters away and \textit{A.
-sagrei}, have previously been shown to be highly mobile \citep{Kamath2018}.
-Moreover, habitat-populations within each island were found to be non-monophyletic and often share identical haplotypes, based on phylogenetic analysis of mitochondrial DNA (van de Schoot et al.
-unpublished thesis).
-This is further evidence of gene flow between habitats.\\
+If our results were an example of sensory drive, we would have expected to see consistent differences between populations from different habitats across islands, given the apparent environmental consistency each of the three habitat types across the islands we sampled.
+In particular, we would have expected divergence in line with increased detectability given local light conditions, such as the high contrasts with background vegetation found in the UV range in \citet{Leal2002} and \citet{Leal2004}.
+We might also have expected mangrove and beach scrub lizards, both inhabiting areas with high light penetration, to have more similar dewlaps, and to differ significantly from lizards from the coppice habitat, where irradiance is low.
+Instead, we found inconsistencies in the way dewlap color differed between habitats across islands.
+While short-wavelengths (UV reflectance) were often involved in color differences, they were not involved on all islands where significant differences were detected.
+On some islands, other or additional variables differed, such as brightness, red reflectance or the reflectance at the ends of the spectrum visible to \textit{Anolis} lizards (UV and red, \citealt{Lazareva2012}) relative to intermediate wavelengths (blue-to-yellow).
+Similar portions of the spectrum were sometimes involved in opposite directions on different islands, such as on Abaco and Cayman Brac, where mangrove lizards had a higher UV-reflectance than beach scrub lizards on the former, but a lower UV-reflectance on the latter.
+Overall, the observed heterogeneity of divergence patterns across islands provides no support to a sensory-drive explanation.\\
 
-It is possible that complex drift scenarios other than simple distance-mediated, diffusion-like processes may be at play \citep{Miles2019}.
-For example,  we did detect a significant signal of isolation-by-distance on Eleuthera, without finding differences in dewlap coloration between habitats, suggesting a possible role of drift on this island.
-Nonetheless, by and large our results align with previously documented cases of highly localized  dewlap color divergence despite gene flow in multiple species of anoles.
-\citet{Ng2012} and \citet{Ng2016} found divergent dewlap coloration despite gene flow between subspecies of \textit{A.
-distichus} on Hispaniola, and proposed this as a mechanism of reproductive isolation in the early stages of speciation \citep{Ng2011, Lambert2013, Ng2017}.
-\citet{Stapley2011} found that dewlap color polymorphism was maintained in the absence of genetic structure between populations of \textit{A.
-apletophallus} from Panama.
-\citet{Thorpe2002a} found that divergence in dewlap coloration matched habitat-type better than mitochondrial lineage in \textit{A.
-roquet} on Martinique, and a convergent pattern was found in \textit{A.
-trinitatis} on the featureless island of St Vincent \citep{Thorpe2002b}.
-Finally, regionally-distinct body coloration, but not dewlap coloration, is present in \textit{A.
-conspersus} on another small island, Grand Cayman, where no physical barriers to gene flow exist \citep{Macedonia2001}.
+It is presently not known if the reported differences in coloration have a genetic basis.
+Yet, we find it unlikely that these differences arose through phenotypic plasticity, as although the carotenoids that partly make up the red and orange colors of anole dewlaps must be found in the diet \citep{Goodwin1984, Hill2002, Hill2006}, studies testing the effect of carotenoid deprivation \citep{Steffen2010, Ng2013} and heritability \citep{Cox2017} of dewlap coloration in \textit{A.
+sagrei} and \textit{A.
+distichus} (another species with a carotenoid-based dewlap), found little support for phenotypic and developmental plasticity in dewlap coloration.
+One exception is a study demonstrating that lizards heavily parasitized by skin mites had duller dewlaps \citep{Cook2013}, but we found no sign of that in our study.\\
 
-\paragraph{A role of phenotypic plasticity is unlikely} Differences in coloration between habitat populations may not be genetically determined and instead may be influenced by environmental factors.
-The yellow, orange and red coloration in anoline dewlaps are produced by pterins and carotenoids \citep{Ortiz1962, Ortiz1962a, Ortiz1963, Ortiz1966, Macedonia2000, Steffen2007, Steffen2009}.
-Animals lack the ability to synthesize carotenoids, and those must therefore be found in the diet, while pterins are synthesized from nucleotides \citep{Goodwin1984, Hill2002, Hill2006}.
-However, experimental manipulation of dietary carotenoid content showed no effect on dewlap coloration in \textit{A.
-sagrei} \citep{Steffen2010} nor in \textit{A.
-distichus} \citep{Ng2013}, which also has an orange-based dewlap.
-(Dewlap color has also been shown to change in relation to parasite load \citep{Cook2013}, though parasites were not investigated in this study.
--- to remove) These laboratory results  suggest a plastic response in dewlap color to differences in diet across habitats unlikely, suggesting the differences we observed could therefore be genetically based.
-This hypothesis is further supported by \citet{Cox2017}, who found a high degree of heritability of dewlap coloration in \textit{A.
-sagrei}.
+\sout{We found no evidence for a role of genetic drift in explaining the observed patterns either.
+First, \textit{A.
+sagrei} was distributed across the islands continuously, usually at relatively high population densities, rather than in small and isolated populations where drift might be expected to have a strong effect.
+Second, we found no evidence of isolation-by-distance except on Abaco, which was sampled at the largest geographical scale, with sites nearly a hundred kilometers apart from each other.
+Hence, while isolation-by-distance may explain long-range differences on this island, most of the differences among habitats across the rest of the sampling region are unlikely to be explained by genetic drift, as habitats were often in close proximity (less than 10km).}\\
 
-\paragraph{Sensory Drive - maybe drop?} One hypothesis that a relationship would exist between dewlap color and habitat stems from the idea that the communication signals evolve adaptively in response to light environment (the sensory drive hypothesis, \citealt{Endler1988, Endler1992, Endler1998}).
-However, we find this explanation unlikely because the differences we observed were both inconsistent among islands and inconsistent with predictions of the sensory drive hypothesis.
-Previous studies have proposed that dewlap coloration may have evolved to be maximally detectable under local light conditions, primarily through UV contrast (i.e.
-UV-brighter dewlaps in UV-dark, mesic habitats and UV-darker dewlaps in UV-bright, xeric habitats), in \textit{ A.
-cristatellus} and \textit{A.
-cooki} from Puerto Rico \citep{Leal2002, Leal2004}.
-On the contrary, we found no apparent habitat-dependent maximization of UV-contrast, or just any contrast in \textit{ A.
-sagrei}.
-Instead, we found for example the darkest dewlaps in the dark, mesic habitat -- primary coppice forest -- on three islands, and dewlaps often differed the most between beach scrub and mangrove forest, two xeric habitats with similar, high irradiance levels \citep{Howard1950, Schoener1968}.
-Studies of Jamaican and Hispaniolan anoles similarly found between-habitat differences in dewlap coloration but no evidence for higher dewlap detectability in different habitats \citep{Fleishman2009, Ng2012}.
-Our data are consistent with those previous results in suggesting that adaptation to local light conditions, or at least broad habitat types, is not a major driver of the within-island variation in dewlap coloration in \textit{A.
-sagrei}.
+\textcolor{olive}{Genetic drift may contribute to some of the observed variation.
+Indeed, while only Abaco showed significant patterns consistent with isolation-by-distance (which may emerge under limited dispersal and drift, \citealt{Wright1943, Kimura1964, Slatkin1987}), there may have been too few sites on most islands to conclusively detect it, and even then, the absence of detectable isolation-by-distance may not necessarily constitute evidence for the absence of drift.
+Besides, spatial autocorrelation was the strongest on islands sampled at the largest scales (e.g.
+some sites on Abaco were nearly 100km apart, and Eleuthera -- the second strongest signal, albeit nonsignificant -- had sites more than 30km apart), such that it is possible that neutral processes and/or dispersal limitations might in fact contribute to shaping variation over long distances.
+That said, many significant differences were found between habitats in close proximity, contrary to what would be expected under isolation-by-distance, including on islands where spatial autocorrelation was detected.
+Moreover, \textit{A.
+sagrei} was distributed across the islands continuously, usually at relatively high population densities, rather than in small and isolated populations particularly prone to drift.
+Together with the fact that isolation-by-distance may not necessarily only emerge from drift (e.g.
+if there is a spatial environmental gradient), this indicates that genetic drift may have limited potential to explain the differences observed between habitats, at least at a local scale.}\\
 
-\paragraph{Fisherian sexual selection - maybe drop?} Selection, however, needs not necessarily be linked to habitat type, and may take the form of arbitrary, "Fisherian" sexual selection, where female preferences differ between localities for reasons unrelated to the environment, driving the divergent evolution of male ornaments \citep{Lande1981, Andersson1994, Higashi1999}.
-This process could account for the idiosyncratic patterns of within-island divergence we report, where initial differences in female preferences could have arisen for nonselective reasons (e.g.
-plasticity or random drift).
-Substantial levels of promiscuity in \textit{A.
-sagrei} suggest ample opportunity for female mate choice \citep{Kamath2018}, and are in line with this scenario.
-However, \citep{Baeckens2018} found no link between \textit{A.
-sagrei} dewlap coloration and size dimorphism (a proxy for sexual selection) in an among-island study of the same archipelagos.
+In this study, we found larger differences among than within islands, a pattern already reported and linked to climatic conditions \citep{Driessens2017} and to densities of predators and of anole congeners \citep{Vanhooydonck2009, Baeckens2018}.
+Differences among habitats within islands, however, are still difficult to account for.
+Remaining hypotheses may include, for example, runaway sexual selection \textcolor{olive}{(i.e.
+arbitrary preferences of females for some colors over others, \citealt{Andersson1994})} \sout{arbitrarily} operating in different directions across islands, but no evidence so far suggests that dewlap is a target of mate choice in anoles \citep{Tokarz2002, Tokarz2005, Lailvaux2006, Nicholson2007}.
+Another hypothesis is that the different genetic constitutions of different islands, perhaps resulting from founder effects (the islands have been colonized independently, \citealt{vandeSchoot2016} unpubl.; \citealt{Driessens2017, Reynolds2020}), may have predisposed populations to adapt differently to similarly selective circumstances.
+\textcolor{olive}{That said, signalling can be a multifacetted behavior, and while we focused on dewlap color here, other potentially important elements of behavioral interactions include dewlap size, dewlap patterning, display activity as well as behaviors associated with dewlap extensions such as headbobs and pushups \citep{Vanhooydonck2005, Driessens2014, Driessens2015, Lailvaux2015}.
+These have been linked in various ways (and not always consistently across studies) not only with environmental variables but also with proxies for sexual selection, predation and species recognition \citep{Vanhooydonck2005, Lailvaux2007, Vanhooydonck2009, Driessens2017, Baeckens2018}, which do vary among islands but may also vary within islands, and possibly in interaction with dewlap color.} Either way, new data would be needed to test these hypotheses.\\
 
-\paragraph{Natural selection may proceed differently across islands and habitats} A further potential explanation for the inconsistencies in dewlap coloration among habitats between islands is that the observed differences in dewlap color are the result of divergent natural selection.
-If that is the case, however, either the habitats we have identified are not as similar as we thought, such that the same habitat type on different islands favors different colors, or, alternatively, the selective pressures causing dewlap divergence are related to some environmental factor other than habitat type that we did not measure.
-The habitats on different islands may, for example, differ in densities of predators or congeners, which have been shown to affect among-island dewlap diversity \cite{Vanhooydonck2009, Baeckens2018}.
-In particular, \citet{Baeckens2018} recently showed that dewlaps with spotted patterns occurred more often in \textit{A.
-sagrei} on islands with more coexisting species of anoles.
-Therefore, if local adaptation occurs, it may be more likely to involve components of the environment that do not encompass our broad habitat categories leading to island-specific differences between habitats rather than parallel differences expected if habitat identity was the primary selective pressure driving divergence.\\
-
-One last possibility to explain idiosyncratic patterns of between-habitat differences we observe is that, if island populations are different in color on average, they may be experiencing different portions of a within-island fitness landscape whose shape changes depending on the starting phenotype.
-In this study, we indeed found that dewlap coloration differed much more between than within islands.
-So, a population may not respond to selection in the same manner depending on its starting conditions.
-This is true not only in terms of phenotype (average dewlap color differs among islands) but also genetic composition (e.g.
-independent colonizations may have important genetic founder effects, \citealt{Reynolds2020}), and so the selective pressures applying to populations with different starting phenotypes may drive these divergent outcomes.
-For example, \citet{Fleishman2020} found that yellow stimuli are less detectable than red stimuli in high-light environments while both colors are readily detectable in low-light environments.
-Consequently, the environmental pressures applying to yellower dewlaps may differ than those for redder dewlaps, as red dewlaps should be always equally or more detectable than yellow dewlaps across light environments, whereas yellow dewlaps could experience more directional selection towards more red when in a high-light environments.
-
-\paragraph{Conclusion} We identified patterns of dewlap color differentiation in populations of \textit{A.
-sagrei} from different habitat-types on multiple small islands of the West Indies.
-However, our main hypothesis to explain our findings -- local adaptation and sensory drive -- received little support from our multiple-island dataset.
-We also found other mechanisms such as drift or plasticity unlikely candidates to explain the observed patterns.
-This suggests that combinations of multiple factors interacting in subtle ways, and not necessarily in parallel across islands, may have contributed to present-day dewlap color diversity.
+Altogether, our results show that dewlap color of \textit{A.
+sagrei} commonly varies between habitat types, even in close geographical proximity, within islands of the West Indies.
+However, coloration differs in different ways across similar habitats from one island to another.
+We discussed several non-mutually exclusive mechanisms that could explain these observations.
+Nevertheless, heterogeneous patterns of divergence across islands do not support an adaptive sensory-drive scenario, and we propose that within-island dewlap color variation may be underlain by a more subtle mosaic of factors.

### 2021-12-10 ms/sections/introduction.tex
answered reviewer comments

@@
-% Animal signals may be shaped by their environment
-
 The staggering diversity of animal communication signals has long been of interest to evolutionary biologists.
@@
 This selective pressure may drive the local adaptation of communication signals.\\
-
-% One main barrier to the maintenance of diversity is gene flow
 
@@
 It has been the subject of numerous studies concerning local adaptation \citep{Losos1994, Losos1997a, Losos2001, Kolbe2012}, biological invasion \citep{Kolbe2008}, and sexual selection \citep{Tokarz2002, Tokarz2005, Tokarz2006, Driessens2014, Steffen2014, Driessens2015} among many other topics.
-Between-island variation in the mainly orange-red color of its dewlap was shown to be better explained by climatic variables \citep{Driessens2017} than by proxies for biotic factors such as sexual selection or predation pressure \citep{Vanhooydonck2009, Baeckens2018}.
+Between-island variation in the mainly orange-red color of its dewlap was shown to be better explained by climatic variables \textcolor{olive}{such as annual precipitation and solar radiation (proposed to affect the average vegetation type on each island and among other things, its ambient light environment, \citealt{Driessens2017})},  than by proxies for biotic factors such as sexual selection or predation pressure \citep{Vanhooydonck2009, Baeckens2018}.
 How intra-island differences in habitat may contribute to the diversity of dewlap coloration, however, remains unexplored, and may reveal new insights into the scale of local differentiation despite gene flow.\\
-
-% What we did
 
@@
 Finally, if habitat characteristics are strong determinants of dewlap color variation, similar patterns should be observed across multiple islands \citep{Harvey1991, Losos2011}.
-
-% Jonathan does not like when we sum up the results at the end of the intro
-
-% We found strong support for fine-scale, within-island differences in coloration between lizards inhabiting the three habitat-types in several color space dimensions, suggesting a potentially strong effect of divergent selection.
-However, the divergence patterns we observed did not match our \textit{a priori} predictions and were inconsistent between islands.
-We found no evidence of isolation-by-distance as an explanation for the observed differences.
-Our results are nevertheless consistent with small-scale adaptive maintenance of signal polymorphism despite presumed considerable opportunity for gene flow, and might suggest that idiosyncracies of local drift, selection, and gene flow contribute to differing outcomes in different populations.
-
-% Bibliography notes:
-
-% Anolis sagrei was shown to be able to rapidly invade available ecological niches upon colonization of new islands through changes in limb proportions (Losos1997a; Losos2001)
-
-% Convergence
-
-% Besides, \textit{Anolis} lizards are well-known examples of eco-morphological evolutionary convergence, where multiple ``ecomorphs'' adapted independently, but in similar ways, to similar ecological niches across islands of the Greater Antilles (Losos2004; Losos2009).
-This extent of evolutionary determinism seems to act within species at present too.
-Populations of \textit{A.
-sagrei} were shown to undergo the same eco-morphological divergence upon colonization of new islands as ecomorphs did millions of years before (Losos1997a; Losos2001; Calsbeek2007; Losos2007).
-Strong developmental constraints on morphology seem to explain this tendency (Sanger2012; McGlothlin2018).
-However, despite the evidence suggesting potential for evolutionary convergence at the intra-specific level in anoles, this pattern is unknown for communication signals such as dewlap coloration.
-Populations of a same species, because they share a common genetic architecture, are expected to evolve convergent adaptations more often than distantly related species (Ord2015), and this could apply to the dewlap of \textit{A.
-sagrei}.
-On the other hand, sexual signals are often evolutionarily labile due to sexual selection (Kraaijeveld2011), but also selection mediated by other recipients of the signal (Endler1981; Endler1988; Endler1984; Endler1993; Endler1993a; Endler1998).
-This is consistent with the high diversity of dewlaps across anole species (Nicholson2007) and could imply evolutionary contingency for this colorful trait.\\

### 2021-12-10 ms/sections/results.tex
backup
answered reviewer comments

--- 2020-07-31 ms/sections/results.tex
@@
- % Little but detectable differences on five islands
+We tested for variation in \textit{A.
+sagrei} dewlap coloration between populations living in three characteristic habitat types across nine islands that span the West Indian range of the brown anole (beach scrub, primary coppice and mangroves).
+We found that most of the variation in coloration was partitioned between islands (two-way PERMANOVA, $F(df = 8) = 43.7$, $P = 0.0001$, explained variance $R^2 = 40.9$\%).
+Nonetheless, we did find evidence for differences in dewlap coloration between habitat types, and those were mostly island-specific (habitat-by-island interaction term, $F(16) = 3.53$, $P = 0.0001$, $R^2 = 6.6$\%), with a significant portion of the variation explained by an habitat effect across all islands, but this effect was relatively small ($F(2) = 4.7$, $P = 0.0001$, $R^2 = 1.1$\%).\\
 
-We tested for variation in \textit{A.
-sagrei} dewlap coloration between populations living in three characteristic habitat types across nine islands that span the West Indian range of the brown anole (Fig \ref{fig:map}, \ref{supfig:map}).
-We found that most of the variation in coloration is partitioned between islands (two-way semi-parametric MANOVA, modified ANOVA-type statistic (MATS) = 2009.6, P < 0.001, Fig.
-\ref{supfig:reflectance}, explained variance $\eta^2 = 44.3$\%, MANOVA approximation).
-Nonetheless, we did find evidence for differences in dewlap coloration between habitat-types, and those were mostly island-specific (habitat-by-island interaction term, MATS = 384.4, P < 0.001, explained variance $\eta^2 = 11.4$\%), with a small but significant portion of the variation explained by an archipelago-wide habitat effect (MATS = 42.5, P = 0.001, $\eta^2 = 4.8$\%).\\
+We subsequently tested for differences in dewlap coloration between habitat-populations within each island, using within-island principal component scores (to maximize the variation captured for each island, see Methods).
+Our within-island random forest classification analyses revealed detectable differences in dewlap coloration on eight out of the nine islands in our sample: Abaco, Bimini, Cayman Brac, Eleuthera, Little Cayman, Long Island, North Andros and South Andros.
+The accuracy of random forest classification exceeded random expectation more often than expected by chance for all these islands (Table \ref{tab:randomforests}).
+Accuracy was as high as 73\% for Cayman Brac.
+We obtained similar results using other machine learning approaches such as support vector machines (Table \ref{tab:ksvms}) and linear discriminant analysis (Table \ref{tab:ldas}), except that these methods did not detect significant differences on Eleuthera and North Andros.
+We describe in details the specific differences detected on each island in the Appendix, and focus here on the general patterns emerging from our data.\\
 
-The small archipelago-wide effect of habitat-type was detected for PC1, PC2 and PC3 (mixed-effect ANOVA with island as a random effect, Table \ref{suptab:anova-pooled}), but this effect was too small for \textit{post hoc} tests to find which habitats differed.
-Archipelago-wide differences in dewlap coloration between habitats were also detected by SVMs trained on pooled data regardless of island identity, both for PCA data and reflectance scores (Fig.
-\ref{supfig:classif-svm-pca-pooled}, \ref{supfig:classif-svm-refl-pooled}).
-This pattern seemed to be driven by mangrove lizards being correctly reassigned more often than predicted by chance.
-Sensitivity analyses on these machines suggest a relatively small role of long wavelengths (red reflectance) in driving this pattern (Fig.
-\ref{supfig:importance-svm-refl-pooled}), but did not reveal strong differences between the PCs in relative importance (Fig.
-\ref{supfig:importance-svm-pca-pooled}).
-Archipelago-wide differences were not detected by LDA classifiers at all (Fig.
-\ref{supfig:classif-lda-pca-pooled}, \ref{supfig:classif-lda-refl-pooled}).\\ %So, the differences could be in something other than means.\\
+Overall, we found significant differences in dewlap coloration between populations that were often in close geographical proximity.
+On Bimini, notably, we found a significant difference between dewlaps from beach scrub and primary coppice forest, at a distance of a few hundred meters, making this contrast the smallest geographical scale at which differences in coloration were found in our study (Fig.
+\ref{fig:Bimini}).
+We also detected significant differences in dewlap coloration at distances below one kilometer on Abaco (Fig.
+\ref{fig:Abaco_supplement}G), and at distances between one and ten kilometers on Bimini (Fig.
+\ref{fig:Bimini}G), Cayman Brac (Fig.
+\ref{fig:CaymanBrac}G), Little Cayman (Fig.
+\ref{fig:LittleCayman}G), Long Island (Fig.
+\ref{fig:LongIsland}G) and South Andros (Fig.
+\ref{fig:SouthAndros}G).\\
 
-Within islands, SVM classifiers correctly assigned individuals to their habitat of origin based solely upon dewlap coloration on five islands: Abaco, Bimini, Cayman Brac, Little Cayman, and Long island (Fig.
-\ref{fig:classif-svm-pca}).
-An LDA approach yielded similar success rates (Fig.
-\ref{supfig:classif-lda-pca}), suggesting robust differences between these populations.
-Of the five islands, Little Cayman was the best discriminated with a mean SVM generalization success of 73.4\% (Table \ref{suptab:classif-svm-pca}).
-The results of the classification analyses on PCA data were very similar to results from SVMs and LDAs trained on reflectance values at 50nm-spaced wavelengths from 300 to 700nm (Fig.
-\ref{supfig:classif-svm-refl} and \ref{supfig:classif-lda-refl}).\\
+We found evidence of spatial autocorrelation in dewlap coloration between the sites within islands for Abaco (Table \ref{tab:autocorrelation}), suggesting that populations from closer sites tend to have more similar dewlaps on this island than expected by chance.
+Abaco was the island we sampled at the largest scale, with some sites nearly a hundred kilometers away from each other (Fig.
+\ref{fig:Abaco}A).
+That said, some sites were also in close proximity, and significant differences in coloration were detected between habitats sometimes less than a kilometer away (Fig.
+\ref{fig:Abaco_supplement}G), suggesting that differences in dewlap coloration between distant sites may be partly attributable to isolation-by-distance, but this may not necessarily be the case for sites in close proximity.
+We did not find evidence for spatial autocorrelation on other islands than Abaco (\textcolor{olive}{although Eleuthera was nearly significant, }Table \ref{tab:autocorrelation}).\\
 
-Differentiation in dewlap coloration occurred in multiple dimensions of color space.
-Moreover, the differences in dewlaps between habitats were not always consistent among islands, thus, we will discuss the habitat-specific variation in dewlap coloration for each island where significant differences were detected in turn (Fig.
-\ref{fig:anova}, Tables \ref{tab:anova}, \ref{suptab:kruskal}).
-Figure \ref{fig:anova}A provides a key to map principal component scores to the underlying wavelengths.\\
-
-On Abaco, dewlaps did not differ in PC1, which represents brightness.
-Mangrove lizards had significantly lower PC2 scores, corresponding to higher ultraviolet reflectance and lower red reflectance.
-Coastal beach scrub lizards had lower scores on PC3, corresponding to lower ultraviolet reflectance and higher blue reflectance.\\
-
-On Bimini, coastal beach scrub lizards had significantly brighter dewlaps than lizards from mangroves (PC1), but mangrove lizards had higher PC2 scores than beach scrub lizards, indicating higher violet and blue reflectance, and lower red reflectance.
-Lizards from primary coppice had higher PC3 scores overall, which correlated very positively with ultraviolet reflectance.\\
-
-On Cayman Brac, coppice-lizard dewlaps were significantly less bright than lizards from the other habitats.
-Coastal beach scrub lizards had dewlaps that scored low on PC2, corresponding to lower violet-blue and more red, while the mangrove lizards exhibited the opposite: relatively higher levels of violet-blue and less red.
-In PC3 space we found that dewlaps from lizards in the coastal habitat had high ultraviolet reflectance, coppice lizards had intermediate levels, and mangrove lizards had relatively low levels.\\
-
-On Little Cayman, the dewlaps of coppice lizards were significantly darker (PC1) than coastal-lizards.
-Mangrove lizards had less ultraviolet and redder dewlaps (PC2).
-The dewlaps of the coastal beach scrub lizards had higher levels of red and ultraviolet reflectance and less blue reflectance than the dewlaps of the other habitat-populations (PC3).\\
-
-On Long Island, lizards from the coppice habitat had darker dewlaps than lizards from the other habitats (PC1).
-Coastal lizards had relatively more ultraviolet and less blue-green reflectance in their dewlaps (PC3).
-These coastal-habitat lizards also scored lower on PC4, corresponding to slightly more violet and green-yellow dewlaps, and less blue dewlaps, than the mangrove lizards on the island.\\
-
-Sensitivity analyses on classifiers suggested an overall higher relative importance for PC2 and PC3 in determining between-group differences on Abaco, both in SVM and LDA classifiers (Fig.
-\ref{supfig:importance-svm-pca}, \ref{supfig:importance-lda-pca}), consistent with our ANOVA results (Fig.
-\ref{fig:anova}B).
-There was no strong signal of differences in relative importance among principal components on the other islands.
-Sensitivity analyses of SVMs trained on reflectance scores rather than principal components revealed, however, a consistently higher importance of ultraviolet reflectance in between-group differences on all islands (Fig.
-\ref{supfig:importance-svm-refl}).
-This pattern was not recovered for LDAs trained on reflectance scores (Fig.
-\ref{supfig:importance-lda-refl}).\\
-
-% No spatial autocorrelation
-
-We did not find significant spatial autocorrelation between the sampling sites on the islands where we detected a significant habitat effect.
-We did, however, detect a significant positive signal of autocorrelation on Eleuthera (Table \ref{suptab:autocor}), suggesting possible color differentiation through isolation-by-distance on this island.\\
-
-In contrast, differences in dewlap coloration between habitats were often detected in close geographical proximity.
-For example, Bimini, Cayman Brac, and Little Cayman were among the smallest islands in our study (Fig.
-\ref{supfig:map}).
-Indeed, most significant differences in dewlap coloration involved sites that were 5-10km apart.
-Our most extreme case of significant differences occurred for PC3 between a beach scrub site and a coppice site, separated from each other by a few hundreds of meters at most on Bimini (multiple pairwise Wilcoxon-Mann-Whitney tests, Fig.
-\ref{supfig:distances}).\\
-
-% Patterns of differentiation were inconsistent across islands
-
-Patterns of differentiation were inconsistent across the five most significant islands.
-Contrasts in principal components between habitats, calculated on pooled data from the whole archipelago, were not similar, for any component, among islands (Fig.
-\ref{supfig:contrasts}; MANOVA, Pillai's trace = 0.354, $F(6, 32) = 1$, P = 0.36).
-No pattern of variation was shared by all five significant islands, along any dimension.
-Some patterns did seem more common however, such as darker dewlaps among coppice lizards (Cayman Brac, Little Cayman, and Long Island, Fig.
-\ref{fig:anova}) or the intermediate position of coppice lizards in chromatic color space (Cayman Brac and Long Island).
-In other cases, patterns of differentiation were reversed from one island to another, with more ultraviolet reflecting dewlaps in mangroves than in coastal habitat on Abaco and Cayman Brac, but the opposite on Little Cayman and Long Island.
-Overall, it seemed that patterns of heterogeneity of variance were often driven by higher variances in coloration within beach scrub lizards (Fig.
-\ref{fig:anova}, Table \ref{tab:anova}).
-Yet other patterns were idiosyncratic, such as the combination of higher red and ultraviolet reflectance in coastal lizards on Little Cayman, where the rule seemed to be a negative correlation between ultraviolet and red reflectance across every other island.\\
+A striking feature of our data was inconsistencies in between-habitat differences among islands, in terms of which habitats differ from which, which dimensions of coloration were involved, and in which direction.
+For example, while on Cayman Brac the random forests could well distinguish between all three habitats (Fig.
+\ref{fig:CaymanBrac}D), on Abaco dewlaps from beach scrub and primary coppice were often mistaken, and on Bimini beach scrub dewlaps were more often classified into primary coppice or mangrove than into beach srub (Fig.
+\ref{fig:Bimini}D).
+In terms of variable importance, for multiple islands the random forests used information in the UV range to discriminate between at least some habitats, particularly on Abaco (Fig.
+\ref{fig:Abaco_supplement}F), Bimini (Fig.
+\ref{fig:Bimini}F), Cayman Brac (Fig.
+\ref{fig:CaymanBrac}F), Little Cayman (Fig.
+\ref{fig:LittleCayman}F) and Long Island (Fig.
+\ref{fig:LongIsland}F), but differences in UV reflectance involved different habitats and were in different directions among these islands.\\

### 2021-12-17 ms/response_JEB_20211024.txt
some changes

@@
 
-(R) We do not have these data unfortunately (do we not? ask Colin).
+(R) We do not have these data unfortunately (do we not?).
 
@@
 Nevertheless, although the species is still classified as territorial, its ability for movement and promiscuity, its continuous distribution at high densities throughout the islands and the polyphyly of individuals from different habitats within islands suggests that long-range gene flow is highly plausible.
-(Maybe the guys will know something about the cruising range of an anolis...)
+(Do you guys know something about the cruising range of an anolis...)
 
@@
 
-(R) We have now precised that these climatic variables include precipitations and solar incidence, among others, and those were interpreted as causing differences in types of habitat, with different light environments, among islands.
+(R) We have now added that these climatic variables include precipitations and solar incidence, among others, and those were interpreted as causing differences in types of habitat, with different light environments, among islands.
 
@@
 
-(R) This is now clarified.
+(R) This is now done.
 

### 2021-12-17 ms/sections/abstract.tex
implemented Graham's comments on the pdf

@@
 We suggest that neutral processes or selection could be involved in several ways, including sexual selection.
-Our results shed new light on the scale at which signal color polymorphism can be maintained in the presence of gene flow, and the relative role of local adaptation and other processes in driving these patterns.
+Our results shed new light on the scale at which signal color polymorphism can be maintained in the presence of gene flow, and the relative role of local adaptation and other processes in driving \sout{these} patterns \textcolor{olive}{of dewlap color variation across islands}.

### 2021-12-17 ms/sections/discussion.tex
implemented Graham's comments on the pdf

@@
 
-\textcolor{olive}{Genetic drift may contribute to some of the observed variation.
+\textcolor{olive}{Genetic drift could contribute to some of the observed variation.
 Indeed, while only Abaco showed significant patterns consistent with isolation-by-distance (which may emerge under limited dispersal and drift, \citealt{Wright1943, Kimura1964, Slatkin1987}), there may have been too few sites on most islands to conclusively detect it, and even then, the absence of detectable isolation-by-distance may not necessarily constitute evidence for the absence of drift.
 Besides, spatial autocorrelation was the strongest on islands sampled at the largest scales (e.g.
-some sites on Abaco were nearly 100km apart, and Eleuthera -- the second strongest signal, albeit nonsignificant -- had sites more than 30km apart), such that it is possible that neutral processes and/or dispersal limitations might in fact contribute to shaping variation over long distances.
+some sites on Abaco were nearly 100km apart, and Eleuthera -- the second strongest signal, albeit nonsignificant -- had sites more than 30km apart), such that it is possible that neutral processes and/or dispersal limitations might contribute to shaping variation over long distances.
 That said, many significant differences were found between habitats in close proximity, contrary to what would be expected under isolation-by-distance, including on islands where spatial autocorrelation was detected.

### 2021-12-17 ms/sections/introduction.tex
implemented Graham's comments on the pdf

@@
 Our sampling design included sites in close proximity; the median distance between two sites within an island was $11.2$km.
-Combining reflectance spectrometry and supervised machine learning, we tested for divergence in dewlap phenotype between habitats within islands and between islands across part of the range of \textit{A.
+\textcolor{olive}{While this species has traditionally been considered territorial, recent study reveals that they are polygynandrous and that gene flow is not impeded by territorial-like behaviors exibited by some males (Kamath and Losos 2018).} Combining reflectance spectrometry and supervised machine learning, we tested for divergence in dewlap phenotype between habitats within islands and between islands across part of the range of \textit{A.
 sagrei}.

### 2021-12-17 ms/sections/results.tex
implemented Graham's comments on the pdf

@@
 
-A striking feature of our data was inconsistencies in between-habitat differences among islands, in terms of which habitats differ from which, which dimensions of coloration were involved, and in which direction.
+A striking feature of our data was inconsistenc\sout{ies}\textcolor{olive}{y} in between-habitat differences among islands, in terms of which habitats differ from which, which dimensions of coloration were involved, and in which direction.
 For example, while on Cayman Brac the random forests could well distinguish between all three habitats (Fig.

### 2021-12-17 ms/response_jeb.tex
latex version of the response letter as proposed by Graham

\documentclass{article}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

\title{Dewlap color variation in \textit{Anolis sagrei} is maintained among habitats within islands of the West Indies}

\date{}

\begin{document}

	\maketitle

	\section*{Reponse letter}

	Dear Editor,\\

	In this response we address the comments we received for our submitted manuscript. Comments are numbered and quotes in \textit{italics}, and responses follow in \textbf{bold font}.

	\pagebreak

	\subsection*{Editor}

	\subsubsection*{Comment 1}

	\textit{While I agree with reviewer 2 that this is a very nice paper with interesting results that could possibly be published as it stands, I also agree with reviewer 1 that an obvious ommision is any analysis of if or how the light environments differ between habitats and islands. How similar are the habitats between islands or could differences in habitat between islands explain some of the variation between islands? Including data and analysis on the differences in light environment between the habitats (and islands), and possibly doing some visual modelling (which I think would require information on the light environment), would move this from being a very good to an excellent paper. I would therefore like to give you to opportunity to consider these suggestions and whether you have or can reasonably obtain the data to do these analyses.}\\

	\textbf{Thank you for this suggestion. We agree with the Editor that this would be a great addition. Unfortunately, we are unable to obtain these data. We do not have light environment data for all our habitats and islands. But, we do have some preliminary data from a single island that show that light environments are different. It is fairly obvious to the observer that light environment differs, say, from coppice forest to coastal. 
	As for visual modeling, this work has not been done such that we could compare \textit{sagrei} with other anoles and their predators, but this could certainly be an interesting future avenue.}

	\subsubsection*{Comment 2}

	\textit{One other more minor point is that lack of evidence for spatial autocorrelation is not the same as evidence for a lack of spatial autocorrelation (indeed there is a reasonably strong, albeit not quite "significant" spatial correlation on Eleuthera), so some of the conclusions about a lack of isolation by distance and genetic drift should be toned down slightly. The reviewers also raise some other minor comments that should be addressed.}\\

	\textbf{We have revised the discussion paragraph on the effect of drift accordingly (new citations added), and also a sentence in the results to point out that Eleuthera had a near-significant spatial correlation. We also toned down that statement in the abstract. We note that the citations have been updated, with new ones accompanying the requested changes by the reviewers.}

	\pagebreak

	\subsection*{Reviewer 1}

	\subsubsection*{Comment 3}

	\textit{Authors did a great job at designing this experiment and at combining diverse statistical analyses that enable testing their hypotheses taking into account potential sources of noise such as geographical proximity. I have one major suggestion/comment. Do you know whether such colour variation nicely described by your analyses, is actually detectable by conspecific, congeners and/or predators, the “agents” of selection?  For instance, do you think all dewlaps will be similarly seen under similar light/background conditions? Or whether observers can tell apart dewlaps coming from different habitats? In the first case, it is likely that the non-directional variation that you describe is not under selection, as to observers’ eyes, all dewlap colours look similar. In the second case, the conclusions will be similar to the ones you get with your current approach. This question can be tested applying vision modelling using your reflectance profiles, and an irradiance profile of at least one of the habitats that you studied. Perhaps this has been already done in other studies. If such is the case, you can clearly mention it.}\\

	\textbf{A very interesting idea. We do not know how detectable the reported differences in coloration are. Indeed, to know this we would need to do some visual modeling, and for this we would also need some habitat irradiance data, which we do not have unfortunately. Visual modeling has been performed in \textit{Anolis} lizards, but to the best of our knowledge it has not been done on our study populations, and so we do not know if the lizards can detect the differences in coloration of conspecifics from different habitats. Please also see our response to Comment 1.}

	\subsubsection*{Comment 4}

	\textit{Other remarks:
	-Among the differences that can be found on the three studied habitats, authors mostly talk about differences in light incidence. Are there any differences also in background colour? Predator community? Anole community? If so, how will they contribute or not to the obtained results?}\\

	\textbf{Interesting idea. Background color definitely varies between habitats, which can be seen with the naked eye from photos, but we did not quantify this precisely (we do not have irradiance profiles for the habitats telling us which wavelengths are different). Anole and predator communities have been shown to vary between islands (at least in density of the main species, which are overall present across all islands, or at least their ecologically similar species), but we do not know whether densities also vary between the habitats within islands.}

	\subsubsection*{Comment 5}

	\textit{-How big is the home range of an Anolis? Sorry of I missed it. This information will help understanding the importance of your experimental design.}\\

	\textbf{The best estimate of \textit{Anolis sagrei} home ranges (Kamath and Losos 2018a) found them to be 225m$^2$ for males and 36m$^2$ for females, but with overlap (and multiple matings for both sexes, a. k. a. polygynandry). Further, space use and movement were dynamic and varied, which increased both overlap among individuals and multiple mating. But these estimates were measured in an introduced population living on a college campus in Florida and may not represent the real territory size on our study islands. 
	Many anoles are considered territorial owing to their defensive displays, but genetic and reproductive data do not bear out the idea that anoles stay put and only mate with individuals in their territories (Kamath and Losos 2018b). Indeed many are polygynandrous. Hence, although the traditional literature suggests that the species is territorial, its ability for movement and promiscuity, its continuous distribution at high densities throughout the islands, and the polyphyly of individuals from different habitats within islands suggests that short and long-range gene flow are highly plausible. We have added some information to this effect in the introduction.}

	\subsubsection*{Comment 6}

	\textit{-In line 91, what do you mean by climatic variables? Light incidence?}\\

	\textbf{We have now added that these climatic variables include precipitation and solar incidence, among others, and those were interpreted as causing differences in types of habitat, with different light environments, among islands.}

	\subsubsection*{Comment 7}

	\textit{-In line 413, perhaps is worth it including a short definition of runaway sexual selection.}\\

	\textbf{This is now done.}

	\pagebreak

	\subsection*{Reviewer 2}

	\subsubsection*{Comment 8}

	\textit{I have only one minor suggestion/comment. The study focuses on the colouration of the dewlap. Of course, also the size and use of the dewlap is essential during signalling (see e.g. papers by Lailvaux, Irschick, Vanhooydonck, Driessens, Husak, Baeckens, to name a few). While I understand this is not the goal of the study, I think it’s still important to invest a couple of sentences or a short paragraph on how other characteristics of the dewlap (aside from colouration) are important too, and that those characteristics (possibly interaction with colouration) might vary within- and among-islands as well.}\\

	\textbf{Thank you for your review and for the suggestion. We have modified the second-to-last paragraph of the discussion to mention those other facets of anole display behavior, and added new references to support them.}

	\pagebreak

	\section*{Literature cited}

	Kamath, A., and Losos, J.B. 2018a. Estimating encounter rates as the first step of sexual selection in the lizard Anolis sagrei. Proceedings of the Royal Society B 286:20172244\\

	Kamath, A. and Losos, J. B., 2018b. Reconsidering territoriality is necessary for understanding Anolis mating systems. Behavioral Ecology and Sociobiology 72: 1-3.\\

\end{document}

### 2021-12-19 ms/sections/discussion.tex
addressing comments

@@
 
-Detectable differences in dewlap color between populations are surprising, as habitats were often in close geographical proximity to each other (as close as a few hundred meters on Bimini and most of the time within ten kilometers).
+Detectable differences in dewlap color between populations are surprising, as habitats were often in close geographical proximity to each other (as close as a few hundred meters on Bimini and most of the time within ten kilometers)\sout{.
 Indeed, given that (1) the populations were continuously distributed between the habitats, (2) populations from different habitats were not monophyletic with respect to mitochondrial haplotypes (\citealt{vandeSchoot2016} unpublished thesis), and (3) \textit{A.
-sagrei} is highly mobile \citep{Kamath2018}, we would have expected more homogeneous distributions of color phenotypes within islands due to gene flow, with fewer differences between populations, especially those in close proximity.\\
+sagrei} is highly mobile \citep{Kamath2018}, we would have expected more homogeneous distributions of color phenotypes within islands due to gene flow, with fewer differences between populations, especially those in close proximity.} \textcolor{olive}{, and we would have expected gene flow to cause a more homogeneous distriution of color phenotypes within islands.
+While little is known about the cruising range of individuals from our study populations (but see \citealt{Kamath2018, Steinberg2017}), \textit{A.
+sagrei} are polygynandrous (where both males and females mate with multiple mates, \citealt{Kamath2018,Kamath2018a}), thus offering opportunity for gene flow, especially given that lizards were distributed continuously and at high densities within the islands we sampled.
+Consistent with that, while populations from different islands were monophyletic, individuals within islands were not monophyletic with respect to habitat based on mitochondrial haplotypes (\citealt{vandeSchoot2016} unpublished thesis).}\\
 
@@
 \textcolor{olive}{That said, signalling can be a multifacetted behavior, and while we focused on dewlap color here, other potentially important elements of behavioral interactions include dewlap size, dewlap patterning, display activity as well as behaviors associated with dewlap extensions such as headbobs and pushups \citep{Vanhooydonck2005, Driessens2014, Driessens2015, Lailvaux2015}.
-These have been linked in various ways (and not always consistently across studies) not only with environmental variables but also with proxies for sexual selection, predation and species recognition \citep{Vanhooydonck2005, Lailvaux2007, Vanhooydonck2009, Driessens2017, Baeckens2018}, which do vary among islands but may also vary within islands, and possibly in interaction with dewlap color.} Either way, new data would be needed to test these hypotheses.\\
+These have been linked in various ways (and not always consistently across studies) not only with environmental variables but also with proxies for sexual selection, predation and species recognition \citep{Vanhooydonck2005, Lailvaux2007, Vanhooydonck2009, Driessens2017, Baeckens2018}, which do vary among islands but may also vary within islands, possibly in interaction with dewlap color.} Either way, new data would be needed to test these hypotheses.
+\textcolor{olive}{Visual modeling, for instance, would be a valuable follow-up analysis as it would indicate whether the differences we detected may be detectable by the organisms themselves (as has been done in \textit{Anolis} in \citealt{Leal2004, Fleishman2020}), but such approach requires irradiance profiles from the habitats in the field, which we presently do not have.}\\
 

### 2021-12-19 ms/sections/discussion.tex
added bit about how different the habitats really are

@@
 \textcolor{olive}{That said, signalling can be a multifacetted behavior, and while we focused on dewlap color here, other potentially important elements of behavioral interactions include dewlap size, dewlap patterning, display activity as well as behaviors associated with dewlap extensions such as headbobs and pushups \citep{Vanhooydonck2005, Driessens2014, Driessens2015, Lailvaux2015}.
-These have been linked in various ways (and not always consistently across studies) not only with environmental variables but also with proxies for sexual selection, predation and species recognition \citep{Vanhooydonck2005, Lailvaux2007, Vanhooydonck2009, Driessens2017, Baeckens2018}, which do vary among islands but may also vary within islands, possibly in interaction with dewlap color.} Either way, new data would be needed to test these hypotheses.
+These have been linked in various ways (and not always consistently across studies) not only with environmental variables but also with proxies for sexual selection, predation and species recognition \citep{Vanhooydonck2005, Lailvaux2007, Vanhooydonck2009, Driessens2017, Baeckens2018}, which do vary among islands but may also vary within islands, possibly in interaction with dewlap color.
+Moreover, while the three habitats were clearly recognizable and consistent across islands to the human eye (Fig.
+\ref{fig:overview}), we did not precisely quantify irradiance within each habitat, and so we may have missed differences between islands.} Either way, new data would be needed to test these hypotheses.
 \textcolor{olive}{Visual modeling, for instance, would be a valuable follow-up analysis as it would indicate whether the differences we detected may be detectable by the organisms themselves (as has been done in \textit{Anolis} in \citealt{Leal2004, Fleishman2020}), but such approach requires irradiance profiles from the habitats in the field, which we presently do not have.}\\

### 2021-12-20 ms/sections/discussion.tex
stuff

@@
 
-Detectable differences in dewlap color between populations are surprising, as habitats were often in close geographical proximity to each other (as close as a few hundred meters on Bimini and most of the time within ten kilometers)\sout{.
+Detectable differences in dewlap color between populations are surprising, as habitats were often in close geographical proximity to each other (as close as a few hundred meters on Bimini and most of the time within ten kilometers)\textcolor{red}{.
 Indeed, given that (1) the populations were continuously distributed between the habitats, (2) populations from different habitats were not monophyletic with respect to mitochondrial haplotypes (\citealt{vandeSchoot2016} unpublished thesis), and (3) \textit{A.
 sagrei} is highly mobile \citep{Kamath2018}, we would have expected more homogeneous distributions of color phenotypes within islands due to gene flow, with fewer differences between populations, especially those in close proximity.} \textcolor{olive}{, and we would have expected gene flow to cause a more homogeneous distriution of color phenotypes within islands.
-While little is known about the cruising range of individuals from our study populations (but see \citealt{Kamath2018, Steinberg2017}), \textit{A.
-sagrei} are polygynandrous (where both males and females mate with multiple mates, \citealt{Kamath2018,Kamath2018a}), thus offering opportunity for gene flow, especially given that lizards were distributed continuously and at high densities within the islands we sampled.
+While little is known about the cruising range of individuals from our study populations (but see \citealt{Steinberg2017, Kamath2018}), \textit{A.
+sagrei} are polygynandrous (where both males and females mate with multiple mates, \citealt{Kamath2017, Kamath2018, Kamath2018a}), thus offering opportunity for gene flow, especially given that lizards were distributed continuously and at high densities within the islands we sampled.
 Consistent with that, while populations from different islands were monophyletic, individuals within islands were not monophyletic with respect to habitat based on mitochondrial haplotypes (\citealt{vandeSchoot2016} unpublished thesis).}\\

### 2021-12-20 ms/sections/abstract.tex
flow is right

--- 2021-12-10 ms/sections/abstract.tex
@@
 We suggest that neutral processes or selection could be involved in several ways, including sexual selection.
-Our results shed new light on the scale at which signal color polymorphism can be maintained in the presence of gene flow, and the relative role of local adaptation and other processes in driving these patterns.
+Our results shed new light on the scale at which signal color polymorphism can be maintained in the presence of gene flow, and the relative role of local adaptation and other processes in driving \textcolor{red}{these} patterns \textcolor{olive}{of dewlap color variation across islands}.

### 2021-12-20 ms/sections/discussion.tex
flow is right

@@
 Indeed, given that (1) the populations were continuously distributed between the habitats, (2) populations from different habitats were not monophyletic with respect to mitochondrial haplotypes (\citealt{vandeSchoot2016} unpublished thesis), and (3) \textit{A.
-sagrei} is highly mobile \citep{Kamath2018}, we would have expected more homogeneous distributions of color phenotypes within islands due to gene flow, with fewer differences between populations, especially those in close proximity.} \textcolor{olive}{, and we would have expected gene flow to cause a more homogeneous distriution of color phenotypes within islands.
-While little is known about the cruising range of individuals from our study populations (but see \citealt{Steinberg2017, Kamath2018}), \textit{A.
-sagrei} are polygynandrous (where both males and females mate with multiple mates, \citealt{Kamath2017, Kamath2018, Kamath2018a}), thus offering opportunity for gene flow, especially given that lizards were distributed continuously and at high densities within the islands we sampled.
+sagrei} is highly mobile \citep{Kamath2018}, we would have expected more homogeneous distributions of color phenotypes within islands due to gene flow, with fewer differences between populations, especially those in close proximity.} \textcolor{olive}{, and we would have expected gene flow to cause a more homogeneous distribution of color phenotypes within islands.
+While little is known about the cruising range of individuals from our study populations (but see \citealt{Steinberg2017, Kamath2018} for other systems), \textit{A.
+sagrei} are polygynandrous (both males and females mate with multiple mates, \citealt{Kamath2017, Kamath2018a, Kamath2018}), thus offering opportunity for gene flow, especially given that lizards were distributed continuously and at high densities within the islands we sampled.
 Consistent with that, while populations from different islands were monophyletic, individuals within islands were not monophyletic with respect to habitat based on mitochondrial haplotypes (\citealt{vandeSchoot2016} unpublished thesis).}\\
@@
 
-\sout{We found no evidence for a role of genetic drift in explaining the observed patterns either.
+\textcolor{red}{We found no evidence for a role of genetic drift in explaining the observed patterns either.
 First, \textit{A.
@@
 Remaining hypotheses may include, for example, runaway sexual selection \textcolor{olive}{(i.e.
-arbitrary preferences of females for some colors over others, \citealt{Andersson1994})} \sout{arbitrarily} operating in different directions across islands, but no evidence so far suggests that dewlap is a target of mate choice in anoles \citep{Tokarz2002, Tokarz2005, Lailvaux2006, Nicholson2007}.
+arbitrary preferences of females for some colors over others, \citealt{Andersson1994})} \textcolor{red}{arbitrarily} operating in different directions across islands, but no evidence so far suggests that dewlap is a target of mate choice in anoles \citep{Tokarz2002, Tokarz2005, Lailvaux2006, Nicholson2007}.
 Another hypothesis is that the different genetic constitutions of different islands, perhaps resulting from founder effects (the islands have been colonized independently, \citealt{vandeSchoot2016} unpubl.; \citealt{Driessens2017, Reynolds2020}), may have predisposed populations to adapt differently to similarly selective circumstances.
-\textcolor{olive}{That said, signalling can be a multifacetted behavior, and while we focused on dewlap color here, other potentially important elements of behavioral interactions include dewlap size, dewlap patterning, display activity as well as behaviors associated with dewlap extensions such as headbobs and pushups \citep{Vanhooydonck2005, Driessens2014, Driessens2015, Lailvaux2015}.
+Either way, new data would be needed to test these hypotheses.
+\textcolor{olive}{Visual modeling, for instance, would be a valuable follow-up analysis as it would indicate whether the differences we detected may be detectable by the organisms themselves (as has been done in \textit{Anolis} in \citealt{Leal2004, Fleishman2020}), but such approach requires irradiance profiles from the habitats in the field, which we presently do not have.}\\
+
+\textcolor{olive}{Last, some patterns may arise owing to factors we did not measure.
+For one, signalling can be a multifacetted behavior, and while we focused on dewlap color here, other potentially important elements of behavioral interactions include dewlap size, dewlap patterning, display activity as well as behaviors associated with dewlap extensions such as headbobs and pushups \citep{Vanhooydonck2005, Driessens2014, Driessens2015, Lailvaux2015}.
 These have been linked in various ways (and not always consistently across studies) not only with environmental variables but also with proxies for sexual selection, predation and species recognition \citep{Vanhooydonck2005, Lailvaux2007, Vanhooydonck2009, Driessens2017, Baeckens2018}, which do vary among islands but may also vary within islands, possibly in interaction with dewlap color.
-Moreover, while the three habitats were clearly recognizable and consistent across islands to the human eye (Fig.
-\ref{fig:overview}), we did not precisely quantify irradiance within each habitat, and so we may have missed differences between islands.} Either way, new data would be needed to test these hypotheses.
-\textcolor{olive}{Visual modeling, for instance, would be a valuable follow-up analysis as it would indicate whether the differences we detected may be detectable by the organisms themselves (as has been done in \textit{Anolis} in \citealt{Leal2004, Fleishman2020}), but such approach requires irradiance profiles from the habitats in the field, which we presently do not have.}\\
+Besides, while the three habitats were clearly recognizable and consistent across islands to the human eye (Fig.
+\ref{fig:overview}), we did not precisely quantify irradiance within each habitat and did not consider any more subtle environmental differences there might have been for a given habitat among islands.}\\
 

### 2021-12-20 ms/sections/introduction.tex
flow is right

@@
 sagrei} dewlaps within nine islands in the Bahamas and Cayman Islands.
-These island systems presently, if not historically, comprise relatively small islands, with no major geographic barriers within islands limiting dispersal for this species \citep{Kamath2018}.
+These island systems presently, if not historically, comprise relatively small islands, with no major geographic barriers within islands limiting dispersal for this species.
 These islands all share three characteristic native West Indian small-island habitat-types -- beach scrub bush, closed-canopy primary coppice forest, and mangrove forest -- that are often spatially intermingled.

### 2021-12-20 ms/sections/discussion.tex
done with revision

@@
 
-Detectable differences in dewlap color between populations are surprising, as habitats were often in close geographical proximity to each other (as close as a few hundred meters on Bimini and most of the time within ten kilometers)\textcolor{red}{.
-Indeed, given that (1) the populations were continuously distributed between the habitats, (2) populations from different habitats were not monophyletic with respect to mitochondrial haplotypes (\citealt{vandeSchoot2016} unpublished thesis), and (3) \textit{A.
-sagrei} is highly mobile \citep{Kamath2018}, we would have expected more homogeneous distributions of color phenotypes within islands due to gene flow, with fewer differences between populations, especially those in close proximity.} \textcolor{olive}{, and we would have expected gene flow to cause a more homogeneous distribution of color phenotypes within islands.
+Detectable differences in dewlap color between populations are surprising, as habitats were often in close geographical proximity to each other (as close as a few hundred meters on Bimini and most of the time within ten kilometers)\sout{.
+Indeed, given that (1) the populations were continuously distributed between the habitats, (2) populations from different habitats were not monophyletic with respect to mitochondrial haplotypes (van de Schoot 2016 unpublished thesis), and (3) \textit{A.
+sagrei} is highly mobile (Kamath and Losos, 2018b), we would have expected more homogeneous distributions of color phenotypes within islands due to gene flow, with fewer differences between populations, especially those in close proximity.} \textcolor{olive}{, and we would have expected gene flow to cause a more homogeneous distribution of color phenotypes within islands.
 While little is known about the cruising range of individuals from our study populations (but see \citealt{Steinberg2017, Kamath2018} for other systems), \textit{A.
@@
 
-\textcolor{red}{We found no evidence for a role of genetic drift in explaining the observed patterns either.
+\sout{We found no evidence for a role of genetic drift in explaining the observed patterns either.
 First, \textit{A.
@@
 Remaining hypotheses may include, for example, runaway sexual selection \textcolor{olive}{(i.e.
-arbitrary preferences of females for some colors over others, \citealt{Andersson1994})} \textcolor{red}{arbitrarily} operating in different directions across islands, but no evidence so far suggests that dewlap is a target of mate choice in anoles \citep{Tokarz2002, Tokarz2005, Lailvaux2006, Nicholson2007}.
+arbitrary preferences of females for some colors over others, \citealt{Andersson1994})} \sout{arbitrarily} operating in different directions across islands, but no evidence so far suggests that dewlap is a target of mate choice in anoles \citep{Tokarz2002, Tokarz2005, Lailvaux2006, Nicholson2007}.
 Another hypothesis is that the different genetic constitutions of different islands, perhaps resulting from founder effects (the islands have been colonized independently, \citealt{vandeSchoot2016} unpubl.; \citealt{Driessens2017, Reynolds2020}), may have predisposed populations to adapt differently to similarly selective circumstances.

### 2022-03-09 ms/response_jeb.tex
accepted

@@
 \usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
+\usepackage[round]{natbib}
+\usepackage{ulem}
+\usepackage{xcolor}
 
@@
 
-	\section*{Reponse letter}
+	\textbf{Object: Response letter}\\\\
 
@@
 
-	In this response we address the comments we received for our submitted manuscript. Comments are numbered and quotes in \textit{italics}, and responses follow in \textbf{bold font}.
+	Thank you for your feedback.
+In this response we address the comments we received for our submitted manuscript.
+Comments are numbered and quoted in \textit{italics}, and our responses follow in \textbf{bold}.
+Please find at the end a list of references mentioned in this letter, including new references that have been added to the the manuscript.
+For each response we indicate where in the manuscript the comment has been addressed.
+Please note that the numbers refer to the lines in the revised version of the manuscript, with changes tracked (deletions are \sout{crossed out} and additions are in \textcolor{olive}{olive}).\\
+
+	In addition to responses to the reviewers' comments, we made a few additional edits to match the formatting requirements of \textit{Journal of Evolutionary Biology}: we placed each table on a separate page, and we added the taxonomic order and family of our study species upon first occurrence in the main text (l.
+56).\\
+
+	We look forward to hearing from you.\\\\
+
+	Kind regards,\\
+
+	Rapha\"{e}l Scherrer
 
@@
 It is fairly obvious to the observer that light environment differs, say, from coppice forest to coastal.
-	As for visual modeling, this work has not been done such that we could compare \textit{sagrei} with other anoles and their predators, but this could certainly be an interesting future avenue.}
+In a newly added paragraph (lines 456-467), we discuss the possibility that differences in habitats between islands may influence the patterns of phenotypic variation.
+	As for visual modeling, this work has not been done such that we could compare \textit{sagrei} with other anoles and their predators, but this could certainly be an interesting future avenue.
+Please also see our response to Comment 3.}
 
@@
 
-	\textbf{We have revised the discussion paragraph on the effect of drift accordingly (new citations added), and also a sentence in the results to point out that Eleuthera had a near-significant spatial correlation.
-We also toned down that statement in the abstract.
-We note that the citations have been updated, with new ones accompanying the requested changes by the reviewers.}
+	\textbf{We have revised the discussion paragraph on the effect of drift accordingly (ll.
+422-437, including new citations by \citealt{Wright1943, Kimura1964, Slatkin1987}), and also a sentence in the results to point out that Eleuthera had a near-significant spatial correlation (ll.
+336).
+We also toned down that statement in the abstract (l.
+18).}
 
@@
 Indeed, to know this we would need to do some visual modeling, and for this we would also need some habitat irradiance data, which we do not have unfortunately.
-Visual modeling has been performed in \textit{Anolis} lizards, but to the best of our knowledge it has not been done on our study populations, and so we do not know if the lizards can detect the differences in coloration of conspecifics from different habitats.
-Please also see our response to Comment 1.}
+Visual modeling has been performed in \textit{Anolis} lizards (e.g.
+\citealt{Leal2004, Fleishman2020}), but to the best of our knowledge it has not been done on our study populations, and so we do not know if the lizards can detect the differences in coloration of conspecifics from different habitats (now discussed on ll.
+450-454).}
 
@@
 Background color definitely varies between habitats, which can be seen with the naked eye from photos, but we did not quantify this precisely (we do not have irradiance profiles for the habitats telling us which wavelengths are different).
-Anole and predator communities have been shown to vary between islands (at least in density of the main species, which are overall present across all islands, or at least their ecologically similar species), but we do not know whether densities also vary between the habitats within islands.}
+Anole and predator communities have been shown to vary between islands (at least in density of the main species, which are overall present across all islands, or at least their ecologically similar species, see e.g.
+\citealt{Baeckens2018}), but we do not know whether densities also vary between the habitats within islands.
+We discuss that in a newly added paragraph (ll.
+456-467).}
 
@@
 
-	\textbf{The best estimate of \textit{Anolis sagrei} home ranges (Kamath and Losos 2018a) found them to be 225m$^2$ for males and 36m$^2$ for females, but with overlap (and multiple matings for both sexes, a.
+	\textbf{The best estimate of \textit{Anolis sagrei} home ranges \citep{Kamath2018} found them to be 225m$^2$ for males and 36m$^2$ for females, but with overlap (and multiple matings for both sexes, a.
 k.
@@
 But these estimates were measured in an introduced population living on a college campus in Florida and may not represent the real territory size on our study islands.
-	Many anoles are considered territorial owing to their defensive displays, but genetic and reproductive data do not bear out the idea that anoles stay put and only mate with individuals in their territories (Kamath and Losos 2018b).
+	Many anoles are considered territorial owing to their defensive displays, but genetic and reproductive data do not bear out the idea that anoles stay put and only mate with individuals in their territories (\citealt{Kamath2017a, Kamath2018a}, added to the manuscript).
 Indeed many are polygynandrous.
 Hence, although the traditional literature suggests that the species is territorial, its ability for movement and promiscuity, its continuous distribution at high densities throughout the islands, and the polyphyly of individuals from different habitats within islands suggests that short and long-range gene flow are highly plausible.
-We have added some information to this effect in the introduction.}
+Besides, we note that \citet{Kamath2018} estimated movement over a few months time, while these ranges could be substantially larger over a lifetime.
+In another trunk-ground anole, \textit{Anolis gundlachi}, makes and females moved 80 meters from their point of capture were able to return to the original location 80\% and 40-60\% of the time, respectively, in less than 24 hours (\citealt{Steinberg2017}, added to the manuscript), suggesting at least the capacity to move very long distances in a lizard lifetime.
+We have added some information to this effect in the introduction (ll.
+112-114) and the discussion (ll.
+363-371).}
 
@@
 
-	\textbf{We have now added that these climatic variables include precipitation and solar incidence, among others, and those were interpreted as causing differences in types of habitat, with different light environments, among islands.}
+	\textbf{We have now added that these climatic variables include precipitation and solar incidence, among others, and those were interpreted as causing differences in types of habitat, with different light environments, among islands (ll.
+92-94).}
 
@@
 
-	\textbf{This is now done.}
+	\textbf{This is now done (l. 443).}
 
@@
 	\textbf{Thank you for your review and for the suggestion.
-We have modified the second-to-last paragraph of the discussion to mention those other facets of anole display behavior, and added new references to support them.}
+We have added a paragraph to the discussion (ll.
+456-467) mentioning those other facets of anole display behavior, and added new references to support them \citep{Vanhooydonck2005, Lailvaux2006, Lailvaux2007, Vanhooydonck2009, Driessens2014, Driessens2015, Lailvaux2015, Driessens2017, Baeckens2018}.
+(Note that \citealt{Lailvaux2006} has been added to the paragraph that comes just before, on ll.
+445-446.)}
 
@@
 
-	\section*{Literature cited}
-
-	Kamath, A., and Losos, J.B. 2018a. Estimating encounter rates as the first step of sexual selection in the lizard Anolis sagrei. Proceedings of the Royal Society B 286:20172244\\
-
-	Kamath, A. and Losos, J. B., 2018b. Reconsidering territoriality is necessary for understanding Anolis mating systems. Behavioral Ecology and Sociobiology 72: 1-3.\\
+	\bibliographystyle{apalike}
+	\bibliography{library.bib}
 

### 2022-03-09 ms/final/final/main.tex
accepted

--- 2020-08-08 ms/main.tex
@@
 
-%% Language and font encodings
+% Language and font encodings
 \usepackage[english]{babel}
@@
 
-%% Sets page size and margins
+% Sets page size and margins
 \usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
 
-%% Useful packages
+% Useful packages
 \usepackage{amsmath}
@@
 \usepackage{booktabs}
-%\restylefloat{figure} % for force placing figures
-%\floatplacement{figure}{H}
-%\floatplacement{table}{H}
-%\floatplacement{longtable}{H}
-%\usepackage[markers,nolists]{endfloat}
 \usepackage{graphicx}
@@
 \usepackage[colorlinks=true, allcolors=blue]{hyperref}
-%\usepackage{apalike}
 \usepackage[round]{natbib}
@@
 \usepackage{multirow}
-% Supp mat
-\newcommand{\beginsupplement}{%
+
+% Supplementary Materials
+\newcommand{\beginsupplement}{
 	\setcounter{table}{0}
-        \renewcommand{\thetable}{S\arabic{table}}%
+    \renewcommand{\thetable}{S\arabic{table}}
     \setcounter{figure}{0}
-        \renewcommand{\thefigure}{S\arabic{figure}}%
+    \renewcommand{\thefigure}{S\arabic{figure}}
 }
@@
 \title{Dewlap color variation in \textit{Anolis sagrei} is maintained among habitats within islands of the West Indies}
+
 \author{
+
     \textsc{Rapha\"{e}l Scherrer$^{1,3}$ \thanks{Corresponding author: r.scherrer@rug.nl}, Colin M. Donihue$^{1,4}$, }\\
 	\textsc{R. Graham Reynolds$^2$, Jonathan B. Losos$^{1,4}$ and Anthony J. Geneva$^{1,5}$} \\[1ex]
-	%\thanks{A thank you} \\[1ex] % Your name
+
 	\normalsize $^1$ Department of Organismic and Evolutionary Biology and Museum of Comparative Zoology \\ \normalsize Harvard University, Cambridge, MA, USA \\
@@
 
+	\pagebreak
+
 	\section*{Methods}
@@
 	\input{sections/methods.tex}
+
+	\pagebreak
 
@@
 
+	\pagebreak
+
 	\section*{Discussion}
@@
 
-\subsection*{Acknowledgements}
+	\pagebreak
+
+	\section*{Acknowledgements}
 
@@
 
-%\section*{Supplementary Figures}
+	\pagebreak
 
-%\input{sections/suppfigures.tex}
+	\section*{Appendix}
+
+	\input{sections/appendix}
 

### 2022-03-09 ms/final/final/sections/abstract.tex
accepted

--- 2021-09-06 ms-jeb/sections/abstract.tex
@@
 Using reflectance spectrometry paired with supervised machine learning, we found significant differences in spectral properties of the dewlap between habitats within small islands, sometimes over very short distances.
-Passive divergence in dewlap phenotype associated with isolation-by-distance did not explain our results.
+Passive divergence in dewlap phenotype associated with isolation-by-distance did not seem to explain our results.
 On the other hand, these habitat-specific dewlap differences varied in magnitude and direction across islands, and thus our primary test for adaptation -- parallel responses across islands -- was not supported.
 We suggest that neutral processes or selection could be involved in several ways, including sexual selection.
-Our results shed new light on the scale at which signal color polymorphism can be maintained in the presence of gene flow, and the relative role of local adaptation and other processes in driving these patterns.
+Our results shed new light on the scale at which signal color polymorphism can be maintained in the presence of gene flow, and the relative role of local adaptation and other processes in driving these patterns of dewlap color variation across islands.

### 2022-03-09 ms/final/final/sections/acknowledgments.tex
accepted

--- 2021-05-23 ms/sections/acknowledgments.tex
@@
 Puente-Rol\'{o}n, Kevin Aviles-Rodriguez, Kristin Winchell, Jason Fredette, Quyhn Quach, Wendy Jesse, Inbar Maayan, Alexis Harrison and Melissa Kemp for assistance in the field and Pratik Gupte, Max Lambert and James Stroud for helpful discussions.
-Funding for this work was provided by the Templeton Foundation (to JBL), NSF DEB~\#1927194 (to JBL and AJG), NSF DEB~\#1500761 (to AJG), NSF DBI~\#1609284 (to CMD), and a Harvard Museum of Comparative Zoology Putnam Expedition Grant (to RGR).
+This project was made possible through the support of a grant from the John Templeton Foundation (to JBL).
+The opinions expressed in this publication are those of the author(s) and do not necessarily reflect the views of the John Templeton Foundation.
+Additional funding for this work was provided by NSF DEB~\#1927194 (to JBL and AJG), NSF DEB~\#1500761 (to AJG), NSF DBI~\#1609284 (to CMD), and a Harvard Museum of Comparative Zoology Putnam Expedition Grant (to RGR).

### 2022-03-09 ms/final/final/sections/appendix.tex
accepted

--- 2021-07-26 ms/sections/appendix.tex
@@
 \ref{fig:Abaco}B).
-Consistent with this, our analyses of variance detected significantly lower PC2 scores in mangrove lizards than in the two other habitats (Fig.
+Consistent with this, our analyses of variance detected significantly higher PC2 scores in mangrove lizards than in the two other habitats (Fig.
 \ref{fig:Abaco}E, Table \ref{tab:anova}), representing a higher UV-reflectance relative to red (Fig.
 \ref{fig:Abaco}C).
-Beach scrub lizards also scored lower on PC3 (Fig.
+Beach scrub lizards also scored higher on PC3 (Fig.
 \ref{fig:Abaco}E, Table \ref{tab:anova}), indicating less curvature of the reflectance profile and relatively higher reflectance at intermediate wavelengths (blue-to-yellow) than at the ends of the range (Fig.
@@
 \ref{fig:Bimini}C, E, Table \ref{tab:anova}).
-The random forests also used UV reflectance to discriminate between coppice and mangrove dewlaps (Fig.
-\ref{fig:Bimini}F), which could reflect the significant difference we detected along PC3 between these two habitats (Fig.
-\ref{fig:Bimini}C, E, Table \ref{tab:anova}).
-Beach scrub lizards were characterized by elevated red reflectance relative to UV (as represented by PC2, (Fig.
-\ref{fig:Bimini}C, E)), and beach scrub and mangrove lizards were characterized by a more even distribution of the reflectance along the spectrum (as represented by PC3, (Fig.
-\ref{fig:Bimini}C, E)), in contrast to coppice lizards which harbored a stronger curvature at intermediate wavelengths (Fig.
-\ref{fig:Bimini}C, E).
+Coppice dewlaps had significantly higher PC2 scores than mangrove dewlaps (Fig.
+\ref{fig:Bimini}E), suggesting a higher curvature (higher UV and red reflectance than more intermediate wavelengths, Fig.
+\ref{fig:Bimini}C).
+For these two habitats the random forests were most sensitive to UV reflectance (Fig.
+\ref{fig:Bimini}F).
+Beach scrub dewlaps had higher PC3 scores than coppice dewlaps but it was not clear what properties of spectral shape this principal component mapped onto (Fig.
+\ref{fig:Bimini}C).
 On this island, the beach scrub and coppice habitats were separated by a few hundred meters, making this contrast the smallest geographical scale at which differences in coloration were found in our study (Fig.
@@
 \ref{fig:CaymanBrac}G), dewlaps in the beach scrub habitat reflected more red light (as represented by PC2, Fig.
-\ref{fig:CaymanBrac}C, E) and more UV (as represented by PC3, Fig.
-\ref{fig:CaymanBrac}C, E) than in the mangrove habitat.
-Coppice lizards were also characterized by a higher UV reflectance than mangrove dewlaps (PC3, Fig.
-\ref{fig:CaymanBrac}C, E, Table \ref{tab:anova}).\\
+\ref{fig:CaymanBrac}C, E) and more UV (as represented by PC3, along which coppice dewlaps were intermediate, Fig.
+\ref{fig:CaymanBrac}C, E) than in the mangrove habitat.\\
 
 On Eleuthera, although random forests detected between-habitat differences in dewlap color, other approaches did not (Tables \ref{tab:ldas} and \ref{tab:ksvms}), suggesting that the differences may be small.
-Consistent with this, the only significant univariate difference detected was for PC2 between beach scrub and mangrove lizards, where beach scrub lizards had higher levels of red reflectance and mangrove lizards higher levels of UV reflectance (Fig.
-\ref{fig:Eleuthera}C, E, Table \ref{tab:anova}).\\
+On Eleuthera, beach scrub and mangrove dewlaps were guessed relatively correctly, but coppice dewlaps were more often mistaken for mangrove dewlaps (Fig.
+\ref{fig:Eleuthera}D).
+Mangrove dewlaps had lower PC2 scores than beach scrub dewlaps (Fig.
+\ref{fig:Eleuthera}E, indicating higher UV relative to red, Fig.
+\ref{fig:Eleuthera}C), and higher PC4 scores than coppice dewlaps (Fig.
+\ref{fig:Eleuthera}E, suggesting more reflectance profiles with more curvature, Fig.
+\ref{fig:Eleuthera}C).
+Random forests did not seem to consistently capture the wavelengths responsible for these differences (Fig.
+\ref{fig:Eleuthera}F).\\
 
-Little Cayman was characterized by a better discrimination of mangrove lizards from the rest than between beach scrub and coppice lizards (Fig.
+Little Cayman was characterized by a better discrimination of mangrove lizards from the rest than between beach scrub and coppice lizards even though all habitats were relatively well discriminated (Fig.
 \ref{fig:LittleCayman}D).
@@
 \ref{fig:LongIsland}F).
-Beach scrub lizards had more curved reflectance profiles than in either of the two other habitats, with higher levels of UV and red reflectance relative to intermediate wavelengths (PC3, Fig.
+Beach scrub lizards had more curved reflectance profiles than in the mangrove, with higher levels of UV and red reflectance relative to intermediate wavelengths (PC3, Fig.
 \ref{fig:LongIsland}C, E, Table \ref{tab:anova}).
-Beach scrub lizards also differed from mangrove lizards along PC4 (Fig.
-\ref{fig:LongIsland}E), which represented a rather small portion of the variance not already explained by the first three principal components, and is therefore difficult to interpret (Fig.
-\ref{fig:LongIsland}C).
 Coppice lizards were significantly darker than mangrove and beach scrub lizards (PC1, Fig.
@@
 
-On North Andros, although the random forest classification was significant ($P = 0.0216$, Table \ref{tab:randomforests}) and the average confusion matrix indicated that lizards from beach scrub were particularly well predicted (Fig.
-\ref{fig:NorthAndros}D), no significant univariate differences were detected along any of the four PCs (Fig.
-\ref{fig:NorthAndros}E, Table \ref{tab:anova}).
-Importance analysis of full-spectrum random forests showed higher importance scores near the UV-end of the spectrum in discriminating beach scrub dewlaps from the rest (Fig.
+On North Andros beach scrub and coppice dewlaps could be discriminated better against each other than with mangrove dewlaps (Fig.
+\ref{fig:NorthAndros}D), with importance profiles supporting UV-reflectance as a predictor of coppice lizards (Fig.
 \ref{fig:NorthAndros}F).
-Besides, reflectance curves of beach scrub dewlaps appeared more similar to each other in the UV range than dewlaps from other habitats (Fig.
-\ref{fig:NorthAndros}B), suggesting that the machines may have used this low within-habitat variance, as opposed to between-habitat differences in means, to correctly classify beach scrub lizards.
-A small sample size on this islands may also have contributed to a lack of power in detecting univariate differences using analyses of variance (Table \ref{tab:counts}).\\
+Coppice lizards had less curved reflectance profiles than beach scrub and mangrove lizards (PC2), and beach scrub dewlaps had the lowest scores on PC4, which was difficult to interpret (Fig.
+\ref{fig:NorthAndros}C, E, Table \ref{tab:anova}).\\
 
-On South Andros beach scrub and coppice dewlaps could be discriminated better against each other than with mangrove dewlaps (Fig.
-\ref{fig:SouthAndros}D), with importance profiles supporting UV-reflectance as a predictor of coppice lizards (Fig.
-\ref{fig:SouthAndros}F).
-Coppice lizards had more curved reflectance profiles than beach scrub lizards (PC3), and lizards from both habitats differed along PC4, which is again more difficult to interpret (Fig.
-\ref{fig:SouthAndros}C, E, Table \ref{tab:anova}).
-Beach scrub lizards also differed from mangrove lizards in PC4 (Fig.
-\ref{fig:SouthAndros}C, E, Table \ref{tab:anova}).\\
-
-Classification success was not significantly better than expected by chance on Ragged Island (Table \ref{tab:randomforests}) where nearly no habitat could be differentiated from any other based on reflectance.
+Classification success was not significantly better than expected by chance on Ragged Island and South Andros (Table \ref{tab:randomforests}, \ref{tab:ldas}, \ref{tab:ksvms}) where nearly no habitat could be differentiated from any other based on reflectance.

### 2022-03-09 ms/final/final/sections/discussion.tex
accepted

--- 2021-12-20 ms/sections/discussion.tex
@@
 
-Detectable differences in dewlap color between populations are surprising, as habitats were often in close geographical proximity to each other (as close as a few hundred meters on Bimini and most of the time within ten kilometers)\sout{.
-Indeed, given that (1) the populations were continuously distributed between the habitats, (2) populations from different habitats were not monophyletic with respect to mitochondrial haplotypes (van de Schoot 2016 unpublished thesis), and (3) \textit{A.
-sagrei} is highly mobile (Kamath and Losos, 2018b), we would have expected more homogeneous distributions of color phenotypes within islands due to gene flow, with fewer differences between populations, especially those in close proximity.} \textcolor{olive}{, and we would have expected gene flow to cause a more homogeneous distribution of color phenotypes within islands.
+Detectable differences in dewlap color between populations are surprising, as habitats were often in close geographical proximity to each other (as close as a few hundred meters on Bimini and most of the time within ten kilometers), and we would have expected gene flow to cause a more homogeneous distribution of color phenotypes within islands.
 While little is known about the cruising range of individuals from our study populations (but see \citealt{Steinberg2017, Kamath2018} for other systems), \textit{A.
-sagrei} are polygynandrous (both males and females mate with multiple mates, \citealt{Kamath2017, Kamath2018a, Kamath2018}), thus offering opportunity for gene flow, especially given that lizards were distributed continuously and at high densities within the islands we sampled.
-Consistent with that, while populations from different islands were monophyletic, individuals within islands were not monophyletic with respect to habitat based on mitochondrial haplotypes (\citealt{vandeSchoot2016} unpublished thesis).}\\
+sagrei} are polygynandrous (both males and females mate with multiple mates, \citealt{Kamath2017a, Kamath2018a, Kamath2018}), thus offering opportunity for gene flow, especially given that lizards were distributed continuously and at high densities within the islands we sampled.
+Consistent with that, while populations from different islands were monophyletic, individuals within islands were not monophyletic with respect to habitat based on mitochondrial haplotypes (\citealt{vandeSchoot2016} unpublished thesis).\\
 
@@
 
-\sout{We found no evidence for a role of genetic drift in explaining the observed patterns either.
-First, \textit{A.
-sagrei} was distributed across the islands continuously, usually at relatively high population densities, rather than in small and isolated populations where drift might be expected to have a strong effect.
-Second, we found no evidence of isolation-by-distance except on Abaco, which was sampled at the largest geographical scale, with sites nearly a hundred kilometers apart from each other.
-Hence, while isolation-by-distance may explain long-range differences on this island, most of the differences among habitats across the rest of the sampling region are unlikely to be explained by genetic drift, as habitats were often in close proximity (less than 10km).}\\
-
-\textcolor{olive}{Genetic drift could contribute to some of the observed variation.
+Genetic drift could contribute to some of the observed variation.
 Indeed, while only Abaco showed significant patterns consistent with isolation-by-distance (which may emerge under limited dispersal and drift, \citealt{Wright1943, Kimura1964, Slatkin1987}), there may have been too few sites on most islands to conclusively detect it, and even then, the absence of detectable isolation-by-distance may not necessarily constitute evidence for the absence of drift.
@@
 Together with the fact that isolation-by-distance may not necessarily only emerge from drift (e.g.
-if there is a spatial environmental gradient), this indicates that genetic drift may have limited potential to explain the differences observed between habitats, at least at a local scale.}\\
+if there is a spatial environmental gradient), this indicates that genetic drift may have limited potential to explain the differences observed between habitats, at least at a local scale.\\
 
@@
 Differences among habitats within islands, however, are still difficult to account for.
-Remaining hypotheses may include, for example, runaway sexual selection \textcolor{olive}{(i.e.
-arbitrary preferences of females for some colors over others, \citealt{Andersson1994})} \sout{arbitrarily} operating in different directions across islands, but no evidence so far suggests that dewlap is a target of mate choice in anoles \citep{Tokarz2002, Tokarz2005, Lailvaux2006, Nicholson2007}.
+Remaining hypotheses may include, for example, runaway sexual selection (i.e.
+arbitrary preferences of females for some colors over others, \citealt{Andersson1994}) operating in different directions across islands, but no evidence so far suggests that dewlap is a target of mate choice in anoles \citep{Tokarz2002, Tokarz2005, Lailvaux2006, Nicholson2007}.
 Another hypothesis is that the different genetic constitutions of different islands, perhaps resulting from founder effects (the islands have been colonized independently, \citealt{vandeSchoot2016} unpubl.; \citealt{Driessens2017, Reynolds2020}), may have predisposed populations to adapt differently to similarly selective circumstances.
 Either way, new data would be needed to test these hypotheses.
-\textcolor{olive}{Visual modeling, for instance, would be a valuable follow-up analysis as it would indicate whether the differences we detected may be detectable by the organisms themselves (as has been done in \textit{Anolis} in \citealt{Leal2004, Fleishman2020}), but such approach requires irradiance profiles from the habitats in the field, which we presently do not have.}\\
+Visual modeling, for instance, would be a valuable follow-up analysis as it would indicate whether the differences we detected may be detectable by the organisms themselves (as has been done in \textit{Anolis} in \citealt{Leal2004, Fleishman2020}), but such approach requires irradiance profiles from the habitats in the field, which we presently do not have.\\
 
-\textcolor{olive}{Last, some patterns may arise owing to factors we did not measure.
+Last, some patterns may arise owing to factors we did not measure.
 For one, signalling can be a multifacetted behavior, and while we focused on dewlap color here, other potentially important elements of behavioral interactions include dewlap size, dewlap patterning, display activity as well as behaviors associated with dewlap extensions such as headbobs and pushups \citep{Vanhooydonck2005, Driessens2014, Driessens2015, Lailvaux2015}.
@@
 Besides, while the three habitats were clearly recognizable and consistent across islands to the human eye (Fig.
-\ref{fig:overview}), we did not precisely quantify irradiance within each habitat and did not consider any more subtle environmental differences there might have been for a given habitat among islands.}\\
+\ref{fig:overview}), we did not precisely quantify irradiance within each habitat and did not consider any more subtle environmental differences there might have been for a given habitat among islands.\\
 

### 2022-03-09 ms/final/final/sections/introduction.tex
accepted

--- 2021-09-06 ms/sections/introduction.tex
@@
 
-% One main barrier to the maintenance of diversity is gene flow
-
 One potential barrier to the maintenance of localized signal divergence is the homogenizing effect of gene flow.
@@
 Therefore, despite evidence that local adaptation may be particularly difficult at small spatial scales where gene flow tends to cause adjoining populations to remain genetically homogeneous, the potential adaptive response of species traits, in particular communication signals, to localized differences in habitats remains relatively unknown \citep{Richardson2014}.
-Lizards of the neotropical genus \textit{Anolis} are an excellent group for studying the eco-evolutionary dynamics of local adaptation and natural selection \citep{Losos2009}.
+Lizards of the neotropical genus \textit{Anolis} (Squamata: Dactyloidae) are an excellent group for studying the eco-evolutionary dynamics of local adaptation and natural selection \citep{Losos2009}.
 A particularly conspicuous trait of anoles is their dewlap, an extensible flap of skin that is typically sexually dimorphic and used as a communication signal in courtship \citep{Sigmund1983, Driessens2014, Driessens2015} and territorial displays \citep{Losos1985, Macedonia1994, Macedonia2013} as well as in predator deterrence \citep{Leal1995, Leal1997, Leal1997a}.
@@
 It has been the subject of numerous studies concerning local adaptation \citep{Losos1994, Losos1997a, Losos2001, Kolbe2012}, biological invasion \citep{Kolbe2008}, and sexual selection \citep{Tokarz2002, Tokarz2005, Tokarz2006, Driessens2014, Steffen2014, Driessens2015} among many other topics.
-Between-island variation in the mainly orange-red color of its dewlap was shown to be better explained by climatic variables \citep{Driessens2017} than by proxies for biotic factors such as sexual selection or predation pressure \citep{Vanhooydonck2009, Baeckens2018}.
+Between-island variation in the mainly orange-red color of its dewlap was shown to be better explained by climatic variables such as annual precipitation and solar radiation (proposed to affect the average vegetation type on each island and among other things, its ambient light environment, \citealt{Driessens2017}),  than by proxies for biotic factors such as sexual selection or predation pressure \citep{Vanhooydonck2009, Baeckens2018}.
 How intra-island differences in habitat may contribute to the diversity of dewlap coloration, however, remains unexplored, and may reveal new insights into the scale of local differentiation despite gene flow.\\
-
-% What we did
 
@@
 
-Our sampling design included sites in close proximity; the median distance between two sites within an island was $11.2$km.
+Our sampling design included sites in close proximity; the median distance between two sites within an island was 8.45km.
+While this species has traditionally been considered territorial, a recent study revealed that they are polygynandrous and that gene flow is not impeded by territorial-like behaviors exibited by some males (Kamath and Losos 2018).
 Combining reflectance spectrometry and supervised machine learning, we tested for divergence in dewlap phenotype between habitats within islands and between islands across part of the range of \textit{A.

### 2022-03-09 ms/final/final/sections/methods.tex
accepted

--- 2021-07-26 ms/sections/methods2.tex
@@
-% Citation for classifiers as a good group comparison tool
-% Yang and Yang on LDA being performed on PC
-
 \subsection*{Data collection}
 
-We sampled 466 male \textit{A.
+We sampled 455 male \textit{A.
 sagrei} from seven islands in the Bahamas Archipelago -- Abaco, North Andros, South Andros, South Bimini, Eleuthera, Long Island, and Ragged Island -- and two in the Cayman Islands -- Cayman Brac and Little Cayman (Figure \ref{fig:maps}A).
@@
 Our sampling design enabled us to test for differences between habitats at a coarse and fine geographical scale.
-The median distance between two localities within an island was $\sim 11$km (Figure \ref{fig:maps}B), and $80.3$\% of all pairwise distances within islands were less than $50$km.
+The median distance between two localities within an island was $~8.5$km (Figure \ref{fig:maps}B), and 79.3\% of all pairwise distances within islands were less than $50$km.
 Additionally, there are no major barriers to dispersal (such as mountains or grassland) on any of the islands that we sampled.
@@
 
-All analyses in this study were performed in R 3.6.1 \citep{RCoreTeam2019} and the source code can be found at https://github.com/rscherrer/dewlap, presently private.
+All analyses in this study were performed in R 3.6.1 \citep{RCoreTeam2019}.
 

### 2022-03-09 ms/final/final/sections/results.tex
accepted

--- 2021-09-06 ms/sections/results2.tex
@@
-% Little but detectable differences on five islands
-
 We tested for variation in \textit{A.
 sagrei} dewlap coloration between populations living in three characteristic habitat types across nine islands that span the West Indian range of the brown anole (beach scrub, primary coppice and mangroves).
-We found that most of the variation in coloration was partitioned between islands (two-way PERMANOVA, $F(df = 8) = 43.7$, $P = 0.0001$, explained variance $R^2 = 40.9$\%).
-Nonetheless, we did find evidence for differences in dewlap coloration between habitat types, and those were mostly island-specific (habitat-by-island interaction term, $F(16) = 3.53$, $P = 0.0001$, $R^2 = 6.6$\%), with a significant portion of the variation explained by an habitat effect across all islands, but this effect was relatively small ($F(2) = 4.7$, $P = 0.0001$, $R^2 = 1.1$\%).\\
+We found that most of the variation in coloration was partitioned between islands (two-way PERMANOVA, $F(df = 8) = 45.38$, $P = 0.001$, explained variance $R^2 = 41.4$\%).
+Nonetheless, we did find evidence for differences in dewlap coloration between habitat types, and those were mostly island-specific (habitat-by-island interaction term, $F(16) = 4.78$, $P = 0.001$, $R^2 = 8.7$\%), with a significant portion of the variation explained by an habitat effect across all islands, but this effect was relatively small ($F(2) = 4.46$, $P = 0.001$, $R^2 = 1$\%).\\
 
 We subsequently tested for differences in dewlap coloration between habitat-populations within each island, using within-island principal component scores (to maximize the variation captured for each island, see Methods).
-Our within-island random forest classification analyses revealed detectable differences in dewlap coloration on eight out of the nine islands in our sample: Abaco, Bimini, Cayman Brac, Eleuthera, Little Cayman, Long Island, North Andros and South Andros.
+Our within-island random forest classification analyses revealed detectable differences in dewlap coloration on seven out of the nine islands in our sample: Abaco, Bimini, Cayman Brac, Eleuthera, Little Cayman, Long Island, and North Andros.
 The accuracy of random forest classification exceeded random expectation more often than expected by chance for all these islands (Table \ref{tab:randomforests}).
-Accuracy was as high as 73\% for Cayman Brac.
-We obtained similar results using other machine learning approaches such as support vector machines (Table \ref{tab:ksvms}) and linear discriminant analysis (Table \ref{tab:ldas}), except that these methods did not detect significant differences on Eleuthera and North Andros.
+Accuracy was as high as 74.8\% for Cayman Brac.
+We obtained similar results using other machine learning approaches such as support vector machines (Table \ref{tab:ksvms}) and linear discriminant analysis (Table \ref{tab:ldas}).
 We describe in details the specific differences detected on each island in the Appendix, and focus here on the general patterns emerging from our data.\\
@@
 \ref{fig:LittleCayman}G), Long Island (Fig.
-\ref{fig:LongIsland}G) and South Andros (Fig.
-\ref{fig:SouthAndros}G).\\
+\ref{fig:LongIsland}G) and North Andros (Fig.
+\ref{fig:NorthAndros}G).\\
 
@@
 \ref{fig:Abaco_supplement}G), suggesting that differences in dewlap coloration between distant sites may be partly attributable to isolation-by-distance, but this may not necessarily be the case for sites in close proximity.
-We did not find evidence for spatial autocorrelation on other islands than Abaco (Table \ref{tab:autocorrelation}).\\
+We did not find evidence for spatial autocorrelation on other islands than Abaco (although Eleuthera was nearly significant, Table \ref{tab:autocorrelation}).\\
 
-A striking feature of our data was inconsistencies in between-habitat differences among islands, in terms of which habitats differ from which, which dimensions of coloration were involved, and in which direction.
+A striking feature of our data was inconsistency in between-habitat differences among islands, in terms of which habitats differ from which, which dimensions of coloration were involved, and in which direction.
 For example, while on Cayman Brac the random forests could well distinguish between all three habitats (Fig.

### 2022-03-09 ms/final/final/sections/supptables.tex
accepted

\begin{table}[H]	
	\caption{Number of lizards sampled in each habitat on each island.}
	\centering
	\input{"./tables/counts.tex"}
	\label{tab:counts}
\end{table}

\pagebreak

\begin{table}[H]
	\caption{Proportion of variance explained by the first four principal components on each island, as well as in the PCA performed on all islands together (last row).}
	\centering
	\input{"./tables/pcavariances.tex"}
	\label{tab:pcavariances}
\end{table}

\pagebreak

\begin{table}[H]
	\caption{Henze-Zirkler's test of multivariate normality, performed on global principal components (i.e. fitted on data from all islands together) in each habitat and on each island. The number of outlier points detected based on the Mahalanobis distance is reported. $HZ$, test statistic. *, $P < 0.05$; **, $P < 0.01$; ***, $P < 0.001$.}
	\centering
	\input{"./tables/multinorm.tex"}
	\label{tab:multinorm}
\end{table}

\pagebreak

\begin{table}[H]
	\caption{Shapiro-Wilk's test of univariate normality of the standardized residuals from OLS and GLS-ANOVAs performed on each island where significant differences were detected by random forest classification. $W$, test statistic. *, $P_{adj} < 0.05$; **, $P_{adj} < 0.01$; ***, $P_{adj} < 0.001$.}
	\centering
	\input{"./tables/normality.tex"}
	\label{tab:normality}
\end{table}

\pagebreak

\begin{table}[H]
	\caption{Locations of the sampling sites across islands, with mean within-island principal component scores per site.}
	\centering
	\input{"./tables/sites.tex"}
	\label{tab:sites}
\end{table}

\pagebreak

\begin{table}[H]
	\caption{Nonparametric Kruskal-Wallis tests performed on each variable on each island where deviations from normality were detected. *, $P < 0.05$; **, $P < 0.01$; ***, $P < 0.001$.}
	\centering
	\input{"./tables/kruskal.tex"}
	\label{tab:kruskal}
\end{table}

\pagebreak

\begin{table}[H]
	\caption{Support vector machine classification results. Legend is as per Table \ref{tab:randomforests}.}
	\centering
	\input{"./tables/ksvms.tex"}
	\label{tab:ksvms}
\end{table}

\pagebreak

\begin{table}[H]
	\caption{Linear discriminant analysis classification results. Legend is as per Table \ref{tab:randomforests}.}
	\centering
	\input{"./tables/ldas.tex"}
	\label{tab:ldas}
\end{table}

### 2022-03-09 ms/sections/appendix_tracked.tex
accepted

--- 2021-07-26 ms/sections/appendix.tex
@@
 \ref{fig:Abaco}B).
-Consistent with this, our analyses of variance detected significantly lower PC2 scores in mangrove lizards than in the two other habitats (Fig.
+Consistent with this, our analyses of variance detected significantly \sout{lower} \textcolor{olive}{higher} PC2 scores in mangrove lizards than in the two other habitats (Fig.
 \ref{fig:Abaco}E, Table \ref{tab:anova}), representing a higher UV-reflectance relative to red (Fig.
 \ref{fig:Abaco}C).
-Beach scrub lizards also scored lower on PC3 (Fig.
+Beach scrub lizards also scored \sout{lower} \textcolor{olive}{higher} on PC3 (Fig.
 \ref{fig:Abaco}E, Table \ref{tab:anova}), indicating less curvature of the reflectance profile and relatively higher reflectance at intermediate wavelengths (blue-to-yellow) than at the ends of the range (Fig.
@@
 \ref{fig:Bimini}C, E, Table \ref{tab:anova}).
-The random forests also used UV reflectance to discriminate between coppice and mangrove dewlaps (Fig.
+\textcolor{olive}{Coppice dewlaps had significantly higher PC2 scores than mangrove dewlaps (Fig.
+\ref{fig:Bimini}E), suggesting a higher curvature (higher UV and red reflectance than more intermediate wavelengths, Fig.
+\ref{fig:Bimini}C).
+For these two habitats the random forests were most sensitive to UV reflectance (Fig.
+\ref{fig:Bimini}F).
+Beach scrub dewlaps had higher PC3 scores than coppice dewlaps but it was not clear what properties of spectral shape this principal component mapped onto (Fig.
+\ref{fig:Bimini}C).} \sout{The random forests also used UV reflectance to discriminate between coppice and mangrove dewlaps (Fig.
 \ref{fig:Bimini}F), which could reflect the significant difference we detected along PC3 between these two habitats (Fig.
@@
 \ref{fig:Bimini}C, E)), in contrast to coppice lizards which harbored a stronger curvature at intermediate wavelengths (Fig.
-\ref{fig:Bimini}C, E).
-On this island, the beach scrub and coppice habitats were separated by a few hundred meters, making this contrast the smallest geographical scale at which differences in coloration were found in our study (Fig.
+\ref{fig:Bimini}C, E).} On this island, the beach scrub and coppice habitats were separated by a few hundred meters, making this contrast the smallest geographical scale at which differences in coloration were found in our study (Fig.
 \ref{fig:Bimini}G).\\
@@
 \ref{fig:CaymanBrac}G), dewlaps in the beach scrub habitat reflected more red light (as represented by PC2, Fig.
-\ref{fig:CaymanBrac}C, E) and more UV (as represented by PC3, Fig.
+\ref{fig:CaymanBrac}C, E) and more UV (as represented by PC3, \textcolor{olive}{along which coppice dewlaps were intermediate,} Fig.
 \ref{fig:CaymanBrac}C, E) than in the mangrove habitat.
-Coppice lizards were also characterized by a higher UV reflectance than mangrove dewlaps (PC3, Fig.
-\ref{fig:CaymanBrac}C, E, Table \ref{tab:anova}).\\
+\sout{Coppice lizards were also characterized by a higher UV reflectance than mangrove dewlaps (PC3, Fig.
+\ref{fig:CaymanBrac}C, E, Table \ref{tab:anova})}.\\
 
 On Eleuthera, although random forests detected between-habitat differences in dewlap color, other approaches did not (Tables \ref{tab:ldas} and \ref{tab:ksvms}), suggesting that the differences may be small.
-Consistent with this, the only significant univariate difference detected was for PC2 between beach scrub and mangrove lizards, where beach scrub lizards had higher levels of red reflectance and mangrove lizards higher levels of UV reflectance (Fig.
-\ref{fig:Eleuthera}C, E, Table \ref{tab:anova}).\\
+\textcolor{olive}{On Eleuthera, beach scrub and mangrove dewlaps were guessed relatively correctly, but coppice dewlaps were more often mistaken for mangrove dewlaps (Fig.
+\ref{fig:Eleuthera}D).
+Mangrove dewlaps had lower PC2 scores than beach scrub dewlaps (Fig.
+\ref{fig:Eleuthera}E, indicating higher UV relative to red, Fig.
+\ref{fig:Eleuthera}C), and higher PC4 scores than coppice dewlaps (Fig.
+\ref{fig:Eleuthera}E, suggesting more reflectance profiles with more curvature, Fig.
+\ref{fig:Eleuthera}C).
+Random forests did not seem to consistently capture the wavelengths responsible for these differences (Fig.
+\ref{fig:Eleuthera}F).} \sout{Consistent with this, the only significant univariate difference detected was for PC2 between beach scrub and mangrove lizards, where beach scrub lizards had higher levels of red reflectance and mangrove lizards higher levels of UV reflectance (Fig.
+\ref{fig:Eleuthera}C, E, Table \ref{tab:anova}).}\\
 
-Little Cayman was characterized by a better discrimination of mangrove lizards from the rest than between beach scrub and coppice lizards (Fig.
+Little Cayman was characterized by a better discrimination of mangrove lizards from the rest than between beach scrub and coppice lizards \textcolor{olive}{even though all habitats were relatively well discriminated} (Fig.
 \ref{fig:LittleCayman}D).
@@
 \ref{fig:LongIsland}F).
-Beach scrub lizards had more curved reflectance profiles than in either of the two other habitats, with higher levels of UV and red reflectance relative to intermediate wavelengths (PC3, Fig.
+Beach scrub lizards had more curved reflectance profiles than in \textcolor{olive}{the mangrove} \sout{either of the two other habitats}, with higher levels of UV and red reflectance relative to intermediate wavelengths (PC3, Fig.
 \ref{fig:LongIsland}C, E, Table \ref{tab:anova}).
-Beach scrub lizards also differed from mangrove lizards along PC4 (Fig.
+\sout{Beach scrub lizards also differed from mangrove lizards along PC4 (Fig.
 \ref{fig:LongIsland}E), which represented a rather small portion of the variance not already explained by the first three principal components, and is therefore difficult to interpret (Fig.
-\ref{fig:LongIsland}C).
-Coppice lizards were significantly darker than mangrove and beach scrub lizards (PC1, Fig.
+\ref{fig:LongIsland}C).} Coppice lizards were significantly darker than mangrove and beach scrub lizards (PC1, Fig.
 \ref{fig:LongIsland}C, E, Table \ref{tab:anova}).\\
 
-On North Andros, although the random forest classification was significant ($P = 0.0216$, Table \ref{tab:randomforests}) and the average confusion matrix indicated that lizards from beach scrub were particularly well predicted (Fig.
+\textcolor{olive}{On North Andros beach scrub and coppice dewlaps could be discriminated better against each other than with mangrove dewlaps (Fig.
+\ref{fig:NorthAndros}D), with importance profiles supporting UV-reflectance as a predictor of coppice lizards (Fig.
+\ref{fig:NorthAndros}F).
+Coppice lizards had less curved reflectance profiles than beach scrub and mangrove lizards (PC2), and beach scrub dewlaps had the lowest scores on PC4, which was difficult to interpret (Fig.
+\ref{fig:NorthAndros}C, E, Table \ref{tab:anova}).}\\
+
+\sout{On North Andros, although the random forest classification was significant ($P = 0.0216$, Table \ref{tab:randomforests}) and the average confusion matrix indicated that lizards from beach scrub were particularly well predicted (Fig.
 \ref{fig:NorthAndros}D), no significant univariate differences were detected along any of the four PCs (Fig.
@@
 \ref{fig:NorthAndros}B), suggesting that the machines may have used this low within-habitat variance, as opposed to between-habitat differences in means, to correctly classify beach scrub lizards.
-A small sample size on this islands may also have contributed to a lack of power in detecting univariate differences using analyses of variance (Table \ref{tab:counts}).\\
+A small sample size on this islands may also have contributed to a lack of power in detecting univariate differences using analyses of variance (Table \ref{tab:counts}).}\\
 
-On South Andros beach scrub and coppice dewlaps could be discriminated better against each other than with mangrove dewlaps (Fig.
+\sout{On South Andros beach scrub and coppice dewlaps could be discriminated better against each other than with mangrove dewlaps (Fig.
 \ref{fig:SouthAndros}D), with importance profiles supporting UV-reflectance as a predictor of coppice lizards (Fig.
@@
 Beach scrub lizards also differed from mangrove lizards in PC4 (Fig.
-\ref{fig:SouthAndros}C, E, Table \ref{tab:anova}).\\
+\ref{fig:SouthAndros}C, E, Table \ref{tab:anova}).}\\
 
-Classification success was not significantly better than expected by chance on Ragged Island (Table \ref{tab:randomforests}) where nearly no habitat could be differentiated from any other based on reflectance.
+Classification success was not significantly better than expected by chance on Ragged Island \textcolor{olive}{and South Andros} (Table \ref{tab:randomforests}\textcolor{olive}{, \ref{tab:ldas}, \ref{tab:ksvms}}) where nearly no habitat could be differentiated from any other based on reflectance.

### 2022-03-09 ms/sections/discussion_tracked.tex
accepted

--- 2022-03-09 ms/final/final/sections/discussion.tex
@@
 Two main insights follow from our results.
-First, we detected significant differences in dewlap coloration between habitats within seven out of the nine islands investigated (excluding North Andros where the follow-up univariate analyses were not significant), suggesting a putatively high potential for local differentiation of dewlap coloration in \textit{Anolis sagrei}.
+First, we detected significant differences in dewlap coloration between habitats within seven out of the nine islands investigated \sout{(excluding North Andros where the follow-up univariate analyses were not significant)}, suggesting a putatively high potential for local differentiation of dewlap coloration in \textit{Anolis sagrei}.
 Second, we found differences in coloration along different dimensions of color space, and in different directions, on different islands.\\

### 2022-03-09 ms/sections/methods_tracked.tex
accepted

--- 2022-03-09 ms/final/final/sections/methods.tex
@@
 
-We sampled 455 male \textit{A.
+We sampled \sout{466} \textcolor{olive}{455} male \textit{A.
 sagrei} from seven islands in the Bahamas Archipelago -- Abaco, North Andros, South Andros, South Bimini, Eleuthera, Long Island, and Ragged Island -- and two in the Cayman Islands -- Cayman Brac and Little Cayman (Figure \ref{fig:maps}A).
@@
 Our sampling design enabled us to test for differences between habitats at a coarse and fine geographical scale.
-The median distance between two localities within an island was $~8.5$km (Figure \ref{fig:maps}B), and 79.3\% of all pairwise distances within islands were less than $50$km.
+The median distance between two localities within an island was \sout{$\sim 11$} \textcolor{olive}{8.5}km (Figure \ref{fig:maps}B), and \sout{80.3} \textcolor{olive}{79.3}\% of all pairwise distances within islands were less than $50$km.
 Additionally, there are no major barriers to dispersal (such as mountains or grassland) on any of the islands that we sampled.

### 2022-03-09 ms/sections/results_tracked.tex
accepted

--- 2022-03-09 ms/final/final/sections/results.tex
@@
 sagrei} dewlap coloration between populations living in three characteristic habitat types across nine islands that span the West Indian range of the brown anole (beach scrub, primary coppice and mangroves).
-We found that most of the variation in coloration was partitioned between islands (two-way PERMANOVA, $F(df = 8) = 45.38$, $P = 0.001$, explained variance $R^2 = 41.4$\%).
-Nonetheless, we did find evidence for differences in dewlap coloration between habitat types, and those were mostly island-specific (habitat-by-island interaction term, $F(16) = 4.78$, $P = 0.001$, $R^2 = 8.7$\%), with a significant portion of the variation explained by an habitat effect across all islands, but this effect was relatively small ($F(2) = 4.46$, $P = 0.001$, $R^2 = 1$\%).\\
+We found that most of the variation in coloration was partitioned between islands (two-way PERMANOVA, \sout{$F(df = 8) = 43.7$, $P = 0.0001$} \textcolor{olive}{$F(df = 8) = 45.38$, $P = 0.001$}, explained variance \sout{$R^2 = 40.9$} \textcolor{olive}{$R^2 = 41.4$}\%).
+Nonetheless, we did find evidence for differences in dewlap coloration between habitat types, and those were mostly island-specific (habitat-by-island interaction term, \sout{$F(16) = 3.53$, $P = 0.0001$, $R^2 = 6.6$} \textcolor{olive}{$F(16) = 4.78$, $P = 0.001$, $R^2 = 8.7$}\%), with a significant portion of the variation explained by an habitat effect across all islands, but this effect was relatively small (\sout{$F(2) = 4.7$, $P = 0.0001$, $R^2 = 1.1$} \textcolor{olive}{$F(2) = 4.46$, $P = 0.001$, $R^2 = 1$}\%).\\
 
 We subsequently tested for differences in dewlap coloration between habitat-populations within each island, using within-island principal component scores (to maximize the variation captured for each island, see Methods).
-Our within-island random forest classification analyses revealed detectable differences in dewlap coloration on seven out of the nine islands in our sample: Abaco, Bimini, Cayman Brac, Eleuthera, Little Cayman, Long Island, and North Andros.
+Our within-island random forest classification analyses revealed detectable differences in dewlap coloration on \sout{eight} \textcolor{olive}{seven} out of the nine islands in our sample: Abaco, Bimini, Cayman Brac, Eleuthera, Little Cayman, Long Island, \textcolor{olive}{and} North Andros \sout{and South Andros}.
 The accuracy of random forest classification exceeded random expectation more often than expected by chance for all these islands (Table \ref{tab:randomforests}).
-Accuracy was as high as 74.8\% for Cayman Brac.
-We obtained similar results using other machine learning approaches such as support vector machines (Table \ref{tab:ksvms}) and linear discriminant analysis (Table \ref{tab:ldas}).
+Accuracy was as high as \sout{73} \textcolor{olive}{74.8}\% for Cayman Brac.
+We obtained similar results using other machine learning approaches such as support vector machines (Table \ref{tab:ksvms}) and linear discriminant analysis (Table \ref{tab:ldas})\sout{, except that these methods did not detect significant differences on Eleuthera and North Andros}.
 We describe in details the specific differences detected on each island in the Appendix, and focus here on the general patterns emerging from our data.\\
@@
 \ref{fig:LittleCayman}G), Long Island (Fig.
-\ref{fig:LongIsland}G) and North Andros (Fig.
+\ref{fig:LongIsland}G) and \sout{South} \textcolor{olive}{North} Andros (Fig.
 \ref{fig:NorthAndros}G).\\

### 2022-03-09 ms/sections/suppfigures_tracked.tex
accepted

--- 2021-05-29 ms/sections/suppfigures.tex
@@
 	\centering
-	\includegraphics[width=\textwidth]{figures/island_maps_white_2.0.png}
-	\caption{Maps of the islands. (A) Map of the West Indies with sampled islands highlighted in black. (B) Sampling sites within islands colored after their respective habitat-types.}
+	\includegraphics[width=\textwidth]{figures/island_maps.png}
+	\caption{Maps of the islands. (A) Map of the West Indies with sampled islands highlighted in black. (B) Sampling sites within islands colored after their respective habitat types.}
 	\label{fig:maps}
@@
 
+\pagebreak
+
 \begin{figure}[H]
 	\centering
-	\includegraphics[width=\textwidth]{figures/Abaco_supplement_1.0b.png}
+	\includegraphics[width=\textwidth]{figures/Abaco_supplement.png}
 	\caption{Comparison of dewlap coloration across habitats on Abaco, with extended results.
@@
 \end{figure}
+
+\pagebreak
 
@@
 
+\pagebreak
+
 \begin{figure}[H]
@@
 	\includegraphics[width=\textwidth]{figures/NorthAndros_supplement.png}
-	\caption{Comparison of dewlap coloration across habitats on North Andros. Legend is as per Figure \ref{fig:Abaco_supplement}, but without panel G.}
+	\caption{Comparison of dewlap coloration across habitats on North Andros. Legend is as per Figure \ref{fig:Abaco_supplement}\sout{, but without panel G}.}
 	\label{fig:NorthAndros}
 \end{figure}
+
+\pagebreak
 
@@
 	\includegraphics[width=\textwidth]{figures/SouthAndros_supplement.png}
-	\caption{Comparison of dewlap coloration across habitats on South Andros. Legend is as per Figure \ref{fig:Abaco_supplement}.}
+	\caption{Comparison of dewlap coloration across habitats on South Andros. Legend is as per Figure \ref{fig:Abaco_supplement}\textcolor{olive}{, but without panel G}.}
 	\label{fig:SouthAndros}
 \end{figure}
+
+\pagebreak
 

### 2022-03-09 scripts/01_process_data.R

@@ -1,0 +1,365 @@
+## ---------------------------
+##
+## Script name: 01_process_data.R
+##
+## Purpose of script: This script extracts, curates, processes and assemble
+## the raw data (reflectance profiles and associated metadata) into a dataset
+## ready for analysis with statistical tools. Among other things, some outliers
+## are removed and the reflectance curves are smoothed using a procedure that
+## removes some very localized artifacts we have identified, without denaturing
+## the signal. Plots are also made to show how processing affects the reflectance
+## curves.
+##
+## Author: Raphael Scherrer
+##
+## Date Created: 2021-02-06
+##
+## This script comes with no guarantee whatsoever.
+##
+## Copyright (c) Raphael Scherrer, 2022
+##
+## Find me on GitHub at https://github.com/rscherrer
+##
+## Email:
+## r.scherrer@rug.nl
+## raphael.scherrer@evobio.eu
+## raph.rjfs@hotmail.fr
+##
+## ---------------------------
+
+rm(list = ls())
+
+library(tidyverse)
+library(pavo)
+library(patchwork)
+
+# Read metadata
+meta <- read_csv("data/raw/master_database.csv")
+
+# Read spectral data
+specs <- map(list.dirs("data/raw")[-1], getspec)
+
+# Or load already saved spectral data to save time
+# specs <- readRDS("data/specs.rds")
+
+# Read specimen tags for each spectrum
+tags <- map(specs, ~ colnames(.x)[-1])
+
+# Remove all double underscores in tags
+fieldtags <- map(tags, str_replace, "__", "_")
+
+# Extract their field tags (slightly different for each dataset)
+fieldtags <- list(
+  str_remove(fieldtags[[1]], "_Refl.*$"),
+  str_remove(str_remove(fieldtags[[2]], "_Refl.*$"), "CB_"),
+  str_remove(str_remove(fieldtags[[3]], "_Refl.*$"), "LC_"),
+  str_remove(fieldtags[[4]], "_Refl.*$"),
+  str_remove(fieldtags[[5]], "_Refl.*$"),
+  str_remove(fieldtags[[6]], "_Refl.*$"),
[…]

### 2022-03-09 scripts/01_run_analyses.R

@@ -1,2054 +1,0 @@
-## ---------------------------
-##
-## Script name: analyses.R
-##
-## Purpose of script: This script contains all the analyses of this study.
-##
-## Author: Raphael Scherrer
-##
-## Date Created: 2021-02-06
-##
-## Copyright (c) Raphael Scherrer, 2021
-##
-## Email:
-## r.scherrer@rug.nl
-## raphael.scherrer@evobio.eu
-## raph.rjfs@hotmail.fr
-##
-## ---------------------------
-
-# Notation used for naming datasets in this script:
-# D = the original dataset
-# X = placeholder within a function
-# P = temporary data frame for plotting or containing plots
-
-rm(list = ls())
-
-library(tidyverse)
-library(ggridges) # ridge plots
-library(broom) # tidy statistical tables
-library(heplots) # Box M test
-library(MVN) # multivariate normality
-library(assertthat) # assertions within functions
-library(rminer) # machine learning
-library(geosphere) # geographical distances
-library(MuMIn) # AICc
-library(nlme) # univariate ANOVAs
-library(PMCMRplus) # posthoc tests
-library(RColorBrewer) # color palette
-library(pairwiseAdonis) # posthoc PERMANOVA
-library(scales) # color palette
-library(vegan) # PERMANOVA
-library(ape) # Mantel test
-
-#### 1. Set-up ####
-
-# Set plotting theme
-theme_set(theme_classic())
-
-# Graphical parameters
-island_colors <- brewer.pal(8, "Set2")
-island_colors <- c(island_colors, "darkgreen")
-habitat_colors <- c("goldenrod", "forestgreen", "mediumseagreen")
-
-# Global settings
-pc_names <- paste0("PC", 1:4)
-wpc_names <- paste0("w", pc_names)
-wl <- 300:700
-wl_names <- paste0("wl", wl)
-
-# Whether to load machines already fitted to whole spectrum data
-# (instead of rerunning the very lengthy classification)
-read_fitted <- TRUE
-
-#### 2. Read reflectance data ####
-
-# Read the data
-D <- read_csv("data/reflectance.csv")
-
-# Useful things to keep in mind
-island_names <- unique(D$island)
-habitat_names <- unique(D$habitat)
-
-# Save sample sizes
-with(D, table(island, habitat)) %>%
[…]

### 2022-03-09 scripts/02_convert_tables.R

@@ -1,268 +1,0 @@
-## ---------------------------
-##
-## Script name: tables.R
-##
-## Purpose of script:
-##
-## This script loads relevant tables from the results and
-## turns them into nice-looking LaTeX tables
-##
-## Author: Raphael Scherrer
-##
-## Date Created: 2021-04-05
-##
-## Copyright (c) Raphael Scherrer, 2021
-##
-## Email:
-## r.scherrer@rug.nl
-## raphael.scherrer@evobio.eu
-## raph.rjfs@hotmail.fr
-##
-## ---------------------------
-
-rm(list = ls())
-
-library(tidyverse)
-library(knitr)
-library(kableExtra)
-
-#### 0. Accessory functions ####
-
-# Function to add a significance column with asterisk symbols to a tibble
-mutate_signif <- function(
-  D, col = "pvalue", levels = c(0.05, 0.01, 0.001), dropname = TRUE
-) {
-
-  D <- D %>%
-    mutate(
-      .signif = ifelse(get(col) < levels[1], "*", ""),
-      .signif = ifelse(get(col) < levels[2], "**", .signif),
-      .signif = ifelse(get(col) < levels[3], "***", .signif)
-    )
-
-  if (dropname) D <- D %>% rename(" " = ".signif")
-
-  return(D)
-
-}
-
-# Round and format P-values
-round_pvalue <- function(pvalue, digits = 4) {
-
-  threshold <- 1 / 10^digits
-  template <- paste0("%.", sprintf("%sf", digits))
-  label <- sprintf(paste("<", template), threshold, digits)
-  rounded <- sprintf(template, round(pvalue, digits), digits)
-  ifelse(pvalue < threshold, label, rounded)
-
-}
-
-#### 1. Random forest table ####
-
-# Random forest results
-read_csv("results/group_comparisons/machine_learning/PC/randomforest/summary.csv") %>%
-  mutate_signif() %>%
-  mutate(pvalue = round_pvalue(pvalue, digits = 4)) %>%
-  kable(
-    "latex",
-    col.names = c("Island", "$N$", "Score", "$P$", ""),
-    digits = c(0, 0, 3, 0, 0),
-    align = "lrrrl",
-    booktabs = TRUE,
-    linesep = "",
[…]

### 2022-03-09 scripts/02_run_analyses.R

@@ -1,0 +1,2068 @@
+## ---------------------------
+##
+## Script name: 02_run_analyses.R
+##
+## Purpose of script: This script contains all the analyses of this study.
+## Make sure to choose read_fitted <- TRUE if you do not want to repeat the
+## whole machine learning fitting: as random forests fitted to reflectance data ---
+## not PCs --- take several hours, we saved the output of the machines in a
+## file called fitted.rds. Note that nevertheless all analyses involving
+## random numbers can be repeated from scratch because we set specific seeds
+## every time.
+##
+## Author: Raphael Scherrer
+##
+## Date Created: 2021-02-06
+##
+## This script comes with no guarantee whatsoever.
+##
+## Copyright (c) Raphael Scherrer, 2022
+##
+## Find me on GitHub at https://github.com/rscherrer
+##
+## Email:
+## r.scherrer@rug.nl
+## raphael.scherrer@evobio.eu
+## raph.rjfs@hotmail.fr
+##
+## ---------------------------
+
+# Notation used for naming datasets in this script:
+# D = the original dataset
+# X = placeholder within a function
+# P = temporary data frame for plotting or containing plots
+
+rm(list = ls())
+
+library(tidyverse)
+library(ggridges) # ridge plots
+library(broom) # tidy statistical tables
+library(heplots) # Box M test
+library(MVN) # multivariate normality
+library(assertthat) # assertions within functions
+library(rminer) # machine learning
+library(geosphere) # geographical distances
+library(MuMIn) # AICc
+library(nlme) # univariate ANOVAs
+library(PMCMRplus) # posthoc tests
+library(RColorBrewer) # color palette
+library(pairwiseAdonis) # posthoc PERMANOVA
+library(scales) # color palette
+library(vegan) # PERMANOVA
+library(ape) # Mantel test
+
+# Function to average several matrices
+mavg <- function(x) Reduce('+', x) / length(x)
+
+#### 1. Set-up ####
+
+# Set plotting theme
+theme_set(theme_classic())
+
+# Graphical parameters
+island_colors <- brewer.pal(8, "Set2")
+island_colors <- c(island_colors, "darkgreen")
[…]

### 2022-03-09 scripts/03_convert_tables.R

@@ -1,0 +1,272 @@
+## ---------------------------
+##
+## Script name: 03_convert_tables.R
+##
+## Purpose of script: This script loads relevant tables from the results and
+## turns them into nice-looking LaTeX tables that can directly be loaded into
+## the manuscript.
+##
+## Author: Raphael Scherrer
+##
+## Date Created: 2021-04-05
+##
+## This script comes with no guarantee whatsoever.
+##
+## Copyright (c) Raphael Scherrer, 2022
+##
+## Find me on GitHub at https://github.com/rscherrer
+##
+## Email:
+## r.scherrer@rug.nl
+## raphael.scherrer@evobio.eu
+## raph.rjfs@hotmail.fr
+##
+## ---------------------------
+
+rm(list = ls())
+
+library(tidyverse)
+library(knitr)
+library(kableExtra)
+
+#### 0. Accessory functions ####
+
+# Function to add a significance column with asterisk symbols to a tibble
+mutate_signif <- function(
+  D, col = "pvalue", levels = c(0.05, 0.01, 0.001), dropname = TRUE
+) {
+
+  D <- D %>%
+    mutate(
+      .signif = ifelse(get(col) < levels[1], "*", ""),
+      .signif = ifelse(get(col) < levels[2], "**", .signif),
+      .signif = ifelse(get(col) < levels[3], "***", .signif)
+    )
+
+  if (dropname) D <- D %>% rename(" " = ".signif")
+
+  return(D)
+
+}
+
+# Round and format P-values
+round_pvalue <- function(pvalue, digits = 4) {
+
+  threshold <- 1 / 10^digits
+  template <- paste0("%.", sprintf("%sf", digits))
+  label <- sprintf(paste("<", template), threshold, digits)
+  rounded <- sprintf(template, round(pvalue, digits), digits)
+  ifelse(pvalue < threshold, label, rounded)
+
+}
+
+#### 1. Random forest table ####
+
+# Random forest results
+read_csv("results/group_comparisons/machine_learning/PC/randomforest/summary.csv") %>%
+  mutate_signif() %>%
+  mutate(pvalue = round_pvalue(pvalue, digits = 4)) %>%
+  kable(
+    "latex",
[…]

### 2022-03-09 scripts/04_median_distance.R

@@ -1,0 +1,47 @@
+## ---------------------------
+##
+## Script name: 04_median_distance.R
+##
+## Purpose of script: This script shows the median within-island distance
+## between two patches, computed across all islands.
+##
+## Author: Raphael Scherrer
+##
+## Date Created: 2022-03-09
+##
+## This script comes with no guarantee whatsoever.
+##
+## Copyright (c) Raphael Scherrer, 2022
+##
+## Find me on GitHub at https://github.com/rscherrer
+##
+## Email:
+## r.scherrer@rug.nl
+## raphael.scherrer@evobio.eu
+## raph.rjfs@hotmail.fr
+##
+## ---------------------------
+
+rm(list = ls())
+
+library(tidyverse)
+library(geosphere)
+
+data <- read_csv("metadata/sites.csv")
+
+data %>%
+  group_by(island) %>%
+  nest() %>%
+  mutate(dist = map(data, function(data) {
+
+    y <- with(data, distm(matrix(c(longitude, latitude), ncol = 2), fun = distGeo))
+    y[upper.tri(y)]
+
+  })) %>%
+  select(-data) %>%
+  unnest(dist) %>%
+  ungroup() %>%
+  select(dist) %>%
+  summarize(median = median(dist))
+
+

### 2022-03-09 ms/run_pandoc.sh
ms readme'

@@ -1,5 +1,0 @@
-# Run this script to convert LaTeX files into ODT and DOCX documents
-# Requires pandoc and pandoc-crossref being installed
-
-pandoc main.tex --filter=pandoc-crossref --bibliography=library.bib --csl=journal-of-evolutionary-biology.csl -o scherrer2021.odt
-pandoc main.tex --filter=pandoc-crossref --bibliography=library.bib --csl=journal-of-evolutionary-biology.csl -o scherrer2021.docx

### 2022-03-09 scripts/01_process_data.R
rm copyright claims

@@ -16,8 +16,6 @@
 ##
 ## This script comes with no guarantee whatsoever.
 ##
-## Copyright (c) Raphael Scherrer, 2022
-##
 ## Find me on GitHub at https://github.com/rscherrer
 ##
 ## Email:

### 2022-03-09 scripts/03_convert_tables.R

@@ -12,8 +12,6 @@
 ##
 ## This script comes with no guarantee whatsoever.
 ##
-## Copyright (c) Raphael Scherrer, 2022
-##
 ## Find me on GitHub at https://github.com/rscherrer
 ##
 ## Email:

### 2022-03-09 scripts/04_median_distance.R

@@ -11,8 +11,6 @@
 ##
 ## This script comes with no guarantee whatsoever.
 ##
-## Copyright (c) Raphael Scherrer, 2022
-##
 ## Find me on GitHub at https://github.com/rscherrer
 ##
 ## Email:

### 2022-03-10 scripts/01_process_data.R
better doc for raw data

@@ -2,13 +2,13 @@
 ##
 ## Script name: 01_process_data.R
 ##
-## Purpose of script: This script extracts, curates, processes and assemble
+## Purpose of script: This script extracts, curates, processes and assembles
 ## the raw data (reflectance profiles and associated metadata) into a dataset
 ## ready for analysis with statistical tools. Among other things, some outliers
 ## are removed and the reflectance curves are smoothed using a procedure that
 ## removes some very localized artifacts we have identified, without denaturing
 ## the signal. Plots are also made to show how processing affects the reflectance
-## curves.
+## curves. Follow the comments for details about each step.
 ##
 ## Author: Raphael Scherrer
 ##
@@ -32,7 +32,7 @@ library(pavo)
 library(patchwork)
 
 # Read metadata
-meta <- read_csv("data/raw/master_database.csv")
+meta <- read_csv("data/master_database.csv")
 
 # Read spectral data
 specs <- map(list.dirs("data/raw")[-1], getspec)

### 2022-03-10 scripts/02_run_analyses.R

@@ -8,7 +8,7 @@
 ## not PCs --- take several hours, we saved the output of the machines in a
 ## file called fitted.rds. Note that nevertheless all analyses involving
 ## random numbers can be repeated from scratch because we set specific seeds
-## every time.
+## every time. Follow the comments for details about each analysis.
 ##
 ## Author: Raphael Scherrer
 ##

### 2022-03-10 scripts/04_median_distance.R

@@ -25,7 +25,7 @@ rm(list = ls())
 library(tidyverse)
 library(geosphere)
 
-data <- read_csv("metadata/sites.csv")
+data <- read_csv("data/sites.csv")
 
 data %>%
   group_by(island) %>%

### 2022-03-10 scripts/01_process_data.R
done with readmes?

@@ -37,9 +37,6 @@ meta <- read_csv("data/master_database.csv")
 # Read spectral data
 specs <- map(list.dirs("data/raw")[-1], getspec)
 
-# Or load already saved spectral data to save time
-# specs <- readRDS("data/specs.rds")
-
 # Read specimen tags for each spectrum
 tags <- map(specs, ~ colnames(.x)[-1])
 
